### 2019-01-03 manuscript/manuscript.tex
committing template for manuscript

\documentclass[11pt]{article} % For LaTeX2e
\usepackage{manuscript,palatino}
\usepackage{graphicx}
\usepackage{amsfonts,amsmath}
\usepackage{algorithm, algpseudocode}%

\title{something something about anxiety}

\author{
Samuel Zorowitz \\
Princeton Neuroscience Institute\\
Princeton University\\
Princeton, NJ 08540 \\
\texttt{zorowitz@princeton.edu} \\
\And
Nathaniel Daw \\
Princeton Neuroscience Institute\\
Princeton University\\
Princeton, NJ 08540 \\
\texttt{ndaw@princeton.edu} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\begin{document}

\maketitle

\begin{abstract}
To be filled in
\end{abstract}

\keywords{
reinforcement learning; avoidance; anxiety
}

\startmain % to start the main 1-4 pages of the submission.

\section{Introduction}

\section{Methods}

\subsection{Markov Decision Processes}

Here, we briefly review the formalism of Markov decision processes (MDPs), which
provide the foundation for our results. For more complete treatments of the
subject, see Sutton \& Barto (1998, 2018) and Bertsekas \& Tsitsiklis (1996).

A MDP is defined by a set of states, $S$, a set of actions, $A$, a reward
function, $R(s,a)$ over state/action pairs, and a state transition distribution,
$P(s'|s,a)$, where $a$ denotes a chosen action. States and rewards occur
sequentially according to these one-step functions, driven by a series of
actions; the goal is to learn to choose a probabilistic policy over actions,
denoted by $\pi$, that maximizes the value function, $V^\pi(s)$, defined as the
expected cumulative discounted reward:

$$ V^\pi(s) = \mathbb{E} \left[ \sum^{\infty}_{k=0} \gamma^k R_{t+k} | S_t = s \right] $$

Here, $\gamma$ is a parameter controlling temporal discounting. The value function
can also be defined recursively as the sum of the immediate reward of the action
chosen in that state, $R(s, a)$, and the value of its successor state $s’$,
averaged over possible actions, $a$, and transitions that would occur if the agent
chose according to $\pi$:

$$ V^\pi(s) = \sum_a \pi(a|s) \left[ R(s,a) + \sum_{s'} P(s'|s,a) \gamma V^\pi (S') \right] $$

The value function under the optimal policy is given by:

$$ V^*(s) = \max_a \mathbb{E} \left[ \sum^{\infty}_{k=0} \gamma^k R_{t+k} | S_t = s \right] $$

$$ = \max_a \left[ R(s,a) + \sum_{s'} P(s'|s,a) \gamma V^\pi (S') \right] $$

Knowledge of the value function can help to guide choices. For instance, we can
define the state-action value function as the value of choosing action $a$ and
following $\pi$ thereafter:

$$ Q^\pi(s,a) = \mathbb{E} \left[ \sum^{\infty}_{k=0} \gamma^k R_{t+k} | S_t = s \right] $$

$$ = R(s,a) + \sum_{s'} P(s'|s,a) \gamma V^\pi (S') $$

Then at any state one could choose the action that maximizes $Q^\pi(s,a)$. Formally
this defines a new policy, which is as good or better than the baseline policy
$\pi$; analogously, Eq 2 can be used to define the optimal state-action value
function, the maximization of which selects optimal actions. Note that it is
possible to write a recursive definition for $Q$ in the same manner as Eq 1, and
work directly with the state-action values, rather than deriving them indirectly
from $V$.

\subsection{Robust Control}

To create the symptoms of anxiety, we need a loss of control. This can come in
two forms. First is through the state transition distribution, $P(s'|s,a)$.
Normally, we assume the world is deterministic; in other words, an agent's chosen
action, $a$ leads the agent to the desired successor state, $s'$. In a stochastic
environment, however, an agent may instead transition to an undesired state, $s^*$.

Alternately, an agent may choose optimally in the present but may be uncertain of
its choice control in the future. In such a case, we deviate from the optimal
policy in the future. We can adopt a different function. There are a number to
choose from (Garcia \& Fernandez, 2015). Here we adopt the beta-pessimism criterion
(Gaskett, 2003):

$$ V(s') = \beta \max_{a'} Q(s',a') + (1 - \beta) \min_{a'} Q(s',a') $$

where $\beta$ controls the degree of pessimism. When $\beta = 1$, an agent expects
complete control to choose the best action (best-case criterion); when $\beta = 0$,
an agent expects the  worst future case; and when $\beta = 0.5$, an agent expects
randomness. (We are not committed to this particular instantiation; it is simply
a convenient notation. Other parameterizations work equally well, see appendix.)

\subsection{Simulations}

In some instances, we solve for Q-values and state values directly through value
iteration. This is a form of Dynamic Programming. In other instances, we estimate
these terms through temporal difference learning. We use beta-pessimism (betamax)
learning for both. We use the former to solve for final (infinite horizon) values
and the latter to identify finite horizon values. We explain when we use both.

\section{Results}

\subsection{Pessimistic Inference}

\subsection{Avoidance}

\subsection{Approach-Avoidance Conflict}

\subsection{Planning}

\subsection{Free Choice Premium}

\subsection{Longitudinal Progression}

\section{Appendix}
\subsection{Algorithms}

\begin{algorithm}
  \caption{Value Iteration}

  \State Algorithm parameter: a small threshold $\theta > 0$ determining accuracy of estimation
  \State Initialize $V(s)$, for all $s \in S$ arbitrarily, except that $V(terminal) = 0$
  \State
  \While{$\Delta > \theta$}
    \State $\Delta \leftarrow 0$
    \Loop \ for each $s \in S$
      \State $v \leftarrow V(s)$
      \State $ V(s) = \max_a \sum_{s',r} p(s',r|s,a) \left[ r + \gamma V(s') \right] $
      \State $\Delta \leftarrow \max(\Delta, |v - V(s)|)$
    \EndLoop
  \EndWhile

\end{algorithm}

\begin{algorithm}
  \caption{Betamax Temporal Difference Learning}

  \State Algorithm parameters: step size $\eta \in (0, 1)$, small $\epsilon > 0$
  \State Initialize $Q(s,a)$, for all $s \in S$, $a \in A$ arbitrarily, except that $Q(terminal,\cdot) = 0$
  \State
  \Loop \ for each episode
    \State Initialize $S$
    \While{$S \notin S(terminal)$}
      \State Choose $A$ from $S$ using policy derived from $Q$
      \State Take action $A$, observe $R, S$
      \State $Q(S, A) \leftarrow Q(S, A) + \eta \left[ R + \gamma \max_a Q(S',a) − Q(S, A) \right] $
      \State $S \leftarrow S'$
    \EndWhile
  \EndLoop

\end{algorithm}

\end{document}

### 2019-01-03 manuscript/manuscript.tex
progress on open field demonstration/text

@@
 
-\title{something something about anxiety}
+\title{Anxiety: a decision-theoretic perspective}
 
@@
 
+First and foremost, pathological anxiety is associated with pessimistic expectations
+about the probability and cost of future negative events (Grupe \& Nitschke, 2013;
+Bishop \& Gagne, 2018). In fear conditioning, individuals with anxiety are more
+likely to exhibit increased threat appraisal for safety signals (CS-) and the
+extinguished fear cue (CS+) (Lissek et al. 2005; Duits et al., 2015). In considering
+hypothetical future negative outcomes, individuals with anxiety are more likely
+to evaluate negative outcomes as more likely and their consequences more severe
+(pick a few references).
+
+Importantly, in the clinical literature, pathological anxiety is also associated
+with a lack of perceived control (Gallagher et al., 2014). Studies suggest that
+this belief mediates anxiety (ref?). This can come in whatever form.
+
+In the first series of simulations, we show directly how lack of perceived control
+can directly account for pessimistic inference. In the open field environment
+(Figure 1a), an agent navigates a simple grid-world. The world consists of almost
+entirely of empty tiles, except for two which yield a reward and punishment,
+respectively.
+
+Under complete control (objective or perceived), an agent in this environment is
+under no threat despite the presence of an aversive state. This is simply because,
+in a full controllable world, a (non-masochistic) agent is never forced to interact
+with the aversive state. When this control breaks down -- through a stochastic
+world or a distrust of future competency -- an agent can no longer guarantee it
+will never encounter the aversive state. In such a scenario, the negative value
+back-propagates to its predecessor states through the recursive formulas defined
+above. The result is an environment that appears worse (i.e. less positive) than
+it otherwise would.
+
+We shoe this above. Using value iteration. Show it for best-case scenario. Show
+it for expected random. Show it for worst-case scenario.
+
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/01_field.png}}%
+  }
+  \caption{\textbf{Open Field Environment}}
+  \par Oooh so pretty.
+\end{figure}
+
 \subsection{Avoidance}

### 2019-01-04 manuscript/template.tex
updated/fixed cliff-walking figure.

\documentclass[11pt]{article} % For LaTeX2e
\usepackage{manuscript,palatino}
\usepackage{graphicx}

\title{something something about anxiety}

\author{
Samuel Zorowitz \\
Princeton Neuroscience Institute\\
Princeton University\\
Princeton, NJ 08540 \\
\texttt{zorowitz@princeton.edu} \\
\And
Nathaniel Daw \\
Princeton Neuroscience Institute\\
Princeton University\\
Princeton, NJ 08540 \\
\texttt{ndaw@princeton.edu} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\begin{document}

\maketitle

\begin{abstract}
To be filled in
\end{abstract}

\keywords{
reinforcement learning; avoidance; anxiety
}

\startmain % to start the main 1-4 pages of the submission.

\section{Submission of papers to RLDM}

RLDM requires electronic submissions.  This year's electronic
submission site is
\begin{center}
   https://cmt3.research.microsoft.com/RLDM2017/
\end{center}

Please read the instructions below, and follow them faithfully. Note
that there is also a template \verb+rldm.rtf+ for Microsoft Word,
which is available from the website below.
\subsection{Style}

Papers to be submitted to RLDM must be prepared according to the
instructions presented here. Papers consist of a \emph{title}, which
has a maximum of 100 characters, an \emph{abstract}, which is a
maximum of 2000 characters, up to five key words, and an
\emph{extended abstract}, which starts on the second page, and can be
between one and four pages. Figures and references should be included
in the latter.

Authors preferring \LaTeX{} are requested to use the RLDM \LaTeX{}
style files obtainable at the RLDM website at
\begin{center}
   http://www.rldm.org/
\end{center}
The file \verb+rldm.pdf+ contains these instructions and illustrates
the various formatting requirements your RLDM paper must
satisfy. There is a \LaTeX{} style file called \verb+rldmsubmit.sty+,
and a \LaTeX{} file \verb+rldm.tex+, which may be used as a ``shell''
for writing your paper. All you have to do is replace the author,
title, abstract, keywords, acknowledgements and text of the paper with
your own. The file
\verb+rldm.rtf+ is provided as an equivalent shell for Microsoft Word users.

\section{General formatting instructions}
\label{gen_inst}

The paper size for RLDM is ``US Letter'' (rather than ``A4''). Margins
are 1.5cm around all sides. Use 11~point type with a vertical spacing
of 12~points. Palatino is the preferred typeface throughout.
Paragraphs are separated by 1/2~line space, with no indentation.

Paper title is 17~point, initial caps/lower case, bold, centered between
2~horizontal rules. Top rule is 4~points thick and bottom rule is 1~point
thick. Allow 0.6cm space above and below title to rules.

The lead author's name is to be listed first (left-most), and
the co-authors' names (if different address) are set to follow. If
there is only one co-author, list both author and co-author side by side.

\section{Preparing PostScript or PDF files}

Please prepare PostScript or PDF files with paper size ``US Letter''.
The -t letter option on dvips will produce US Letter files.

\end{document}

### 2019-01-04 sisyphus/__init__.py

@@ -7,3 +7,4 @@ from .envs import (CliffWalking, FlightInitiationDistance, FreeChoice,
 from .dp import (ValueIteration)
 from .td import (ModelFree)
 from .misc import (softmax,betamax)
+from .viz import (plot_policy, plot_cliff)

### 2019-01-04 sisyphus/viz.py

@@ -1,52 +1,94 @@
 import numpy as np
 import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
 import seaborn as sns
-from mpl_toolkits.axes_grid1 import make_axes_locatable
-from .agents._routines import softmax
+from matplotlib.colors import ListedColormap
 
-def CanvasPolicy():
+def plot_policy(ax, gym, pi, color='w', head_width=0.25, head_length=0.25):
+    """Plot agent policy on grid world.
     
-    ## Initialize canvas.
-    fig = plt.figure(figsize=(21,5))
-
-    ## Define boundaries of environment heatmap.
-    gs = gridspec.GridSpec(1, 1)
-    gs.update(left=0.05, right=0.24, hspace=0)
-    axes = [plt.subplot(gs[0])]
-
-    ## Define boundaries of policy heatmaps.
-    gs = gridspec.GridSpec(1, 3)
-    gs.update(left=0.25, right=0.95, hspace=0)
-    axes += [plt.subplot(gs[i]) for i in range(3)]
-    
-    return fig, axes
-
-def plot_policy(env, Q, visits=None, beta=None, n_steps=100, color='0.6', cmap=None, 
-                center=None, vmin=None, vmax=None, cbar=True, ax=None, cbar_ax=None):
+    Parameters
+    ----------
+    ax : matplotlib Axes
+        Axes in which to draw the plot.
+    gym : GridWorld instance
+        Simulation environment.
+    py : array
+        Agent policy, i.e. ordered visitation of states.
+    color : str (default = white)
+        Color of arrow.
+    head_width : float (default=0.25)
+        Width of the arrow head.
+    head_length : float (default=0.25)
+        Length of the arrow head.
         
-    ## Prepare Q-values for visualization.
-    if beta is None: V = Q.max(axis=1)
-    else: V = np.sum(np.apply_along_axis(softmax, 1, Q) * Q, axis=1)
-    V[env.terminal] = np.nan
-    V = V.reshape(env.shape)
-    
-    ## Define boundaries of colobar.
-    if ax is not None and cbar_ax is None:
-        divider = make_axes_locatable(ax)
-        cbar_ax = divider.append_axes("right", size="5%", pad=0.01)
-    
-    ## Plot heatmap.
-    ax = sns.heatmap(V, cmap=cmap, center=center, vmin=vmin, vmax=vmax, cbar=cbar,
[…]

### 2019-01-05 sisyphus/__init__.py
reorganized module. updated visualizations. decision tree plots much better.

@@ -2,9 +2,6 @@
 
 __version__ = '0.1'
 
-from .envs import (CliffWalking, FlightInitiationDistance, FreeChoice, 
-                   OpenField, DecisionTree)
-from .dp import (ValueIteration)
-from .td import (ModelFree)
-from .misc import (softmax,betamax)
-from .viz import (plot_policy, plot_cliff)
+from . import envs
+from . import mdp
+from . import viz

### 2019-01-05 sisyphus/envs/__init__.py

@@ -1,7 +1,7 @@
-"""MDP environments code"""
+"""MDP environments submodule"""
 
-from .choice import FreeChoice
-from .cliff import CliffWalking
-from .fid import FlightInitiationDistance
-from .field import OpenField
-from .tree import DecisionTree
+from ._choice import FreeChoice
+from ._cliff import CliffWalking
+from ._fid import FlightInitiationDistance
+from ._field import OpenField
+from ._tree import DecisionTree

### 2019-01-05 sisyphus/envs/_choice.py

@@ -1,6 +1,6 @@
 import numpy as np
 from scipy.stats import norm
-from .base import GraphWorld
+from ._base import GraphWorld
 
 class FreeChoice(GraphWorld):
     """Instrumental variant of the Free Choice task.

### 2019-01-05 sisyphus/envs/_cliff.py

@@ -1,5 +1,5 @@
 import numpy as np
-from .base import GraphWorld, grid_to_adj
+from ._base import GraphWorld, grid_to_adj
 
 class CliffWalking(GraphWorld):
     """Cliff-walking task environment.

### 2019-01-05 sisyphus/envs/_fid.py

@@ -1,6 +1,6 @@
 import numpy as np
 from scipy.stats import norm
-from .base import GraphWorld
+from ._base import GraphWorld
 
 class FlightInitiationDistance(GraphWorld):
     """Flight initiation distance environment. 

### 2019-01-05 sisyphus/envs/_field.py

@@ -1,5 +1,5 @@
 import numpy as np
-from .base import GraphWorld, grid_to_adj
+from ._base import GraphWorld, grid_to_adj
 
 class OpenField(GraphWorld):
     """Open field task environment."""

### 2019-01-05 sisyphus/envs/_tree.py

@@ -1,6 +1,6 @@
 import numpy as np
 from scipy.stats import norm
-from .base import GraphWorld
+from ._base import GraphWorld
 
 class DecisionTree(GraphWorld):
     """Variant of the decision tree game.

### 2019-01-05 sisyphus/mdp/__init__.py

@@ -1,0 +1,4 @@
+"""Markov decision process algorithms"""
+
+from ._dp import ValueIteration
+from ._td import ModelFree

### 2019-01-05 sisyphus/mdp/_dp.py

@@ -1,7 +1,7 @@
 """Dynamic programming module"""
 
 import numpy as np
-from .misc import check_params, softmax, betamax
+from ._misc import check_params, softmax, betamax
 from warnings import warn
 
 class ValueIteration(object):

### 2019-01-05 sisyphus/mdp/_td.py

@@ -1,7 +1,7 @@
 """Temporal difference module"""
 
 import numpy as np
-from .misc import check_params, softmax, betamax, categorical
+from ._misc import check_params, softmax, betamax, categorical
 
 def greedy_choice(arr, epsilon):
     if np.random.binomial(1,1-epsilon): 

### 2019-01-05 sisyphus/viz/__init__.py

@@ -1,0 +1,5 @@
+"""Visualization functions"""
+
+from ._base import plot_policy
+from ._cliff import plot_cliff
+from ._tree import plot_decision_tree

### 2019-01-05 sisyphus/viz/_base.py

@@ -1,0 +1,43 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.colors import ListedColormap
+
+def plot_policy(ax, gym, pi, color='w', head_width=0.25, head_length=0.25):
+    """Plot agent policy on grid world.
+    
+    Parameters
+    ----------
+    ax : matplotlib Axes
+        Axes in which to draw the plot.
+    gym : GridWorld instance
+        Simulation environment.
+    py : array
+        Agent policy, i.e. ordered visitation of states.
+    color : str (default = white)
+        Color of arrow.
+    head_width : float (default=0.25)
+        Width of the arrow head.
+    head_length : float (default=0.25)
+        Length of the arrow head.
+        
+    Returns
+    -------
+    ax : matplotlib Axes
+        Axes in which to draw the plot.
+    """
+        
+    ## Initialize grid.
+    grid = gym.states.reshape(gym.shape)
+    
+    ## Iteratively plot arrows.
+    for i in range(len(pi)-1):
+        
+        ## Identify S, S' coordinates.
+        y1, x1 = np.where(grid==pi[i])
+        y2, x2 = np.where(grid==pi[i+1])
+
+        ## Plot.
+        ax.arrow(int(x1)+0.5, int(y1)+0.5, 0.5*int(x2-x1), 0.5*int(y2-y1), 
+                 color=color, head_width=head_width, head_length=head_length)
+        
+    return ax

### 2019-01-05 sisyphus/viz/_cliff.py

@@ -3,46 +3,6 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 from matplotlib.colors import ListedColormap
 
-def plot_policy(ax, gym, pi, color='w', head_width=0.25, head_length=0.25):
-    """Plot agent policy on grid world.
-    
-    Parameters
-    ----------
-    ax : matplotlib Axes
-        Axes in which to draw the plot.
-    gym : GridWorld instance
-        Simulation environment.
-    py : array
-        Agent policy, i.e. ordered visitation of states.
-    color : str (default = white)
-        Color of arrow.
-    head_width : float (default=0.25)
-        Width of the arrow head.
-    head_length : float (default=0.25)
-        Length of the arrow head.
-        
-    Returns
-    -------
-    ax : matplotlib Axes
-        Axes in which to draw the plot.
-    """
-        
-    ## Initialize grid.
-    grid = gym.states.reshape(gym.shape)
-    
-    ## Iteratively plot arrows.
-    for i in range(len(pi)-1):
-        
-        ## Identify S, S' coordinates.
-        y1, x1 = np.where(grid==pi[i])
-        y2, x2 = np.where(grid==pi[i+1])
-
-        ## Plot.
-        ax.arrow(int(x1)+0.5, int(y1)+0.5, 0.5*int(x2-x1), 0.5*int(y2-y1), 
-                 color=color, head_width=head_width, head_length=head_length)
-        
-    return ax
-
 def plot_cliff(ax, cbar=False, viable='0.8', cliff='0.1', start='#1f77b4', end='#2ca02c'):
     """Plot cliff-walking environment.
     

### 2019-01-05 sisyphus/viz/_tree.py

@@ -1,0 +1,116 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+from matplotlib.cm import get_cmap
+from matplotlib.colors import ListedColormap, Normalize
+
+def _draw_nodes(ax, xpos, ypos, s=1000, color=None, cmap=None, vmin=None, vmax=None, 
+                alpha=1.0, linewidth=0):
+    """Draw decision tree nodes. See plot_decision tree for details."""
+    
+    ## Define colors.
+    if color is None: 
+        colors = np.repeat('#1f77b4', len(xpos))
+    elif isinstance(color, str):
+        colors = np.repeat(color, len(xpos))
+    elif np.issubdtype(np.array(color).dtype, np.number):
+        assert np.equal(len(color), len(xpos))        
+        if not isinstance(cmap, ListedColormap): cmap = get_cmap(cmap)
+        colors = cmap(Normalize(vmin, vmax)(np.array(color)))
+    else:
+        assert np.equal(len(color), len(xpos))        
+        colors = np.copy(color)
+        
+    ## Define transparency.
+    if isinstance(alpha, float): 
+        alphas = np.repeat(alpha, len(xpos))
+    else: 
+        assert np.equal(len(alpha), len(xpos))
+        alphas = np.copy(alpha)
+        
+    ## Iteratively plot.
+    for i, (x, y, color, alpha) in enumerate(zip(xpos, ypos, colors, alphas)):
+        ax.scatter(x, y, s=s, color='w', alpha=1)
+        ax.scatter(x, y, s=s, color=color, alpha=alpha, linewidth=linewidth, 
+                   edgecolor='k')
+                
+    return ax
+
+def _draw_node_labels(ax, xpos, ypos, fontsize=14):
+    """Draw decision tree node labels. See plot_decision tree for details."""
+   
+    for i, (x, y) in enumerate(zip(xpos, ypos)):
+        ax.text(x, y, '%0.0f' %(i+1), ha='center', va='center', fontsize=fontsize)
+        
+    return ax
+
+def _draw_edges(ax, xpos, ypos, edges, color='0.5'):
[…]

### 2019-01-06 sisyphus/viz/_tree.py
making progress on decision tree figure but still not happy

@@ -44,11 +44,17 @@ def _draw_node_labels(ax, xpos, ypos, fontsize=14):
         
     return ax
 
-def _draw_edges(ax, xpos, ypos, edges, color='0.5'):
+def _draw_edges(ax, xpos, ypos, edges, linewidth=None, color='0.5'):
     """Draw decision tree edges. See plot_decision tree for details."""
     
-    for i, j in edges:
-        ax.plot([xpos[i], xpos[j]], [ypos[i], ypos[j]], color=color, zorder=0)
+    ## Define line widths.
+    if linewidth is None: linewidth = np.repeat(0.5, len(edges))
+    assert np.equal(len(linewidth), len(edges)) 
+    linewidth = Normalize(0, 1)(np.array(linewidth)) * 5 + 0.5
+        
+    for i, (s1, s2) in enumerate(edges):
+        ax.plot([xpos[s1], xpos[s2]], [ypos[s1], ypos[s2]], color=color, 
+                lw=linewidth[i], zorder=0)
 
     return ax
 
@@ -80,7 +86,8 @@ def _draw_path_sums(ax, xpos, linewidth=5, fontsize=14):
     return ax
     
 def plot_decision_tree(ax, s=1000, color=None, cmap=None, vmin=None, vmax=None, alpha=1.0, 
-                       node_width=None, node_labels=False, edge_labels=False, path_sums=False):
+                       node_width=None, node_labels=False, edge_labels=False, edge_width=None, 
+                       path_sums=False):
 
     ## Define decision tree.
     T = np.zeros((15,15))
@@ -100,7 +107,7 @@ def plot_decision_tree(ax, s=1000, color=None, cmap=None, vmin=None, vmax=None, 
     ypos = [3, 2, 2, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0]
     
     ## Draw DAG.
-    ax = _draw_edges(ax, xpos, ypos, edges=edges)
+    ax = _draw_edges(ax, xpos, ypos, edges=edges, linewidth=edge_width)
     ax = _draw_nodes(ax, xpos, ypos, s=s, color=color, cmap=cmap, vmin=vmin, vmax=vmax,
                         alpha=alpha, linewidth=node_width)
     

### 2019-01-06 sisyphus/viz/__init__.py
first pass at a real free choice image

@@ -1,5 +1,6 @@
 """Visualization functions"""
 
 from ._base import plot_policy
+from ._choice import plot_free_choice
 from ._cliff import plot_cliff
 from ._tree import plot_decision_tree

### 2019-01-06 sisyphus/viz/_choice.py

@@ -1,0 +1,79 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+from matplotlib.lines import Line2D
+
+def _draw_nodes(ax, s, free_color, fixed_color):
+    
+    ## Draw initial choice.
+    xpos = [0.0]; ypos = [0.0]
+    ax.plot(xpos, ypos, marker='o', markersize=s, fillstyle='top', linestyle='none',
+            color=free_color, markerfacecoloralt=fixed_color, markeredgecolor='k')
+
+    ## Draw free choice.
+    xpos = [1.0, 2.0, 2.0]; ypos = [0.25, 0.5, 0.0]
+    ax.plot(xpos, ypos, marker='o', markersize=s, fillstyle='full', linestyle='none',
+            color=free_color, markerfacecoloralt=fixed_color, markeredgecolor='k')
+    
+    ## Draw fixed choice.
+    xpos = [1.0, 2.0]; ypos = [-0.25, -0.5]
+    ax.plot(xpos, ypos, marker='o', markersize=s, fillstyle='full', linestyle='none',
+            color=fixed_color, markerfacecoloralt=free_color, markeredgecolor='k')
+    
+    return ax
+
+def _draw_edges(ax):
+
+    ## Draw edges.
+    ax.plot([0.0, 2.0], [0.0, 0.5], color='k', zorder=0)
+    ax.plot([0.0, 2.0], [0.0,-0.5], color='k', zorder=0)
+    ax.plot([1.0, 2.0], [0.25,0.0], color='k', zorder=0)
+
+    return ax
+
+def _draw_edge_labels(ax, fontsize=14):
+    
+    ax.text(0.5,  0.15, '0', ha='center', va='bottom', fontsize=fontsize)
+    ax.text(0.5, -0.15, '0', ha='center', va='top', fontsize=fontsize)
+    ax.text(1.5,  0.40, r'$R$', ha='center', va='bottom', fontsize=fontsize)
+    ax.text(1.5,  0.10, r'$R$', ha='center', va='top', fontsize=fontsize)
+    ax.text(1.5, -0.40, r'$R$', ha='center', va='top', fontsize=fontsize)
+    ax.annotate(r'$R = \{0.33: -1, 0.33: 0, 0.33: 1\}$', (0,0), (0,0), 
+                xycoords='axes fraction', ha='left', va='bottom', fontsize=fontsize-4)
+    return ax
+
[…]

### 2019-01-06 sisyphus/viz/_tree.py
updated decision tree figure. better, still not perfect

@@ -5,7 +5,7 @@ from matplotlib.cm import get_cmap
 from matplotlib.colors import ListedColormap, Normalize
 
 def _draw_nodes(ax, xpos, ypos, s=1000, color=None, cmap=None, vmin=None, vmax=None, 
-                alpha=1.0, linewidth=0):
+                alpha=1.0, linewidth=1.0):
     """Draw decision tree nodes. See plot_decision tree for details."""
     
     ## Define colors.
@@ -44,50 +44,57 @@ def _draw_node_labels(ax, xpos, ypos, fontsize=14):
         
     return ax
 
-def _draw_edges(ax, xpos, ypos, edges, linewidth=None, color='0.5'):
+def _draw_edges(ax, xpos, ypos, edges, linewidth=1, color='0.5'):
     """Draw decision tree edges. See plot_decision tree for details."""
     
     ## Define line widths.
-    if linewidth is None: linewidth = np.repeat(0.5, len(edges))
-    assert np.equal(len(linewidth), len(edges)) 
-    linewidth = Normalize(0, 1)(np.array(linewidth)) * 5 + 0.5
+    if isinstance(linewidth, (int, float)): 
+        linewidth = np.repeat(linewidth, len(edges))
         
+    ## Iteratively draw.
     for i, (s1, s2) in enumerate(edges):
         ax.plot([xpos[s1], xpos[s2]], [ypos[s1], ypos[s2]], color=color, 
                 lw=linewidth[i], zorder=0)
 
     return ax
 
-def _draw_edge_labels(ax, T, fontsize=14):
+def _draw_edge_labels(ax, labels, fontsize=14, alpha=1):
     
     ## Define label positions.
     xpos = [-0.6,0.6,-1.3,-0.7,0.7,1.3,-1.67,-1.33,-0.67,-0.33,0.33,0.67,1.33,1.67]
     ypos = [2.5] * 2 + [1.5] * 4 + [0.5] * 8
     halign = ['right','left'] * 7
     
+    ## Define label transparency.
+    if isinstance(alpha, (int, float)): alpha = np.repeat(alpha, len(xpos))
+    
     ## Draw edge labels.
-    for x, y, r, ha in zip(xpos, ypos, T[T.nonzero()], halign):
-        ax.text(x,y,'%0.0f' %r, va='center', ha=ha, fontsize=fontsize)
+    for x, y, r, ha, a in zip(xpos, ypos, labels, halign, alpha):
[…]

### 2019-01-06 manuscript/manuscript.tex
updated/fixed cliff-walking figure.
caught up latex to gdocs

@@
 
-\startmain % to start the main 1-4 pages of the submission.
+\startmain
 
 \section{Introduction}
+Many psychiatric disorders involve distortions in models of the external world
+and, correspondingly, aberrations of learning and decision making. As a normative
+framework, decision theory allows us to describe and pose quantita­tive questions
+about optimal and approximately optimal choice behavior \cite{DayanDaw2008} and
+is, therefore, a critical tool for modeling, understanding, and predicting the
+psychological and neurobiological mechanisms responsible for the abnormalities
+observed in psychiatric behavior \cite{HuysDawDayan2015}. Many recent investigations
+employing this approach have provided important insights into psychiatric symptoms,
+including anhedonia in major depression \cite{Rutledge2017}, mood instability
+in bipolar disorder \cite{EldarNiv2015, EldarDolanNiv2016},  habitual decision
+making in disorders of compulsivity \cite{gillan2016}, and hallucinations in
+schizophrenia \cite{powers2017, corlett2018}.
+
+Anxiety disorders (generalized anxiety disorder, social anxiety disorder, panic
+disorder, obsessive-compulsive disorder) represent the most common family of
+psychiatric disorders (citation). Despite the differences in the content of worry,
+all anxiety disorders share two core features: pessimistic inference and excessive
+avoidance. Pessimistic inference describes a tendency to appraise threatening
+situations as more frequent and/or more severe than they objectively are. This has
+been reliably observed experimentally, as in fear conditioning wherein individuals
+with anxiety exhibit exaggerated fear towards safe cues (CS-) and extinguished
+threatening cues \cite{lissek2005, Duits2015}, and clinically, through self-report
+(pick 3 citaitons). Excessive avoidance describes the corresponding tendency for
+individuals with anxiety to consistently flee or avoid threatening situations.
+Avoidance is often chosen even at the expense of potential reward (approch-avoidance
+conflict), and is frequently a major impediment to recovery \cite{Arnaudova2017}.
+
+Recent studies have attempted to provide some computational models of this behavior
+\cite{Moutoussis2008, Maia2010}. For example, \cite{Maia2010} was able to show
+how reinforcement learning could explain sustained avoidance in the absence of
+repeated reinforcement, observed in rodents \cite{servatius2008} and something
+not explained by two-factor theory \cite{Krypotos2015}. As has been discussed
+elsewhere \cite{Moutoussis2017}, such accounts only explain the maintenance of
+avoidance in short timescales, but do not explain why resistant to change and
+why they do not go away in general. Such theories also do not explain the magnitude
+of pessimistic inference, or the fear generalization. A broader account is needed.
+
+In this article, we put forward a decision theoretic explanation of the core
+symptoms of anxiety. In sequential decision environments, we show how a loss of
+optimistic inference results in pessimistic inference, chronic avoidance, and
+generalization of fear.
 
@@
   }
-  \caption{\textbf{Open Field Environment}}
+  \caption{\textbf{Open Field}}
   \par Oooh so pretty.
@@
 
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/02_cliff.png}}%
+  }
+  \caption{\textbf{Cliff Walking}}
+  \par You better watch out YOU BETTER WATCH OUT
+\end{figure}
+
+Discuss similar observations in other tasks (Bach et al., 2014; Sheynin et al., 2015).
+
 \subsection{Approach-Avoidance Conflict}
+
+Discuss Bach behavioral inhibition.
 
@@
 \subsection{Longitudinal Progression}
+
+\section{Discussion}
+
+In this article, we showed...
+
+Discuss previous computational approaches: This has also been previously been
+investigated in previous computational modeling \cite{Huys2009}. There
+the authors found that beliefs about a lack of controllability, mechanistically
+modeled as corrupted priors about action-outcome contingencies, found deficient
+learning in novel environments. The present results are no doubt related, but
+address a different environment and a different behavior.
+
+We discussed two mechanisms by which negative values can spread. There is a third
+that we did not address here: that of the state itself. Here we assumed that states
+were fully observable; in other words, there was no uncertainty about the state
+an agent was currently occupying. In partially-observable Markov decision problems,
+the state itself must be inferred from external and internal variables. This is
+yet another means by which bad value can spread if undue likelihood is granted
+to undesirable or aversive states. This ideas has previously been proposed to
+explain some symptoms in anxiety \cite{Paulus2012}, and not without good reason.
+Anxiety disorders involve negatively biased interpretation of ambiguous percepts
+\cite{Hartley2012}, which may result in aberrant planning and increased avoidance
+for similar reasons to those described above. This is an area for future research.
+
+In this article, we have explored a computational principle of approach and
+avoidance learning in sequential decision environments. We have not discussed,
+however, the particular psychological or neurobiological mechanisms by which this
+sort of learning may take place. This issue has been discussed in detail elsewhere
+\cite{Bishop2018}, but we briefly discuss a few possibilities here. Mechanically,
+we understand that complex decision problems are solved by mental processes that
+exist on a spectrum between model-based and model-free (Daw et al., 2005). The
+difference between these two families of methods hinges on the reliance of an
+internal model of the world used to make predictions about expected long-run
+outcomes of particular choices.
+
+For model-based decision making, one possibility is biased planning. In large
+state spaces, online sequential decision making suffers from the curve of
+dimensionality: in planning a series of actions to take, comparing between
+multiple options rapidly becomes computationally intractable when the number of
+options and depth of search becomes even moderately large. Thus, heuristics for
+reducing the size of this search can be useful. Biased pruning of search paths
+(such as immediately discarding paths that require some minor loss in pursuit of
+larger gains) may result in maladaptive decision making (Huys et al., 2012).
+Alternately, if planning relies on internal sampling from previously experienced
+episodes in order to make predictions about future value, then biased sampling
+ may similarly result in maladaptive decision making. Recent work has shown how
+ finite sampling during planning can result in risk aversion and the overrepresentation
+ of low probability but strongly negative outcomes (Lieder et al., 2018). Such a
+ mechanism is prima facie in line with results showing the availability in memory
+ of negative outcomes being higher for patients with anxiety (Borkovec et al., 1999;
+ Miranda and Mennin, 2007).
+
+Within the realm of model-free decision making, several studies have found evidence
+suggesting increased learning rates for negative reward prediction errors in
+individuals with elevated levels of anxiety \cite{Aylward, Huang2017, Harle2017
+garrett2018}. Though an asymmetry in sensitivity to
+positive and negative reward prediction errors bears superficial similarity, for
+the reasons described above merely attending more to negative information does not
+necessarily predict increased avoidance insofar that avoidance can successfully
+isolate negative outcomes. One possibility, and topic of future research, is how
+prior beliefs about the reward statistics of the current environment dictate
+learning rates. A second possibility is prioritized replay (Russek et al., 2017;
+Mattar and Daw, 2018). Offline replay of previous experiences (through hippocampal
+mechanisms) are known to facilitate learning. It has been proposed that biased
+replay could bias value estimates (Gagne et al., 2018). This is another exciting
+area for future research.
+
+\section{References}
+
+\bibliographystyle{authordate1}
+\bibliography{citations}
 

### 2019-01-07 sisyphus/viz/__init__.py
first pass at images unified under single color scheme

@@ -3,4 +3,5 @@
 from ._base import plot_policy
 from ._choice import plot_free_choice
 from ._cliff import plot_cliff
+from ._field import plot_field
 from ._tree import plot_decision_tree

### 2019-01-07 sisyphus/viz/_cliff.py

@@ -48,7 +48,9 @@ def plot_cliff(ax, cbar=False, viable='0.8', cliff='0.1', start='#1f77b4', end='
     ax.hlines(np.arange(1,11),0,12,lw=0.1)
     
     ## Add text.
-    ax.text(0.5,10.5,'S',ha='center',va='center',fontsize=18,weight='semibold')
-    ax.text(11.5,10.5,'G',ha='center',va='center',fontsize=18,weight='semibold')
+    ax.annotate('S', (0, 0), (0.041, 0.04), 'axes fraction', ha='center', 
+                va='center', fontsize=18, weight='semibold')
+    ax.annotate('G', (0, 0), (0.954, 0.04), 'axes fraction', ha='center', 
+                va='center', fontsize=18, weight='semibold')
     
     return ax

### 2019-01-07 sisyphus/viz/_field.py

@@ -1,0 +1,43 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+from matplotlib.colors import ListedColormap
+
+def plot_field(ax, cbar=False, viable='0.8', reward='#f3e1db', shock='#1c142a'):
+    """Plot cliff-walking environment.
+    
+    Parameters
+    ----------
+    ax : matplotlib Axes
+        Axes in which to draw the plot.
+    cbar : bool
+        Whether to draw a colorbar.
+    viable : str
+        Color of viable squares.
+    reward : str
+        Color of rewarding tile.
+    end : str
+        Color of ending tile.
+        
+    Returns
+    -------
+    ax : matplotlib Axes
+        Axes in which to draw the plot.
+    """
+    
+    ## Define grid.
+    grid = np.zeros((11, 11))    # Viable states
+    grid[5,[2,-3]] = [1, 2]      # Reward/shock states
+    
+    ## Define colormap.
+    cmap = ListedColormap([viable, reward, shock])
+
+    ## Plot cliff.
+    ax = sns.heatmap(grid, cmap=cmap, cbar=cbar)
+    ax.set(xticklabels=[], yticklabels=[], title='Cliff Walking')  
+    
+    ## Add outline.
+    ax.vlines(np.arange(1,11),0,11,lw=0.1)
+    ax.hlines(np.arange(1,11),0,11,lw=0.1)
+    
+    return ax

### 2019-01-10 manuscript/manuscript.tex
first pass at images unified under single color scheme
first pass at longidutinal figure

@@
 \documentclass[11pt]{article} % For LaTeX2e
-\usepackage{manuscript,palatino}
+\usepackage{manuscript, palatino}
 \usepackage{graphicx}
-\usepackage{amsfonts,amsmath}
+\usepackage{amsfonts, amsmath}
 \usepackage{algorithm, algpseudocode}%
@@
 }
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to \LaTeX{} to determine where to break
-% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
-% puts 3 of 4 authors names on the first line, and the last on the second
-% line, try using \AND instead of \And before the third author name.
 
@@
 policy in the future. We can adopt a different function. There are a number to
-choose from (Garcia \& Fernandez, 2015). Here we adopt the beta-pessimism criterion
-(Gaskett, 2003):
+choose from \cite{Garcia2015}. Here we adopt the beta-pessimism criterion
+\cite{Gaskett2003}:
 
@@
   \caption{\textbf{Open Field}}
-  \par Oooh so pretty.
+  \par Current parameters: Reward = 10, Shock = -10, gamma = 0.95.
 \end{figure}
@@
 \subsection{Avoidance}
+
+1. Open with definition of avoidance. Follow \cite{Arnaudova2017} in definitions, i.e.
+the basic idea that avoidance is the distancing of oneself from real or perceived
+threat.
+2. Describe cliff walking environment. Need to discuss that we are expanding
+on the result originally described in \cite{Gaskett2003}.
+3. With increasing pessimism, agent further distances self from cliff. Note that
+this is the optimal policy under those set of beliefs about the world.
+4. This can explain how avoidance is maintained. Even with luck and chance in the
+the world, these sets of beliefs actually lead to and sustain these behavioral
+policies.
+
+In discussion of section, need to also mention other relevant findings. Specifically,
+need to describe the approach-avoidance findings of Bach (2014, 2017) and the
+complex task findings of Sheynin et al., 2015.
 
@@
   \caption{\textbf{Cliff Walking}}
-  \par You better watch out YOU BETTER WATCH OUT
+  \par Cliff = -100, Cost = -1, gamma = 0.99
 \end{figure}
-
-Discuss similar observations in other tasks (Bach et al., 2014; Sheynin et al., 2015).
 
@@
 
-Discuss Bach behavioral inhibition.
+In the third section of the results, we discuss approach-avoidance conflict.
+This covers tasks such as the balloon analog risk task (BART) and the flight
+initiation distance task, which has been found to correlate with anxiety. The
+hallmark of these sorts of tasks is the tradeoff between increased reward vs.
+the possibility of an aversive outcome (e.g. balloon pop, predation). Discuss
+previous findings, i.e. earlier leave times.
+
+We can show that this is easily explained by our model. A lack of belief of
+future control predicts exiting while things are still good. (We see this in
+Figure 3). The optimal policy in these environments is an earlier leaving time.
+
+This can also be used to explain behavioral inhibition \cite{bach2015, khemka2017}.
+The idea is slightly different than what is advocated. The claim here is that
+in anxious individuals, the Q-values should be closer together, and thus there
+should be longer waiting time. This requires demonstrating that decisions are
+not made in advance; only at the time of decision making. This is something
+I think we will have to argue. (Need to double check the literature here.)
+
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/03_fid.png}}%
+  }
+  \caption{\textbf{Flight Initiation Distance}}
+  \par remember to add parameters
+\end{figure}
 
@@
 
+Basically reviewing the arguments of Huys here \cite{Huys2012, Lally2017}.
+Need to get facts straight before going through all of this.
+
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/04_tree.png}}%
+  }
+  \caption{\textbf{Decision Tree}}
+  \par remember to add parameters
+\end{figure}
+
 \subsection{Free Choice Premium}
+
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/05_choice.png}}%
+  }
+  \caption{\textbf{Free Choice Premium}}
+  \par remember to add parameters
+\end{figure}
 

### 2019-01-10 sisyphus/envs/__init__.py

@@ -4,4 +4,5 @@ from ._choice import FreeChoice
 from ._cliff import CliffWalking
 from ._fid import FlightInitiationDistance
 from ._field import OpenField
+from ._lh import Helplessness
 from ._tree import DecisionTree

### 2019-01-10 sisyphus/envs/_lh.py

@@ -1,0 +1,30 @@
+import numpy as np
+from ._base import GraphWorld
+
+class Helplessness(GraphWorld):
+    
+    def __init__(self, rewards=[-1,1]):
+        
+        ## Error catching.
+        assert np.equal(len(rewards), 2)
+            
+        ## Define one-step transition matrix.
+        T = np.zeros((5, 5)) * np.nan
+        T[0, [1, 4]] = 1
+        T[1, [2, 3]] = 1
+        T[np.arange(2,5),np.arange(2,5)] = 1
+
+        ## Define rewards.
+        R = np.copy(T)
+        R[np.where(~np.isnan(R))] = 0
+        R[1,[2, 3]] = rewards
+        
+        ## Define start/terminal states.
+        start = 0
+        terminal = np.arange(2,5)
+        
+        ## Initialize GraphWorld.
+        GraphWorld.__init__(self, T, R, start, terminal, 0)
+        
+    def __repr__(self):
+        return '<GraphWorld | Learned Helplessness>'

### 2019-01-10 sisyphus/mdp/_td.py

@@ -1,16 +1,16 @@
 """Temporal difference module"""
 
 import numpy as np
-from ._misc import check_params, softmax, betamax, categorical
+from copy import deepcopy
+from ._misc import check_params, betamax, categorical
+from ._misc import softmax as _softmax
 
-def greedy_choice(arr, epsilon):
-    if np.random.binomial(1,1-epsilon): 
-        return np.argmax(arr)
-    else: 
-        return np.random.choice(np.arange(len(arr)),1,replace=False)
+def egreedy(arr, epsilon):
+    if np.random.binomial(1,1-epsilon): return np.argmax(arr)
+    else: return np.random.choice(np.arange(len(arr)),1,replace=False)[0]
     
-def soft_choice(arr, beta):
-    theta = softmax(arr * beta)
+def softmax(arr, beta):
+    theta = _softmax(arr * beta)
     return categorical(theta)  
 
 class ModelFree(object):
@@ -38,7 +38,7 @@ class ModelFree(object):
         self.policy = policy
         if policy == 'max': self._policy = np.max
         elif policy == 'min': self._policy = np.min
-        elif policy == 'softmax': self._policy = lambda arr: arr @ softmax(arr * self.beta)
+        elif policy == 'softmax': self._policy = lambda arr: arr @ _softmax(arr * self.beta)
         elif policy == 'betamax': self._policy = lambda arr: betamax(arr, self.beta)
         else: raise ValueError('Policy "%s" not valid!' %self.policy)
         
@@ -51,6 +51,10 @@ class ModelFree(object):
     def __repr__(self):
         return '<Model Free Agent>'
             
+    def copy(self):
+        """Return copy of agent."""
+        return deepcopy(self)
+        
     def _run_episode(self, Q, gym, choice, epsilon, n_steps=100):
         
         ## Define starting state.
@@ -64,7 +68,8 @@ class ModelFree(object):
                 
             ## Select action.
             copy['Q'] = Q.copy()
-            a = copy[copy.S==s].index[choice(copy.loc[copy.S==s,'Q'], epsilon)]
+            i = choice(copy.loc[copy.S==s,'Q'].values, epsilon)
+            a = copy[copy.S==s].index[i]
                         
             ## Observe next state and reward.
             i = categorical(copy.loc[a,'T'])
@@ -136,24 +141,22 @@ class ModelFree(object):
         if choice == 'greedy': 
 
             ## Define choice rule.
-            _choice = greedy_choice
+            choice = egreedy
 
             ## Define schedule.
[…]

### 2019-01-10 sisyphus/viz/__init__.py

@@ -4,4 +4,5 @@ from ._base import plot_policy
 from ._choice import plot_free_choice
 from ._cliff import plot_cliff
 from ._field import plot_field
+from ._lh import plot_helplessness
 from ._tree import plot_decision_tree

### 2019-01-10 sisyphus/viz/_lh.py

@@ -1,0 +1,60 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+from numpy import radians as rad
+from matplotlib.patches import Arc, RegularPolygon
+
+def _draw_nodes(ax, s):
+    
+    ## Define node positions.
+    xpos = [-1, 0, 1, 1]
+    ypos = [0.0, 0.0, 0.5,-0.5]
+    
+    ## Draw nodes.
+    ax.scatter(xpos, ypos, s)
+    
+    return ax
+
+def _draw_edges(ax):
+
+    ## Draw directed edges.
+    ax.plot([-1.0, 0.0], [0.0, 0.0], color='k', zorder=0)
+    ax.plot([ 0.0, 1.0], [0.0, 0.5], color='k', zorder=0)
+    ax.plot([ 0.0, 1.0], [0.0,-0.5], color='k', zorder=0)
+    
+    ## Draw self-loop.
+    ax = _draw_self_loop(ax, 0.4, -1.25, 0, 70, -155)
+    
+    return ax
+
+def _draw_self_loop(ax, radius, centX, centY, angle_, theta2_, color_='black'):
+    
+    ## Construct line.
+    arc = Arc([centX,centY],radius,radius,angle=angle_,
+          theta1=0,theta2=theta2_,capstyle='round',linestyle='-',lw=1,color=color_)
+    ax.add_patch(arc)
+
+    ## Construct arrow head.
+    endX=centX+(radius/2)*np.cos(rad(theta2_+angle_)) #Do trig to determine end position
+    endY=centY+(radius/2)*np.sin(rad(theta2_+angle_))
+
+    ax.add_patch(                    #Create triangle as arrow head
+        RegularPolygon(
+            (endX, endY),            # (x,y)
+            3,                       # number of vertices
+            radius/9,                # radius
+            rad(angle_+theta2_),     # orientation
+            color=color_
+        )
+    )
+    
+    return ax
+
+def plot_helplessness(ax):
+    
+    _draw_edges(ax)
+    _draw_nodes(ax, 2500)
+
+    ## Clean up.
+    ax.set(xticks=[], yticks=[], ylim=(-1.5, 1.5), xlim=(-1.6, 1.6))
+    sns.despine(top=True, bottom=True, left=True, right=True, ax=ax)

### 2019-01-13 manuscript/manuscript.tex
sunday writing (pt 1)

@@
 \usepackage{algorithm, algpseudocode}%
+
+\usepackage{natbib}
 
@@
 \texttt{zorowitz@princeton.edu} \\
+\And
+Ida Momennejad \\
+columbia University\\
+New York, NY 10027 \\
+\texttt{ida.m@columbia.edu} \\
 \And
@@
 framework, decision theory allows us to describe and pose quantita­tive questions
-about optimal and approximately optimal choice behavior \cite{DayanDaw2008} and
+about optimal and approximately optimal choice behavior \citep{DayanDaw2008} and
 is, therefore, a critical tool for modeling, understanding, and predicting the
 psychological and neurobiological mechanisms responsible for the abnormalities
-observed in psychiatric behavior \cite{HuysDawDayan2015}. Many recent investigations
+observed in psychiatric behavior \citep{HuysDawDayan2015}. Many recent investigations
 employing this approach have provided important insights into psychiatric symptoms,
-including anhedonia in major depression \cite{Rutledge2017}, mood instability
-in bipolar disorder \cite{EldarNiv2015, EldarDolanNiv2016},  habitual decision
-making in disorders of compulsivity \cite{gillan2016}, and hallucinations in
-schizophrenia \cite{powers2017, corlett2018}.
+including anhedonia in major depression \citep{Rutledge2017}, mood instability
+in bipolar disorder \citep{EldarNiv2015, EldarDolanNiv2016},  habitual decision
+making in disorders of compulsivity \citep{gillan2016}, and hallucinations in
+schizophrenia \citep{powers2017, corlett2018}.
 
@@
 with anxiety exhibit exaggerated fear towards safe cues (CS-) and extinguished
-threatening cues \cite{lissek2005, Duits2015}, and clinically, through self-report
+threatening cues \citep{lissek2005, Duits2015}, and clinically, through self-report
 (pick 3 citaitons). Excessive avoidance describes the corresponding tendency for
@@
 Avoidance is often chosen even at the expense of potential reward (approch-avoidance
-conflict), and is frequently a major impediment to recovery \cite{Arnaudova2017}.
+conflict), and is frequently a major impediment to recovery \citep{Arnaudova2017}.
 
 Recent studies have attempted to provide some computational models of this behavior
-\cite{Moutoussis2008, Maia2010}. For example, \cite{Maia2010} was able to show
+\citep{Moutoussis2008, Maia2010}. For example, \citep{Maia2010} was able to show
 how reinforcement learning could explain sustained avoidance in the absence of
-repeated reinforcement, observed in rodents \cite{servatius2008} and something
-not explained by two-factor theory \cite{Krypotos2015}. As has been discussed
-elsewhere \cite{Moutoussis2017}, such accounts only explain the maintenance of
+repeated reinforcement, observed in rodents \citep{servatius2008} and something
+not explained by two-factor theory \citep{Krypotos2015}. As has been discussed
+elsewhere \citep{Moutoussis2017}, such accounts only explain the maintenance of
 avoidance in short timescales, but do not explain why resistant to change and
@@
 
+Importantly, in the clinical literature, pathological anxiety is also associated
+with a lack of perceived control (Gallagher et al., 2014). Studies suggest that
+this belief mediates anxiety (ref?). This can come in whatever form.
+
 Alternately, an agent may choose optimally in the present but may be uncertain of
@@
 policy in the future. We can adopt a different function. There are a number to
-choose from \cite{Garcia2015}. Here we adopt the beta-pessimism criterion
-\cite{Gaskett2003}:
+choose from \citep{Garcia2015}. Here we adopt the beta-pessimism criterion
+\citep{Gaskett2003}:
 
@@
 
-First and foremost, pathological anxiety is associated with pessimistic expectations
-about the probability and cost of future negative events (Grupe \& Nitschke, 2013;
-Bishop \& Gagne, 2018). In fear conditioning, individuals with anxiety are more
-likely to exhibit increased threat appraisal for safety signals (CS-) and the
-extinguished fear cue (CS+) (Lissek et al. 2005; Duits et al., 2015). In considering
-hypothetical future negative outcomes, individuals with anxiety are more likely
-to evaluate negative outcomes as more likely and their consequences more severe
-(pick a few references).
+One symptom central to all of the anxiety disorders is pessimistic inference,
+or a tendency to appraise the likelihood and/or severity of threatening situations
+out of proportion to their actual environmental contingencies \citep{dsm5, BeckClark1997,
+ClarkBeck2011}. For example, an individual with specific phobia may vastly exaggerate
+the probability of encountering their fear (e.g. a venomous spider) in their
+everyday lives. The first studies of pessimistic inference in clinically anxious
+populations involved individuals judging the perceived likelihood and severity
+of different aversive situations, with the general finding that anxious individuals
+rated these as more likely than non-clinical individuals \citep{ButlerMathews1983,
+ButlerMathews1987, Foa1996, MacLeod1996, MacLeod1997, Luten1997, Stober1997,
+Borkovec1999}; a finding that has been replicated more recently \citep{Maner2006,
+Mitte2007, Miranda2007}.
 
-Importantly, in the clinical literature, pathological anxiety is also associated
-with a lack of perceived control (Gallagher et al., 2014). Studies suggest that
-this belief mediates anxiety (ref?). This can come in whatever form.
+Experimentally, aberrations in threat appraisal have been repeatedly observed
+in a litany of associative learning studies. Common to all of these studies,
+participants learn that a particular stimulus (CS+) predicts the occurrence
+of a negative outcome (e.g. shock, loud noise, aversive image). In simple conditioning,
+there is some evidence to suggest that individuals with anxiety exhibit elevated
+levels of response to threatening stimuli \citep{lissek2005, MinekaOehlberg2008,
+Duits2015}. Importantly, conditioning studies also suggest impaired safety learning
+in anxiety, wherein a stimulus never paired with an aversive outcome (CS-) is also
+feared \citep{Duits2015}. Other notable findings include impaired blocking,
+conditioned discrimination (AX+/BX-) (citations), and conditioned inhibition
+(A+, AB-) (citations).
 
-In the first series of simulations, we show directly how lack of perceived control
-can directly account for pessimistic inference. In the open field environment
-(Figure 1a), an agent navigates a simple grid-world. The world consists of almost
-entirely of empty tiles, except for two which yield a reward and punishment,
-respectively.
-
-Under complete control (objective or perceived), an agent in this environment is
-under no threat despite the presence of an aversive state. This is simply because,
-in a full controllable world, a (non-masochistic) agent is never forced to interact
-with the aversive state. When this control breaks down -- through a stochastic
-world or a distrust of future competency -- an agent can no longer guarantee it
-will never encounter the aversive state. In such a scenario, the negative value
-back-propagates to its predecessor states through the recursive formulas defined
-above. The result is an environment that appears worse (i.e. less positive) than
-it otherwise would.
-
-We shoe this above. Using value iteration. Show it for best-case scenario. Show
-it for expected random. Show it for worst-case scenario.
+% I think when you come back to edit this, you need to highlight these findings
+% more in the context of what you are about to explain. Pessimistic inference
+% is both making bad things worse and, most importantly, that non-bad things
+% become bad. Although one question is how much to include the other less
+% studied paradigms
 
@@
 
+In our first result, we demonstrate one simple mechanism by which pessimistic
+inference and exaggerated threat appraisals may arise. In the *open field* environment,
+an agent is placed in a large grid world entirely empty except for two states,
+leading to positive and negative outcomes, respectively. Under complete control
+(objective or perceived), an agent in this environment in under no threat despite
+the presence of negative states. This is simply because, in a fully controllable
+world, an agent need not ever interact with the aversive state. In fact, an agent
+can decide to walk straight up to the aversive state and, insofar it does not
+choose to take one step further, will never be in any actual danger.
+
+This is not so when deterministic control cannot be assumed. If an agent believes
+it is not in full control of its environment, then it cannot rule out the possibility
+that it may inadvertently encounter the negative state at some point in the future,
+especially if it veers to close to the state. This is true irrespective of whether
+the agent believes it cannot guarantee its own future competency or the benevolence
+of the environment. Importantly, this belief need not match the true environmental
+statistics; belief is enough.
+
+The long-run state values under different beliefs are plotted in Figure 1. As
+expected, under optimistic beliefs (Figure 1b) the entire environment sans negative
+state takes on a positive value (close to the positive state itself). As this belief
+breaks down, we observe a different pattern in the long-run state values. When
+future occupancy of the negative state becomes a non-zero probability, the undesired
+action transition exerts influence on action-outcome calculations, back-propagating,
+and thereby infecting antecedent states. As the agent's belief becomes increasingly
+pessimistic, the more the worst possible action exerts influence and the more
+negative, on average, the state-space grows.
+
+In the breakdown of perceived control then, negative states can grow worse (i.e.
+more predictive of future negative outcome) and otherwise positive states can too.
+This occurs as a simple result that if control not guaranteed, cannot guaranteed
+bad things wont happen.
+
 \subsection{Avoidance}
 
-1. Open with definition of avoidance. Follow \cite{Arnaudova2017} in definitions, i.e.
+1. Open with definition of avoidance. Follow \citep{Arnaudova2017} in definitions, i.e.
 the basic idea that avoidance is the distancing of oneself from real or perceived
@@
 2. Describe cliff walking environment. Need to discuss that we are expanding
-on the result originally described in \cite{Gaskett2003}.
+on the result originally described in \citep{Gaskett2003}.
 3. With increasing pessimism, agent further distances self from cliff. Note that
@@
 
-This can also be used to explain behavioral inhibition \cite{bach2015, khemka2017}.
+This can also be used to explain behavioral inhibition \citep{bach2015, khemka2017}.
 The idea is slightly different than what is advocated. The claim here is that
@@
 
-Basically reviewing the arguments of Huys here \cite{Huys2012, Lally2017}.
+Basically reviewing the arguments of Huys here \citep{Huys2012, Lally2017}.
 Need to get facts straight before going through all of this.
@@
 Discuss previous computational approaches: This has also been previously been
-investigated in previous computational modeling \cite{Huys2009}. There
+investigated in previous computational modeling \citep{Huys2009}. There
 the authors found that beliefs about a lack of controllability, mechanistically
@@
 to undesirable or aversive states. This ideas has previously been proposed to
-explain some symptoms in anxiety \cite{Paulus2012}, and not without good reason.
+explain some symptoms in anxiety \citep{Paulus2012}, and not without good reason.
 Anxiety disorders involve negatively biased interpretation of ambiguous percepts
-\cite{Hartley2012}, which may result in aberrant planning and increased avoidance
+\citep{Hartley2012}, which may result in aberrant planning and increased avoidance
 for similar reasons to those described above. This is an area for future research.
@@
 sort of learning may take place. This issue has been discussed in detail elsewhere
-\cite{Bishop2018}, but we briefly discuss a few possibilities here. Mechanically,
+\citep{Bishop2018}, but we briefly discuss a few possibilities here. Mechanically,
 we understand that complex decision problems are solved by mental processes that
@@
 suggesting increased learning rates for negative reward prediction errors in
-individuals with elevated levels of anxiety \cite{Aylward, Huang2017, Harle2017
+individuals with elevated levels of anxiety \citep{Aylward, Huang2017, Harle2017
 garrett2018}. Though an asymmetry in sensitivity to
@@
 
-\bibliographystyle{authordate1}
+\bibliographystyle{vancouver-authoryear}
 \bibliography{citations}

### 2019-01-13 manuscript/manuscript.tex
sunday writing (pt 2)

@@
 
-1. Open with definition of avoidance. Follow \citep{Arnaudova2017} in definitions, i.e.
-the basic idea that avoidance is the distancing of oneself from real or perceived
-threat.
-2. Describe cliff walking environment. Need to discuss that we are expanding
-on the result originally described in \citep{Gaskett2003}.
-3. With increasing pessimism, agent further distances self from cliff. Note that
-this is the optimal policy under those set of beliefs about the world.
-4. This can explain how avoidance is maintained. Even with luck and chance in the
-the world, these sets of beliefs actually lead to and sustain these behavioral
-policies.
+A second symptom central to all anxiety disorders is avoidance \citep{dsm5,
+Krypotos2015, Arnaudova2017}, or actions taken to increase or maintain the temporal,
+physical and/or mental distance between an individual and a threatening situation.
+Although avoidance is a natural response to danger, excessive avoidance as is
+observed in anxiety disorders is can be highly disruptive to everyday functioning
+\cite{Salter2004} and may work to sustain anxious appraisals and behavior. For
+example, an individual with social anxiety may avoid attending large social events
+for fear of public embarrassment only to never learn that such worry is not justified.
 
-In discussion of section, need to also mention other relevant findings. Specifically,
-need to describe the approach-avoidance findings of Bach (2014, 2017) and the
-complex task findings of Sheynin et al., 2015.
+In the laboratory, avoidance has been measured in a variety of ways. In animal
+models of anxiety, anxious rodents are found to exhibit avoidance learning more
+quickly and maintain avoidance for longer \citep{servatius2008}. In studies of
+human avoidance, anxiety has been found to correlate with the maintenance of
+increased distance from threat, such as virtual predators \citep{Bach2014, Bach2017,
+Sheynin2014}.
+
+% Arnaudova2017 makes reference to some spider approach studies that might be
+% interesting to try and track down. Essential idea is that even though
+% participants are totally safe, may not touch the glass. The task is the
+% behavioral avoidance task. (e.g. Siegel & Weinberger, 2012, emotion).
+% the nice part about this is that participants are objectively safe.
+
+There is a long history of models attempting to explain avoidance learning. The
+fundamental challenge is explaining how an action can be reinforced in the
+absence of an outcome. Two-factor theory was the explanation for a long while,
+but suffered from a number of noteworthy criticisms \citep{Krypotos2015}.
+Recent work has revitalized two-factor theory and integrated it with reinforcement
+learning \citep{Moutoussis2008, Maia2010}, and in the process addressing several
+criticisms of two-factor theory. As has been discussed elsewhere \citep{Moutoussis2017},
+such accounts only explain the maintenance of avoidance in short timescales, but
+do not explain why resistant to change and why they do not go away in general.
+
+% Need to go into more detail than this, need to actually make this a coherent
+% explanation, but this can wait for now.
 
@@
 
+In our second result, we demonstrate how a breakdown in control can lead to
+sustained policies of avoidance. We do so in the cliff-walking environment
+\citep{SuttonBarto1998, SuttonBarto2018, Gaskett2003}. Here an agent learns to navigate from
+an initial starting position to an end state. The shortest distance is along a
+cliff edge. If the agent steps off the cliff, it incurs a large penalty. Under
+complete control (objective or perceived), agent in this environment is under no
+threat despite the presence of the cliff's edge. As in the first simulation, this
+is simply because, in a fully controllable world, an agent need not ever interact
+with the aversive state. In fact, an agent can decide to walk straight up to the
+aversive state and, insofar it does not choose to take one step further, will
+never be in any actual danger.
+
+This is not so when deterministic control cannot be assumed. If an agent believes
+it is not in full control of its environment, then it cannot rule out the possibility
+that it may inadvertently encounter the negative state at some point in the future,
+especially if it veers to close to the state. This is true irrespective of whether
+the agent believes it cannot guarantee its own future competency or the benevolence
+of the environment. Importantly, this belief need not match the true environmental
+statistics; belief is enough.
+
+The long-run state values under different beliefs are plotted in Figure 2. As
+expected, under optimistic beliefs (Figure 1b) the entire environment sans negative
+state takes on a positive value (close to the positive state itself). As this belief
+breaks down, we observe a different pattern in the long-run state values. When
+future occupancy of the negative state becomes a non-zero probability, the undesired
+action transition exerts influence on action-outcome calculations, back-propagating,
+and thereby infecting antecedent states. As the agent's belief becomes increasingly
+pessimistic, the more the worst possible action exerts influence and the more
+negative, on average, the state-space grows. Most importantly, because these are
+the long-run state values, the policy is the optimal policy; in other words, this
+is the long-run policy an agent will take.
+
 \subsection{Approach-Avoidance Conflict}
 
-In the third section of the results, we discuss approach-avoidance conflict.
-This covers tasks such as the balloon analog risk task (BART) and the flight
-initiation distance task, which has been found to correlate with anxiety. The
-hallmark of these sorts of tasks is the tradeoff between increased reward vs.
-the possibility of an aversive outcome (e.g. balloon pop, predation). Discuss
-previous findings, i.e. earlier leave times.
+Next, we turn our attention towards a related phenomenon: a bias towards avoidance
+in approach-avoidance conflict. In environments with correlated risk and rewards,
+a bias towards avoidance can ultimately result in fewer expected rewards and worse
+long-run returns. Returning to the example above, an individual with social anxiety
+may forgo social situations to avoid the risk of social embarrassment at the expense
+of friendships, etc. This is another way in which pathological anxiety can be
+disruptive to everyday functioning.
 
-We can show that this is easily explained by our model. A lack of belief of
-future control predicts exiting while things are still good. (We see this in
-Figure 3). The optimal policy in these environments is an earlier leaving time.
-
-This can also be used to explain behavioral inhibition \citep{bach2015, khemka2017}.
-The idea is slightly different than what is advocated. The claim here is that
-in anxious individuals, the Q-values should be closer together, and thus there
-should be longer waiting time. This requires demonstrating that decisions are
-not made in advance; only at the time of decision making. This is something
-I think we will have to argue. (Need to double check the literature here.)
+Experimentally, there is a long history of measuring behavior during
+approach-avoidance conflict in individuals with anxiety (citation, maybe yin-yang anxiety paper).
+One popular paradigm is the balloon analog risk task (BART) \citep{Lejuez2002},
+which has been shown to correlate with anxiety \cite{Maner2007, Giorgetta2012}.
+In the task, a participant pumps a balloon with air. For each pump, the participant
+earns points. If she pumps too much, the balloon pops and she loses all points.
+The participant must decide within each episode when to stop pumping. Leave time
+is the measure of behavior, with earlier leave times in anxiety suggesting a
+greater tendency towards avoidance.
 
@@
   }
-  \caption{\textbf{Flight Initiation Distance}}
+  \caption{\textbf{BART}}
   \par remember to add parameters
 \end{figure}
+
+Our sequential model of decision making easily explains these results. A schematic
+of the task is presented in Figure 3. Under optimistic beliefs, points should
+travel back from the mean. Under pessimistic assumptions, however, different
+degrees of the worst outcome backpropagate. At a certain point, the perceived cost
+of staying outweighs the guaranteed gains of leaving. Thus, patients with anxiety
+leave earlier.
+
+We can explain similar findings with the flight initiation distance \citep{Mobbs2018,
+Mobbs2019}. One thing our model cannot explain is the neural difference between
+fast and slow escape. Indeed, our model cannot capture all but perhaps is relevant
+for certain calculations. One could imagine these beliefs interacting in such a
+way to change the distance calculations that are necessary for mediating between
+passive and active defense responses.
+
+This can also be used to explain behavioral inhibition \citep{bach2015, khemka2017}.
+The idea is slightly different than what is advocated. The claim here is that
+in anxious individuals, the Q-values should be closer together, and thus there
+should be longer waiting time.
 

### 2019-01-28 manuscript/manuscript.tex
overhaul of figures. need to finish few more

@@
 
+% I think you want to explain second order conditioning in PTSD in this context.
+
 \subsection{Avoidance}
@@
 
+Review the original findings of the papers. Aversive pruning as a means of
+limiting computational demands in forward planning in complex sequential
+environments. Remark that this is a sensible theory and we are not attempting to
+argue against it.
+
+Rather we are trying to show that our theory makes similar predictions in the
+sorts of tasks. Combine that with expectations of one's own incompetency, and
+maybe its not all that surprising. Combine this with working memory deficits
+in anxiety, and maybe its doubly not surprising. So... Yeah.
+
+There are ways to teasing this apart.
+
 \begin{figure}
@@
 Discuss previous computational approaches: This has also been previously been
-investigated in previous computational modeling \citep{Huys2009}. There
+investigated in previous computational modeling \citep{HuysDayan2009}. There
 the authors found that beliefs about a lack of controllability, mechanistically

### 2019-01-28 sisyphus/envs/_cliff.py

@@ -4,10 +4,10 @@ from ._base import GraphWorld, grid_to_adj
 class CliffWalking(GraphWorld):
     """Cliff-walking task environment.
     
-    
     References
     ----------
-    1. Gaskett, C. (2003). Reinforcement learning under circumstances beyond its control.
+    1. Sutton, R. S., & Barto, A. G. (1998). Reinforcement learning: An introduction. MIT press.
+    2. Gaskett, C. (2003). Reinforcement learning under circumstances beyond its control.
     """
     
     def __init__(self, epsilon=0):
@@ -35,3 +35,99 @@ class CliffWalking(GraphWorld):
         
     def __repr__(self):
         return '<GraphWorld | Cliff-Walking Task>'
+    
+    def plot_cliff(self, annot=True, grid_color='0.8', cliff_color='0.1', start_color='0.8', 
+                   goal_color='0.8', cbar=False, annot_kws=None, ax=None):
+        """Plot cliff-walking environment.
+
+        Parameters
+        ----------
+        annot : bool
+            Annotate states.
+        grid_color : str
+            Color of grid tiles.
+        cliff_color : str
+            Color of cliff tiles.
+        start_color : str
+            Color of starting tile.
+        goal_color : str
+            Color of goal tile.
+        cbar : bool
+            Whether to draw a colorbar.
+        annot_kws : dict of key, value mappings, optional
+            Keyword arguments for ax.text when annot is True. 
+        ax : matplotlib Axes
+            Axes in which to draw the plot.
+
+        Returns
+        -------
+        ax : matplotlib Axes
+            Axes in which to draw the plot.
+        """
+
+        import matplotlib.pyplot as plt
+        import seaborn as sns
+        from matplotlib.colors import ListedColormap
+        
+        ## Initialize canvas.
+        if ax is None: fig, ax = plt.subplots(1,1,figsize=(5,5))
+        
+        ## Define grid.
+        grid = np.zeros((11, 12))    # Grid titles
+        grid[-1, 1:-1] = 1           # Cliff edge
+        grid[-1,0] = 2               # Start tile
+        grid[-1,-1] = 3              # Goal tile
+
+        ## Define colormap.
+        cmap = ListedColormap([grid_color, cliff_color, start_color, goal_color])
+
+        ## Plot cliff.
+        ax = sns.heatmap(grid, cmap=cmap, cbar=cbar)
[…]

### 2019-01-28 sisyphus/envs/_field.py

@@ -29,3 +29,65 @@ class OpenField(GraphWorld):
         
     def __repr__(self):
         return '<GraphWorld | Open Field Task>'
+    
+    def plot_field(self, reward=10, shock=-10, annot=True, grid_color='0.8',  
+                   reward_color='#f3e1db', shock_color='#1c142a', 
+                   cbar=False, annot_kws=None, ax=None):
+        """Plot cliff-walking environment.
+
+        Parameters
+        ----------
+        reward : float
+            Reward value.
+        shock : float
+            Shock value.
+        annot : bool
+            Annotate states.
+        grid_color : str
+            Color of grid tiles.
+        reward_color : str
+            Color of rewarding tile.
+        shock_color : str
+            Color of punishing tile.
+        cbar : bool
+            Whether to draw a colorbar.
+        annot_kws : dict of key, value mappings, optional
+            Keyword arguments for ax.text when annot is True. 
+        ax : matplotlib Axes
+            Axes in which to draw the plot.
+
+        Returns
+        -------
+        ax : matplotlib Axes
+            Axes in which to draw the plot.
+        """
+        
+        import matplotlib.pyplot as plt
+        import seaborn as sns
+        from matplotlib.colors import ListedColormap
+        
+        ## Initialize canvas.
+        if ax is None: fig, ax = plt.subplots(1,1,figsize=(5,5))
+
+        ## Define grid.
+        grid = np.zeros((11, 11))    # Viable states
+        grid[5,[2,-3]] = [1, 2]      # Reward/shock states
+
+        ## Define colormap.
+        cmap = ListedColormap([grid_color, reward_color, shock_color])
+
+        ## Plot cliff.
+        ax = sns.heatmap(grid, cmap=cmap, cbar=cbar, ax=ax)
+        ax.set(xticklabels=[], yticklabels=[], title='Cliff Walking')  
+
+        ## Add outline.
+        ax.vlines(np.arange(1,11),0,11,lw=0.1)
+        ax.hlines(np.arange(1,11),0,11,lw=0.1)
+
+        ## Annotate.
+        if annot:
+            if annot_kws is None: annot_kws = dict()
+            ax.text(2.5,5.5,reward,ha='center',va='center',**annot_kws)
+            ax.text(8.45,5.5,shock,ha='center',va='center',**annot_kws)
+        
+        return ax

### 2019-01-28 sisyphus/envs/_lh.py

@@ -3,15 +3,15 @@ from ._base import GraphWorld, grid_to_adj
 
 class Helplessness(GraphWorld):
     
-    def __init__(self, outcomes=[10,-10,5,0], epsilon=0):
+    def __init__(self, outcomes=[10,-10], epsilon=0):
         
         ## Define gridworld.
-        self.grid = np.arange(5*21).reshape(5,21)
+        self.grid = np.arange(5*15).reshape(5,15)
         self.shape = self.grid.shape
         
         ## Define start/terminal states.
-        start = 56
-        terminal = np.array([44,48,52,60])
+        start = 44
+        terminal = np.array([30, 37])
 
         ## Define one-step transition matrix.
         T = grid_to_adj(self.grid, terminal)
@@ -27,5 +27,78 @@ class Helplessness(GraphWorld):
         GraphWorld.__init__(self, T, R, start, terminal, epsilon)
         self.R = R
         
+        ## Update start.
+        self.terminal = np.append(self.terminal, start)
+        self.info = self.info.loc[self.info.S != 44]
+        for i in range(4):
+            d = {"S":44, "S'":np.roll([29, 43, 44, 59],i), "T":np.array([1,0,0,0]), "R":np.zeros(4)}
+            self.info = self.info.append(d, ignore_index=True)
+        self.info = self.info.sort_values('S').reset_index(drop=True)
+        
     def __repr__(self):
         return '<GraphWorld | Learned Helplessness>'
+    
+    def plot_lh(self, reward=10, shock=-10, annot=True, grid_color='0.8',  
+                   reward_color='#f3e1db', shock_color='#1c142a', 
+                   cbar=False, annot_kws=None, ax=None):
+        """Plot cliff-walking environment.
+
+        Parameters
+        ----------
+        reward : float
+            Reward value.
+        shock : float
+            Shock value.
+        annot : bool
+            Annotate states.
+        grid_color : str
+            Color of grid tiles.
+        reward_color : str
+            Color of rewarding tile.
+        shock_color : str
+            Color of punishing tile.
+        cbar : bool
+            Whether to draw a colorbar.
+        annot_kws : dict of key, value mappings, optional
+            Keyword arguments for ax.text when annot is True. 
+        ax : matplotlib Axes
+            Axes in which to draw the plot.
+
+        Returns
+        -------
[…]

### 2019-01-28 sisyphus/envs/_tree.py

@@ -70,3 +70,175 @@ class DecisionTree(GraphWorld):
         
     def __repr__(self):
         return '<GraphWorld | Decision Tree>'
+    
+    def _draw_nodes(self, ax, xpos, ypos, s=1000, color=None, cmap=None, vmin=None, vmax=None, 
+                alpha=1.0, linewidth=1.0):
+        """Draw decision tree nodes. See plot_decision tree for details."""
+
+        from matplotlib.cm import get_cmap
+        from matplotlib.colors import ListedColormap, Normalize
+        
+        ## Define colors.
+        if color is None: 
+            colors = np.repeat('#1f77b4', len(xpos))
+        elif isinstance(color, str):
+            colors = np.repeat(color, len(xpos))
+        elif np.issubdtype(np.array(color).dtype, np.number):
+            assert np.equal(len(color), len(xpos))        
+            if not isinstance(cmap, ListedColormap): cmap = get_cmap(cmap)
+            colors = cmap(Normalize(vmin, vmax)(np.array(color)))
+        else:
+            assert np.equal(len(color), len(xpos))        
+            colors = np.copy(color)
+
+        ## Define transparency.
+        if isinstance(alpha, float): 
+            alphas = np.repeat(alpha, len(xpos))
+        else: 
+            assert np.equal(len(alpha), len(xpos))
+            alphas = np.copy(alpha)
+
+        ## Iteratively plot.
+        for i, (x, y, color, alpha) in enumerate(zip(xpos, ypos, colors, alphas)):
+            ax.scatter(x, y, s=s, color='w', alpha=1)
+            ax.scatter(x, y, s=s, color=color, alpha=alpha, linewidth=linewidth, 
+                       edgecolor='k')
+
+        return ax
+
+    def _draw_node_labels(self, ax, xpos, ypos, fontsize=14):
+        """Draw decision tree node labels. See plot_decision tree for details."""
+
+        for i, (x, y) in enumerate(zip(xpos, ypos)):
+            ax.text(x, y, '%0.0f' %(i+1), ha='center', va='center', fontsize=fontsize)
+
+        return ax
+
+    def _draw_edges(self, ax, xpos, ypos, edges, linewidth=1, color='0.5'):
+        """Draw decision tree edges. See plot_decision tree for details."""
+
+        ## Define line widths.
[…]

### 2019-01-28 sisyphus/viz/__init__.py

@@ -1,8 +1,3 @@
 """Visualization functions"""
 
-from ._base import plot_policy
 from ._choice import plot_free_choice
-from ._cliff import plot_cliff
-from ._field import plot_field
-from ._lh import plot_helplessness
-from ._tree import plot_decision_tree

### 2019-01-28 sisyphus/viz/_base.py

@@ -1,43 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.colors import ListedColormap
-
-def plot_policy(ax, gym, pi, color='w', head_width=0.25, head_length=0.25):
-    """Plot agent policy on grid world.
-    
-    Parameters
-    ----------
-    ax : matplotlib Axes
-        Axes in which to draw the plot.
-    gym : GridWorld instance
-        Simulation environment.
-    py : array
-        Agent policy, i.e. ordered visitation of states.
-    color : str (default = white)
-        Color of arrow.
-    head_width : float (default=0.25)
-        Width of the arrow head.
-    head_length : float (default=0.25)
-        Length of the arrow head.
-        
-    Returns
-    -------
-    ax : matplotlib Axes
-        Axes in which to draw the plot.
-    """
-        
-    ## Initialize grid.
-    grid = gym.states.reshape(gym.shape)
-    
-    ## Iteratively plot arrows.
-    for i in range(len(pi)-1):
-        
-        ## Identify S, S' coordinates.
-        y1, x1 = np.where(grid==pi[i])
-        y2, x2 = np.where(grid==pi[i+1])
-
-        ## Plot.
-        ax.arrow(int(x1)+0.5, int(y1)+0.5, 0.5*int(x2-x1), 0.5*int(y2-y1), 
-                 color=color, head_width=head_width, head_length=head_length)
-        
-    return ax

### 2019-01-28 sisyphus/viz/_cliff.py

@@ -1,56 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-from matplotlib.colors import ListedColormap
-
-def plot_cliff(ax, cbar=False, viable='0.8', cliff='0.1', start='#1f77b4', end='#2ca02c'):
-    """Plot cliff-walking environment.
-    
-    Parameters
-    ----------
-    ax : matplotlib Axes
-        Axes in which to draw the plot.
-    cbar : bool
-        Whether to draw a colorbar.
-    viable : str
-        Color of viable squares.
-    cliff : str
-        Color of cliff edge.
-    start : str (deprecated)
-        Color of starting tile.
-    end : str (deprecated)
-        Color of ending tile.
-        
-    Returns
-    -------
-    ax : matplotlib Axes
-        Axes in which to draw the plot.
-    """
-    
-    ## Define grid.
-    grid = np.zeros((11, 12))    # Viable states
-    grid[-1, 1:-1] = 1           # Cliff edge
-    
-    ## Define colormap.
-    cmap = ListedColormap([viable, cliff])
-    
-    ## Alternate coloring.
-    # grid[-1, 0] = 2              # Start
-    # grid[-1,-1] = 3              # Finish
-    # cmap = ListedColormap([viable, cliff, start, end])
-
-    ## Plot cliff.
-    ax = sns.heatmap(grid, cmap=cmap, cbar=cbar)
-    ax.set(xticklabels=[], yticklabels=[], title='Cliff Walking')  
-    
-    ## Add outline.
-    ax.vlines(np.arange(1,12),0,10,lw=0.1)
-    ax.hlines(np.arange(1,11),0,12,lw=0.1)
-    
-    ## Add text.
-    ax.annotate('S', (0, 0), (0.041, 0.04), 'axes fraction', ha='center', 
-                va='center', fontsize=18, weight='semibold')
-    ax.annotate('G', (0, 0), (0.954, 0.04), 'axes fraction', ha='center', 
-                va='center', fontsize=18, weight='semibold')
-    
-    return ax

### 2019-01-28 sisyphus/viz/_field.py

@@ -1,43 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-from matplotlib.colors import ListedColormap
-
-def plot_field(ax, cbar=False, viable='0.8', reward='#f3e1db', shock='#1c142a'):
-    """Plot cliff-walking environment.
-    
-    Parameters
-    ----------
-    ax : matplotlib Axes
-        Axes in which to draw the plot.
-    cbar : bool
-        Whether to draw a colorbar.
-    viable : str
-        Color of viable squares.
-    reward : str
-        Color of rewarding tile.
-    end : str
-        Color of ending tile.
-        
-    Returns
-    -------
-    ax : matplotlib Axes
-        Axes in which to draw the plot.
-    """
-    
-    ## Define grid.
-    grid = np.zeros((11, 11))    # Viable states
-    grid[5,[2,-3]] = [1, 2]      # Reward/shock states
-    
-    ## Define colormap.
-    cmap = ListedColormap([viable, reward, shock])
-
-    ## Plot cliff.
-    ax = sns.heatmap(grid, cmap=cmap, cbar=cbar)
-    ax.set(xticklabels=[], yticklabels=[], title='Cliff Walking')  
-    
-    ## Add outline.
-    ax.vlines(np.arange(1,11),0,11,lw=0.1)
-    ax.hlines(np.arange(1,11),0,11,lw=0.1)
-    
-    return ax

### 2019-01-28 sisyphus/viz/_lh.py

@@ -1,60 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-from numpy import radians as rad
-from matplotlib.patches import Arc, RegularPolygon
-
-def _draw_nodes(ax, s, labels=True):
-    
-    ## Define node positions.
-    xpos = [-1, 0, 1, 1]
-    ypos = [0.0, 0.0, 0.5,-0.5]
-    
-    ## Draw nodes.
-    ax.scatter(xpos, ypos, s)
-    
-    return ax
-
-def _draw_edges(ax):
-
-    ## Draw directed edges.
-    ax.plot([-1.0, 0.0], [0.0, 0.0], color='k', zorder=0)
-    ax.plot([ 0.0, 1.0], [0.0, 0.5], color='k', zorder=0)
-    ax.plot([ 0.0, 1.0], [0.0,-0.5], color='k', zorder=0)
-    
-    ## Draw self-loop.
-    ax = _draw_self_loop(ax, 0.4, -1.25, 0, 70, -155)
-    
-    return ax
-
-def _draw_self_loop(ax, radius, centX, centY, angle_, theta2_, color_='black'):
-    
-    ## Construct line.
-    arc = Arc([centX,centY],radius,radius,angle=angle_,
-          theta1=0,theta2=theta2_,capstyle='round',linestyle='-',lw=1,color=color_)
-    ax.add_patch(arc)
-
-    ## Construct arrow head.
-    endX=centX+(radius/2)*np.cos(rad(theta2_+angle_)) #Do trig to determine end position
-    endY=centY+(radius/2)*np.sin(rad(theta2_+angle_))
-
-    ax.add_patch(                    #Create triangle as arrow head
-        RegularPolygon(
-            (endX, endY),            # (x,y)
-            3,                       # number of vertices
-            radius/9,                # radius
-            rad(angle_+theta2_),     # orientation
-            color=color_
-        )
-    )
-    
-    return ax
-
-def plot_helplessness(ax):
-    
-    _draw_edges(ax)
-    _draw_nodes(ax, 2500)
-
-    ## Clean up.
-    ax.set(xticks=[], yticks=[], ylim=(-1.5, 1.5), xlim=(-1.7, 1.5))
-    sns.despine(top=True, bottom=True, left=True, right=True, ax=ax)

### 2019-01-28 sisyphus/viz/_tree.py

@@ -1,133 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-from matplotlib.cm import get_cmap
-from matplotlib.colors import ListedColormap, Normalize
-
-def _draw_nodes(ax, xpos, ypos, s=1000, color=None, cmap=None, vmin=None, vmax=None, 
-                alpha=1.0, linewidth=1.0):
-    """Draw decision tree nodes. See plot_decision tree for details."""
-    
-    ## Define colors.
-    if color is None: 
-        colors = np.repeat('#1f77b4', len(xpos))
-    elif isinstance(color, str):
-        colors = np.repeat(color, len(xpos))
-    elif np.issubdtype(np.array(color).dtype, np.number):
-        assert np.equal(len(color), len(xpos))        
-        if not isinstance(cmap, ListedColormap): cmap = get_cmap(cmap)
-        colors = cmap(Normalize(vmin, vmax)(np.array(color)))
-    else:
-        assert np.equal(len(color), len(xpos))        
-        colors = np.copy(color)
-        
-    ## Define transparency.
-    if isinstance(alpha, float): 
-        alphas = np.repeat(alpha, len(xpos))
-    else: 
-        assert np.equal(len(alpha), len(xpos))
-        alphas = np.copy(alpha)
-        
-    ## Iteratively plot.
-    for i, (x, y, color, alpha) in enumerate(zip(xpos, ypos, colors, alphas)):
-        ax.scatter(x, y, s=s, color='w', alpha=1)
-        ax.scatter(x, y, s=s, color=color, alpha=alpha, linewidth=linewidth, 
-                   edgecolor='k')
-                
-    return ax
-
-def _draw_node_labels(ax, xpos, ypos, fontsize=14):
-    """Draw decision tree node labels. See plot_decision tree for details."""
-   
-    for i, (x, y) in enumerate(zip(xpos, ypos)):
-        ax.text(x, y, '%0.0f' %(i+1), ha='center', va='center', fontsize=fontsize)
-        
-    return ax
-
-def _draw_edges(ax, xpos, ypos, edges, linewidth=1, color='0.5'):
[…]

### 2019-01-31 manuscript/manuscript.tex
sections 4,5

--- 2019-01-13 manuscript/manuscript.tex
@@
 
-Basically reviewing the arguments of Huys here \citep{Huys2012, Lally2017}.
-Need to get facts straight before going through all of this.
+In this next section, we review the literature on aversive pruning. In large,
+multi-step sequential decision environments, the space of all possible sequences
+of choices grows exponentially larger with increasing sequence length. As such,
+it is infeasible for decision making agents to exhaustively explore all possible
+sequences select that which maximizes reward from the entire set. Heuristics for
+narrowing the search space then are not only plausible but beneficial for a
+decision making agent.
+
+One such heuristic that has been proposed in the literature is aversive pruning
+\citep{Huys2012}. Aversive pruning is defined as a Pavlovian response to encountering
+a large loss in planning such that sequences involving large losses are discarded
+from further evaluation. An example of scenario is presented in Figure 4. The prediction
+in such an environment is that an agent would discard the branch of the decision
+tree involving an immediate large negative loss even though this branch objectively
+contains the reward maximizing (loss minimizing) sequence of choices.
+
+\cite{Huys2012} (and later \cite{Lally2017}) tested this prediction in the
+choice envrionment displayed in Figure 4. Participants received extensive training
+in this decision tree environment (first without outcomes and then later with
+rewards) in order to facilitate planning. In free-planning trials, many participants
+exhibited this Pavlovian bias, rejecting choice sequences which involved the large
+loss even when the sequence was objectively reward maximizing (loss minimizing).
+Importantly \cite{Huys2012} and \cite{Lally2017} found the degree of aversive
+pruning was correlated with depressive and anxiety symptoms, respectively.
+
+The aversive pruning principle makes tremendous sense and draws on a robust literature
+of heuristics and computational shortcuts for circumventing computationally intractable
+cognitive processes. It is important to note, however, that a control interpretation
+also makes similar predictions. Figure 4b-d shows the preferred routes of simulated
+agents in the same environments under increasingly pessimistic expectations of
+future actions. As can be observed, pessimistic agents are similarly likely to avoid
+the branch with the large loss, though for different reasons. Under the pessimistic case,
+the agent is increasingly less confident that the large gains later in
+sequence will be realized; as such, the agent is less confident that the initial
+large loss will be offset. Consequently, pessimistic agents would prefer the branch
+with smaller losses (even if this is objectively suboptimal). This prediction
+is similarly consistent with the finding that avoidance of the initial large loss
+is correlated with anxiety \cite{Lally2017}.
+
+An interpretation like this would be consistent with reduced beliefs in one's
+own competency in such an environment. This is not strictly surprising. Planning
+multi-choice sequences even in relatively simple environments like those used
+in these experiments still ostensibly require working memory ability to keep in
+mind the possible consequences along each branch. Insofar that working memory
+ability is disrupted in anxiety disorders \citep{Moran2016}, a pessimistic belief
+may be justified and possibly optimal.
+
+It is also worht noting that the two theories can be easily disentangled with a
+simple manipulation of the choice environment. The two theories predict
+different choices when the large loss comes later in sequence. For aversive pruning,
+late large losses should have little effect on choice behavior. For pessimistic
+learning, however, the value of large losses should back-propagate and contaminate
+its associated branch. Future experiments can tease these apart as tests of the theory.
 
@@
 \subsection{Free Choice Premium}
+
+If an agent is confident that their compentency in decision-making in intact,
+then it follows that situations in which an individual is able to make a choice
+should be preferable to situations in which an individual is not. At worst,
+an agent is no better off than if they had not made a choice; at best, an agent
+is better able to steer themselves toward desirable outcomes. Such is the logic
+that underlies an increasingly robust literature suggesting choice is inherently
+valuable \citep{Leotti2010}.
+
+This free choice premium has been demonstrated multiple times in humans \citep{Suzuki1997,Leotti2011,Leotti2014,Cockburn2014} using a variety of decision
+paradigms. For the present purposes, we will describe the experiments presented in
+\citep{Leotti2011,Leotti2014}. In these experiments, participants complete a
+two-stage decision making task (Figure 5). In the first stage, participants are
+make a choice between two cues: one leading to free choice and the other leading
+to a fixed choice. In the second stage, participants are able to choose between
+a second set of cues (free choice) or are forced to choose a cue (fixed choice).
+Unbeknownst to participants, all cues yield reward following the same outcome
+distribution; thus, no cue is advantageous in the long-run. The free choice
+premium is measured as a preference for the cue leading to free choice. \cite{Leotti2011,Leotti2014}
+find participants still prefer the free choice cue, despite this conferring no
+objective benefit to participants.
+
+Insofar that this effect is contingent on the belief of one's own competency,
+it stands to reason that this free choice premium should be absent in individuals
+with anxiety and anxious control beliefs. We demonstrate this effect in Figure 5.
+In each trial, the reward outcomes are equal for all options (-1, 1). In individuals
+with increasingly pessimistic beliefs, the worst outcome in free choice is increasingly
+back-propagated to the free-choice cue, thereby making it less attractive than
+the fixed choice cue. Hence, pessimistic learners, on average, do not exhibit
+this choice bias.
+
+To our knowledge this prediction has not been empirically tested.
 

### 2019-01-31 manuscript/manuscript.tex
section 6

@@
 
+In this final section, we want to discuss the relationship between anxiety and
+depression. Anxiety and depression are highly comorbid, with roughly 45% of individuals
+with lifetime depression diagnosis also diagnosed with at least one anxiety disorder
+\citep{kessler2015}. This fact has long been recognized in the clinical literature,
+and prominent theories of depression have suggested that anxious control is a necessary
+precursor to at least some types of depression \citep{alloy1990}. Recent large-scale
+epidemiological studies have provided  some empirical support for this claim, finding that
+anxiety symptoms precede and predict the onset of depression \citep{mathew2011, jacobson2014,
+kessler2015} (though also see \cite{jacobson2017, plana2019}).
+
+An emerging idea in this literature is that pathological avoidance as a result
+of anxiety mediates the transition to depression, perhaps by reduce the likelihood
+of experiencing positive events and activities \citep{moitra2008, jacobson2014}.
+This finding is perfectly in line with the model presented so far. A sketch of
+how this is possible is shown in Figure 6. Under unperturbed learning, avoidable
+threat does not impede behavior. Under pessimistic learning, threat can block
+future reward. This avoidance can manifest as a failure of behavioral initiation,
+or a lack of overall mobility (here represented by an agent self-transitioning
+at the starting state). Thus, the model can accomodate the deficits in positive
+affect observed in depression as a function of behavioral avoidance, as predicted
+by these epidemiological studies.
+
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/06_lh.png}}%
+  }
+  \caption{\textbf{Free Choice Premium}}
+  \par remember to add parameters
+\end{figure}
+
 \section{Discussion}

### 2019-02-02 manuscript/manuscript.tex
minor discussion updates

@@
 or a lack of overall mobility (here represented by an agent self-transitioning
-at the starting state). Thus, the model can accomodate the deficits in positive
+at the starting state). Thus, the model can accommodate the deficits in positive
 affect observed in depression as a function of behavioral avoidance, as predicted
@@
 
-In this article, we showed...
+All of the anxiety disorders share core symptoms including pessimistic inference,
+second-order conditioning, and sustained avoidance. In this article, we offered
+a simple computational account of how these aberrations may arise and be maintained.
+Following the literature in Markov decision processes, we showed how a simple
+breakdown in expectations about the likelihood of encountering future negative
+outcomes (as a result of an unpredictable environment or unreliable policy) can
+effectively backpropagate negative expectations through an agent's internal value
+map. The result of which yields expectations and behaviors resembling that observed
+in pathological anxiety. Importantly, we also showed that such a model can account
+for aberrant behavior previously observed in individuals with anxiety in several
+assays of psychiatric behavior. By no means a complete theory of anxiety, this
+account helps to at least tie together several core symptoms of anxiety in one
+account.
 
-Discuss previous computational approaches: This has also been previously been
-investigated in previous computational modeling \citep{HuysDayan2009}. There
-the authors found that beliefs about a lack of controllability, mechanistically
-modeled as corrupted priors about action-outcome contingencies, found deficient
-learning in novel environments. The present results are no doubt related, but
-address a different environment and a different behavior.
+This computational account draws upon a longstanding recognition of the importance
+of perceived control to the development and maintenance of anxiety disorders. Several
+prominent theories of anxiety in the clinical anxiety focus on the importance of
+a lack of perceived control, including helplessness-hopelessness theories, perceived
+efficacy, and the triple vulnerability model. It is worth noting that that these
+theories differ, among other things, in where a lack of control arises. For learned
+helplessness, an agent believes to occupy an uncontrollable world; in other words,
+there is no reliable mapping between action and outcome. For self-efficacy theory,
+an agent believes to be limited or incapable of exerting beneficial influence on
+the environment, though this is otherwise possible. Note that both can easily be
+modeled in the current approach, either in the state transition function or in
+future policy function. Insofar that they are multiplicative, they should result
+in the same effect. Future work may tease these apart, as they may be interesting
+for guiding treatment.
+
+It is worth noting that we are not the first to attempt to formalize a theory
+of control in MDP. \cite{HuysDayan2009} provided a first computational account
+of learned helplessness through simple models of action-outcome contingencies.
+\cite{HuysDayan2009} found that training an agent first in an unreliable environment
+led to later impaired learning, akin to early learned helplessness experiments,
+by virtue of the prior of controllability. This provides a nice demonstration
+of learning impairment of action-outcome contingencies. Importantly, our accounts
+in at least three respects. We handle sequential decision environments, which is
+important as it is in sequential decision environments where the consequences
+of pessimistic inference become most pronounced. Second, our account claims
+that the worst outcomes are believed, wherein their account focuses on randomness.
+This seems closer to the symptoms described in anxiety. Finally, our account
+provides insights into maintenance. With an otherwise unmodified Bayesian learner,
+their agents should eventually learn and be ok.
 
@@
 
-\section{References}
-
 \bibliographystyle{vancouver-authoryear}

### 2019-02-03 manuscript/manuscript.tex
many methods updates

@@
 Ida Momennejad \\
-columbia University\\
+Columbia University\\
 New York, NY 10027 \\
@@
 \section{Introduction}
-Many psychiatric disorders involve distortions in models of the external world
-and, correspondingly, aberrations of learning and decision making. As a normative
-framework, decision theory allows us to describe and pose quantita­tive questions
-about optimal and approximately optimal choice behavior \citep{DayanDaw2008} and
-is, therefore, a critical tool for modeling, understanding, and predicting the
-psychological and neurobiological mechanisms responsible for the abnormalities
-observed in psychiatric behavior \citep{HuysDawDayan2015}. Many recent investigations
-employing this approach have provided important insights into psychiatric symptoms,
-including anhedonia in major depression \citep{Rutledge2017}, mood instability
-in bipolar disorder \citep{EldarNiv2015, EldarDolanNiv2016},  habitual decision
-making in disorders of compulsivity \citep{gillan2016}, and hallucinations in
-schizophrenia \citep{powers2017, corlett2018}.
 
-Anxiety disorders (generalized anxiety disorder, social anxiety disorder, panic
-disorder, obsessive-compulsive disorder) represent the most common family of
-psychiatric disorders (citation). Despite the differences in the content of worry,
-all anxiety disorders share two core features: pessimistic inference and excessive
-avoidance. Pessimistic inference describes a tendency to appraise threatening
-situations as more frequent and/or more severe than they objectively are. This has
-been reliably observed experimentally, as in fear conditioning wherein individuals
-with anxiety exhibit exaggerated fear towards safe cues (CS-) and extinguished
-threatening cues \citep{lissek2005, Duits2015}, and clinically, through self-report
-(pick 3 citaitons). Excessive avoidance describes the corresponding tendency for
-individuals with anxiety to consistently flee or avoid threatening situations.
-Avoidance is often chosen even at the expense of potential reward (approch-avoidance
-conflict), and is frequently a major impediment to recovery \citep{Arnaudova2017}.
+Central to many psychiatric disorders is distortion to representations of the
+external world and, correspondingly, pronounced changes to learning and decision
+making. As a normative framework, decision theory allows us to describe and pose
+quantita­tive questions about optimal choice behavior \citep{DayanDaw2008} and is,
+therefore, a critical tool for modeling, understanding, and predicting the psychological
+and neurobiological mechanisms underlying psychiatric behavior \citep{maia2011, HuysDawDayan2015}.
+Recent investigations employing this framework have provided novel and important
+insights into psychiatric symptoms and disorders, including anhedonia in major depression \citep{Rutledge2017},
+mood instability in bipolar disorder \citep{EldarNiv2015, EldarDolanNiv2016},
+habitual behavior in obsessive compulsive and addictive disorders \citep{gillan2016}, and
+hallucinations in schizophrenia \citep{powers2017, corlett2018}.
 
-Recent studies have attempted to provide some computational models of this behavior
-\citep{Moutoussis2008, Maia2010}. For example, \citep{Maia2010} was able to show
-how reinforcement learning could explain sustained avoidance in the absence of
-repeated reinforcement, observed in rodents \citep{servatius2008} and something
-not explained by two-factor theory \citep{Krypotos2015}. As has been discussed
-elsewhere \citep{Moutoussis2017}, such accounts only explain the maintenance of
-avoidance in short timescales, but do not explain why resistant to change and
-why they do not go away in general. Such theories also do not explain the magnitude
-of pessimistic inference, or the fear generalization. A broader account is needed.
+The family of anxiety disorders (which includes generalized anxiety disorder, social
+anxiety disorder, panic disorder, and specific phobia) are characterized by excessive
+fear and worry about future negative outcomes, and represent the most common
+group of psychiatric disorders \citep{kessler2009}. Though the disorders typically
+differ in the types of objects or situations that induce fear or anxiety, all share
+a set of core disturbances to cognition and behavior. First, the anxiety disorders
+involve pessimistic inference, or a tendency to appraise threatening objects or
+situations as more likely and/or more severe than they objectively are. Clinical
+anxiety is also associated with second-order conditioning, or learning to fear
+neutral cues indirectly associated with threat cues. Finally, anxiety is associated
+with sustained avoidance, or the tendency for individuals to flee or avoid threatening
+situations. As an example, an individual is mugged in an alleyway. In a pathological
+response, this individual may exhibit persistent worry that it will happen again
+(pessimistic inference), increased anxiety towards the street or time of day
+at which it occurred (second-order conditioning), and increased effort to stay
+away from any cue which reminds them of the event (avoidance).
 
-In this article, we put forward a decision theoretic explanation of the core
-symptoms of anxiety. In sequential decision environments, we show how a loss of
-optimistic inference results in pessimistic inference, chronic avoidance, and
-generalization of fear.
+Several studies have attempted to provide some insight into why some of this
+may be in anxiety disorders \cite{Maia2010, norbury2018}, though no account has
+tied these symptoms together. For example, \citep{Maia2010} provided a simple
+account of how simple actor-critic reinforcement learning models could account
+for sustained avoidance in rodent shuttle-box experiments \citep{servatius2008}.
+As has been discussed elsewhere \citep{Moutoussis2017}, such an account fails to account for why avoidance
+is maintained even with non-threatening outcomes and do not account for why exposure
+therapy only works for some individuals. Other accounts have attempted to explain
+pessimistic inference (cite learning rate), but do not mention where from.
+
+In the remainder of this article, we present a simple decision theoretic account
+of how the core symptoms of anxiety may arise and be maintained. Specifically, we
+demonstrate that the loss of optimistic inference in the form of reduced expectations
+of future reward can yield negatively biased estimates of value in sequential
+decision environments. In such an event, pessimistic inference, second-order
+conditioning, and chronic avoidance naturally follow.
 
@@
 
-Here, we briefly review the formalism of Markov decision processes (MDPs), which
-provide the foundation for our results. For more complete treatments of the
-subject, see Sutton \& Barto (1998, 2018) and Bertsekas \& Tsitsiklis (1996).
+Before moving forward, we first briefly review the formalism of Markov decision
+processes, which provide the framework for our results. For complete treatments,
+see \cite{SuttonBarto1998, SuttonBarto2018} and \cite{bertsekas2005}.
 
 A MDP is defined by a set of states, $S$, a set of actions, $A$, a reward
-function, $R(s,a)$ over state/action pairs, and a state transition distribution,
-$P(s'|s,a)$, where $a$ denotes a chosen action. States and rewards occur
-sequentially according to these one-step functions, driven by a series of
-actions; the goal is to learn to choose a probabilistic policy over actions,
-denoted by $\pi$, that maximizes the value function, $V^\pi(s)$, defined as the
-expected cumulative discounted reward:
+function defined over state/action pairs, $R(s,a)$, and a state transition distribution,
+$P(s'|s,a)$, where $a$ denotes a chosen action. In a MDP, states and rewards occur
+sequentially according to the one-step transition structure where transitions follow
+actions. The goal for an agent in a MDP is to learn to choose a policy, $\pi$, or
+mapping of state to action, that maximizes long-run expected reward. Here we define
+the value of a state under a policy, $V^\pi(s)$, as the expected cumulative discounted reward:
 
@@
 
-Here, $\gamma$ is a parameter controlling temporal discounting. The value function
-can also be defined recursively as the sum of the immediate reward of the action
-chosen in that state, $R(s, a)$, and the value of its successor state $s’$,
-averaged over possible actions, $a$, and transitions that would occur if the agent
-chose according to $\pi$:
+where $\gamma$ is the temporal discounting parameter, controlling the weighting
+of distant future rewards. The value function can be defined recursively as the
+sum of the immediate reward following an action chosen in a state, $R(s, a)$, and
+the value of its successor state $s’$, averaged over potential future actions, $a$,
+and transitions according to the current policy $\pi$:
 
-$$ V^\pi(s) = \sum_a \pi(a|s) \left[ R(s,a) + \sum_{s'} P(s'|s,a) \gamma V^\pi (S') \right] $$
+$$ V^\pi(s) = \sum_a \pi(a|s) \sum_{s',r}p(s',r|s,a) \left[ r + \gamma v(s') \right] $$
 
-The value function under the optimal policy is given by:
+Under the optimal policy, the value function is given by:
 
@@
 
-$$ = \max_a \left[ R(s,a) + \sum_{s'} P(s'|s,a) \gamma V^\pi (S') \right] $$
+$$ = \max_a \sum_{s',r}p(s',r|s,a) \left[ r + \gamma v_*(s') \right] $$
 
-Knowledge of the value function can help to guide choices. For instance, we can
-define the state-action value function as the value of choosing action $a$ and
-following $\pi$ thereafter:
+Using the formalisms above, the value of one particular action in a state under
+the optimal policy is defined as:
 
@@
 
-$$ = R(s,a) + \sum_{s'} P(s'|s,a) \gamma V^\pi (S') $$
+$$ = \sum_{s',r}p(s',r|s,a) \left[ r + \gamma v_*(s') \right] $$
 
-Then at any state one could choose the action that maximizes $Q^\pi(s,a)$. Formally
-this defines a new policy, which is as good or better than the baseline policy
-$\pi$; analogously, Eq 2 can be used to define the optimal state-action value
-function, the maximization of which selects optimal actions. Note that it is
-possible to write a recursive definition for $Q$ in the same manner as Eq 1, and
-work directly with the state-action values, rather than deriving them indirectly
-from $V$.
+Insofar that the value of a state under the optimal policy is defined as the best
+action under that state, we can re-express the above equivalently as:
+
+$$ q_*(s,a) = \sum_{s',r}p(s',r|s,a) \left[ r + \gamma \max_{a'} q_*(s',a') \right] $$
+
+Thus in a conventional MDP, the value of states in the environment under the optimal
+policy assumes the agent is able to act according to the optimal policy in the future.
 
 \subsection{Robust Control}
+In the previous section, we observed that the value of a state and its associated
+state-action pairs are a function of the environment transition distribution and
+the best future action. In standard MDP problems, it is typically assumed that the
+transition distribution is known and accurate, and that the agent is consistently
+able to operate under the optimal policy in the future. In this section, we introduce
+a formalism deviating from these standard assumptions.
 
-To create the symptoms of anxiety, we need a loss of control. This can come in
-two forms. First is through the state transition distribution, $P(s'|s,a)$.
-Normally, we assume the world is deterministic; in other words, an agent's chosen
-action, $a$ leads the agent to the desired successor state, $s'$. In a stochastic
-environment, however, an agent may instead transition to an undesired state, $s^*$.
+Central to many theories of clinical anxiety is a perceived loss of control (see
+discussion for elaboration). In the MDP framework described above, this may come
+in one of several forms. First, an agent may perceive the transition structure
+of the environment to be less reliable than it actually is. Put another way, an
+agent may have a biased expectation that, despite choosing an action $a$ with the
+intent of transition to state $s'$, the agent may instead transition to an alternate,
+undesired state $s^$. In the case of clinical anxiety, the expectation may be more
+dire, that the environment is "adversarial" is likely to transition an agent, not
+randomly, but to the worse states.
 
-Importantly, in the clinical literature, pathological anxiety is also associated
-with a lack of perceived control (Gallagher et al., 2014). Studies suggest that
-this belief mediates anxiety (ref?). This can come in whatever form.
+Second, an agent may doubt its own agentive control. An agent may expect to be
+able to choose the best possible action for itself in the present, but may not be
+confident it will still be able to decide optimally for itself in the future. In
+either event, following the value computations defined above, we should expect
+the value estimates for such an agent to diminish. If an agent believes itself
+or the environment to be unreliable, then the agent cannot be optimistic about
+its future and should instead adopt a different policy.
 
-Alternately, an agent may choose optimally in the present but may be uncertain of
-its choice control in the future. In such a case, we deviate from the optimal
-policy in the future. We can adopt a different function. There are a number to
-choose from \citep{Garcia2015}. Here we adopt the beta-pessimism criterion
-\citep{Gaskett2003}:
+These sorts of problems have been explored in the safe reinforcement learning
+and robust control literature \citep{Garcia2015}. There, we are concerned
+with programming agents that identify the optimal policy under pessimistic
+assumptions. A number of approaches have been proposed in that literature. Here,
+we focus on one particular simple implemented proposed by \cite{Gaskett2003}, namely
+*w-pessimism*:
 
-$$ V(s') = \beta \max_{a'} Q(s',a') + (1 - \beta) \min_{a'} Q(s',a') $$
+$$ V(s') = w \max_{a'} q(s',a') + (1 - w) \min_{a'} q(s',a') $$
 
-where $\beta$ controls the degree of pessimism. When $\beta = 1$, an agent expects
-complete control to choose the best action (best-case criterion); when $\beta = 0$,
-an agent expects the  worst future case; and when $\beta = 0.5$, an agent expects
-randomness. (We are not committed to this particular instantiation; it is simply
-a convenient notation. Other parameterizations work equally well, see appendix.)
+such that the state value under the optimal policy is defined as:
+
+$$ V(s) = \max_a \sum_{s',r}p(s',r|s,a) \left[ r + \gamma \left( w \max_{a'} q(s',a') + (1 - w) \min_{a'} q(s',a') \right) \right] $$
+
+In this formalism, $w$ controls the degree of pessimism. When $\w = 1$, an agent
+is maximally optimistic and fully expects to act according to the best possible actions
+in the future. When $w = 0$, the agent is maximally pessimistic and fully expects
+to act according to the worst possible actions in the future. When $w = 0.5$,
+the weighs the best and worst possible future actions equally.
+
+We note that we are not committed to this particular implementation of a pessimistic
+agent. As mentioned above, an alternative parameterization would have been to
+program an agent with a belief in an adversarial environment, preferentially
+transitioning it to worse states. Insofar that these two terms are multiplicative
+in value calculations, either implementation should yield similar results. Similarly,
+the w-pessimism function could easily be replaced with other functions (e.g. softmax
+with negative inverse temperature). This formalism is simply convenient and is
+useful for demonstrating the points we demonstrate below. It is not a substantive
+claim about anxiety.
 
@@
 Mitte2007, Miranda2007}.
+
+In self-report,
+individuals with anxiety rate hypothetical future negative outcomes as
+being more likely to occur and worse to experience than non-anxious counterparts
+\citep{ButlerMathews1983, ButlerMathews1987, Foa1996, MacLeod1996, MacLeod1997,
+Luten1997, Stober1997, Borkovec1999, Maner2006, Mitte2007, Miranda2007}. In the
+laboratory, individuals with anxiety demonstrate excessive fear in fear conditioning
+paradigms. For example, anxiety is correlated with exaggerated fear towards
+non-threatening cues during conditioning and extinguished cues during extinction
+\citep{lissek2005, Duits2015}.
 
@@
 for fear of public embarrassment only to never learn that such worry is not justified.
+
+Avoidance is often chosen even at the expense of potential reward (approch-avoidance
+conflict), and is frequently a major impediment to recovery \citep{Arnaudova2017}.
 

### 2019-02-03 manuscript/manuscript.tex
making progress, making progress

@@
-\documentclass[11pt]{article} % For LaTeX2e
+environment\documentclass[11pt]{article} % For LaTeX2e
 \usepackage{manuscript, palatino}
@@
 
-\subsection{Robust Control}
+\subsection{Pessimistic Learning}
+
 In the previous section, we observed that the value of a state and its associated
@@
 
-In some instances, we solve for Q-values and state values directly through value
-iteration. This is a form of Dynamic Programming. In other instances, we estimate
-these terms through temporal difference learning. We use beta-pessimism (betamax)
-learning for both. We use the former to solve for final (infinite horizon) values
-and the latter to identify finite horizon values. We explain when we use both.
+In the following sections, we present the results of simulations of pessimistic
+learning in a variety of MDP environments. In all simulations, the state-action
+values under the optimal policy, $Q_*(s,a)$, were solved for directly through
+Q-value iteration \citep{SuttonBarto1998,SuttonBarto1998,bertsekas2005}. The
+details of this algorithm are presented in the appendix. Model parameters are
+reported for each environment in their corresponding sections. All simulations
+were carried out using the scientific python distribution and are publicly available
+at on Github (https://github.com/szorowi1/SecretFunTimes).
 
@@
 \subsection{Pessimistic Inference}
-
-One symptom central to all of the anxiety disorders is pessimistic inference,
-or a tendency to appraise the likelihood and/or severity of threatening situations
-out of proportion to their actual environmental contingencies \citep{dsm5, BeckClark1997,
-ClarkBeck2011}. For example, an individual with specific phobia may vastly exaggerate
-the probability of encountering their fear (e.g. a venomous spider) in their
-everyday lives. The first studies of pessimistic inference in clinically anxious
-populations involved individuals judging the perceived likelihood and severity
-of different aversive situations, with the general finding that anxious individuals
-rated these as more likely than non-clinical individuals \citep{ButlerMathews1983,
-ButlerMathews1987, Foa1996, MacLeod1996, MacLeod1997, Luten1997, Stober1997,
-Borkovec1999}; a finding that has been replicated more recently \citep{Maner2006,
-Mitte2007, Miranda2007}.
-
-In self-report,
-individuals with anxiety rate hypothetical future negative outcomes as
-being more likely to occur and worse to experience than non-anxious counterparts
-\citep{ButlerMathews1983, ButlerMathews1987, Foa1996, MacLeod1996, MacLeod1997,
-Luten1997, Stober1997, Borkovec1999, Maner2006, Mitte2007, Miranda2007}. In the
-laboratory, individuals with anxiety demonstrate excessive fear in fear conditioning
-paradigms. For example, anxiety is correlated with exaggerated fear towards
-non-threatening cues during conditioning and extinguished cues during extinction
-\citep{lissek2005, Duits2015}.
-
-Experimentally, aberrations in threat appraisal have been repeatedly observed
-in a litany of associative learning studies. Common to all of these studies,
-participants learn that a particular stimulus (CS+) predicts the occurrence
-of a negative outcome (e.g. shock, loud noise, aversive image). In simple conditioning,
-there is some evidence to suggest that individuals with anxiety exhibit elevated
-levels of response to threatening stimuli \citep{lissek2005, MinekaOehlberg2008,
-Duits2015}. Importantly, conditioning studies also suggest impaired safety learning
-in anxiety, wherein a stimulus never paired with an aversive outcome (CS-) is also
-feared \citep{Duits2015}. Other notable findings include impaired blocking,
-conditioned discrimination (AX+/BX-) (citations), and conditioned inhibition
-(A+, AB-) (citations).
-
-% I think when you come back to edit this, you need to highlight these findings
-% more in the context of what you are about to explain. Pessimistic inference
-% is both making bad things worse and, most importantly, that non-bad things
-% become bad. Although one question is how much to include the other less
-% studied paradigms
 
@@
 
-In our first result, we demonstrate one simple mechanism by which pessimistic
-inference and exaggerated threat appraisals may arise. In the *open field* environment,
-an agent is placed in a large grid world entirely empty except for two states,
-leading to positive and negative outcomes, respectively. Under complete control
-(objective or perceived), an agent in this environment in under no threat despite
-the presence of negative states. This is simply because, in a fully controllable
-world, an agent need not ever interact with the aversive state. In fact, an agent
-can decide to walk straight up to the aversive state and, insofar it does not
-choose to take one step further, will never be in any actual danger.
+In this first section, we demonstrate how learning under an assumption of a lack
+of future control naturally leads to pessimistic inference and second-order conditioning.
+A symptom central to all of the anxiety disorders is pessimistic inference,
+or a tendency to appraise the likelihood and/or severity of threatening situations
+out of proportion to their actual environmental contingencies \citep{dsm5, BeckClark1997,
+ClarkBeck2011}. For example, an individual with specific phobia may vastly exaggerate
+the probability of encountering their fear (e.g. a venomous spider) in their
+everyday lives. Anxious individuals will also tend to associate fear with cues
+tangentially related to signals of threat. For example, the same individual may
+begin to fear locations where they may encounter spider webs (e.g. basements).
+
+Pessimistic inference in anxiety is well-documented. In studies of self-reported beliefs,
+individuals with anxiety rate
+hypothetical future negative outcomes as being more likely to occur and worse to
+experience than non-anxious counterparts \citep{ButlerMathews1983, ButlerMathews1987,
+Foa1996, MacLeod1996, MacLeod1997, Luten1997, Stober1997, Borkovec1999, Maner2006,
+Mitte2007, Miranda2007}. In the laboratory, exaggerated threat appraisal in anxiety
+has traditionally been measured through fear conditioning. The results of these
+studies demonstrate that anxious individuals exhibit greater fear towards threatening
+cues, non-threatening cues, and extinguished threatening cues as compared to non-anxious
+counterparts \citep{lissek2005, MinekaOehlberg2008, Duits2015}. Importantly, threatening
+cues can become associated with secondary neutral cues, allowing negative value to
+spread citep{wessa2007}. In sum, anxiety is associated with inflated estimates of
+threat, both likelihood and value, and these estimates can spread from the source
+to distantly associated cues.
+
+In our first simulation, we demonstrate how this may come about. In the open field
+environment (Figure 1a), an agent is free to navigate a large grid world devoid
+of value except for two states: one rewarding and one aversive red. In this environment,
+the one-step transition structure is deterministic such that the environment
+fully respects the agent's choice. As such, the agent is under no threat despite
+the presence of a harmful state. Insofar that the agent never acts under a
+masochistic policy and intentionally approaches harm, the agent need not ever
+encounter the harmful state.
+
+The long-run value estimates learned by an optimistic agent (i.e. an agent expecting
+to be able to follow a reward-maximizing policy now and in the future) reflect this
+(Figure 1b). For an optimistic agent, all states in the open field environment,
+sans the harmful state itself, take on a positive value. This is unsurprising
+insofar that, except for the harmful state, the agent will always encounter
+a reward in the long-run no matter its starting position in the field.
 

### 2019-02-03 manuscript/manuscript.tex
rewrote results 3.1

@@
-environment\documentclass[11pt]{article} % For LaTeX2e
+\documentclass[11pt]{article} % For LaTeX2e
 \usepackage{manuscript, palatino}
@@
 reported for each environment in their corresponding sections. All simulations
-were carried out using the scientific python distribution and are publicly available
-at on Github (https://github.com/szorowi1/SecretFunTimes).
+were carried out with the python programming language and are publicly available
+on Github (https://github.com/szorowi1/SecretFunTimes).
 
@@
   }
-  \caption{\textbf{Open Field}}
-  \par Current parameters: Reward = 10, Shock = -10, gamma = 0.95.
+  \par Figure 1: The open field environment is a large deterministic gridworld sans one rewarding
+  (blue) and aversive (red) state (A). Under optimistic learning conditions (B), all states except
+  for the harmful one become positive. Under pessimistic conditions (C), negative value spreads
+  from the source to its antecedent states. The extent of the spread is worse for increased
+  pessimism (D). (Parameters: $\gamma = 0.95$)
 \end{figure}
@@
 
-This is not so when deterministic control cannot be assumed. If an agent believes
-it is not in full control of its environment, then it cannot rule out the possibility
-that it may inadvertently encounter the negative state at some point in the future,
-especially if it veers to close to the state. This is true irrespective of whether
-the agent believes it cannot guarantee its own future competency or the benevolence
-of the environment. Importantly, this belief need not match the true environmental
-statistics; belief is enough.
+The same cannot be said for a the pessimistic agent. If an agent doubts its future
+ability to act in according with the optimal policy, then it cannot rule out that,
+in the future, it may inadvertently encounter the harmful state. This is more so
+likely if the agent is already proximal to the harmful state. Under this pessimistic
+belief, the possibility of future harm is no longer zero, as it is for the optimistic
+agent, and thus negative value can propagate from the states antecedent to harm,
+back to their respective antecedents and so on. The extent of this spread will depend
+on the strength of the pessimistic belief (Figures 1c, 1d). With increasing pessimism,
+harmful outcomes are assigned greater credence and thus exert greater and wider
+influence on antecedent states.
 
-The long-run state values under different beliefs are plotted in Figure 1. As
-expected, under optimistic beliefs (Figure 1b) the entire environment sans negative
-state takes on a positive value (close to the positive state itself). As this belief
-breaks down, we observe a different pattern in the long-run state values. When
-future occupancy of the negative state becomes a non-zero probability, the undesired
-action transition exerts influence on action-outcome calculations, back-propagating,
-and thereby infecting antecedent states. As the agent's belief becomes increasingly
-pessimistic, the more the worst possible action exerts influence and the more
-negative, on average, the state-space grows.
-
-In the breakdown of perceived control then, negative states can grow worse (i.e.
-more predictive of future negative outcome) and otherwise positive states can too.
-This occurs as a simple result that if control not guaranteed, cannot guaranteed
-bad things wont happen.
-
-% I think you want to explain second order conditioning in PTSD in this context.
+In sequential decision environments then, a perceived lack of self-efficacy can
+result in the propagation of negative value from potential future harms. Such a
+pessimistic agent may correspondingly report an increased likelihood of threat
+(i.e. a greater number of threatening states), increased severity of threat (i.e.
+states with more negative value), and fear of states antecedent to the fear source.
+As such, pessimistic learning can account for two of the hallmark symptoms of
+the anxiety disorders.
 
@@
 \section{Appendix}
-\subsection{Algorithms}
 
@@
 
-\begin{algorithm}
-  \caption{Betamax Temporal Difference Learning}
-
-  \State Algorithm parameters: step size $\eta \in (0, 1)$, small $\epsilon > 0$
-  \State Initialize $Q(s,a)$, for all $s \in S$, $a \in A$ arbitrarily, except that $Q(terminal,\cdot) = 0$
-  \State
-  \Loop \ for each episode
-    \State Initialize $S$
-    \While{$S \notin S(terminal)$}
-      \State Choose $A$ from $S$ using policy derived from $Q$
-      \State Take action $A$, observe $R, S$
-      \State $Q(S, A) \leftarrow Q(S, A) + \eta \left[ R + \gamma \max_a Q(S',a) − Q(S, A) \right] $
-      \State $S \leftarrow S'$
-    \EndWhile
-  \EndLoop
-
-\end{algorithm}
-
 \end{document}

### 2019-02-03 manuscript/manuscript.tex
figure updates for manuscript

@@
   }
-  \par Figure 1: The open field environment is a large deterministic gridworld sans one rewarding
-  (blue) and aversive (red) state (A). Under optimistic learning conditions (B), all states except
-  for the harmful one become positive. Under pessimistic conditions (C), negative value spreads
-  from the source to its antecedent states. The extent of the spread is worse for increased
-  pessimism (D). (Parameters: $\gamma = 0.95$)
+  \par \textbf{Figure 1:} The open field is a deterministic gridworld with only one rewarding
+  (blue) and aversive (red) state (A). For an optimistic agent ($w=1$), all states (other than
+  the harmful state) take on positive value (B). For a pessimistic agent ($w=0.5$), negative value spreads
+  from the source to antecedent states (C). With increasing pessimism ($w=0$), the extent of the spread
+  grows worse (D). (Parameters: $\gamma = 0.95$)
 \end{figure}
@@
 cues can become associated with secondary neutral cues, allowing negative value to
-spread citep{wessa2007}. In sum, anxiety is associated with inflated estimates of
+spread \citep{wessa2007}. In sum, anxiety is associated with inflated estimates of
 threat, both likelihood and value, and these estimates can spread from the source
@@
 
-The same cannot be said for a the pessimistic agent. If an agent doubts its future
+The same cannot be said for a pessimistic agent. If an agent doubts its future
 ability to act in according with the optimal policy, then it cannot rule out that,
@@
 
-A second symptom central to all anxiety disorders is avoidance \citep{dsm5,
-Krypotos2015, Arnaudova2017}, or actions taken to increase or maintain the temporal,
-physical and/or mental distance between an individual and a threatening situation.
-Although avoidance is a natural response to danger, excessive avoidance as is
-observed in anxiety disorders is can be highly disruptive to everyday functioning
-\cite{Salter2004} and may work to sustain anxious appraisals and behavior. For
-example, an individual with social anxiety may avoid attending large social events
-for fear of public embarrassment only to never learn that such worry is not justified.
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/02_cliff.png}}%
+  }
+  \par \textbf{Figure 2:} Cliff-walking is a deterministic gridworld where an agent
+  must navigate from one side (S) to the goal on the other side (G) (A). Every
+  step costs the agent ($r=-1$) but stepping off the cliff results in great harm ($r=-100$).
+  An optimistic agent learns to travel the shortest (loss-minimizing) path along the cliff (B).
+  For a pessimistic agent ($w=0.8$), the potential for harm yields an optimal policy that
+  maintains distance from the cliff's edge (C). With increasing pessimism ($w=0.6$), the agent
+  learns an optimal path off the cliff so as to minimize unnecessary steps prior to its
+  perceived inevitable harm (D). (Parameters: $\gamma = 1$)
+\end{figure}
 
-Avoidance is often chosen even at the expense of potential reward (approch-avoidance
-conflict), and is frequently a major impediment to recovery \citep{Arnaudova2017}.
+In addition to pessimistic inference, all anxiety disorders share pathological
+avoidance as core symptom \citep{dsm5, Krypotos2015, Arnaudova2017}. Avoidance
+behaviors seekto increase or maintain the temporal, physical and/or mental distance
+between an individual and a threatening situation. Although avoidance is a natural response to danger, excessive avoidance as is observed in anxiety disorders can be highly disruptive to everyday
+functioning \cite{Salter2004} and frequently function so as maintain anxiety
+by preventing extinction \citep{Arnaudova2017}. For example, an individual with social anxiety may avoid
+attending large social events for fear of public embarrassment only to never learn
+that such worry is not justified.
 
-In the laboratory, avoidance has been measured in a variety of ways. In animal
-models of anxiety, anxious rodents are found to exhibit avoidance learning more
-quickly and maintain avoidance for longer \citep{servatius2008}. In studies of
-human avoidance, anxiety has been found to correlate with the maintenance of
-increased distance from threat, such as virtual predators \citep{Bach2014, Bach2017,
-Sheynin2014}.
-
-% Arnaudova2017 makes reference to some spider approach studies that might be
-% interesting to try and track down. Essential idea is that even though
-% participants are totally safe, may not touch the glass. The task is the
-% behavioral avoidance task. (e.g. Siegel & Weinberger, 2012, emotion).
-% the nice part about this is that participants are objectively safe.
+Avoidance behaviors in anxiety have been studied using a variety of methods. Perhaps
+the most classic paradigm is the behavioral avoidance test \citep{bandura1977}, wherein participants
+complete a number of tasks involving a feared object or situation and the number of
+tasks completed (or lack thereof) is a measure of avoidance. Computerized avoidance
+tasks have become increasingly common, and increased distance from virtual threat
+has been found to correlate with anxiety \citep{Bach2014, Bach2017, Sheynin2014}.
 
@@
 
-% Need to go into more detail than this, need to actually make this a coherent
-% explanation, but this can wait for now.
+In our second simulation, we demonstrate how pessimistic learning can yield
+sustained avoidance. In the cliff-walking environment \citep{SuttonBarto1998,
+SuttonBarto2018, Gaskett2003}, an agent must navigate from a start on side of the
+map to the other (Figure 2a). The agent experiences a small penalty for every
+step it takes. The shortest (least costly) path is along the cliff's edge; if the
+agent steps off the cliff, however, it incurs a much larger penalty. As in the first
+environment, we assume a deterministic transition structure.
 
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/02_cliff.png}}%
-  }
-  \caption{\textbf{Cliff Walking}}
-  \par Cliff = -100, Cost = -1, gamma = 0.99
-\end{figure}
+Just as in the first environment, the long-run value estimates by an optimistic
+agent do not reflect the potential harm of the cliff (Figure 2b). Because this agent (rightfully)
+expects to act according to the optimal policy in the future (i.e. never step off
+the cliff), the agent perceives no risk of future disaster and thus the large
+penalties of stepping off the cliff do not pollute the agent's value estimates.
+The optimistic agent learns the loss-minimizing policy and walks along the cliff's edge.
 
-In our second result, we demonstrate how a breakdown in control can lead to
-sustained policies of avoidance. We do so in the cliff-walking environment
-\citep{SuttonBarto1998, SuttonBarto2018, Gaskett2003}. Here an agent learns to navigate from
-an initial starting position to an end state. The shortest distance is along a
-cliff edge. If the agent steps off the cliff, it incurs a large penalty. Under
-complete control (objective or perceived), agent in this environment is under no
-threat despite the presence of the cliff's edge. As in the first simulation, this
-is simply because, in a fully controllable world, an agent need not ever interact
-with the aversive state. In fact, an agent can decide to walk straight up to the
-aversive state and, insofar it does not choose to take one step further, will
-never be in any actual danger.
+This is not the case for the pessimistic agent. As above, when the agent doubts
+its future ability to act in a reward-maximizing (loss-minimizing) manor, it
+cannot rule out the possibility of accidentally stepping off the cliff. Thus,
+the potential for future harm is no longer zero and the large penalties for stepping
+off the cliff may back-propagate to their antecedent states, and so on. The extent
+of this spread is dependent on the strength of the pessimistic belief.
+With increasing pessimism, the optimal policy for the agent is increasingly
+distance from the cliff's edge (Figure 2c). Surprisingly, at high levels of pessimistic
+belief, the optimal policy is to immediately step off the cliff (Figure 2d). The
+reason for this is simple: with high levels of pessimism, the likelihood of disaster
+grows large enough that any step is simply viewed as adding onto the pain.
 
-This is not so when deterministic control cannot be assumed. If an agent believes
-it is not in full control of its environment, then it cannot rule out the possibility
-that it may inadvertently encounter the negative state at some point in the future,
-especially if it veers to close to the state. This is true irrespective of whether
-the agent believes it cannot guarantee its own future competency or the benevolence
-of the environment. Importantly, this belief need not match the true environmental
-statistics; belief is enough.
-
-The long-run state values under different beliefs are plotted in Figure 2. As
-expected, under optimistic beliefs (Figure 1b) the entire environment sans negative
-state takes on a positive value (close to the positive state itself). As this belief
-breaks down, we observe a different pattern in the long-run state values. When
-future occupancy of the negative state becomes a non-zero probability, the undesired
-action transition exerts influence on action-outcome calculations, back-propagating,
-and thereby infecting antecedent states. As the agent's belief becomes increasingly
-pessimistic, the more the worst possible action exerts influence and the more
-negative, on average, the state-space grows. Most importantly, because these are
-the long-run state values, the policy is the optimal policy; in other words, this
-is the long-run policy an agent will take.
+Under pessimistic learning then, a perceived lack of self-efficacy can yield the
+propagation of negative value from potential future harms. In such a scenario,
+the optimal policy for an agent is to maintain a distance from threat far enough
+so as to optimally balance the risk of catastrophe with foregone losses. Importantly,
+these simulations demonstrate the long-run value estimates rather than any momentary
+value estimates. In other words, the behavioral policy these agents demonstrate
+will be maintained insofar beliefs about self-efficacy are maintained, irrespective
+of any momentary noisiness in actual outcomes. Thus, pessimistic learning can
+account for pathological avoidance in anxiety disorders.
 

### 2019-02-03 sisyphus/__init__.py

@@ -4,4 +4,3 @@ __version__ = '0.1'
 
 from . import envs
 from . import mdp
-from . import viz

### 2019-02-03 sisyphus/envs/_choice.py

@@ -61,11 +61,87 @@ class FreeChoice(GraphWorld):
         ## Initialize GraphWorld.
         GraphWorld.__init__(self, T, R, start, terminal, 0)
         
-        ## Remove duplicate actions.
-        self.info = self.info.drop([5,6,8,9,11,12]).reset_index(drop=True)
-        
-        ## Update transition probabilities.
-        for i in [4,5,6]: self.info.at[i,'T'] = probs
+        ## Update info.
+        for s in [2,3,4]:
+            ix, = np.where(self.info.S == s)
+            self.info = self.info.drop(ix[1:]).reset_index(drop=True)
+            self.info.at[ix[0],'T'] = probs
         
     def __repr__(self):
         return '<GraphWorld | Instrumental Free Choice>'
+    
+    def _draw_nodes(self, ax, s, free_color, fixed_color):
+    
+        ## Draw initial choice.
+        xpos = [0.0]; ypos = [0.0]
+        ax.plot(xpos, ypos, marker='o', markersize=s, fillstyle='top', linestyle='none',
+                color=free_color, markerfacecoloralt=fixed_color, markeredgecolor='k')
+
+        ## Draw free choice.
+        xpos = [1.0, 2.0, 2.0]; ypos = [0.25, 0.5, 0.0]
+        ax.plot(xpos, ypos, marker='o', markersize=s, fillstyle='full', linestyle='none',
+                color=free_color, markerfacecoloralt=fixed_color, markeredgecolor='k')
+
+        ## Draw fixed choice.
+        xpos = [1.0, 2.0]; ypos = [-0.25, -0.5]
+        ax.plot(xpos, ypos, marker='o', markersize=s, fillstyle='full', linestyle='none',
+                color=fixed_color, markerfacecoloralt=free_color, markeredgecolor='k')
+
+        return ax
+
+    def _draw_edges(self, ax):
+
+        ## Draw edges.
+        ax.plot([0.0, 2.0], [0.0, 0.5], color='k', zorder=0)
+        ax.plot([0.0, 2.0], [0.0,-0.5], color='k', zorder=0)
+        ax.plot([1.0, 2.0], [0.25,0.0], color='k', zorder=0)
+
+        return ax
+
+    def _draw_edge_labels(self, ax, fontsize=14):
+
+        ax.text(0.5,  0.15, '0', ha='center', va='bottom', fontsize=fontsize)
+        ax.text(0.5, -0.15, '0', ha='center', va='top', fontsize=fontsize)
[…]

### 2019-02-03 sisyphus/envs/_cliff.py

@@ -83,7 +83,7 @@ class CliffWalking(GraphWorld):
 
         ## Plot cliff.
         ax = sns.heatmap(grid, cmap=cmap, cbar=cbar)
-        ax.set(xticklabels=[], yticklabels=[], title='Cliff Walking')  
+        ax.set(xticklabels=[], yticklabels=[])  
 
         ## Add outline.
         ax.vlines(np.arange(1,12),0,10,lw=0.1)

### 2019-02-03 sisyphus/envs/_field.py

@@ -33,7 +33,7 @@ class OpenField(GraphWorld):
     def plot_field(self, reward=10, shock=-10, annot=True, grid_color='0.8',  
                    reward_color='#f3e1db', shock_color='#1c142a', 
                    cbar=False, annot_kws=None, ax=None):
-        """Plot cliff-walking environment.
+        """Plot open field environment.
 
         Parameters
         ----------
@@ -76,9 +76,9 @@ class OpenField(GraphWorld):
         ## Define colormap.
         cmap = ListedColormap([grid_color, reward_color, shock_color])
 
-        ## Plot cliff.
+        ## Plot open field.
         ax = sns.heatmap(grid, cmap=cmap, cbar=cbar, ax=ax)
-        ax.set(xticklabels=[], yticklabels=[], title='Cliff Walking')  
+        ax.set(xticklabels=[], yticklabels=[])  
 
         ## Add outline.
         ax.vlines(np.arange(1,11),0,11,lw=0.1)

### 2019-02-03 sisyphus/envs/_lh.py

@@ -84,9 +84,9 @@ class Helplessness(GraphWorld):
         ## Define colormap.
         cmap = ListedColormap([grid_color, reward_color, shock_color])
 
-        ## Plot cliff.
+        ## Plot environment.
         ax = sns.heatmap(grid, cmap=cmap, cbar=cbar, ax=ax)
-        ax.set(xticklabels=[], yticklabels=[], title='Learned Helplessness')  
+        ax.set(xticklabels=[], yticklabels=[])  
 
         ## Add outline.
         x,y = self.shape
@@ -101,4 +101,39 @@ class Helplessness(GraphWorld):
             annot_kws['color'] = 'k'
             ax.text(14.5,2.5,'S',ha='center',va='center',**annot_kws)
         
+        return ax
+    
+    def plot_policy(self, ax, pi, color='w', head_width=0.25, head_length=0.25):
+        """Plot agent policy on grid world.
+
+        Parameters
+        ----------
+        ax : matplotlib Axes
+            Axes in which to draw the plot.
+        pi : array
+            Agent policy, i.e. ordered visitation of states.
+        color : str (default = white)
+            Color of arrow.
+        head_width : float (default=0.25)
+            Width of the arrow head.
+        head_length : float (default=0.25)
+            Length of the arrow head.
+
+        Returns
+        -------
+        ax : matplotlib Axes
+            Axes in which to draw the plot.
+        """
+
+        ## Iteratively plot arrows.
+        for i in range(len(pi)-1):
+
+            ## Identify S, S' coordinates.
+            y1, x1 = np.where(self.grid==pi[i])
+            y2, x2 = np.where(self.grid==pi[i+1])
+
+            ## Plot.
+            ax.arrow(int(x1)+0.5, int(y1)+0.5, 0.5*int(x2-x1), 0.5*int(y2-y1), 
+                     color=color, head_width=head_width, head_length=head_length)
+
         return ax

### 2019-02-03 sisyphus/viz/__init__.py

@@ -1,3 +1,0 @@
-"""Visualization functions"""
-
-from ._choice import plot_free_choice

### 2019-02-03 sisyphus/viz/_choice.py

@@ -1,79 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-from matplotlib.lines import Line2D
-
-def _draw_nodes(ax, s, free_color, fixed_color):
-    
-    ## Draw initial choice.
-    xpos = [0.0]; ypos = [0.0]
-    ax.plot(xpos, ypos, marker='o', markersize=s, fillstyle='top', linestyle='none',
-            color=free_color, markerfacecoloralt=fixed_color, markeredgecolor='k')
-
-    ## Draw free choice.
-    xpos = [1.0, 2.0, 2.0]; ypos = [0.25, 0.5, 0.0]
-    ax.plot(xpos, ypos, marker='o', markersize=s, fillstyle='full', linestyle='none',
-            color=free_color, markerfacecoloralt=fixed_color, markeredgecolor='k')
-    
-    ## Draw fixed choice.
-    xpos = [1.0, 2.0]; ypos = [-0.25, -0.5]
-    ax.plot(xpos, ypos, marker='o', markersize=s, fillstyle='full', linestyle='none',
-            color=fixed_color, markerfacecoloralt=free_color, markeredgecolor='k')
-    
-    return ax
-
-def _draw_edges(ax):
-
-    ## Draw edges.
-    ax.plot([0.0, 2.0], [0.0, 0.5], color='k', zorder=0)
-    ax.plot([0.0, 2.0], [0.0,-0.5], color='k', zorder=0)
-    ax.plot([1.0, 2.0], [0.25,0.0], color='k', zorder=0)
-
-    return ax
-
-def _draw_edge_labels(ax, fontsize=14):
-    
-    ax.text(0.5,  0.15, '0', ha='center', va='bottom', fontsize=fontsize)
-    ax.text(0.5, -0.15, '0', ha='center', va='top', fontsize=fontsize)
-    ax.text(1.5,  0.40, r'$R$', ha='center', va='bottom', fontsize=fontsize)
-    ax.text(1.5,  0.10, r'$R$', ha='center', va='top', fontsize=fontsize)
-    ax.text(1.5, -0.40, r'$R$', ha='center', va='top', fontsize=fontsize)
-    ax.annotate(r'$R = \{0.33: -1, 0.33: 0, 0.33: 1\}$', (0,0), (0,0), 
-                xycoords='axes fraction', ha='left', va='bottom', fontsize=fontsize-4)
-    return ax
-
[…]

### 2019-02-03 manuscript/manuscript.tex
future unwrapping to be done

@@
 
-Central to many psychiatric disorders is distortion to representations of the
-external world and, correspondingly, pronounced changes to learning and decision
-making. As a normative framework, decision theory allows us to describe and pose
-quantita­tive questions about optimal choice behavior \citep{DayanDaw2008} and is,
-therefore, a critical tool for modeling, understanding, and predicting the psychological
-and neurobiological mechanisms underlying psychiatric behavior \citep{maia2011, HuysDawDayan2015}.
-Recent investigations employing this framework have provided novel and important
-insights into psychiatric symptoms and disorders, including anhedonia in major depression \citep{Rutledge2017},
-mood instability in bipolar disorder \citep{EldarNiv2015, EldarDolanNiv2016},
-habitual behavior in obsessive compulsive and addictive disorders \citep{gillan2016}, and
-hallucinations in schizophrenia \citep{powers2017, corlett2018}.
+Many psychiatric disorders involve distortions to representations of the external world and, as a likely consequence, pronounced changes to learning and decision making.
+As a normative framework, decision theory allows us to describe and pose quantita­tive questions about optimal choice behavior \citep{DayanDaw2008} and is, therefore, a critical tool for modeling, understanding, and predicting the psychological and neurobiological mechanisms underlying psychiatric behavior \citep{maia2011, HuysDawDayan2015}.
+Recent investigations employing this framework have provided novel and important insights into psychiatric symptoms and disorders, including anhedonia in major depression \citep{Rutledge2017}, mood instability in bipolar disorder \citep{EldarNiv2015, EldarDolanNiv2016}, habitual behavior in obsessive compulsive and addictive disorders \citep{gillan2016}, and hallucinations in schizophrenia \citep{powers2017, corlett2018}.
 
-The family of anxiety disorders (which includes generalized anxiety disorder, social
-anxiety disorder, panic disorder, and specific phobia) are characterized by excessive
-fear and worry about future negative outcomes, and represent the most common
-group of psychiatric disorders \citep{kessler2009}. Though the disorders typically
-differ in the types of objects or situations that induce fear or anxiety, all share
-a set of core disturbances to cognition and behavior. First, the anxiety disorders
-involve pessimistic inference, or a tendency to appraise threatening objects or
-situations as more likely and/or more severe than they objectively are. Clinical
-anxiety is also associated with second-order conditioning, or learning to fear
-neutral cues indirectly associated with threat cues. Finally, anxiety is associated
-with sustained avoidance, or the tendency for individuals to flee or avoid threatening
-situations. As an example, an individual is mugged in an alleyway. In a pathological
-response, this individual may exhibit persistent worry that it will happen again
-(pessimistic inference), increased anxiety towards the street or time of day
-at which it occurred (second-order conditioning), and increased effort to stay
-away from any cue which reminds them of the event (avoidance).
+The family of anxiety disorders (which includes generalized anxiety disorder, social anxiety disorder, panic disorder, and specific phobia) are characterized by excessive fear and worry about future negative outcomes, and represent the most common group of psychiatric disorders \citep{kessler2009}.
+Though the disorders typically differ in the types of objects or situations that induce fear or anxiety, all share a set of core disturbances to cognition and behavior.
+First, the anxiety disorders involve pessimistic inference, or a tendency to appraise threatening objects or situations as more likely and/or more severe than they objectively are.
+Clinical anxiety is also associated with second-order conditioning, or learning to fear neutral cues indirectly associated with threat cues.
+Finally, anxiety is associated with avoidance, or the tendency to flee or avoid threatening situations.
+As an example, an individual with a clinical phobia of dogs may show chronic worry about coming into contact with aggressive dogs (pessimistic inference), may develop fear towards parks or other public spaces where they may encounter dogs (second-order conditioning), and might exert greater effort to avoid contact with dogs or locations associated with dogs (avoidance).
 
-Several studies have attempted to provide some insight into why some of this
-may be in anxiety disorders \cite{Maia2010, norbury2018}, though no account has
-tied these symptoms together. For example, \citep{Maia2010} provided a simple
-account of how simple actor-critic reinforcement learning models could account
-for sustained avoidance in rodent shuttle-box experiments \citep{servatius2008}.
-As has been discussed elsewhere \citep{Moutoussis2017}, such an account fails to account for why avoidance
-is maintained even with non-threatening outcomes and do not account for why exposure
-therapy only works for some individuals. Other accounts have attempted to explain
-pessimistic inference (cite learning rate), but do not mention where from.
+Recent work has begun to shine some light into the mechanisms underlying these symptoms.
+For example, a number of studies have found that anxiety is correlated with increased learning from worse-than-predicted outcomes \citep{Harle2017, Garrett2018, Aylward2019}.
+Systematic biases in learning could result in pessimistic inference by virtue of negatively biasing those same cached values used to make predictions at the time of retrieval.
+This begs the question as to why learning from negative prediction errors is greater in anxiety, and others have suggested this may reflect prior expectations in anxiety; that pessimistic expectations bias learning from expectation-congruent information.
+This raises the possibility of circular arguments.
 
-In the remainder of this article, we present a simple decision theoretic account
-of how the core symptoms of anxiety may arise and be maintained. Specifically, we
-demonstrate that the loss of optimistic inference in the form of reduced expectations
-of future reward can yield negatively biased estimates of value in sequential
-decision environments. In such an event, pessimistic inference, second-order
-conditioning, and chronic avoidance naturally follow.
+Other studies have attempted to provide novel insights into avoidance.
+\cite{Maia2010} offered a reinforcement learning-based account of maintained avoidance in rodent conditioning tasks, an empirical finding that previous theories of avoidance could not account for \citep{Krypotos2015}.
+They argue that sustained avoidance is possible insofar that a well-established avoidance response generates no reward prediction error, such that no new learning takes place, and avoidance continues.
+As is discussed elsewhere \citep{Moutoussis2017}, the framework cannot account for sustained avoidance when an agent is accidentally exposed to an extinguished outcome.
+Returning to the previous example, the framework cannot account for why dog phobia remains for an individual in spite of an unplanned exposure to a friend's benign dog.
+More broadly, the framework cannot for why avoidance, and anxiety in general, is so persistent even with exposure therapy \citep{Craske2014}.
+
+In the remainder of this article, we present a simple decision theoretic account of how the core symptoms of anxiety may arise and be maintained.
+Specifically, we demonstrate that how a loss of confidence in the reliability of the self or the environment can result in the catastrophic spread of negative value in sequential decision environments.
+We show how, under these assumptions, pessimistic inference, second-order conditioning, and chronic avoidance naturally follow, as well as a number of other behaviors previously found to be correlated with anxiety.
 
@@
 
-Before moving forward, we first briefly review the formalism of Markov decision
-processes, which provide the framework for our results. For complete treatments,
-see \cite{SuttonBarto1998, SuttonBarto2018} and \cite{bertsekas2005}.
+Before moving forward, we first briefly review the formalism of Markov decision processes, which provide the framework for our results.
+For complete treatments, see \cite{SuttonBarto1998, SuttonBarto2018} and \cite{bertsekas2005}.
 
-A MDP is defined by a set of states, $S$, a set of actions, $A$, a reward
-function defined over state/action pairs, $R(s,a)$, and a state transition distribution,
-$P(s'|s,a)$, where $a$ denotes a chosen action. In a MDP, states and rewards occur
-sequentially according to the one-step transition structure where transitions follow
-actions. The goal for an agent in a MDP is to learn to choose a policy, $\pi$, or
-mapping of state to action, that maximizes long-run expected reward. Here we define
-the value of a state under a policy, $V^\pi(s)$, as the expected cumulative discounted reward:
+A MDP is defined by a set of states, $S$, a set of actions, $A$, a reward function defined over state/action pairs, $R(s,a)$, and a state transition distribution, $P(s'|s,a)$, where $a$ denotes a chosen action.
+In a MDP, states and rewards occur sequentially according to the one-step transition structure where transitions follow actions.
+The goal for an agent in a MDP is to learn to choose a policy, $\pi$, or mapping of state to action, that maximizes long-run expected reward.
+Here we define the value of a state under a policy, $V^\pi(s)$, as the expected cumulative discounted reward:
 
@@
 
-where $\gamma$ is the temporal discounting parameter, controlling the weighting
-of distant future rewards. The value function can be defined recursively as the
-sum of the immediate reward following an action chosen in a state, $R(s, a)$, and
-the value of its successor state $s’$, averaged over potential future actions, $a$,
-and transitions according to the current policy $\pi$:
+where $\gamma$ is the temporal discounting parameter, controlling the weighting of distant future rewards.
+The value function can be defined recursively as the sum of the immediate reward following an action chosen in a state, $R(s, a)$, and the value of its successor state $s’$, averaged over potential future actions, $a$, and transitions according to the current policy $\pi$:
 
@@
 
-Using the formalisms above, the value of one particular action in a state under
-the optimal policy is defined as:
+Using the formalisms above, the value of one particular action in a state under the optimal policy is defined as:
 
@@
 
-Insofar that the value of a state under the optimal policy is defined as the best
-action under that state, we can re-express the above equivalently as:
+Insofar that the value of a state under the optimal policy is defined as the best action under that state, we can re-express the above equivalently as:
 
@@
 
-Thus in a conventional MDP, the value of states in the environment under the optimal
-policy assumes the agent is able to act according to the optimal policy in the future.
+Thus in a conventional MDP, the value of states in the environment under the optimal policy assumes the agent is able to act according to the optimal policy in the future.
 
@@
 
-In the previous section, we observed that the value of a state and its associated
-state-action pairs are a function of the environment transition distribution and
-the best future action. In standard MDP problems, it is typically assumed that the
-transition distribution is known and accurate, and that the agent is consistently
-able to operate under the optimal policy in the future. In this section, we introduce
-a formalism deviating from these standard assumptions.
+In the previous section, we observed that the value of a state and its associated state-action pairs are a function of the environment transition distribution and the best future action.
+In standard MDP problems, it is typically assumed that the transition distribution is known and accurate, and that the agent is consistently able to operate under the optimal policy in the future.
+In this section, we introduce a formalism deviating from these standard assumptions.
 
-Central to many theories of clinical anxiety is a perceived loss of control (see
-discussion for elaboration). In the MDP framework described above, this may come
-in one of several forms. First, an agent may perceive the transition structure
-of the environment to be less reliable than it actually is. Put another way, an
-agent may have a biased expectation that, despite choosing an action $a$ with the
-intent of transition to state $s'$, the agent may instead transition to an alternate,
-undesired state $s^$. In the case of clinical anxiety, the expectation may be more
-dire, that the environment is "adversarial" is likely to transition an agent, not
-randomly, but to the worse states.
+Central to many theories of clinical anxiety is a perceived loss of control (see discussion for elaboration).
+In the MDP framework described above, this may come in one of several forms.
+First, an agent may perceive the transition structure of the environment to be less reliable than it actually is.
+Put another way, an agent may have a biased expectation that, despite choosing an action $a$ with the intent of transition to state $s'$, the agent may instead transition to an alternate, undesired state $s^$.
+In the case of clinical anxiety, the expectation may be more dire, that the environment is "adversarial" is likely to transition an agent, not randomly, but to the worse states.
 
-Second, an agent may doubt its own agentive control. An agent may expect to be
-able to choose the best possible action for itself in the present, but may not be
-confident it will still be able to decide optimally for itself in the future. In
-either event, following the value computations defined above, we should expect
-the value estimates for such an agent to diminish. If an agent believes itself
-or the environment to be unreliable, then the agent cannot be optimistic about
-its future and should instead adopt a different policy.
+Second, an agent may doubt its own agentive control.
+An agent may expect to be able to choose the best possible action for itself in the present, but may not be confident it will still be able to decide optimally for itself in the future.
+In either event, following the value computations defined above, we should expect the value estimates for such an agent to diminish.
+If an agent believes itself or the environment to be unreliable, then the agent cannot be optimistic about its future and should instead adopt a different policy.
 
-These sorts of problems have been explored in the safe reinforcement learning
-and robust control literature \citep{Garcia2015}. There, we are concerned
-with programming agents that identify the optimal policy under pessimistic
-assumptions. A number of approaches have been proposed in that literature. Here,
-we focus on one particular simple implemented proposed by \cite{Gaskett2003}, namely
-*w-pessimism*:
+These sorts of problems have been explored in the safe reinforcement learning and robust control literature \citep{Garcia2015}.
+There, we are concerned with programming agents that identify the optimal policy under pessimistic assumptions.
+A number of approaches have been proposed in that literature.
+Here, we focus on one particular simple implemented proposed by \cite{Gaskett2003}:
 
@@
 
-In this formalism, $w$ controls the degree of pessimism. When $\w = 1$, an agent
-is maximally optimistic and fully expects to act according to the best possible actions
-in the future. When $w = 0$, the agent is maximally pessimistic and fully expects
-to act according to the worst possible actions in the future. When $w = 0.5$,
-the weighs the best and worst possible future actions equally.
+In this formalism, $w$ controls the degree of pessimism.
+When $\w = 1$, an agent is maximally optimistic and fully expects to act according to the best possible actions in the future.
+When $w = 0$, the agent is maximally pessimistic and fully expects to act according to the worst possible actions in the future.
+When $w = 0.5$, the weighs the best and worst possible future actions equally.
 
-We note that we are not committed to this particular implementation of a pessimistic
-agent. As mentioned above, an alternative parameterization would have been to
-program an agent with a belief in an adversarial environment, preferentially
-transitioning it to worse states. Insofar that these two terms are multiplicative
-in value calculations, either implementation should yield similar results. Similarly,
-the w-pessimism function could easily be replaced with other functions (e.g. softmax
-with negative inverse temperature). This formalism is simply convenient and is
-useful for demonstrating the points we demonstrate below. It is not a substantive
-claim about anxiety.
+We note that we are not committed to this particular implementation of a pessimistic agent.
+As mentioned above, an alternative parameterization would have been to program an agent with a belief in an adversarial environment, preferentially transitioning it to worse states.
+Insofar that these two terms are multiplicative in value calculations, either implementation should yield similar results.
+Similarly, the w-pessimism function could easily be replaced with other functions (e.g.
+softmax with negative inverse temperature).
+This formalism is simply convenient and is useful for demonstrating the points we demonstrate below.
+It is not a substantive claim about anxiety.
 
@@
 
-In the following sections, we present the results of simulations of pessimistic
-learning in a variety of MDP environments. In all simulations, the state-action
-values under the optimal policy, $Q_*(s,a)$, were solved for directly through
-Q-value iteration \citep{SuttonBarto1998,SuttonBarto1998,bertsekas2005}. The
-details of this algorithm are presented in the appendix. Model parameters are
-reported for each environment in their corresponding sections. All simulations
-were carried out with the python programming language and are publicly available
-on Github (https://github.com/szorowi1/SecretFunTimes).
+In the following sections, we present the results of simulations of pessimistic learning in a variety of MDP environments.
+In all simulations, the state-action values under the optimal policy, $Q_*(s,a)$, were solved for directly through Q-value iteration \citep{SuttonBarto1998, SuttonBarto1998, bertsekas2005}.
+The details of this algorithm are presented in the appendix.
+Model parameters are reported for each environment in their corresponding sections.
+All simulations were carried out with the python programming language and are publicly available on Github (https://github.com/szorowi1/SecretFunTimes).
 
@@
   }
-  \par \textbf{Figure 1:} The open field is a deterministic gridworld with only one rewarding
-  (blue) and aversive (red) state (A). For an optimistic agent ($w=1$), all states (other than
-  the harmful state) take on positive value (B). For a pessimistic agent ($w=0.5$), negative value spreads
-  from the source to antecedent states (C). With increasing pessimism ($w=0$), the extent of the spread
-  grows worse (D). (Parameters: $\gamma = 0.95$)
+  \par \textbf{Figure 1:} The open field is a deterministic gridworld with only one rewarding (blue) and aversive (red) state (A).
+For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value (B).
+For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states (C).
+With increasing pessimism ($w=0$), the extent of the spread grows worse (D).
+(Parameters: $\gamma = 0.95$)
 \end{figure}
 
-In this first section, we demonstrate how learning under an assumption of a lack
-of future control naturally leads to pessimistic inference and second-order conditioning.
-A symptom central to all of the anxiety disorders is pessimistic inference,
-or a tendency to appraise the likelihood and/or severity of threatening situations
-out of proportion to their actual environmental contingencies \citep{dsm5, BeckClark1997,
-ClarkBeck2011}. For example, an individual with specific phobia may vastly exaggerate
-the probability of encountering their fear (e.g. a venomous spider) in their
-everyday lives. Anxious individuals will also tend to associate fear with cues
-tangentially related to signals of threat. For example, the same individual may
-begin to fear locations where they may encounter spider webs (e.g. basements).
+In this first section, we demonstrate how learning under an assumption of a lack of future control naturally leads to pessimistic inference and second-order conditioning.
+A symptom central to all of the anxiety disorders is pessimistic inference, or a tendency to appraise the likelihood and/or severity of threatening situations out of proportion to their actual environmental contingencies \citep{dsm5, BeckClark1997, ClarkBeck2011}.
+For example, an individual with specific phobia may vastly exaggerate the probability of encountering their fear (e.g.
+a venomous spider) in their everyday lives.
+Anxious individuals will also tend to associate fear with cues tangentially related to signals of threat.
+For example, the same individual may begin to fear locations where they may encounter spider webs (e.g.
+basements).
 
-Pessimistic inference in anxiety is well-documented. In studies of self-reported beliefs,
-individuals with anxiety rate
-hypothetical future negative outcomes as being more likely to occur and worse to
-experience than non-anxious counterparts \citep{ButlerMathews1983, ButlerMathews1987,
-Foa1996, MacLeod1996, MacLeod1997, Luten1997, Stober1997, Borkovec1999, Maner2006,
-Mitte2007, Miranda2007}. In the laboratory, exaggerated threat appraisal in anxiety
-has traditionally been measured through fear conditioning. The results of these
-studies demonstrate that anxious individuals exhibit greater fear towards threatening
-cues, non-threatening cues, and extinguished threatening cues as compared to non-anxious
-counterparts \citep{lissek2005, MinekaOehlberg2008, Duits2015}. Importantly, threatening
-cues can become associated with secondary neutral cues, allowing negative value to
-spread \citep{wessa2007}. In sum, anxiety is associated with inflated estimates of
-threat, both likelihood and value, and these estimates can spread from the source
-to distantly associated cues.
+Pessimistic inference in anxiety is well-documented.
+In studies of self-reported beliefs, individuals with anxiety rate hypothetical future negative outcomes as being more likely to occur and worse to experience than non-anxious counterparts \citep{ButlerMathews1983, ButlerMathews1987, Foa1996, MacLeod1996, MacLeod1997, Luten1997, Stober1997, Borkovec1999, Maner2006, Mitte2007, Miranda2007}.
+In the laboratory, exaggerated threat appraisal in anxiety has traditionally been measured through fear conditioning.
+The results of these studies demonstrate that anxious individuals exhibit greater fear towards threatening cues, non-threatening cues, and extinguished threatening cues as compared to non-anxious counterparts \citep{lissek2005, MinekaOehlberg2008, Duits2015}.
+Importantly, threatening cues can become associated with secondary neutral cues, allowing negative value to spread \citep{wessa2007}.
+In sum, anxiety is associated with inflated estimates of threat, both likelihood and value, and these estimates can spread from the source to distantly associated cues.
 
-In our first simulation, we demonstrate how this may come about. In the open field
-environment (Figure 1a), an agent is free to navigate a large grid world devoid
-of value except for two states: one rewarding and one aversive red. In this environment,
-the one-step transition structure is deterministic such that the environment
-fully respects the agent's choice. As such, the agent is under no threat despite
-the presence of a harmful state. Insofar that the agent never acts under a
-masochistic policy and intentionally approaches harm, the agent need not ever
-encounter the harmful state.
+In our first simulation, we demonstrate how this may come about.
+In the open field environment (Figure 1a), an agent is free to navigate a large grid world devoid of value except for two states: one rewarding and one aversive red.
+In this environment, the one-step transition structure is deterministic such that the environment fully respects the agent's choice.
+As such, the agent is under no threat despite the presence of a harmful state.
+Insofar that the agent never acts under a masochistic policy and intentionally approaches harm, the agent need not ever encounter the harmful state.
 
-The long-run value estimates learned by an optimistic agent (i.e. an agent expecting
-to be able to follow a reward-maximizing policy now and in the future) reflect this
-(Figure 1b). For an optimistic agent, all states in the open field environment,
-sans the harmful state itself, take on a positive value. This is unsurprising
-insofar that, except for the harmful state, the agent will always encounter
-a reward in the long-run no matter its starting position in the field.
+The long-run value estimates learned by an optimistic agent (i.e.
+an agent expecting to be able to follow a reward-maximizing policy now and in the future) reflect this (Figure 1b).
+For an optimistic agent, all states in the open field environment, sans the harmful state itself, take on a positive value.
+This is unsurprising insofar that, except for the harmful state, the agent will always encounter a reward in the long-run no matter its starting position in the field.
 
-The same cannot be said for a pessimistic agent. If an agent doubts its future
-ability to act in according with the optimal policy, then it cannot rule out that,
-in the future, it may inadvertently encounter the harmful state. This is more so
-likely if the agent is already proximal to the harmful state. Under this pessimistic
-belief, the possibility of future harm is no longer zero, as it is for the optimistic
-agent, and thus negative value can propagate from the states antecedent to harm,
-back to their respective antecedents and so on. The extent of this spread will depend
-on the strength of the pessimistic belief (Figures 1c, 1d). With increasing pessimism,
-harmful outcomes are assigned greater credence and thus exert greater and wider
-influence on antecedent states.
+The same cannot be said for a pessimistic agent.
+If an agent doubts its future ability to act in according with the optimal policy, then it cannot rule out that, in the future, it may inadvertently encounter the harmful state.
+This is more so likely if the agent is already proximal to the harmful state.
+Under this pessimistic belief, the possibility of future harm is no longer zero, as it is for the optimistic agent, and thus negative value can propagate from the states antecedent to harm, back to their respective antecedents and so on.
+The extent of this spread will depend on the strength of the pessimistic belief (Figures 1c, 1d).
+With increasing pessimism, harmful outcomes are assigned greater credence and thus exert greater and wider influence on antecedent states.
 
-In sequential decision environments then, a perceived lack of self-efficacy can
-result in the propagation of negative value from potential future harms. Such a
-pessimistic agent may correspondingly report an increased likelihood of threat
-(i.e. a greater number of threatening states), increased severity of threat (i.e.
+In sequential decision environments then, a perceived lack of self-efficacy can result in the propagation of negative value from potential future harms.
+Such a pessimistic agent may correspondingly report an increased likelihood of threat (i.e.
+a greater number of threatening states), increased severity of threat (i.e.
 states with more negative value), and fear of states antecedent to the fear source.
-As such, pessimistic learning can account for two of the hallmark symptoms of
-the anxiety disorders.
+As such, pessimistic learning can account for two of the hallmark symptoms of the anxiety disorders.
 
@@
   }
-  \par \textbf{Figure 2:} Cliff-walking is a deterministic gridworld where an agent
-  must navigate from one side (S) to the goal on the other side (G) (A). Every
-  step costs the agent ($r=-1$) but stepping off the cliff results in great harm ($r=-100$).
+  \par \textbf{Figure 2:} Cliff-walking is a deterministic gridworld where an agent must navigate from one side (S) to the goal on the other side (G) (A).
+Every step costs the agent ($r=-1$) but stepping off the cliff results in great harm ($r=-100$).
 An optimistic agent learns to travel the shortest (loss-minimizing) path along the cliff (B).
-  For a pessimistic agent ($w=0.8$), the potential for harm yields an optimal policy that
-  maintains distance from the cliff's edge (C). With increasing pessimism ($w=0.6$), the agent
-  learns an optimal path off the cliff so as to minimize unnecessary steps prior to its
-  perceived inevitable harm (D). (Parameters: $\gamma = 1$)
+For a pessimistic agent ($w=0.8$), the potential for harm yields an optimal policy that maintains distance from the cliff's edge (C).
+With increasing pessimism ($w=0.6$), the agent learns an optimal path off the cliff so as to minimize unnecessary steps prior to its perceived inevitable harm (D).
+(Parameters: $\gamma = 1$)
 \end{figure}
 
-In addition to pessimistic inference, all anxiety disorders share pathological
-avoidance as core symptom \citep{dsm5, Krypotos2015, Arnaudova2017}. Avoidance
-behaviors seekto increase or maintain the temporal, physical and/or mental distance
-between an individual and a threatening situation. Although avoidance is a natural response to danger, excessive avoidance as is observed in anxiety disorders can be highly disruptive to everyday
-functioning \cite{Salter2004} and frequently function so as maintain anxiety
-by preventing extinction \citep{Arnaudova2017}. For example, an individual with social anxiety may avoid
-attending large social events for fear of public embarrassment only to never learn
-that such worry is not justified.
+In addition to pessimistic inference, all anxiety disorders share pathological avoidance as core symptom \citep{dsm5, Krypotos2015, Arnaudova2017}.
+Avoidance behaviors seek to increase or maintain the temporal, physical and/or mental distance between an individual and a threatening situation.
+Although avoidance is a natural response to danger, excessive avoidance as is observed in anxiety disorders can be highly disruptive to everyday functioning \citep{Salter2004} and frequently function so as maintain anxiety by preventing extinction \citep{Arnaudova2017}.
+For example, an individual with social anxiety may avoid attending large social events for fear of public embarrassment only to never learn that such worry is not justified.
 
-Avoidance behaviors in anxiety have been studied using a variety of methods. Perhaps
-the most classic paradigm is the behavioral avoidance test \citep{bandura1977}, wherein participants
-complete a number of tasks involving a feared object or situation and the number of
-tasks completed (or lack thereof) is a measure of avoidance. Computerized avoidance
-tasks have become increasingly common, and increased distance from virtual threat
-has been found to correlate with anxiety \citep{Bach2014, Bach2017, Sheynin2014}.
+Avoidance behaviors in anxiety have been studied using a variety of methods.
+Perhaps the most classic paradigm is the behavioral avoidance test \citep{bandura1977}, wherein participants complete a number of tasks involving a feared object or situation and the number of tasks completed (or lack thereof) is a measure of avoidance.
+Computerized avoidance tasks have become increasingly common, and increased distance from virtual threat has been found to correlate with anxiety \citep{Bach2014, Bach2017, Sheynin2014}.
 
-There is a long history of models attempting to explain avoidance learning. The
-fundamental challenge is explaining how an action can be reinforced in the
-absence of an outcome. Two-factor theory was the explanation for a long while,
-but suffered from a number of noteworthy criticisms \citep{Krypotos2015}.
-Recent work has revitalized two-factor theory and integrated it with reinforcement
-learning \citep{Moutoussis2008, Maia2010}, and in the process addressing several
-criticisms of two-factor theory. As has been discussed elsewhere \citep{Moutoussis2017},
-such accounts only explain the maintenance of avoidance in short timescales, but
-do not explain why resistant to change and why they do not go away in general.
+There is a long history of models attempting to explain avoidance learning.
+The fundamental challenge is explaining how an action can be reinforced in the absence of an outcome.
+Two-factor theory was the explanation for a long while, but suffered from a number of noteworthy criticisms \citep{Krypotos2015}.
+Recent work has revitalized two-factor theory and integrated it with reinforcement learning \citep{Moutoussis2008, Maia2010}, and in the process addressing several criticisms of two-factor theory.
+As has been discussed elsewhere \citep{Moutoussis2017}, such accounts only explain the maintenance of avoidance in short timescales, but do not explain why resistant to change and why they do not go away in general.
 
-In our second simulation, we demonstrate how pessimistic learning can yield
-sustained avoidance. In the cliff-walking environment \citep{SuttonBarto1998,
-SuttonBarto2018, Gaskett2003}, an agent must navigate from a start on side of the
-map to the other (Figure 2a). The agent experiences a small penalty for every
-step it takes. The shortest (least costly) path is along the cliff's edge; if the
-agent steps off the cliff, however, it incurs a much larger penalty. As in the first
-environment, we assume a deterministic transition structure.
+In our second simulation, we demonstrate how pessimistic learning can yield sustained avoidance.
+In the cliff-walking environment \citep{SuttonBarto1998, SuttonBarto2018, Gaskett2003}, an agent must navigate from a start on side of the map to the other (Figure 2a).
+The agent experiences a small penalty for every step it takes.
+The shortest (least costly) path is along the cliff's edge; if the agent steps off the cliff, however, it incurs a much larger penalty.
+As in the first environment, we assume a deterministic transition structure.
 
-Just as in the first environment, the long-run value estimates by an optimistic
-agent do not reflect the potential harm of the cliff (Figure 2b). Because this agent (rightfully)
-expects to act according to the optimal policy in the future (i.e. never step off
-the cliff), the agent perceives no risk of future disaster and thus the large
-penalties of stepping off the cliff do not pollute the agent's value estimates.
+Just as in the first environment, the long-run value estimates by an optimistic agent do not reflect the potential harm of the cliff (Figure 2b).
+Because this agent (rightfully) expects to act according to the optimal policy in the future (i.e.
+never step off the cliff), the agent perceives no risk of future disaster and thus the large penalties of stepping off the cliff do not pollute the agent's value estimates.
 The optimistic agent learns the loss-minimizing policy and walks along the cliff's edge.
 
-This is not the case for the pessimistic agent. As above, when the agent doubts
-its future ability to act in a reward-maximizing (loss-minimizing) manor, it
-cannot rule out the possibility of accidentally stepping off the cliff. Thus,
-the potential for future harm is no longer zero and the large penalties for stepping
-off the cliff may back-propagate to their antecedent states, and so on. The extent
-of this spread is dependent on the strength of the pessimistic belief.
-With increasing pessimism, the optimal policy for the agent is increasingly
-distance from the cliff's edge (Figure 2c). Surprisingly, at high levels of pessimistic
-belief, the optimal policy is to immediately step off the cliff (Figure 2d). The
-reason for this is simple: with high levels of pessimism, the likelihood of disaster
-grows large enough that any step is simply viewed as adding onto the pain.
+This is not the case for the pessimistic agent.
+As above, when the agent doubts its future ability to act in a reward-maximizing (loss-minimizing) manor, it cannot rule out the possibility of accidentally stepping off the cliff.
+Thus, the potential for future harm is no longer zero and the large penalties for stepping off the cliff may back-propagate to their antecedent states, and so on.
+The extent of this spread is dependent on the strength of the pessimistic belief.
+With increasing pessimism, the optimal policy for the agent is increasingly distance from the cliff's edge (Figure 2c).
+Surprisingly, at high levels of pessimistic belief, the optimal policy is to immediately step off the cliff (Figure 2d).
+The reason for this is simple: with high levels of pessimism, the likelihood of disaster grows large enough that any step is simply viewed as adding onto the pain.
 
-Under pessimistic learning then, a perceived lack of self-efficacy can yield the
-propagation of negative value from potential future harms. In such a scenario,
-the optimal policy for an agent is to maintain a distance from threat far enough
-so as to optimally balance the risk of catastrophe with foregone losses. Importantly,
-these simulations demonstrate the long-run value estimates rather than any momentary
-value estimates. In other words, the behavioral policy these agents demonstrate
-will be maintained insofar beliefs about self-efficacy are maintained, irrespective
-of any momentary noisiness in actual outcomes. Thus, pessimistic learning can
-account for pathological avoidance in anxiety disorders.
+Under pessimistic learning then, a perceived lack of self-efficacy can yield the propagation of negative value from potential future harms.
+In such a scenario, the optimal policy for an agent is to maintain a distance from threat far enough so as to optimally balance the risk of catastrophe with foregone losses.
+Importantly, these simulations demonstrate the long-run value estimates rather than any momentary value estimates.
+In other words, the behavioral policy these agents demonstrate will be maintained insofar beliefs about self-efficacy are maintained, irrespective of any momentary noisiness in actual outcomes.
+Thus, pessimistic learning can account for pathological avoidance in anxiety disorders.
 
@@
 
-Next, we turn our attention towards a related phenomenon: a bias towards avoidance
-in approach-avoidance conflict. In environments with correlated risk and rewards,
-a bias towards avoidance can ultimately result in fewer expected rewards and worse
-long-run returns. Returning to the example above, an individual with social anxiety
-may forgo social situations to avoid the risk of social embarrassment at the expense
-of friendships, etc. This is another way in which pathological anxiety can be
-disruptive to everyday functioning.
+Next, we turn our attention towards a related phenomenon: a bias towards avoidance in approach-avoidance conflict.
+In environments with correlated risk and rewards, a bias towards avoidance can ultimately result in fewer expected rewards and worse long-run returns.
+Returning to the example above, an individual with social anxiety may forgo social situations to avoid the risk of social embarrassment at the expense of friendships, etc.
+This is another way in which pathological anxiety can be disruptive to everyday functioning.
 
-Experimentally, there is a long history of measuring behavior during
-approach-avoidance conflict in individuals with anxiety (citation, maybe yin-yang anxiety paper).
-One popular paradigm is the balloon analog risk task (BART) \citep{Lejuez2002},
-which has been shown to correlate with anxiety \cite{Maner2007, Giorgetta2012}.
-In the task, a participant pumps a balloon with air. For each pump, the participant
-earns points. If she pumps too much, the balloon pops and she loses all points.
-The participant must decide within each episode when to stop pumping. Leave time
-is the measure of behavior, with earlier leave times in anxiety suggesting a
-greater tendency towards avoidance.
+Experimentally, there is a long history of measuring behavior during approach-avoidance conflict in individuals with anxiety (citation, maybe yin-yang anxiety paper).
+One popular paradigm is the balloon analog risk task (BART) \citep{Lejuez2002}, which has been shown to correlate with anxiety \cite{Maner2007, Giorgetta2012}.
+In the task, a participant pumps a balloon with air.
+For each pump, the participant earns points.
+If she pumps too much, the balloon pops and she loses all points.
+The participant must decide within each episode when to stop pumping.
+Leave time is the measure of behavior, with earlier leave times in anxiety suggesting a greater tendency towards avoidance.
 
@@
 
-Our sequential model of decision making easily explains these results. A schematic
-of the task is presented in Figure 3. Under optimistic beliefs, points should
-travel back from the mean. Under pessimistic assumptions, however, different
-degrees of the worst outcome backpropagate. At a certain point, the perceived cost
-of staying outweighs the guaranteed gains of leaving. Thus, patients with anxiety
-leave earlier.
+Our sequential model of decision making easily explains these results.
+A schematic of the task is presented in Figure 3.
+Under optimistic beliefs, points should travel back from the mean.
+Under pessimistic assumptions, however, different degrees of the worst outcome backpropagate.
+At a certain point, the perceived cost of staying outweighs the guaranteed gains of leaving.
+Thus, patients with anxiety leave earlier.
 
-We can explain similar findings with the flight initiation distance \citep{Mobbs2018,
-Mobbs2019}. One thing our model cannot explain is the neural difference between
-fast and slow escape. Indeed, our model cannot capture all but perhaps is relevant
-for certain calculations. One could imagine these beliefs interacting in such a
-way to change the distance calculations that are necessary for mediating between
-passive and active defense responses.
+We can explain similar findings with the flight initiation distance \citep{Mobbs2018, Mobbs2019}.
+One thing our model cannot explain is the neural difference between fast and slow escape.
+Indeed, our model cannot capture all but perhaps is relevant for certain calculations.
+One could imagine these beliefs interacting in such a way to change the distance calculations that are necessary for mediating between passive and active defense responses.
 
 This can also be used to explain behavioral inhibition \citep{bach2015, khemka2017}.
-The idea is slightly different than what is advocated. The claim here is that
-in anxious individuals, the Q-values should be closer together, and thus there
-should be longer waiting time.
+The idea is slightly different than what is advocated.
+The claim here is that in anxious individuals, the Q-values should be closer together, and thus there should be longer waiting time.
 
@@
 
-In this next section, we review the literature on aversive pruning. In large,
-multi-step sequential decision environments, the space of all possible sequences
-of choices grows exponentially larger with increasing sequence length. As such,
-it is infeasible for decision making agents to exhaustively explore all possible
-sequences select that which maximizes reward from the entire set. Heuristics for
-narrowing the search space then are not only plausible but beneficial for a
-decision making agent.
+In this next section, we review the literature on aversive pruning.
+In large, multi-step sequential decision environments, the space of all possible sequences of choices grows exponentially larger with increasing sequence length.
+As such, it is infeasible for decision making agents to exhaustively explore all possible sequences select that which maximizes reward from the entire set.
+Heuristics for narrowing the search space then are not only plausible but beneficial for a decision making agent.
 
-One such heuristic that has been proposed in the literature is aversive pruning
-\citep{Huys2012}. Aversive pruning is defined as a Pavlovian response to encountering
-a large loss in planning such that sequences involving large losses are discarded
-from further evaluation. An example of scenario is presented in Figure 4. The prediction
-in such an environment is that an agent would discard the branch of the decision
-tree involving an immediate large negative loss even though this branch objectively
-contains the reward maximizing (loss minimizing) sequence of choices.
+One such heuristic that has been proposed in the literature is aversive pruning \citep{Huys2012}.
+Aversive pruning is defined as a Pavlovian response to encountering a large loss in planning such that sequences involving large losses are discarded from further evaluation.
+An example of scenario is presented in Figure 4.
+The prediction in such an environment is that an agent would discard the branch of the decision tree involving an immediate large negative loss even though this branch objectively contains the reward maximizing (loss minimizing) sequence of choices.
 
-\cite{Huys2012} (and later \cite{Lally2017}) tested this prediction in the
-choice envrionment displayed in Figure 4. Participants received extensive training
-in this decision tree environment (first without outcomes and then later with
-rewards) in order to facilitate planning. In free-planning trials, many participants
-exhibited this Pavlovian bias, rejecting choice sequences which involved the large
-loss even when the sequence was objectively reward maximizing (loss minimizing).
-Importantly \cite{Huys2012} and \cite{Lally2017} found the degree of aversive
-pruning was correlated with depressive and anxiety symptoms, respectively.
+\cite{Huys2012} (and later \cite{Lally2017}) tested this prediction in the choice envrionment displayed in Figure 4.
+Participants received extensive training in this decision tree environment (first without outcomes and then later with rewards) in order to facilitate planning.
+In free-planning trials, many participants exhibited this Pavlovian bias, rejecting choice sequences which involved the large loss even when the sequence was objectively reward maximizing (loss minimizing).
+Importantly \cite{Huys2012} and \cite{Lally2017} found the degree of aversive pruning was correlated with depressive and anxiety symptoms, respectively.
 
-The aversive pruning principle makes tremendous sense and draws on a robust literature
-of heuristics and computational shortcuts for circumventing computationally intractable
-cognitive processes. It is important to note, however, that a control interpretation
-also makes similar predictions. Figure 4b-d shows the preferred routes of simulated
-agents in the same environments under increasingly pessimistic expectations of
-future actions. As can be observed, pessimistic agents are similarly likely to avoid
-the branch with the large loss, though for different reasons. Under the pessimistic case,
-the agent is increasingly less confident that the large gains later in
-sequence will be realized; as such, the agent is less confident that the initial
-large loss will be offset. Consequently, pessimistic agents would prefer the branch
-with smaller losses (even if this is objectively suboptimal). This prediction
-is similarly consistent with the finding that avoidance of the initial large loss
-is correlated with anxiety \cite{Lally2017}.
+The aversive pruning principle makes tremendous sense and draws on a robust literature of heuristics and computational shortcuts for circumventing computationally intractable cognitive processes.
+It is important to note, however, that a control interpretation also makes similar predictions.
+Figure 4b-d shows the preferred routes of simulated agents in the same environments under increasingly pessimistic expectations of future actions.
+As can be observed, pessimistic agents are similarly likely to avoid the branch with the large loss, though for different reasons.
+Under the pessimistic case, the agent is increasingly less confident that the large gains later in sequence will be realized; as such, the agent is less confident that the initial large loss will be offset.
+Consequently, pessimistic agents would prefer the branch with smaller losses (even if this is objectively suboptimal).
+This prediction is similarly consistent with the finding that avoidance of the initial large loss is correlated with anxiety \cite{Lally2017}.
 
-An interpretation like this would be consistent with reduced beliefs in one's
-own competency in such an environment. This is not strictly surprising. Planning
-multi-choice sequences even in relatively simple environments like those used
-in these experiments still ostensibly require working memory ability to keep in
-mind the possible consequences along each branch. Insofar that working memory
-ability is disrupted in anxiety disorders \citep{Moran2016}, a pessimistic belief
-may be justified and possibly optimal.
+An interpretation like this would be consistent with reduced beliefs in one's own competency in such an environment.
+This is not strictly surprising.
+Planning multi-choice sequences even in relatively simple environments like those used in these experiments still ostensibly require working memory ability to keep in mind the possible consequences along each branch.
+Insofar that working memory ability is disrupted in anxiety disorders \citep{Moran2016}, a pessimistic belief may be justified and possibly optimal.
 
-It is also worht noting that the two theories can be easily disentangled with a
-simple manipulation of the choice environment. The two theories predict
-different choices when the large loss comes later in sequence. For aversive pruning,
-late large losses should have little effect on choice behavior. For pessimistic
-learning, however, the value of large losses should back-propagate and contaminate
-its associated branch. Future experiments can tease these apart as tests of the theory.
+It is also worht noting that the two theories can be easily disentangled with a simple manipulation of the choice environment.
+The two theories predict different choices when the large loss comes later in sequence.
+For aversive pruning, late large losses should have little effect on choice behavior.
+For pessimistic learning, however, the value of large losses should back-propagate and contaminate its associated branch.
+Future experiments can tease these apart as tests of the theory.
 
@@
 
-If an agent is confident that their compentency in decision-making in intact,
-then it follows that situations in which an individual is able to make a choice
-should be preferable to situations in which an individual is not. At worst,
-an agent is no better off than if they had not made a choice; at best, an agent
-is better able to steer themselves toward desirable outcomes. Such is the logic
-that underlies an increasingly robust literature suggesting choice is inherently
-valuable \citep{Leotti2010}.
+If an agent is confident that their compentency in decision-making in intact, then it follows that situations in which an individual is able to make a choice should be preferable to situations in which an individual is not.
+At worst, an agent is no better off than if they had not made a choice; at best, an agent is better able to steer themselves toward desirable outcomes.
+Such is the logic that underlies an increasingly robust literature suggesting choice is inherently valuable \citep{Leotti2010}.
 
-This free choice premium has been demonstrated multiple times in humans \citep{Suzuki1997,Leotti2011,Leotti2014,Cockburn2014} using a variety of decision
-paradigms. For the present purposes, we will describe the experiments presented in
-\citep{Leotti2011,Leotti2014}. In these experiments, participants complete a
-two-stage decision making task (Figure 5). In the first stage, participants are
-make a choice between two cues: one leading to free choice and the other leading
-to a fixed choice. In the second stage, participants are able to choose between
-a second set of cues (free choice) or are forced to choose a cue (fixed choice).
-Unbeknownst to participants, all cues yield reward following the same outcome
-distribution; thus, no cue is advantageous in the long-run. The free choice
-premium is measured as a preference for the cue leading to free choice. \cite{Leotti2011,Leotti2014}
-find participants still prefer the free choice cue, despite this conferring no
-objective benefit to participants.
+This free choice premium has been demonstrated multiple times in humans \citep{Suzuki1997, Leotti2011, Leotti2014, Cockburn2014} using a variety of decision paradigms.
+For the present purposes, we will describe the experiments presented in \citep{Leotti2011, Leotti2014}.
+In these experiments, participants complete a two-stage decision making task (Figure 5).
+In the first stage, participants are make a choice between two cues: one leading to free choice and the other leading to a fixed choice.
+In the second stage, participants are able to choose between a second set of cues (free choice) or are forced to choose a cue (fixed choice).
+Unbeknownst to participants, all cues yield reward following the same outcome distribution; thus, no cue is advantageous in the long-run.
+The free choice premium is measured as a preference for the cue leading to free choice.
+\cite{Leotti2011, Leotti2014} find participants still prefer the free choice cue, despite this conferring no objective benefit to participants.
 
-Insofar that this effect is contingent on the belief of one's own competency,
-it stands to reason that this free choice premium should be absent in individuals
-with anxiety and anxious control beliefs. We demonstrate this effect in Figure 5.
-In each trial, the reward outcomes are equal for all options (-1, 1). In individuals
-with increasingly pessimistic beliefs, the worst outcome in free choice is increasingly
-back-propagated to the free-choice cue, thereby making it less attractive than
-the fixed choice cue. Hence, pessimistic learners, on average, do not exhibit
-this choice bias.
+Insofar that this effect is contingent on the belief of one's own competency, it stands to reason that this free choice premium should be absent in individuals with anxiety and anxious control beliefs.
+We demonstrate this effect in Figure 5.
+In each trial, the reward outcomes are equal for all options (-1, 1).
+In individuals with increasingly pessimistic beliefs, the worst outcome in free choice is increasingly back-propagated to the free-choice cue, thereby making it less attractive than the fixed choice cue.
+Hence, pessimistic learners, on average, do not exhibit this choice bias.
 
@@
 
-In this final section, we want to discuss the relationship between anxiety and
-depression. Anxiety and depression are highly comorbid, with roughly 45% of individuals
-with lifetime depression diagnosis also diagnosed with at least one anxiety disorder
-\citep{kessler2015}. This fact has long been recognized in the clinical literature,
-and prominent theories of depression have suggested that anxious control is a necessary
-precursor to at least some types of depression \citep{alloy1990}. Recent large-scale
-epidemiological studies have provided  some empirical support for this claim, finding that
-anxiety symptoms precede and predict the onset of depression \citep{mathew2011, jacobson2014,
-kessler2015} (though also see \cite{jacobson2017, plana2019}).
+In this final section, we want to discuss the relationship between anxiety and depression.
+Anxiety and depression are highly comorbid, with roughly 45% of individuals with lifetime depression diagnosis also diagnosed with at least one anxiety disorder \citep{kessler2015}.
+This fact has long been recognized in the clinical literature, and prominent theories of depression have suggested that anxious control is a necessary precursor to at least some types of depression \citep{alloy1990}.
+Recent large-scale epidemiological studies have provided  some empirical support for this claim, finding that anxiety symptoms precede and predict the onset of depression \citep{mathew2011, jacobson2014, kessler2015} (though also see \cite{jacobson2017, plana2019}).
 
-An emerging idea in this literature is that pathological avoidance as a result
-of anxiety mediates the transition to depression, perhaps by reduce the likelihood
-of experiencing positive events and activities \citep{moitra2008, jacobson2014}.
-This finding is perfectly in line with the model presented so far. A sketch of
-how this is possible is shown in Figure 6. Under unperturbed learning, avoidable
-threat does not impede behavior. Under pessimistic learning, threat can block
-future reward. This avoidance can manifest as a failure of behavioral initiation,
-or a lack of overall mobility (here represented by an agent self-transitioning
-at the starting state). Thus, the model can accommodate the deficits in positive
-affect observed in depression as a function of behavioral avoidance, as predicted
-by these epidemiological studies.
+An emerging idea in this literature is that pathological avoidance as a result of anxiety mediates the transition to depression, perhaps by reduce the likelihood of experiencing positive events and activities \citep{moitra2008, jacobson2014}.
+This finding is perfectly in line with the model presented so far.
+A sketch of how this is possible is shown in Figure 6.
+Under unperturbed learning, avoidable threat does not impede behavior.
+Under pessimistic learning, threat can block future reward.
+This avoidance can manifest as a failure of behavioral initiation, or a lack of overall mobility (here represented by an agent self-transitioning at the starting state).
+Thus, the model can accommodate the deficits in positive affect observed in depression as a function of behavioral avoidance, as predicted by these epidemiological studies.
 
@@
 
-All of the anxiety disorders share core symptoms including pessimistic inference,
-second-order conditioning, and sustained avoidance. In this article, we offered
-a simple computational account of how these aberrations may arise and be maintained.
-Following the literature in Markov decision processes, we showed how a simple
-breakdown in expectations about the likelihood of encountering future negative
-outcomes (as a result of an unpredictable environment or unreliable policy) can
-effectively backpropagate negative expectations through an agent's internal value
-map. The result of which yields expectations and behaviors resembling that observed
-in pathological anxiety. Importantly, we also showed that such a model can account
-for aberrant behavior previously observed in individuals with anxiety in several
-assays of psychiatric behavior. By no means a complete theory of anxiety, this
-account helps to at least tie together several core symptoms of anxiety in one
-account.
+All of the anxiety disorders share core symptoms including pessimistic inference, second-order conditioning, and sustained avoidance.
+In this article, we offered a simple computational account of how these aberrations may arise and be maintained.
+Following the literature in Markov decision processes, we showed how a simple breakdown in expectations about the likelihood of encountering future negative outcomes (as a result of an unpredictable environment or unreliable policy) can effectively backpropagate negative expectations through an agent's internal value map.
+The result of which yields expectations and behaviors resembling that observed in pathological anxiety.
+Importantly, we also showed that such a model can accountfor aberrant behavior previously observed in individuals with anxiety in several assays of psychiatric behavior.
+By no means a complete theory of anxiety, this account helps to at least tie together several core symptoms of anxiety in one account.
 
-This computational account draws upon a longstanding recognition of the importance
-of perceived control to the development and maintenance of anxiety disorders. Several
-prominent theories of anxiety in the clinical anxiety focus on the importance of
-a lack of perceived control, including helplessness-hopelessness theories, perceived
-efficacy, and the triple vulnerability model. It is worth noting that that these
-theories differ, among other things, in where a lack of control arises. For learned
-helplessness, an agent believes to occupy an uncontrollable world; in other words,
-there is no reliable mapping between action and outcome. For self-efficacy theory,
-an agent believes to be limited or incapable of exerting beneficial influence on
-the environment, though this is otherwise possible. Note that both can easily be
-modeled in the current approach, either in the state transition function or in
-future policy function. Insofar that they are multiplicative, they should result
-in the same effect. Future work may tease these apart, as they may be interesting
-for guiding treatment.
+This computational account draws upon a longstanding recognition of the importance of perceived control to the development and maintenance of anxiety disorders.
+Several prominent theories of anxiety in the clinical anxiety focus on the importance of a lack of perceived control, including helplessness-hopelessness theories, perceived efficacy, and the triple vulnerability model.
+It is worth noting that that these theories differ, among other things, in where a lack of control arises.
+For learned helplessness, an agent believes to occupy an uncontrollable world; in other words, there is no reliable mapping between action and outcome.
+For self-efficacy theory, an agent believes to be limited or incapable of exerting beneficial influence on the environment, though this is otherwise possible.
+Note that both can easily be modeled in the current approach, either in the state transition function or in future policy function.
+Insofar that they are multiplicative, they should result in the same effect.
+Future work may tease these apart, as they may be interesting for guiding treatment.
 
-It is worth noting that we are not the first to attempt to formalize a theory
-of control in MDP. \cite{HuysDayan2009} provided a first computational account
-of learned helplessness through simple models of action-outcome contingencies.
-\cite{HuysDayan2009} found that training an agent first in an unreliable environment
-led to later impaired learning, akin to early learned helplessness experiments,
-by virtue of the prior of controllability. This provides a nice demonstration
-of learning impairment of action-outcome contingencies. Importantly, our accounts
-in at least three respects. We handle sequential decision environments, which is
-important as it is in sequential decision environments where the consequences
-of pessimistic inference become most pronounced. Second, our account claims
-that the worst outcomes are believed, wherein their account focuses on randomness.
-This seems closer to the symptoms described in anxiety. Finally, our account
-provides insights into maintenance. With an otherwise unmodified Bayesian learner,
-their agents should eventually learn and be ok.
+It is worth noting that we are not the first to attempt to formalize a theory of control in MDP.
+\cite{HuysDayan2009} provided a first computational account of learned helplessness through simple models of action-outcome contingencies.
+\cite{HuysDayan2009} found that training an agent first in an unreliable environment led to later impaired learning, akin to early learned helplessness experiments, by virtue of the prior of controllability.
+This provides a nice demonstration of learning impairment of action-outcome contingencies.
+Importantly, our accounts in at least three respects.
+We handle sequential decision environments, which is important as it is in sequential decision environments where the consequences of pessimistic inference become most pronounced.
+Second, our account claims that the worst outcomes are believed, wherein their account focuses on randomness.
+This seems closer to the symptoms described in anxiety.
+Finally, our account provides insights into maintenance.
+With an otherwise unmodified Bayesian learner, their agents should eventually learn and be ok.
 
-We discussed two mechanisms by which negative values can spread. There is a third
-that we did not address here: that of the state itself. Here we assumed that states
-were fully observable; in other words, there was no uncertainty about the state
-an agent was currently occupying. In partially-observable Markov decision problems,
-the state itself must be inferred from external and internal variables. This is
-yet another means by which bad value can spread if undue likelihood is granted
-to undesirable or aversive states. This ideas has previously been proposed to
-explain some symptoms in anxiety \citep{Paulus2012}, and not without good reason.
-Anxiety disorders involve negatively biased interpretation of ambiguous percepts
-\citep{Hartley2012}, which may result in aberrant planning and increased avoidance
-for similar reasons to those described above. This is an area for future research.
+We discussed two mechanisms by which negative values can spread.
+There is a third that we did not address here: that of the state itself.
+Here we assumed that states were fully observable; in other words, there was no uncertainty about the state an agent was currently occupying.
+In partially-observable Markov decision problems, the state itself must be inferred from external and internal variables.
+This is yet another means by which bad value can spread if undue likelihood is granted to undesirable or aversive states.
+This ideas has previously been proposed to explain some symptoms in anxiety \citep{Paulus2012}, and not without good reason.
+Anxiety disorders involve negatively biased interpretation of ambiguous percepts \citep{Hartley2012}, which may result in aberrant planning and increased avoidance for similar reasons to those described above.
+This is an area for future research.
 
-In this article, we have explored a computational principle of approach and
-avoidance learning in sequential decision environments. We have not discussed,
-however, the particular psychological or neurobiological mechanisms by which this
-sort of learning may take place. This issue has been discussed in detail elsewhere
-\citep{Bishop2018}, but we briefly discuss a few possibilities here. Mechanically,
-we understand that complex decision problems are solved by mental processes that
-exist on a spectrum between model-based and model-free (Daw et al., 2005). The
-difference between these two families of methods hinges on the reliance of an
-internal model of the world used to make predictions about expected long-run
-outcomes of particular choices.
+In this article, we have explored a computational principle of approach and avoidance learning in sequential decision environments.
+We have not discussed, however, the particular psychological or neurobiological mechanisms by which this sort of learning may take place.
+This issue has been discussed in detail elsewhere \citep{Bishop2018}, but we briefly discuss a few possibilities here.
+Mechanically, we understand that complex decision problems are solved by mental processes that exist on a spectrum between model-based and model-free (Daw et al., 2005).
+The difference between these two families of methods hinges on the reliance of an internal model of the world used to make predictions about expected long-run outcomes of particular choices.
 
-For model-based decision making, one possibility is biased planning. In large
-state spaces, online sequential decision making suffers from the curve of
-dimensionality: in planning a series of actions to take, comparing between
-multiple options rapidly becomes computationally intractable when the number of
-options and depth of search becomes even moderately large. Thus, heuristics for
-reducing the size of this search can be useful. Biased pruning of search paths
-(such as immediately discarding paths that require some minor loss in pursuit of
-larger gains) may result in maladaptive decision making (Huys et al., 2012).
-Alternately, if planning relies on internal sampling from previously experienced
-episodes in order to make predictions about future value, then biased sampling
-may similarly result in maladaptive decision making. Recent work has shown how
-finite sampling during planning can result in risk aversion and the overrepresentation
-of low probability but strongly negative outcomes (Lieder et al., 2018). Such a
-mechanism is prima facie in line with results showing the availability in memory
-of negative outcomes being higher for patients with anxiety (Borkovec et al., 1999;
-Miranda and Mennin, 2007).
+For model-based decision making, one possibility is biased planning.
+In large state spaces, online sequential decision making suffers from the curve of dimensionality: in planning a series of actions to take, comparing between multiple options rapidly becomes computationally intractable when the number of options and depth of search becomes even moderately large.
+Thus, heuristics for reducing the size of this search can be useful.
+Biased pruning of search paths (such as immediately discarding paths that require some minor loss in pursuit of larger gains) may result in maladaptive decision making \citep{Huys2012}.
+Alternately, if planning relies on internal sampling from previously experienced episodes in order to make predictions about future value, then biased sampling may similarly result in maladaptive decision making.
+Recent work has shown how finite sampling during planning can result in risk aversion and the overrepresentation of low probability but strongly negative outcomes (Lieder et al., 2018).
+Such a mechanism is prima facie in line with results showing the availability in memory of negative outcomes being higher for patients with anxiety \citep{Borkovec1999, Miranda2007}.
 
-Within the realm of model-free decision making, several studies have found evidence
-suggesting increased learning rates for negative reward prediction errors in
-individuals with elevated levels of anxiety \citep{Aylward, Huang2017, Harle2017
-garrett2018}. Though an asymmetry in sensitivity to
-positive and negative reward prediction errors bears superficial similarity, for
-the reasons described above merely attending more to negative information does not
-necessarily predict increased avoidance insofar that avoidance can successfully
-isolate negative outcomes. One possibility, and topic of future research, is how
-prior beliefs about the reward statistics of the current environment dictate
-learning rates. A second possibility is prioritized replay (Russek et al., 2017;
-Mattar and Daw, 2018). Offline replay of previous experiences (through hippocampal
-mechanisms) are known to facilitate learning. It has been proposed that biased
-replay could bias value estimates (Gagne et al., 2018). This is another exciting
-area for future research.
+Within the realm of model-free decision making, several studies have found evidence suggesting increased learning rates for negative reward prediction errors in individuals with elevated levels of anxiety \citep{Harle2017, Garrett2018}.
+Though an asymmetry in sensitivity to positive and negative reward prediction errors bears superficial similarity, for the reasons described above merely attending more to negative information does not necessarily predict increased avoidance insofar that avoidance can successfully isolate negative outcomes.
+One possibility, and topic of future research, is how prior beliefs about the reward statistics of the current environment dictate learning rates.
+A second possibility is prioritized replay (Russek et al., 2017; Mattar and Daw, 2018).
+Offline replay of previous experiences (through hippocampal mechanisms) are known to facilitate learning.
+It has been proposed that biased replay could bias value estimates (Gagne et al., 2018).
+This is another exciting area for future research.
 

### 2019-02-04 manuscript/manuscript.tex
major edits to manuscript

@@
 
-\title{Anxiety: a decision-theoretic perspective}
+\title{Title Pending}
 
@@
 \begin{abstract}
-To be filled in
+Pending
 \end{abstract}
@@
 \keywords{
-reinforcement learning; avoidance; anxiety
+decision theory; computational psychiatry; anxiety; value iteration
 }
@@
 
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/01_field.png}}%
-  }
-  \par \textbf{Figure 1:} The open field is a deterministic gridworld with only one rewarding (blue) and aversive (red) state (A).
-For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value (B).
-For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states (C).
-With increasing pessimism ($w=0$), the extent of the spread grows worse (D).
-(Parameters: $\gamma = 0.95$)
-\end{figure}
-
 In this first section, we demonstrate how learning under an assumption of a lack of future control naturally leads to pessimistic inference and second-order conditioning.
@@
 
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/01_field.png}}%
+  }
+  \par \textbf{Figure 1:} The open field is a deterministic gridworld with only one rewarding (blue) and aversive (red) state (A).
+For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value (B).
+For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states (C).
+With increasing pessimism ($w=0$), the extent of the spread grows worse (D).
+(Parameters: $\gamma = 0.95$)
+\end{figure}
+
 The long-run value estimates learned by an optimistic agent (i.e.
@@
 
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/02_cliff.png}}%
-  }
-  \par \textbf{Figure 2:} Cliff-walking is a deterministic gridworld where an agent must navigate from one side (S) to the goal on the other side (G) (A).
-Every step costs the agent ($r=-1$) but stepping off the cliff results in great harm ($r=-100$).
-An optimistic agent learns to travel the shortest (loss-minimizing) path along the cliff (B).
-For a pessimistic agent ($w=0.8$), the potential for harm yields an optimal policy that maintains distance from the cliff's edge (C).
-With increasing pessimism ($w=0.6$), the agent learns an optimal path off the cliff so as to minimize unnecessary steps prior to its perceived inevitable harm (D).
-(Parameters: $\gamma = 1$)
-\end{figure}
-
 In addition to pessimistic inference, all anxiety disorders share pathological avoidance as core symptom \citep{dsm5, Krypotos2015, Arnaudova2017}.
@@
 As has been discussed elsewhere \citep{Moutoussis2017}, such accounts only explain the maintenance of avoidance in short timescales, but do not explain why resistant to change and why they do not go away in general.
+
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/02_cliff.png}}%
+  }
+  \par \textbf{Figure 2:} Cliff-walking is a deterministic gridworld where an agent must navigate from one side (S) to the goal on the other side (G) (A).
+Every step costs the agent ($r=-1$) but stepping off the cliff results in great harm ($r=-100$).
+An optimistic agent learns to travel the shortest (loss-minimizing) path along the cliff (B).
+For a pessimistic agent ($w=0.8$), the potential for harm yields an optimal policy that maintains distance from the cliff's edge (C).
+With increasing pessimism ($w=0.6$), the agent learns an optimal path off the cliff so as to minimize unnecessary steps prior to its perceived inevitable harm (D).
+(Parameters: $\gamma = 1$)
+\end{figure}
 
@@
 
-Next, we turn our attention towards a related phenomenon: a bias towards avoidance in approach-avoidance conflict.
-In environments with correlated risk and rewards, a bias towards avoidance can ultimately result in fewer expected rewards and worse long-run returns.
-Returning to the example above, an individual with social anxiety may forgo social situations to avoid the risk of social embarrassment at the expense of friendships, etc.
-This is another way in which pathological anxiety can be disruptive to everyday functioning.
+Next, we turn our attention to a related phenomenon observed in anxiety: a bias towards avoidance during approach-avoidance conflict \citep{aupperle2010}.
+In situations where the potential for reward and harm are correlated,
+anxiety predicts the tendency to forego reward in favor of minimizing the risk of threat.
+As a consequence, individuals with anxiety may experience fewer rewards in their environment as compared to non-anxious individuals.
+For example, an individual with social anxiety may avoid a party for fear of social embarrassment at the expense of positive social interactions with friends.
+This is a primary mechanism by which clinical anxiety can disrupt everyday functioning.
 
-Experimentally, there is a long history of measuring behavior during approach-avoidance conflict in individuals with anxiety (citation, maybe yin-yang anxiety paper).
-One popular paradigm is the balloon analog risk task (BART) \citep{Lejuez2002}, which has been shown to correlate with anxiety \cite{Maner2007, Giorgetta2012}.
-In the task, a participant pumps a balloon with air.
-For each pump, the participant earns points.
-If she pumps too much, the balloon pops and she loses all points.
-The participant must decide within each episode when to stop pumping.
-Leave time is the measure of behavior, with earlier leave times in anxiety suggesting a greater tendency towards avoidance.
+A relationship between anxiety and biased decision making in approach-avoidance conflict is well-documented in the empirical literature.
+One common measure is the balloon analog risk task (BART) \citep{Lejuez2002}.
+The task is straightforward: a participant inflates a virtual balloon, one pump at a time.
+The more inflated the balloon, the more points the participant earns.
+With increasing pumps, however, comes the risk of the balloon popping and the loss of all earned points so far.
+Thus, there is a conflict: a participant can choose to pump the balloon more to earn more points, at the risk of the balloon popping; alternately, the participant can choose to end the trial earlier with fewer points earned.
+Previous studies have found that anxiety is correlated with earlier leaving times, or a greater tendency to avoid risk \cite{Maner2007, Giorgetta2012}.
 
@@
   }
-  \caption{\textbf{BART}}
-  \par remember to add parameters
+  \par \textbf{Figure 3:} The Balloon Analog Risk Task (BART).
+Risk of the balloon popping increases with each pump (top row).
+An optimistic agent ($w=1$) continues to pump until the risk of losing all points earned outweighs the marginal value of one additional pump (second row).
+A pessimistic agent ($w=0.8$) is less confident that it will make the optimal decision in the future, and thus leaves earlier (third row).
+Leaving time becomes sooner with increasing pessimism ($w=0.6$) (bottom row).
+(Parameters: $\gamma = 1$.)
 \end{figure}
 
-Our sequential model of decision making easily explains these results.
-A schematic of the task is presented in Figure 3.
-Under optimistic beliefs, points should travel back from the mean.
-Under pessimistic assumptions, however, different degrees of the worst outcome backpropagate.
-At a certain point, the perceived cost of staying outweighs the guaranteed gains of leaving.
-Thus, patients with anxiety leave earlier.
+This result can be accounted for by a loss of confidence in future reliability.
+A simulation of performance on the BART is shown in Figure 3.
+For an optimistic agent ($w=1$), the optimal policy is to continue pumping until the risk of popping is all but guaranteed.
+The optimal policy for pessimistic agents differ.
+Under a loss of confidence for future action, pessimistic agents cannot be certain they will make optimal choices in the future.
+As such, negative value propagates backwards resulting in earlier leaving times.
+The latency of leaving is dependent on the strength of the pessimistic belief.
 
-We can explain similar findings with the flight initiation distance \citep{Mobbs2018, Mobbs2019}.
-One thing our model cannot explain is the neural difference between fast and slow escape.
-Indeed, our model cannot capture all but perhaps is relevant for certain calculations.
-One could imagine these beliefs interacting in such a way to change the distance calculations that are necessary for mediating between passive and active defense responses.
+Interestingly, these results are similar to recent findings regarding flight initiation distance and anxiety \citep{Mobbs2018, Mobbs2019}.
+Flight initiation distance refers to a measure from behavioral ecology that assesses the leaving time of a foraging animal at risk of predation.
+In classic flight initiation scenarios, an animal can continue to forage and exploit resources, but the longer it does so the greater the risk it will encounter a predator.
+Thus, flight initiation scenarios are similar in structure to approach-avoidance conflict and the BART.
+Recently, \cite{Mobbs2019} found that anxiety is correlated with earlier leaving times in virtual flight initiation tasks.
+Interestingly, the authors make a distinction between "fast-" and "slow-attacking" predators, and find that trait anxiety is only correlated only in behavioral response to the latter.
+The findings here draw upon an emerging literature of distinct neural systems mediating "cognitive" vs.
+"reactive" threat responses, where anxiety evidently impacts the deliberative, cognitive responses.
+Thus, it is important to note the present theory can account for anxious approach-avoidance behavior in some, but certainly not all, threatening scenarios.
 
-This can also be used to explain behavioral inhibition \citep{bach2015, khemka2017}.
-The idea is slightly different than what is advocated.
-The claim here is that in anxious individuals, the Q-values should be closer together, and thus there should be longer waiting time.
+These results may also be related to recent work on the computational mechanisms underlying behavioral inhibition, a temperament linked to anxiety \citep{bach2015, khemka2017}.
+\cite{bach2015} argues that delayed behavioral responses under threat are normative under the assumption that a pause in behavior (e.g.
+freezing) can be useful for surveying a threatening environment, avoiding detection by predators, and potentially waiting out a nearby threat.
+Behavioral inhibition, it is argued, is a useful Pavlovian heuristic that has evolved over time.
+The current models may predict similar results.
+Insofar that a loss of confidence in future action yields more negative estimates of action values, then the best action may often be brought closer in value to its alternatives.
+As such, a pessimistic agent may face increased decision conflict as compared to a non-anxious counterpart and therefore show increased response times.
+This is not to argue against this particular model of behavioral inhibition, but just to say that our account may yield similar behaviors.
+
+\subsection{Relation to Depression}
+
+As an extension of approach-avoidance conflict, we next want to discuss the relationship between anxiety and depression.
+Anxiety and depression are highly comorbid, with roughly 45\% of individuals with a lifetime depression diagnosis also diagnosed with at least one anxiety disorder \citep{kessler2015}.
+Given this, many have clinical theories assume that depression and anxiety are slightly variant manifestations of the same underlying mechanisms involving negative affect (for review, see \cite{jacobson2014}).
+
+Where the two disorders can be distinguished is in relation to rewards and anhedonia; depressive disorders, but not anxiety disorders, are associated with disrupted reward processing.
+One interesting proposal in the literature is that at least certain types of anhedonic depression are preceded by anxiety \citep{alloy1990, moitra2008, jacobson2014}.
+Specifically, persistent avoidance has the long-run effect of reducing the likelihood of anxious individuals encountering positive events and activities in their everyday lives.
+As a consequence, anxious individuals come to expect an absence of reward in their environments and subsequently develop diminished reward motivation and behavioral activation.
+Recent large-scale epidemiological studies have provided  some empirical support for this claim, finding that anxiety symptoms and behavioral avoidance precede and predict the longitudinal onset of depression (\cite{mathew2011, jacobson2014, kessler2015}; though also see \cite{jacobson2017, plana2019}).
+
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/06_lh.png}}%
+  }
+  \par \textbf{Figure 4:} Progression from anxiety to depression.
+In this open environment, an agent starts opposite a reward (blue) with harm separating them (A).
+For an optimistic agent ($w=1$), positive value spreads from the reward to the starting tile (Panels B-D, left-to-right).
+Harm is mitigated through the belief of future self-efficacy.
+For a pessimistic agent ($w=0.6$), both positive and negative value spread (Panels E-G, left-to-right).
+Given the agent's closer proximity to harm than reward, the optimal policy for this agent is to stay at the initial state.
+(Parameters: $\gamma = 0.95$)
+\end{figure}
+
+This prediction is easily accommodated by the present model.
+Simulations of how this might come back is presented in Figure 4.
+In this environment, there is only one rewarding and harmful state.
+Just as in the open field environment, the world is deterministic and the harmful state poses no true threat to an agent.
+An optimistic agent ($w=1$), with a belief in its ability to maintain an optimistic policy, is unfazed by the presence of the harmful state and develops over time an overall positive representation of the value of the environment (Figure 4b-d).
+
+In contrast, a pessimistic agent ($w=0.6$) cannot be certain of its ability to steer clear of future harm.
+As such, negative value spreads from the harmful state to its antecedent states (Figure 4e-g).
+Because in this environment risk and reward are correlated (i.e.
+the agent must approach the harmful state in order to reach the reward given its starting position), the consequence of negatively spreading value is that the optimal policy is to stay in the initial position.
+Thus, the pessimistic agent adopts a behavioral policy akin to depressive immobility.
+Moreover, reward is inaccessible to this agent such that its positive expectations of the environment are also diminished.
+In summary, a loss of confidence in future reliability can result in behavioral policies that resemble depressive-like states in environments with correlated risk and reward.
 
@@
 Aversive pruning is defined as a Pavlovian response to encountering a large loss in planning such that sequences involving large losses are discarded from further evaluation.
-An example of scenario is presented in Figure 4.
-The prediction in such an environment is that an agent would discard the branch of the decision tree involving an immediate large negative loss even though this branch objectively contains the reward maximizing (loss minimizing) sequence of choices.
+An example of such a scenario is presented in Figure 5.
+The prediction in this decision tree environment is that an agent would discard the branch of the decision tree involving an immediate large negative loss, even though this branch objectively contains the reward maximizing (loss minimizing) sequence of choices.
 
-\cite{Huys2012} (and later \cite{Lally2017}) tested this prediction in the choice envrionment displayed in Figure 4.
+\cite{Huys2012} (and later \cite{Lally2017}) tested this prediction in this particular choice environment.
 Participants received extensive training in this decision tree environment (first without outcomes and then later with rewards) in order to facilitate planning.
@@
 
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/04_tree.png}}%
+  }
+  \par \textbf{Figure 5:} The decision tree environment from \cite{Huys2012} and \cite{Lally2017} (A).
+\end{figure}
+
 The aversive pruning principle makes tremendous sense and draws on a robust literature of heuristics and computational shortcuts for circumventing computationally intractable cognitive processes.
 It is important to note, however, that a control interpretation also makes similar predictions.
-Figure 4b-d shows the preferred routes of simulated agents in the same environments under increasingly pessimistic expectations of future actions.
+Figure 5b-d shows the preferred routes of simulated agents in the same environments under increasingly pessimistic expectations of future actions.
 As can be observed, pessimistic agents are similarly likely to avoid the branch with the large loss, though for different reasons.
@@
 
-An interpretation like this would be consistent with reduced beliefs in one's own competency in such an environment.
+%% An interpretation like this would be consistent with reduced beliefs in one's own competency in such an environment.
 This is not strictly surprising.
@@
 
-It is also worht noting that the two theories can be easily disentangled with a simple manipulation of the choice environment.
-The two theories predict different choices when the large loss comes later in sequence.
-For aversive pruning, late large losses should have little effect on choice behavior.
-For pessimistic learning, however, the value of large losses should back-propagate and contaminate its associated branch.
-Future experiments can tease these apart as tests of the theory.
-
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/04_tree.png}}%
-  }
-  \caption{\textbf{Decision Tree}}
-  \par remember to add parameters
-\end{figure}
+It is worth noting that these two alternate accounts can be disentangled through slight manipulations of the choice environment presented above.
+Aversive pruning and loss of confidence make similar predictions when a large loss begins a sequence, but not necessarily when a large potential loss comes later in a sequence.
+In the loss of confidence view, the potential for large loss late in a sequence negatively impacts the estimated value of actions early in the sequence; this is not true for aversive pruning.
+Identifying such choice sequences in the decision tree task may help identify whether this sort of inference is present in individuals with anxiety performing the task.
 
@@
 
-If an agent is confident that their compentency in decision-making in intact, then it follows that situations in which an individual is able to make a choice should be preferable to situations in which an individual is not.
+In this final section, we propose based on the present theory that anxiety should be inversely related to the free choice bias.
+If an agent is confident in their future decision making abilities, then it follows that situations in which an individual is able to make a future choice should be preferable to situations in which an individual is not.
 At worst, an agent is no better off than if they had not made a choice; at best, an agent is better able to steer themselves toward desirable outcomes.
-Such is the logic that underlies an increasingly robust literature suggesting choice is inherently valuable \citep{Leotti2010}.
+Such is the logic that underlies an emerging literature suggesting that choice is inherently valuable \citep{Leotti2010}.
 
-This free choice premium has been demonstrated multiple times in humans \citep{Suzuki1997, Leotti2011, Leotti2014, Cockburn2014} using a variety of decision paradigms.
+This free choice premium has been observed across several behavioral experiments \citep{Suzuki1997, Leotti2011, Leotti2014, Cockburn2014} using a variety of decision paradigms.
 For the present purposes, we will describe the experiments presented in \citep{Leotti2011, Leotti2014}.
-In these experiments, participants complete a two-stage decision making task (Figure 5).
+In these experiments, participants complete a two-stage decision making task (Figure 6).
 In the first stage, participants are make a choice between two cues: one leading to free choice and the other leading to a fixed choice.
@@
 
-Insofar that this effect is contingent on the belief of one's own competency, it stands to reason that this free choice premium should be absent in individuals with anxiety and anxious control beliefs.
-We demonstrate this effect in Figure 5.
-In each trial, the reward outcomes are equal for all options (-1, 1).
-In individuals with increasingly pessimistic beliefs, the worst outcome in free choice is increasingly back-propagated to the free-choice cue, thereby making it less attractive than the fixed choice cue.
-Hence, pessimistic learners, on average, do not exhibit this choice bias.
-
+Insofar that the free choice premimum is contingent on an agent's belief in its future ability, it stands to reason that this free choice premium should be absent in individuals with anxiety and a loss of confidence.
+Simulations of this effect are presented in Figure 6.
+Simulated agents complete an experiment similar to that used in \cite{Leotti2011, Leotti2014}.
+In contrast to optimistic agents, the value of the worst action in the free choice arm back-propagated to the free-choice cue in pessimistic agents.
+With increasing pessimism then, the free choice option is less preferable than the fixed choice arm.
+As a result, pessimistic learners, on average, do not exhibit a bias towards the free choice arm.
 To our knowledge this prediction has not been empirically tested.
@@
   }
-  \caption{\textbf{Free Choice Premium}}
-  \par remember to add parameters
-\end{figure}
-
-\subsection{Longitudinal Progression}
-
-In this final section, we want to discuss the relationship between anxiety and depression.
-Anxiety and depression are highly comorbid, with roughly 45% of individuals with lifetime depression diagnosis also diagnosed with at least one anxiety disorder \citep{kessler2015}.
-This fact has long been recognized in the clinical literature, and prominent theories of depression have suggested that anxious control is a necessary precursor to at least some types of depression \citep{alloy1990}.
-Recent large-scale epidemiological studies have provided  some empirical support for this claim, finding that anxiety symptoms precede and predict the onset of depression \citep{mathew2011, jacobson2014, kessler2015} (though also see \cite{jacobson2017, plana2019}).
-
-An emerging idea in this literature is that pathological avoidance as a result of anxiety mediates the transition to depression, perhaps by reduce the likelihood of experiencing positive events and activities \citep{moitra2008, jacobson2014}.
-This finding is perfectly in line with the model presented so far.
-A sketch of how this is possible is shown in Figure 6.
-Under unperturbed learning, avoidable threat does not impede behavior.
-Under pessimistic learning, threat can block future reward.
-This avoidance can manifest as a failure of behavioral initiation, or a lack of overall mobility (here represented by an agent self-transitioning at the starting state).
-Thus, the model can accommodate the deficits in positive affect observed in depression as a function of behavioral avoidance, as predicted by these epidemiological studies.
-
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/06_lh.png}}%
-  }
-  \caption{\textbf{Free Choice Premium}}
-  \par remember to add parameters
+  \par \textbf{Figure 6:} Schematic of the free choice bias tasks used in \cite{Leotti2011, Leotti2014} (A). The free choice bias is diminished in increasingly pessimistic agents (B).
 \end{figure}

### 2019-02-04 manuscript/manuscript.tex
edits to discussion

@@
   \par \textbf{Figure 5:} The decision tree environment from \cite{Huys2012} and \cite{Lally2017} (A).
+An optimistic agent ($w=1$) learns the optimal loss-minimizing policy through the initial large loss (B).
+A moderately pessimistic agent ($w=0.75$) will find both arms equally preferable insofar that it is less confident it will be able to execute optimal future choices (C).
+More pessimistic agents comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss (D).
 \end{figure}
@@
 
-All of the anxiety disorders share core symptoms including pessimistic inference, second-order conditioning, and sustained avoidance.
-In this article, we offered a simple computational account of how these aberrations may arise and be maintained.
-Following the literature in Markov decision processes, we showed how a simple breakdown in expectations about the likelihood of encountering future negative outcomes (as a result of an unpredictable environment or unreliable policy) can effectively backpropagate negative expectations through an agent's internal value map.
-The result of which yields expectations and behaviors resembling that observed in pathological anxiety.
-Importantly, we also showed that such a model can accountfor aberrant behavior previously observed in individuals with anxiety in several assays of psychiatric behavior.
-By no means a complete theory of anxiety, this account helps to at least tie together several core symptoms of anxiety in one account.
+Central to the acquisition and maintenance of anxiety disorders are symptoms including pessimistic inference, second-order conditioning, and persistent avoidance.
+In this article, we presented a simple computational account of these aberrations in learning and decision making.
+Specifically, we showed how a loss of confidence in the reliability of the future self and/or the environment can effectively backpropagate negative value across states of the environment.
+The result of this process are a series of inferences and behaviors resembling those observed in clinical anxiety.
+Though this is by no means a complete account of anxiety, the present model helps to explain the core symptoms of anxiety under one unifying framework.
 
-This computational account draws upon a longstanding recognition of the importance of perceived control to the development and maintenance of anxiety disorders.
-Several prominent theories of anxiety in the clinical anxiety focus on the importance of a lack of perceived control, including helplessness-hopelessness theories, perceived efficacy, and the triple vulnerability model.
-It is worth noting that that these theories differ, among other things, in where a lack of control arises.
-For learned helplessness, an agent believes to occupy an uncontrollable world; in other words, there is no reliable mapping between action and outcome.
-For self-efficacy theory, an agent believes to be limited or incapable of exerting beneficial influence on the environment, though this is otherwise possible.
-Note that both can easily be modeled in the current approach, either in the state transition function or in future policy function.
-Insofar that they are multiplicative, they should result in the same effect.
+Importantly, this computational account draws upon a longstanding recognition of the importance of perceived control to the development and maintenance of anxiety disorders.
+Central to many prominent theories of anxiety in the clinical literature highlight is a perceived lack of control.
+For example, learned helplessness theory and its successors claim that pathological anxiety results from a stable belief that the environment is uncontrollable, such that threat cannot be effectively mitigated.
+In contrast, Bandura's self-efficacy theory of anxiety posited that a lack of confidence in one's own ability to effectively handle threat was central to the pathogenesis of anxiety.
+Later theories, such as the triple vulnerability model, are equivocal as to whether an unreliable future environment or self is the root cause of anxiety, but highlights control beliefs as being a necessary compontent for developing anxiety.
+As we note above, the present model can accommodate either effect.
+Insofar that the state transition function and future state-action function are multiplicative, a pessimistic change to either should make similar predictions.
+It is possible that under certain conditions they may predict different outcomes.
 Future work may tease these apart, as they may be interesting for guiding treatment.
@@
 This provides a nice demonstration of learning impairment of action-outcome contingencies.
-Importantly, our accounts in at least three respects.
-We handle sequential decision environments, which is important as it is in sequential decision environments where the consequences of pessimistic inference become most pronounced.
-Second, our account claims that the worst outcomes are believed, wherein their account focuses on randomness.
-This seems closer to the symptoms described in anxiety.
-Finally, our account provides insights into maintenance.
-With an otherwise unmodified Bayesian learner, their agents should eventually learn and be ok.
+Importantly, our accounts differ in at least three respects.
+First, we consider sequential decision environments, which is necessary for more accurately characterizing the symptoms of anxiety disorders.
+Second, our account specifies a mechanism by which the worst outcomes are realized; in their account, the world simply becomes random.
+This pessimism seems closer to what is observed in anxiety disorders.
+Finally, our account accounts for the persistence of these symptoms.
+Presumably, their model would eventually learn to escape with additional training; our account shows aberrant behaviors in the optimal (long-run) policy.
 
-We discussed two mechanisms by which negative values can spread.
+In this article, we discussed two mechanisms by which negative values can spread, through the state transition distribution and future state-action action.
 There is a third that we did not address here: that of the state itself.
@@
 This issue has been discussed in detail elsewhere \citep{Bishop2018}, but we briefly discuss a few possibilities here.
-Mechanically, we understand that complex decision problems are solved by mental processes that exist on a spectrum between model-based and model-free (Daw et al., 2005).
-The difference between these two families of methods hinges on the reliance of an internal model of the world used to make predictions about expected long-run outcomes of particular choices.
-
-For model-based decision making, one possibility is biased planning.
-In large state spaces, online sequential decision making suffers from the curve of dimensionality: in planning a series of actions to take, comparing between multiple options rapidly becomes computationally intractable when the number of options and depth of search becomes even moderately large.
-Thus, heuristics for reducing the size of this search can be useful.
-Biased pruning of search paths (such as immediately discarding paths that require some minor loss in pursuit of larger gains) may result in maladaptive decision making \citep{Huys2012}.
-Alternately, if planning relies on internal sampling from previously experienced episodes in order to make predictions about future value, then biased sampling may similarly result in maladaptive decision making.
+One possibility is in biased sampling.
+If decision making and planning may rely on internal sampling from previously experienced episodes in order to make predictions about future value, then biased sampling may result in maladaptive decision making.
 Recent work has shown how finite sampling during planning can result in risk aversion and the overrepresentation of low probability but strongly negative outcomes (Lieder et al., 2018).
 Such a mechanism is prima facie in line with results showing the availability in memory of negative outcomes being higher for patients with anxiety \citep{Borkovec1999, Miranda2007}.
+Another possibility are an overreliance on Pavlovian heuristics, such as aversive pruning \citep{Huys2012,Lally2017}.
+As discussed above, future work should tease apart the predictions of our model and the aversive pruning heuristic.
 
-Within the realm of model-free decision making, several studies have found evidence suggesting increased learning rates for negative reward prediction errors in individuals with elevated levels of anxiety \citep{Harle2017, Garrett2018}.
-Though an asymmetry in sensitivity to positive and negative reward prediction errors bears superficial similarity, for the reasons described above merely attending more to negative information does not necessarily predict increased avoidance insofar that avoidance can successfully isolate negative outcomes.
-One possibility, and topic of future research, is how prior beliefs about the reward statistics of the current environment dictate learning rates.
-A second possibility is prioritized replay (Russek et al., 2017; Mattar and Daw, 2018).
+Several recent studies have found evidence suggesting increased learning rates for negative reward prediction errors in individuals with elevated levels of anxiety \citep{Harle2017, Garrett2018, Aylward2019}.
+Though incearned learning from negative reward prediction errors may result in anxiety-like behaviors, these are likely not the sole cause of anxiety disorders for the reasons discussed above.
+Future research might investigate the relationship of prior beliefs about the reward statistics of the environment and learning to positive and negative prediction errors.
+A separate interestisng possibility is prioritized replay (Russek et al., 2017; Mattar and Daw, 2018).
 Offline replay of previous experiences (through hippocampal mechanisms) are known to facilitate learning.
 It has been proposed that biased replay could bias value estimates (Gagne et al., 2018).
-This is another exciting area for future research.
 

### 2019-02-17 manuscript/manuscript.tex
updated introduction

@@
 
-Many psychiatric disorders involve distortions to representations of the external world and, as a likely consequence, pronounced changes to learning and decision making.
-As a normative framework, decision theory allows us to describe and pose quantita­tive questions about optimal choice behavior \citep{DayanDaw2008} and is, therefore, a critical tool for modeling, understanding, and predicting the psychological and neurobiological mechanisms underlying psychiatric behavior \citep{maia2011, HuysDawDayan2015}.
-Recent investigations employing this framework have provided novel and important insights into psychiatric symptoms and disorders, including anhedonia in major depression \citep{Rutledge2017}, mood instability in bipolar disorder \citep{EldarNiv2015, EldarDolanNiv2016}, habitual behavior in obsessive compulsive and addictive disorders \citep{gillan2016}, and hallucinations in schizophrenia \citep{powers2017, corlett2018}.
+Central to the experience and maintenance of anxiety disorders are the generalization of fear and persistent avoidance.
+In fear generalization, the set of perceived threats expands to include secondary cues or situations associated with the primary danger.
+For example, a dog phobic individual may first develop a fear of dog parks, then parks in general, and finally any public space where a dog may be plausibly encountered.
+As a consequence, avoidance behavior becomes more prolific often at considerable cost to the anxious individual, either in foregone rewards or in effort expenditure.
+The same dog phobic individual may take lengthy detours to avoid proximity to feared locations and may even abandon social obligations for fear of encountering a dog.
+Importantly, anxious avoidance is persistent despite the many safety precautions often in place that limit or negate the possibility of danger (e.g.
+fences, leash laws).
+These symptoms are puzzling then insofar the objective risk of danger is often so low and the cost of anxious behavior so high (as is often acknowledged by the patients themselves).
 
-The family of anxiety disorders (which includes generalized anxiety disorder, social anxiety disorder, panic disorder, and specific phobia) are characterized by excessive fear and worry about future negative outcomes, and represent the most common group of psychiatric disorders \citep{kessler2009}.
-Though the disorders typically differ in the types of objects or situations that induce fear or anxiety, all share a set of core disturbances to cognition and behavior.
-First, the anxiety disorders involve pessimistic inference, or a tendency to appraise threatening objects or situations as more likely and/or more severe than they objectively are.
-Clinical anxiety is also associated with second-order conditioning, or learning to fear neutral cues indirectly associated with threat cues.
-Finally, anxiety is associated with avoidance, or the tendency to flee or avoid threatening situations.
-As an example, an individual with a clinical phobia of dogs may show chronic worry about coming into contact with aggressive dogs (pessimistic inference), may develop fear towards parks or other public spaces where they may encounter dogs (second-order conditioning), and might exert greater effort to avoid contact with dogs or locations associated with dogs (avoidance).
+From the perspective of decision theory, present choice (e.g.
+to approach or avoid) is contingent on assumptions about the future, such as the reliability of the environment or the set of available future choices.
+A common finding in clinical research is that anxiety is correlated with pessimistic expectations about the future, such that anxious individuals are more likely to endorse a belief in a lack of control over future danger.
+As has been noted elsewhere \cite{Krypotos2015}, a great many studies of anxious decision making have focused on avoidance behavior in response to immediate danger rather than on early avoidance of potential future threat.
+The latter are arguably more important to the understanding and treatment of anxiety, as much of its disruption to everyday functioning stems from repeated avoidance of anticipated threat.
 
-Recent work has begun to shine some light into the mechanisms underlying these symptoms.
-For example, a number of studies have found that anxiety is correlated with increased learning from worse-than-predicted outcomes \citep{Harle2017, Garrett2018, Aylward2019}.
-Systematic biases in learning could result in pessimistic inference by virtue of negatively biasing those same cached values used to make predictions at the time of retrieval.
-This begs the question as to why learning from negative prediction errors is greater in anxiety, and others have suggested this may reflect prior expectations in anxiety; that pessimistic expectations bias learning from expectation-congruent information.
-This raises the possibility of circular arguments.
+Here, we explore these issues by offering a decision theoretic account of sequential choice under pessimistic future expectations.
+We consider optimal behavior in deterministic, finite Markov decision processes (MDP), in which agents learn reward-maximizing (or loss-minimizing) policies in grid-world environments requiring series of choices.
+Specifically, our model treats pessimistic expectations as the aberrant assumption that an agent will be unable to act consistently with its preferences in the future.
+In other words, our model assumes that agents are reward-maximizing manner in present choice but expect that such choices will be unavailable to them in the future.
+We show that this minor deviation from standard temporal difference learning models captures many features of anxious behavior, including exaggerated threat appraisal, fear generalization, and persistent avoidance.
 
-Other studies have attempted to provide novel insights into avoidance.
-\cite{Maia2010} offered a reinforcement learning-based account of maintained avoidance in rodent conditioning tasks, an empirical finding that previous theories of avoidance could not account for \citep{Krypotos2015}.
-They argue that sustained avoidance is possible insofar that a well-established avoidance response generates no reward prediction error, such that no new learning takes place, and avoidance continues.
-As is discussed elsewhere \citep{Moutoussis2017}, the framework cannot account for sustained avoidance when an agent is accidentally exposed to an extinguished outcome.
-Returning to the previous example, the framework cannot account for why dog phobia remains for an individual in spite of an unplanned exposure to a friend's benign dog.
-More broadly, the framework cannot for why avoidance, and anxiety in general, is so persistent even with exposure therapy \citep{Craske2014}.
-
-In the remainder of this article, we present a simple decision theoretic account of how the core symptoms of anxiety may arise and be maintained.
-Specifically, we demonstrate that how a loss of confidence in the reliability of the self or the environment can result in the catastrophic spread of negative value in sequential decision environments.
-We show how, under these assumptions, pessimistic inference, second-order conditioning, and chronic avoidance naturally follow, as well as a number of other behaviors previously found to be correlated with anxiety.
+In the remaining sections, we also demonstrate that this simple model can account for the findings reported in the handful of empirical studies measuring anxious choice in sequential choice environments.
+We then offer a speculative prediction from the model that, consistent with negative future expectations, anxious individuals should not exhibit the free choice preference observed in non-anxious individuals.
+Finally, we discuss the relationship of the present model to clinical theories postulating a role for control and helplessness in anxiety disorders.
 

### 2019-02-17 manuscript/manuscript.tex
considerable rewrites.

@@
 
-\section{Methods}
+\section{The model}
 
-\subsection{Markov Decision Processes}
-
-Before moving forward, we first briefly review the formalism of Markov decision processes, which provide the framework for our results.
+We first briefly review the formalism of Markov decision processes, which will provide the framework for our results.
 For complete treatments, see \cite{SuttonBarto1998, SuttonBarto2018} and \cite{bertsekas2005}.
 
-A MDP is defined by a set of states, $S$, a set of actions, $A$, a reward function defined over state/action pairs, $R(s,a)$, and a state transition distribution, $P(s'|s,a)$, where $a$ denotes a chosen action.
-In a MDP, states and rewards occur sequentially according to the one-step transition structure where transitions follow actions.
-The goal for an agent in a MDP is to learn to choose a policy, $\pi$, or mapping of state to action, that maximizes long-run expected reward.
-Here we define the value of a state under a policy, $V^\pi(s)$, as the expected cumulative discounted reward:
+A MDP is defined by a set of states, $S$, a set of actions, $A$, a reward function defined over state-action pairs, $R(s,a)$, and a state transition distribution, $P(s'|s,a)$.
+In a MDP, states and rewards are experienced sequentially according to chosen actions and the one-step transition structure.
+The goal for an agent in a MDP is to learn a policy, or mapping of actions to states, that maximizes expected cumulative (discounted) reward.
+The value of a state can be defined recursively as the sum of the immediate reward received following an action, $R(s, a)$, and the value of its successor state $s’$, averaged over potential future transitions and actions according to the current policy $\pi$.
+Under Bellman optimality, the value of a state under the optimal policy is equal to the expected return for the best action from that state.
 
-$$ V^\pi(s) = \mathbb{E} \left[ \sum^{\infty}_{k=0} \gamma^k R_{t+k} | S_t = s \right] $$
+$$ V^*(s) = \max_a \sum_{s',r} p(s',r|s,a) \left[ r + \gamma v(s') \right] $$
 
-where $\gamma$ is the temporal discounting parameter, controlling the weighting of distant future rewards.
-The value function can be defined recursively as the sum of the immediate reward following an action chosen in a state, $R(s, a)$, and the value of its successor state $s’$, averaged over potential future actions, $a$, and transitions according to the current policy $\pi$:
+where $\gamma$ is the temporal discounting parameter, which controls the influence of distant future rewards.
 
-$$ V^\pi(s) = \sum_a \pi(a|s) \sum_{s',r}p(s',r|s,a) \left[ r + \gamma v(s') \right] $$
+Crucially then, the value of a state under the optimal policy is contingent on an agent's particular expectations about the future.
+First, it is a function of the state transition distribution, or beliefs about the likelihood of encountering particular states conditioned on undertaking some action.
+Second, it is also a function of the actions available in those successor states, as those in turn define the value of these states.
+Under Bellman optimality, it is normative for an agent to assume that the value of a successor state is equivalent to the best action from that state.
+This can be expressed as:
 
-Under the optimal policy, the value function is given by:
+$$ V^*(s) = \max_a \sum_{s',r}p(s',r|s,a) \left[ r + \gamma \max_{a'} q_*(s',a') \right] $$
 
-$$ V^*(s) = \max_a \mathbb{E} \left[ \sum^{\infty}_{k=0} \gamma^k R_{t+k} | S_t = s \right] $$
+where $q(s',a')$ denotes the a state-action value. As we will demonstrate, this assumption has profound consequences for learning about approach and avoidance behaviors.
 
-$$ = \max_a \sum_{s',r}p(s',r|s,a) \left[ r + \gamma v_*(s') \right] $$
+In a MDP, the discovery of a rewarding action by an agent has an effect not only on the associated state but also on its antecedent states.
+This follows naturally from the recursive definition of state values defined above.
+So when an agent learns a particular state predicts reward, the states which immediately precede it also acquire value (as do their antecedent states, and so on) thereby propagating positive value across the state space so as to reinforce reward-maximizing policies.
+This process does not occur, however, upon the discovery of a suboptimal or aversive action.
+Because under Bellman optimality the value of a state is dependent in part on the best possible future action, the discovery of harmful actions will not necessarily impact state value estimates or the optimal policy.
+Insofar that suboptimal or aversive actions can be avoided, then Bellman optimality predicts a fundamental asymmetry: reward opportunities propagate recursively to antecedent states, but danger does not.
 
-Using the formalisms above, the value of one particular action in a state under the optimal policy is defined as:
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/01_field.png}}%
+  }
+  \par \textbf{Figure 1:} The open field is a deterministic gridworld with only one rewarding (blue) and aversive (red) state (A).
+For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value (B).
+For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states (C).
+With increasing pessimism ($w=0$), the extent of the spread grows worse (D).
+(Parameters: $\gamma = 0.95$)
+\end{figure}
 
-$$ Q^\pi(s,a) = \mathbb{E} \left[ \sum^{\infty}_{k=0} \gamma^k R_{t+k} | S_t = s \right] $$
+This principle is highlighted in Figure 1.
+In an open field environment, an agent is free to navigate a discrete, deterministic grid-world devoid of rewards except for two states: one rewarding (blue) and one harmful (red, Figure 1a).
+In this environment, the one-step transition distribution respects the agent's choice; that is, the agent comes into experiences an aversive outcome only if it decides to approach danger.
+As can be observed, the state value estimates learned by an agent under Bellman optimality (Figure 1b) reflect the asymmetry described above.
+Despite the presence of danger, the agent learns to associate all states (except the harmful state itself) with positive value as the reward opportunity propagates outwards to all antecedent states.
+Insofar that danger can be successfully avoided, all states (even those adjacent to the danger) come to predict expected future reward.
 
-$$ = \sum_{s',r}p(s',r|s,a) \left[ r + \gamma v_*(s') \right] $$
+What happens, however, when an agent believes that danger is not perfectly avoidable and safety not guaranteed?
+Such a belief may manifest as inaccurate assumptions about the state transition distribution.
+For example, an agent may wrongly believe it possible that it could transition to a state proximal to danger in spite of its choices.
+Alternately, this belief may manifest as pessimistic expectations about the availability of future actions.
+An agent may believe it has the capacity to act in accordance with the (reward-maximizing) optimal policy in the present, but may be unable to do so in the future.
+Importantly, such beliefs about the unreliability of the environment or of the self are commonly endorsed in anxiety disorders.
 
-Insofar that the value of a state under the optimal policy is defined as the best action under that state, we can re-express the above equivalently as:
-
-$$ q_*(s,a) = \sum_{s',r}p(s',r|s,a) \left[ r + \gamma \max_{a'} q_*(s',a') \right] $$
-
-Thus in a conventional MDP, the value of states in the environment under the optimal policy assumes the agent is able to act according to the optimal policy in the future.
-
-\subsection{Pessimistic Learning}
-
-In the previous section, we observed that the value of a state and its associated state-action pairs are a function of the environment transition distribution and the best future action.
-In standard MDP problems, it is typically assumed that the transition distribution is known and accurate, and that the agent is consistently able to operate under the optimal policy in the future.
-In this section, we introduce a formalism deviating from these standard assumptions.
-
-Central to many theories of clinical anxiety is a perceived loss of control (see discussion for elaboration).
-In the MDP framework described above, this may come in one of several forms.
-First, an agent may perceive the transition structure of the environment to be less reliable than it actually is.
-Put another way, an agent may have a biased expectation that, despite choosing an action $a$ with the intent of transition to state $s'$, the agent may instead transition to an alternate, undesired state $s^$.
-In the case of clinical anxiety, the expectation may be more dire, that the environment is "adversarial" is likely to transition an agent, not randomly, but to the worse states.
-
-Second, an agent may doubt its own agentive control.
-An agent may expect to be able to choose the best possible action for itself in the present, but may not be confident it will still be able to decide optimally for itself in the future.
-In either event, following the value computations defined above, we should expect the value estimates for such an agent to diminish.
-If an agent believes itself or the environment to be unreliable, then the agent cannot be optimistic about its future and should instead adopt a different policy.
-
-These sorts of problems have been explored in the safe reinforcement learning and robust control literature \citep{Garcia2015}.
-There, we are concerned with programming agents that identify the optimal policy under pessimistic assumptions.
-A number of approaches have been proposed in that literature.
-Here, we focus on one particular simple implemented proposed by \cite{Gaskett2003}:
+We can explicitly model this type of pessimism about future expectations through a minor change to the state value function defined above.
+Here, we focus on one particular formulation, previously introduced by \cite{Gaskett2003}, that captures the expectation that future action may deviate from the optimal reward-maximizing policy.
+We define the value of the successor state as:
 
@@
 
-In this formalism, $w$ controls the degree of pessimism.
+where the parameter $w$ controls the degree of pessimism.
 When $\w = 1$, an agent is maximally optimistic and fully expects to act according to the best possible actions in the future.
@@
 When $w = 0.5$, the weighs the best and worst possible future actions equally.
+We note we are not committed to this particular implementation as anything more than a computational convenience.
+Many alternate functions could be defined over the space of possible future actions.
+(As mentioned above, we could also modify the transition distribution.) This formalism highlights an important computational principle; it does not constitute a mechanistic claim about anxiety disorders.
 
-We note that we are not committed to this particular implementation of a pessimistic agent.
-As mentioned above, an alternative parameterization would have been to program an agent with a belief in an adversarial environment, preferentially transitioning it to worse states.
-Insofar that these two terms are multiplicative in value calculations, either implementation should yield similar results.
-Similarly, the w-pessimism function could easily be replaced with other functions (e.g.
-softmax with negative inverse temperature).
-This formalism is simply convenient and is useful for demonstrating the points we demonstrate below.
-It is not a substantive claim about anxiety.
-
-\subsection{Simulations}
+\section{Simulations}
 
@@
 All simulations were carried out with the python programming language and are publicly available on Github (https://github.com/szorowi1/SecretFunTimes).
-
-\section{Results}
 
@@
 In sum, anxiety is associated with inflated estimates of threat, both likelihood and value, and these estimates can spread from the source to distantly associated cues.
-
-In our first simulation, we demonstrate how this may come about.
-In the open field environment (Figure 1a), an agent is free to navigate a large grid world devoid of value except for two states: one rewarding and one aversive red.
-In this environment, the one-step transition structure is deterministic such that the environment fully respects the agent's choice.
-As such, the agent is under no threat despite the presence of a harmful state.
-Insofar that the agent never acts under a masochistic policy and intentionally approaches harm, the agent need not ever encounter the harmful state.
-
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/01_field.png}}%
-  }
-  \par \textbf{Figure 1:} The open field is a deterministic gridworld with only one rewarding (blue) and aversive (red) state (A).
-For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value (B).
-For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states (C).
-With increasing pessimism ($w=0$), the extent of the spread grows worse (D).
-(Parameters: $\gamma = 0.95$)
-\end{figure}
-
-The long-run value estimates learned by an optimistic agent (i.e.
-an agent expecting to be able to follow a reward-maximizing policy now and in the future) reflect this (Figure 1b).
-For an optimistic agent, all states in the open field environment, sans the harmful state itself, take on a positive value.
-This is unsurprising insofar that, except for the harmful state, the agent will always encounter a reward in the long-run no matter its starting position in the field.
 

### 2019-02-18 manuscript/manuscript.tex
reorganization of section 2

@@
 
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/01_field.png}}%
-  }
-  \par \textbf{Figure 1:} The open field is a deterministic gridworld with only one rewarding (blue) and aversive (red) state (A).
-For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value (B).
-For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states (C).
-With increasing pessimism ($w=0$), the extent of the spread grows worse (D).
-(Parameters: $\gamma = 0.95$)
-\end{figure}
-
 This principle is highlighted in Figure 1.
@@
 
-\section{Simulations}
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/01_field.png}}%
+  }
+  \par \textbf{Figure 1:} The open field is a deterministic gridworld with only one rewarding (blue) and aversive (red) state (A).
+For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value (B).
+For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states (C).
+With increasing pessimism ($w=0$), the extent of the spread grows worse (D).
+(Parameters: $\gamma = 0.95$)
+\end{figure}
 
-In the following sections, we present the results of simulations of pessimistic learning in a variety of MDP environments.
-In all simulations, the state-action values under the optimal policy, $Q_*(s,a)$, were solved for directly through Q-value iteration \citep{SuttonBarto1998, SuttonBarto1998, bertsekas2005}.
-The details of this algorithm are presented in the appendix.
-Model parameters are reported for each environment in their corresponding sections.
-All simulations were carried out with the python programming language and are publicly available on Github (https://github.com/szorowi1/SecretFunTimes).
+The consequences of this manipulation for the open field environment are presented in Figure 1c,d.
+When Bellman optimality is violated (i.e.
+an agent no longer assumes the value of a successor state is equal the best available action), more than reward opportunities are able to propagate value through the state space.
+Indeed when $w$ is non-zero, the agent treats the future potential of danger as possible and thus negative value is able to spread from its source recursively through antecedent states thereby degrading the value of distal states (the degree of which determined by the extent of pessimism, $w$, and temporal discounting, $\gamma$).
+With increasing certainty that the worst policy will be followed in the future, the larger the extent to which threat value can permeate.
 
-\subsection{Pessimistic Inference}
-
-In this first section, we demonstrate how learning under an assumption of a lack of future control naturally leads to pessimistic inference and second-order conditioning.
-A symptom central to all of the anxiety disorders is pessimistic inference, or a tendency to appraise the likelihood and/or severity of threatening situations out of proportion to their actual environmental contingencies \citep{dsm5, BeckClark1997, ClarkBeck2011}.
-For example, an individual with specific phobia may vastly exaggerate the probability of encountering their fear (e.g.
-a venomous spider) in their everyday lives.
-Anxious individuals will also tend to associate fear with cues tangentially related to signals of threat.
-For example, the same individual may begin to fear locations where they may encounter spider webs (e.g.
-basements).
-
-Pessimistic inference in anxiety is well-documented.
-In studies of self-reported beliefs, individuals with anxiety rate hypothetical future negative outcomes as being more likely to occur and worse to experience than non-anxious counterparts \citep{ButlerMathews1983, ButlerMathews1987, Foa1996, MacLeod1996, MacLeod1997, Luten1997, Stober1997, Borkovec1999, Maner2006, Mitte2007, Miranda2007}.
-In the laboratory, exaggerated threat appraisal in anxiety has traditionally been measured through fear conditioning.
-The results of these studies demonstrate that anxious individuals exhibit greater fear towards threatening cues, non-threatening cues, and extinguished threatening cues as compared to non-anxious counterparts \citep{lissek2005, MinekaOehlberg2008, Duits2015}.
-Importantly, threatening cues can become associated with secondary neutral cues, allowing negative value to spread \citep{wessa2007}.
-In sum, anxiety is associated with inflated estimates of threat, both likelihood and value, and these estimates can spread from the source to distantly associated cues.
-
-The same cannot be said for a pessimistic agent.
-If an agent doubts its future ability to act in according with the optimal policy, then it cannot rule out that, in the future, it may inadvertently encounter the harmful state.
-This is more so likely if the agent is already proximal to the harmful state.
-Under this pessimistic belief, the possibility of future harm is no longer zero, as it is for the optimistic agent, and thus negative value can propagate from the states antecedent to harm, back to their respective antecedents and so on.
-The extent of this spread will depend on the strength of the pessimistic belief (Figures 1c, 1d).
-With increasing pessimism, harmful outcomes are assigned greater credence and thus exert greater and wider influence on antecedent states.
-
-In sequential decision environments then, a perceived lack of self-efficacy can result in the propagation of negative value from potential future harms.
-Such a pessimistic agent may correspondingly report an increased likelihood of threat (i.e.
-a greater number of threatening states), increased severity of threat (i.e.
-states with more negative value), and fear of states antecedent to the fear source.
-As such, pessimistic learning can account for two of the hallmark symptoms of the anxiety disorders.
-
-\subsection{Avoidance}
+In addition to drawing inspiration from clinical theories of anxiety, this pessimistic model also captures several core symptoms of anxiety disorders.
+The hallmark symptom of anxiety disorders is exaggerated threat appraisal, or the tendency to perceive threat as disproportionately greater in likelihood and severity than is justified.
+This is evident in the anxious state value estimates presented in Figure 1c,d.
+Indeed when safety is perceived to be uncertain, threat in the environment is exaggerated in magnitude and prevalence.
+As should be apparent, this very same mechanism captures the generalization of fear that often accompanies anxiety disorders.
 
@@
 Computerized avoidance tasks have become increasingly common, and increased distance from virtual threat has been found to correlate with anxiety \citep{Bach2014, Bach2017, Sheynin2014}.
-
-There is a long history of models attempting to explain avoidance learning.
-The fundamental challenge is explaining how an action can be reinforced in the absence of an outcome.
-Two-factor theory was the explanation for a long while, but suffered from a number of noteworthy criticisms \citep{Krypotos2015}.
-Recent work has revitalized two-factor theory and integrated it with reinforcement learning \citep{Moutoussis2008, Maia2010}, and in the process addressing several criticisms of two-factor theory.
-As has been discussed elsewhere \citep{Moutoussis2017}, such accounts only explain the maintenance of avoidance in short timescales, but do not explain why resistant to change and why they do not go away in general.
-
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/02_cliff.png}}%
-  }
-  \par \textbf{Figure 2:} Cliff-walking is a deterministic gridworld where an agent must navigate from one side (S) to the goal on the other side (G) (A).
-Every step costs the agent ($r=-1$) but stepping off the cliff results in great harm ($r=-100$).
-An optimistic agent learns to travel the shortest (loss-minimizing) path along the cliff (B).
-For a pessimistic agent ($w=0.8$), the potential for harm yields an optimal policy that maintains distance from the cliff's edge (C).
-With increasing pessimism ($w=0.6$), the agent learns an optimal path off the cliff so as to minimize unnecessary steps prior to its perceived inevitable harm (D).
-(Parameters: $\gamma = 1$)
-\end{figure}
-
-In our second simulation, we demonstrate how pessimistic learning can yield sustained avoidance.
-In the cliff-walking environment \citep{SuttonBarto1998, SuttonBarto2018, Gaskett2003}, an agent must navigate from a start on side of the map to the other (Figure 2a).
-The agent experiences a small penalty for every step it takes.
-The shortest (least costly) path is along the cliff's edge; if the agent steps off the cliff, however, it incurs a much larger penalty.
-As in the first environment, we assume a deterministic transition structure.
-
-Just as in the first environment, the long-run value estimates by an optimistic agent do not reflect the potential harm of the cliff (Figure 2b).
-Because this agent (rightfully) expects to act according to the optimal policy in the future (i.e.
-never step off the cliff), the agent perceives no risk of future disaster and thus the large penalties of stepping off the cliff do not pollute the agent's value estimates.
-The optimistic agent learns the loss-minimizing policy and walks along the cliff's edge.
-
 This is not the case for the pessimistic agent.
@@
 
-Under pessimistic learning then, a perceived lack of self-efficacy can yield the propagation of negative value from potential future harms.
-In such a scenario, the optimal policy for an agent is to maintain a distance from threat far enough so as to optimally balance the risk of catastrophe with foregone losses.
-Importantly, these simulations demonstrate the long-run value estimates rather than any momentary value estimates.
-In other words, the behavioral policy these agents demonstrate will be maintained insofar beliefs about self-efficacy are maintained, irrespective of any momentary noisiness in actual outcomes.
-Thus, pessimistic learning can account for pathological avoidance in anxiety disorders.
+\section{Simulations}
+
+In the following sections, we present the results of simulations of pessimistic learning in a variety of MDP environments.
+In all simulations, the state-action values under the optimal policy, $Q_*(s,a)$, were solved for directly through Q-value iteration \citep{SuttonBarto1998, SuttonBarto1998, bertsekas2005}.
+The details of this algorithm are presented in the appendix.
+Model parameters are reported for each environment in their corresponding sections.
+All simulations were carried out with the python programming language and are publicly available on Github (https://github.com/szorowi1/SecretFunTimes).
 

### 2019-02-18 manuscript/manuscript.tex
next round of edits ready for Nathaniel

@@
 \keywords{
-decision theory; computational psychiatry; anxiety; value iteration
+anxiety; avoidance; fear generalization; decision theory; computational psychiatry
 }
@@
 
-Central to the experience and maintenance of anxiety disorders are the generalization of fear and persistent avoidance.
-In fear generalization, the set of perceived threats expands to include secondary cues or situations associated with the primary danger.
-For example, a dog phobic individual may first develop a fear of dog parks, then parks in general, and finally any public space where a dog may be plausibly encountered.
-As a consequence, avoidance behavior becomes more prolific often at considerable cost to the anxious individual, either in foregone rewards or in effort expenditure.
-The same dog phobic individual may take lengthy detours to avoid proximity to feared locations and may even abandon social obligations for fear of encountering a dog.
-Importantly, anxious avoidance is persistent despite the many safety precautions often in place that limit or negate the possibility of danger (e.g.
-fences, leash laws).
-These symptoms are puzzling then insofar the objective risk of danger is often so low and the cost of anxious behavior so high (as is often acknowledged by the patients themselves).
+Anxiety disorders are characterized by aberrations in the cognitive processing of and response to threat.
+The hallmark symptom of anxiety disorders is exaggerated threat appraisal, or the tendency to perceive threat as disproportionately greater in likelihood and/or severity than is warranted.
+Clinical anxiety often also involves the generalization of fear, wherein an individual comes to perceive as threatening cues or situations secondarily associated with the primary fear.
+For example, a dog phobic individual may first fear dog parks, then parks in general, and finally any public space where a dog may plausibly be encountered.
+Unsurprisingly then, anxiety disorders involve persistent avoidance behavior.
+Avoidance behaviors in anxiety are especially pernicious as they prevent potential exposure to disconfirming evidence about perceived threats, thereby maintaining the disorder, and because they are often highly disruptive to everyday functioning.
+Avoidance of perceived threat often comes at considerable expense, either in foregone rewards or in the expenditure of effort.
+On their walk to work, the same dog phobic individual may take increasingly circuitous paths in order to avoid certain public spaces and may abandon social obligations for fear of encountering a dog.
+These behaviors are puzzling to theories of learning and choice behavior, not in the least because they are still observed even in safe situations posing little to no objective threat.
 
-From the perspective of decision theory, present choice (e.g.
-to approach or avoid) is contingent on assumptions about the future, such as the reliability of the environment or the set of available future choices.
-A common finding in clinical research is that anxiety is correlated with pessimistic expectations about the future, such that anxious individuals are more likely to endorse a belief in a lack of control over future danger.
-As has been noted elsewhere \cite{Krypotos2015}, a great many studies of anxious decision making have focused on avoidance behavior in response to immediate danger rather than on early avoidance of potential future threat.
-The latter are arguably more important to the understanding and treatment of anxiety, as much of its disruption to everyday functioning stems from repeated avoidance of anticipated threat.
+Some insight into anxious cognition might possibly be gleaned from a decision theoretic perspective.
+In decision theory, the value of potential actions (e.g.
+to approach or avoid) is defined in part as a function of expectations about the future.
+The best action is that which yields the greatest cumulative reward, contingent on current beliefs of what will likely be encountered in the future.
+Interestingly, anxiety disorders are associated with pessimistic beliefs about the future; anxious individuals are more likely to endorse beliefs like that they will surely face danger in the future and be unable to respond effectively to it.
+Despite this common finding, most studies of anxious decision making interrogate behavior in response to immediate but not future threat.
+Studying anxious appraisal of future threat may be more important for understanding anxiety disorders as symptoms like fear generalization and avoidance show worry for would-be but not immediate danger.
 
 Here, we explore these issues by offering a decision theoretic account of sequential choice under pessimistic future expectations.
-We consider optimal behavior in deterministic, finite Markov decision processes (MDP), in which agents learn reward-maximizing (or loss-minimizing) policies in grid-world environments requiring series of choices.
-Specifically, our model treats pessimistic expectations as the aberrant assumption that an agent will be unable to act consistently with its preferences in the future.
-In other words, our model assumes that agents are reward-maximizing manner in present choice but expect that such choices will be unavailable to them in the future.
-We show that this minor deviation from standard temporal difference learning models captures many features of anxious behavior, including exaggerated threat appraisal, fear generalization, and persistent avoidance.
+We consider optimal behavior in deterministic, finite Markov decision processes (MDP), in which agents learn reward-maximizing (or loss-minimizing) policies in grid world environments requiring a series of choices.
+We first describe a fundamental asymmetry in the differential influence of reward and danger on value estimation under normative models of learning, namely that reward propagates recursively but danger does not.
+We then demonstrate how this asymmetry is eliminated when agents instead learn with anxious-like expectations of future action.
+We find that a minor deviation from normativity captures many features of anxious behavior including exaggerated threat appraisals, fear generalization, and persistent avoidance.
 
-In the remaining sections, we also demonstrate that this simple model can account for the findings reported in the handful of empirical studies measuring anxious choice in sequential choice environments.
-We then offer a speculative prediction from the model that, consistent with negative future expectations, anxious individuals should not exhibit the free choice preference observed in non-anxious individuals.
-Finally, we discuss the relationship of the present model to clinical theories postulating a role for control and helplessness in anxiety disorders.
+In the remaining sections, we also show that this simple model can account for a number of findings reported in the minority of studies that have probed anxious behavior in sequential choice environments.
+We then offer a novel prediction based on the the model that, consistent with negative future expectations, anxious individuals should not exhibit a free choice premium.
+We then close with a discussion of the relationship of our decision theoretic account to prominent clinical theories of anxiety, potential extensions to the model, and possible neural correlates.
 
@@
 Crucially then, the value of a state under the optimal policy is contingent on an agent's particular expectations about the future.
-First, it is a function of the state transition distribution, or beliefs about the likelihood of encountering particular states conditioned on undertaking some action.
-Second, it is also a function of the actions available in those successor states, as those in turn define the value of these states.
+First, it is a function of the state transition distribution, or beliefs about the likelihood of encountering particular successor states conditioned on performing some action.
+Second, it is also a function of the actions available in those successor states, as future actions in turn define their value.
 Under Bellman optimality, it is normative for an agent to assume that the value of a successor state is equivalent to the best action from that state.
@@
 
-where $q(s',a')$ denotes the a state-action value. As we will demonstrate, this assumption has profound consequences for learning about approach and avoidance behaviors.
+where $q(s',a')$ denotes the a state-action value. As we will demonstrate, this assumption has profound consequences for learning about reward and danger.
 
@@
 This follows naturally from the recursive definition of state values defined above.
-So when an agent learns a particular state predicts reward, the states which immediately precede it also acquire value (as do their antecedent states, and so on) thereby propagating positive value across the state space so as to reinforce reward-maximizing policies.
-This process does not occur, however, upon the discovery of a suboptimal or aversive action.
-Because under Bellman optimality the value of a state is dependent in part on the best possible future action, the discovery of harmful actions will not necessarily impact state value estimates or the optimal policy.
-Insofar that suboptimal or aversive actions can be avoided, then Bellman optimality predicts a fundamental asymmetry: reward opportunities propagate recursively to antecedent states, but danger does not.
+When an agent learns a particular state predicts reward, the states which immediately precede it also acquire value (as do their antecedent states, and so on).
+The discovery of a rewarding action thus propagates positive value across the state space, so as to motivate the future visitation of rewarding states.
+A similar spread of value does not occur, however, upon the discovery of a suboptimal or harmful action.
+Under the optimal policy, only the best action is chosen; a reward-maximizing agent will avoid performing a harmful action insofar that another option is available.
+When avoidance is possible, the discovery of a harmful action will not impact the value of its associated state.
+Moreover, an avoidable harmful action will not influence the value of its antecedent states, as is reflected by the max operator in the equation above.
+Thus, the optimistic assumptions about future behavior under Bellman optimality yields a fundamental asymmetry in the treatment of reward and danger: whereas rewarding opportunity propagates recursively to antecedent states, danger does not whenever avoidance is possible.
 
-This principle is highlighted in Figure 1.
-In an open field environment, an agent is free to navigate a discrete, deterministic grid-world devoid of rewards except for two states: one rewarding (blue) and one harmful (red, Figure 1a).
-In this environment, the one-step transition distribution respects the agent's choice; that is, the agent comes into experiences an aversive outcome only if it decides to approach danger.
-As can be observed, the state value estimates learned by an agent under Bellman optimality (Figure 1b) reflect the asymmetry described above.
-Despite the presence of danger, the agent learns to associate all states (except the harmful state itself) with positive value as the reward opportunity propagates outwards to all antecedent states.
-Insofar that danger can be successfully avoided, all states (even those adjacent to the danger) come to predict expected future reward.
+This principle is highlight in a toy MDP, the open field environment (Figure 1a).
+The open field is a grid world that entirely empty except for two states, one rewarding and one harmful.
+Importantly, the open field MDP is deterministic, such that the one-step transition distribution respects the agent's choice.
+An agent in this environment then will only experience an aversive outcome if it chooses to approach the dangerous state.
+The state value estimates learned by an agent under Bellman optimality reflect the asymmetry described above (Figure 1b).
+In the open field, reward propagates recursively from the source to antecedent states whereas danger does not.
+Thus the agent learns to associate all states (except the harmful state itself) with positive value.
+Because harm is avoidable in this environment, all states (even those proximal to threat) come to represent the opportunity for reward.
 
-What happens, however, when an agent believes that danger is not perfectly avoidable and safety not guaranteed?
-Such a belief may manifest as inaccurate assumptions about the state transition distribution.
-For example, an agent may wrongly believe it possible that it could transition to a state proximal to danger in spite of its choices.
-Alternately, this belief may manifest as pessimistic expectations about the availability of future actions.
-An agent may believe it has the capacity to act in accordance with the (reward-maximizing) optimal policy in the present, but may be unable to do so in the future.
-Importantly, such beliefs about the unreliability of the environment or of the self are commonly endorsed in anxiety disorders.
+What happens then when an agent believes (falsely) that danger is not avoidable?
+In the MDP formalism, such an errant  belief may manifest in numerous ways.
+First, an agent may possess inaccurate assumptions about the state transition distribution.
+For example, an agent may wrongly believe it possible that it could transition to a state closer to danger in spite of an avoidance action.
+Alternately, an agent may operate under pessimistic expectations about the availability of future actions.
+It may believe it has the capacity to act in accordance with the optimal (reward-maximizing) policy in the present, but may be unable to do so in the future.
+We note here that similar beliefs about the unreliability of self or the environment in the presence of perceived threat are often endorsed in anxiety disorders.
 
-We can explicitly model this type of pessimism about future expectations through a minor change to the state value function defined above.
-Here, we focus on one particular formulation, previously introduced by \cite{Gaskett2003}, that captures the expectation that future action may deviate from the optimal reward-maximizing policy.
-We define the value of the successor state as:
+We can easily model pessimistic expectations about future action through a minor change to the definition of state value.
+Previously introduced by \cite{Gaskett2003}, we define pessimistic learning as the expectation of future deviation from the optimal policy such that the value of a successor state is equal to:
 
@@
 
-where the parameter $w$ controls the degree of pessimism.
+Here, the parameter $w$ controls the degree of pessimism.
 When $\w = 1$, an agent is maximally optimistic and fully expects to act according to the best possible actions in the future.
@@
 When $w = 0.5$, the weighs the best and worst possible future actions equally.
-We note we are not committed to this particular implementation as anything more than a computational convenience.
-Many alternate functions could be defined over the space of possible future actions.
-(As mentioned above, we could also modify the transition distribution.) This formalism highlights an important computational principle; it does not constitute a mechanistic claim about anxiety disorders.
+We note we are not committed to this particular implementation.
+Pessimistic learning could be introduced through a variety of other manipulations.
+We have selected this particular formalism because out of convenience; it does not constitute a mechanistic claim about anxious cognition.
 
@@
 
+For a pessimistic agent ($w > 0$), the expectation that future action will always be optimal (reward-maximizing) is violated; instead, the agent believes that it may act with some probability in direct contrast to its preferences.
+Under these conditions, an agent cannot expect to invariably avoid choosing a harmful action.
+Therefore, the presence of danger now poses threat and safety cannot be guaranteed.
 The consequences of this manipulation for the open field environment are presented in Figure 1c,d.
-When Bellman optimality is violated (i.e.
-an agent no longer assumes the value of a successor state is equal the best available action), more than reward opportunities are able to propagate value through the state space.
-Indeed when $w$ is non-zero, the agent treats the future potential of danger as possible and thus negative value is able to spread from its source recursively through antecedent states thereby degrading the value of distal states (the degree of which determined by the extent of pessimism, $w$, and temporal discounting, $\gamma$).
-With increasing certainty that the worst policy will be followed in the future, the larger the extent to which threat value can permeate.
+The deviation from Bellman optimality, wherein the value of antecedent states are contingent on both the best and worst possible future actions, means that now more than reward opportunities are able to propagate recursively through state space.
+Indeed when $w$ is non-zero, threat value from danger also recursively propagates from its source through adjacent states.
+The extent of this spread is contingent on the degree of pessimistic belief.
+As an agent grows increasingly certain of future deleterious action (i.e.
+$w \rightarrow 1$), the greater the extent to which threat value permeates the state space.
 
-In addition to drawing inspiration from clinical theories of anxiety, this pessimistic model also captures several core symptoms of anxiety disorders.
-The hallmark symptom of anxiety disorders is exaggerated threat appraisal, or the tendency to perceive threat as disproportionately greater in likelihood and severity than is justified.
-This is evident in the anxious state value estimates presented in Figure 1c,d.
-Indeed when safety is perceived to be uncertain, threat in the environment is exaggerated in magnitude and prevalence.
-As should be apparent, this very same mechanism captures the generalization of fear that often accompanies anxiety disorders.
+We also note that this simple demonstration produces behaviors similar to the core symptoms of anxiety disorders.
+As is evidence in Figure 1c,d, the pessimistic agent exhibits exaggerated threat appraisals.
+It estimates danger at disproportionately more states than is warranted (likelihood bias) and perceives those states as more threatening than they objectively are (severity bias).
+Similarly, the propagation of threat value is reminiscent of the fear generalization observed in clinical anxiety.
+Through pessimistic learning the agent comes to regard as threatening states with increasingly distant associations to the danger source.
 
-In addition to pessimistic inference, all anxiety disorders share pathological avoidance as core symptom \citep{dsm5, Krypotos2015, Arnaudova2017}.
-Avoidance behaviors seek to increase or maintain the temporal, physical and/or mental distance between an individual and a threatening situation.
-Although avoidance is a natural response to danger, excessive avoidance as is observed in anxiety disorders can be highly disruptive to everyday functioning \citep{Salter2004} and frequently function so as maintain anxiety by preventing extinction \citep{Arnaudova2017}.
-For example, an individual with social anxiety may avoid attending large social events for fear of public embarrassment only to never learn that such worry is not justified.
-
-Avoidance behaviors in anxiety have been studied using a variety of methods.
-Perhaps the most classic paradigm is the behavioral avoidance test \citep{bandura1977}, wherein participants complete a number of tasks involving a feared object or situation and the number of tasks completed (or lack thereof) is a measure of avoidance.
-Computerized avoidance tasks have become increasingly common, and increased distance from virtual threat has been found to correlate with anxiety \citep{Bach2014, Bach2017, Sheynin2014}.
-This is not the case for the pessimistic agent.
-As above, when the agent doubts its future ability to act in a reward-maximizing (loss-minimizing) manor, it cannot rule out the possibility of accidentally stepping off the cliff.
-Thus, the potential for future harm is no longer zero and the large penalties for stepping off the cliff may back-propagate to their antecedent states, and so on.
-The extent of this spread is dependent on the strength of the pessimistic belief.
-With increasing pessimism, the optimal policy for the agent is increasingly distance from the cliff's edge (Figure 2c).
-Surprisingly, at high levels of pessimistic belief, the optimal policy is to immediately step off the cliff (Figure 2d).
-The reason for this is simple: with high levels of pessimism, the likelihood of disaster grows large enough that any step is simply viewed as adding onto the pain.
+The model also predicts persistent avoidance behaviors similar to those observed in clinical anxiety.
+Importantly, the avoidance behavior exhibited by the model begins early in the chain of sequential choices.
+In other words, the agent performs avoidance before it is proximal to danger.
+This is because agents in this model act in response to anticipated future threat.
+Specifically, the agent cannot be certain it will not act against its preferences in the future.
+To minimize the likelihood of inadvertently encountering danger, the agent therefore places as much distance between itself and the possible danger through early avoidance actions.
+As discussed in the introduction, this is what is observed both clinically and empirically in studies of anxious individuals responding to threat \citep{Bach2014, Bach2017, Sheynin2014}.
+Thus our model not only predicts persistent avoidance behavior, but also captures the excessive, overly cautious profile of avoidance behavior exhibited in clinical anxiety.
 
@@
 
-In the following sections, we present the results of simulations of pessimistic learning in a variety of MDP environments.
-In all simulations, the state-action values under the optimal policy, $Q_*(s,a)$, were solved for directly through Q-value iteration \citep{SuttonBarto1998, SuttonBarto1998, bertsekas2005}.
+In the following sections, we show that our model is capable of recreating the results of several empirical studies of anxious cognition in sequential choice environments.
+We represent these experiments again using the formalism of MDPs.
+In all simulations, the values of environment states under the optimal policy were solved for directly through Q-value iteration \citep{SuttonBarto1998, SuttonBarto1998, bertsekas2005}.
 The details of this algorithm are presented in the appendix.
-Model parameters are reported for each environment in their corresponding sections.
 All simulations were carried out with the python programming language and are publicly available on Github (https://github.com/szorowi1/SecretFunTimes).

### 2019-02-23 sisyphus/envs/_field.py
updates to figure 1

@@ -11,16 +11,16 @@ class OpenField(GraphWorld):
         self.shape = self.grid.shape
 
         ## Define start/terminal states.
-        start = 5
-        terminal = np.array([57,63])
+        start = 115
+        terminal = np.array([13,19])
 
         ## Define one-step transition matrix.
         T = grid_to_adj(self.grid, terminal)
 
         ## Define rewards.
         R = 0 * np.ones_like(T)               # Majority transitions
-        R[:,57] =  10                         # Reward transition
-        R[:,63] = -10                         # Punishment transition
+        R[:,13] =  10                         # Reward transition
+        R[:,19] = -10                         # Punishment transition
         R[terminal,terminal] = 0              # Terminal transitions
         R *= T
 
@@ -71,7 +71,7 @@ class OpenField(GraphWorld):
 
         ## Define grid.
         grid = np.zeros((11, 11))    # Viable states
-        grid[5,[2,-3]] = [1, 2]      # Reward/shock states
+        grid[1,[2,-3]] = [1, 2]      # Reward/shock states
 
         ## Define colormap.
         cmap = ListedColormap([grid_color, reward_color, shock_color])
@@ -87,7 +87,49 @@ class OpenField(GraphWorld):
         ## Annotate.
         if annot:
             if annot_kws is None: annot_kws = dict()
-            ax.text(2.5,5.5,reward,ha='center',va='center',**annot_kws)
-            ax.text(8.45,5.5,shock,ha='center',va='center',**annot_kws)
+            ax.text(2.5,1.5,reward,ha='center',va='center',**annot_kws)
+            ax.text(8.45,1.5,shock,ha='center',va='center',**annot_kws)
         
+        return ax
+    
+    def plot_policy(self, ax, pi, color='w', head_width=0.25, head_length=0.25):
+        """Plot agent policy on grid world.
+
+        Parameters
+        ----------
+        ax : matplotlib Axes
+            Axes in which to draw the plot.
+        pi : array
+            Agent policy, i.e. ordered visitation of states.
+        color : str, list
+            Color(s) of arrow.
+        head_width : float (default=0.25)
+            Width of the arrow head.
+        head_length : float (default=0.25)
+            Length of the arrow head.
+
+        Returns
+        -------
+        ax : matplotlib Axes
+            Axes in which to draw the plot.
+        """
+
+        ## Error-catching.
+        if isinstance(color, str):
[…]

### 2019-02-24 sisyphus/envs/__init__.py
renaming. updated bart and free choice figures

@@ -2,7 +2,7 @@
 
 from ._choice import FreeChoice
 from ._cliff import CliffWalking
-from ._fid import FlightInitiationDistance
+from ._bart import BART
 from ._field import OpenField
 from ._lh import Helplessness
 from ._tree import DecisionTree

### 2019-02-24 sisyphus/envs/_bart.py

@@ -2,15 +2,15 @@ import numpy as np
 from scipy.stats import norm
 from ._base import GraphWorld
 
-class FlightInitiationDistance(GraphWorld):
-    """Flight initiation distance environment. 
+class BART(GraphWorld):
+    """Balloon analog risk task environment.
     
     Parameters
     ----------
-    runway : int
-        Number of states in 2-D runway.
+    pumps : int
+        Maximum number of balloon pumps.
     mu : float
-        Average state at which predator strikes.
+        Average state at which balloon pops.
     sd : float
         Deviation around mean.
     
@@ -31,16 +31,18 @@ class FlightInitiationDistance(GraphWorld):
                 
     References
     ----------
-    1. Qi, S., Hassabis, D., Sun, J., Guo, F., Daw, N., & Mobbs, D. (2018). 
+    1. Lejuez, C. W., Read, J. P., Kahler, C. W., Richards, J. B., Ramsey, S. E., Stuart, 
+       G. L., ... & Brown, R. A. (2002). Evaluation of a behavioral measure of risk taking: 
+       the Balloon Analogue Risk Task (BART). Journal of Experimental Psychology: Applied, 8(2), 75.
+    2. Qi, S., Hassabis, D., Sun, J., Guo, F., Daw, N., & Mobbs, D. (2018). 
        How cognitive and reactive fear circuits optimize escape decisions in humans. 
        Proceedings of the National Academy of Sciences, 115(12), 3186-3191. 
     """
     
-    def __init__(self, runway=10, mu=5, sd=1, shock=-1):
-    
+    def __init__(self, pumps=10, mu=5, sd=1):
         
         ## Define one-step transition matrix.
-        n = runway
+        n = pumps
         T = np.zeros((n+2,n+2)) * np.nan
         T[np.arange(n),np.arange(n)+1] = 1   # Corridor transitions
         T[:n,n] = 1                          # Safety transition
@@ -51,7 +53,7 @@ class FlightInitiationDistance(GraphWorld):
         R = np.copy(T)
         R[np.arange(n),np.arange(n)+1] = 0   # Corridor transitions
         R[:n,n] = np.arange(n) + 1           # Safety transition
-        R[:n,n+1] = shock                    # Danger transition
+        R[:n,n+1] = -np.arange(n)            # Danger transition
         R[[n,n+1],[n,n+1]] = 0               # Terminal states 
         
         ## Define start/terminal states.
[…]

### 2019-02-24 sisyphus/envs/_choice.py

@@ -35,7 +35,7 @@ class FreeChoice(GraphWorld):
        over monetary gains and losses. Psychological science, 25(2), 596-604.
     """
     
-    def __init__(self, rewards=[-1,0,1], probs=None):
+    def __init__(self, rewards=[-1,1], probs=None):
     
         ## Error-catching.
         if probs is None: probs = np.ones_like(rewards) / len(rewards)

### 2019-02-25 sisyphus/envs/_lh.py
updated learned helplessness figures

@@ -2,22 +2,52 @@ import numpy as np
 from ._base import GraphWorld, grid_to_adj
 
 class Helplessness(GraphWorld):
+    """Learned helplessness environment.
     
-    def __init__(self, outcomes=[10,-10], epsilon=0):
+    Parameters
+    ----------
+    reward : float
+        Value of reward.
+    shock : float
+        Value of punishment.
+    epsilon : float
+        Randomness of state transitions (deterministic if 0).
+    
+    Attributes
+    ----------
+    states : array, shape = (n,)
+        Indices of states.
+    n_states : int
+        Total number of states.
+    viable_states : array
+        Indices of viable states.
+    n_viable_states : int
+        Number of viable states.
+    info : DataFrame
+        Pandas DataFrame  storing the dynamics of the Markov decision process.
+        Rows correspond to each viable Q-value, whereas each column contains
+        its associated information.
+    """
+    
+    def __init__(self, reward=10, shock=-10, epsilon=0):
         
         ## Define gridworld.
-        self.grid = np.arange(5*15).reshape(5,15)
+        grid = np.zeros((5,16),dtype=float)
+        grid[[0, 1, 3, 4], [-1, -1, -1, -1]] = np.nan
+        grid[np.where(~np.isnan(grid))] = np.arange(np.invert(np.isnan(grid)).sum())
+        self.grid = grid
         self.shape = self.grid.shape
         
         ## Define start/terminal states.
         start = 44
-        terminal = np.array([30, 37])
+        terminal = np.array([30, 37, 45])
 
         ## Define one-step transition matrix.
         T = grid_to_adj(self.grid, terminal)
         self.T = T
         
         ## Define rewards.
+        outcomes = [reward, shock, 0]
         R = np.zeros_like(T) 
         for s, r in zip(terminal, outcomes): R[:,s] = r
         R[terminal,terminal] = 0
@@ -26,14 +56,6 @@ class Helplessness(GraphWorld):
         ## Initialize GraphWorld.
         GraphWorld.__init__(self, T, R, start, terminal, epsilon)
         self.R = R
-        
-        ## Update start.
-        self.terminal = np.append(self.terminal, start)
-        self.info = self.info.loc[self.info.S != 44]
-        for i in range(4):
-            d = {"S":44, "S'":np.roll([29, 43, 44, 59],i), "T":np.array([1,0,0,0]), "R":np.zeros(4)}
[…]

### 2019-02-25 setup.py
documentation updates. probably ready for public viewing

@@ -11,10 +11,10 @@ path = os.path.abspath(os.path.dirname(__file__))
 DISTNAME = 'sisyphus'
 MAINTAINER = 'Sam Zorowitz'
 MAINTAINER_EMAIL = 'zorowitz@princeton.edu'
-DESCRIPTION = 'Code associated with manuscript'
+DESCRIPTION = 'Evaluating sequential choice biases in anxiety disorders'
 URL = 'https://dawlab.princeton.edu/'
 LICENSE = 'MIT'
-DOWNLOAD_URL = 'https://github.com/szorowi1/SecretFunTimes'
+DOWNLOAD_URL = 'https://github.com/ndawlab/seqanx'
 
 with open(os.path.join(path, 'README.rst'), encoding='utf-8') as readme_file:
     README = readme_file.read()

### 2019-02-25 sisyphus/envs/_bart.py

@@ -25,7 +25,7 @@ class BART(GraphWorld):
     n_viable_states : int
         Number of viable states.
     info : DataFrame
-        Pandas DataFrame  storing the dynamics of the Markov decision process.
+        Pandas DataFrame storing the dynamics of the Markov decision process.
         Rows correspond to each viable Q-value, whereas each column contains
         its associated information.
                 
@@ -34,9 +34,6 @@ class BART(GraphWorld):
     1. Lejuez, C. W., Read, J. P., Kahler, C. W., Richards, J. B., Ramsey, S. E., Stuart, 
        G. L., ... & Brown, R. A. (2002). Evaluation of a behavioral measure of risk taking: 
        the Balloon Analogue Risk Task (BART). Journal of Experimental Psychology: Applied, 8(2), 75.
-    2. Qi, S., Hassabis, D., Sun, J., Guo, F., Daw, N., & Mobbs, D. (2018). 
-       How cognitive and reactive fear circuits optimize escape decisions in humans. 
-       Proceedings of the National Academy of Sciences, 115(12), 3186-3191. 
     """
     
     def __init__(self, pumps=10, mu=5, sd=1):
@@ -64,7 +61,7 @@ class BART(GraphWorld):
             if np.all(s_prime == s): terminal.append(s)
 
         ## Initialize GraphWorld.
-        GraphWorld.__init__(self, T, R, start, terminal, 0)
+        GraphWorld.__init__(self, T, R, start, terminal, epsilon=0)
             
         ## Remove masochistic Q-values (i.e. agent cannot elect to be eaten).
         bps = self.n_states - 1

### 2019-02-25 sisyphus/envs/_base.py

@@ -3,7 +3,25 @@ from pandas import DataFrame
 from scipy.spatial.distance import cdist
 
 def grid_to_adj(grid, terminal=False):
-    """Grid to adjacency matrix."""
+    """Convert grid world to adjacency matrix.
+    
+    Parameters
+    ----------
+    grid : array, shape (i,j)
+        Grid world.
+    terminal : array
+        List of terminal states.
+        
+    Returns
+    -------
+    T : array, shape (n_states, n_states)
+        Adjacency matrix of states.
+        
+    Notes
+    -----
+    The initial grid can contain any value. Grid states defined as NaNs 
+    are treated as nonviable states and excluded from further processing.        
+    """
     
     ## Identify coordinates of viable states.
     rr = np.array(np.where(~np.isnan(grid))).T
@@ -12,15 +30,46 @@ def grid_to_adj(grid, terminal=False):
     A = (cdist(rr,rr)==1).astype(int)
 
     ## Define one-step transition matrix.
-    T = np.where(A, 1, np.nan)            # Non-terminal transitions
+    T = np.where(A, 1, np.nan)
     
+    ## Update terminal states.
     if np.any(terminal):
-        T[terminal] = np.nan              # Terminal transitions
-        T[terminal,terminal] = 1          # Terminal transitions
+        T[terminal] = np.nan
+        T[terminal,terminal] = 1
     
     return T
 
 class GraphWorld(object):
+    """Base graph world object.
+    
+    Parameters
+    ----------
+    T : array, shape (n_states, n_states)
+        Graph adjacency matrix.
+    R : array, shape (n_states, n_states)
+        One-step reward function.
+    start : int
+        Starting state.
+    terminal : int | list
+        Terminal states.
+    epsilon : int
+        Randomness parameter. If zero, transitions are deterministic.
+        
+    Attributes
+    ----------
+    states : array, shape = (n,)
+        Indices of states.
+    n_states : int
+        Total number of states.
+    viable_states : array
+        Indices of viable states.
+    n_viable_states : int
+        Number of viable states.
+    info : DataFrame
+        Pandas DataFrame storing the dynamics of the Markov decision process.
+        Rows correspond to each viable Q-value, whereas each column contains
+        its associated information.
+    """
     
     def __init__(self, T, R, start, terminal, epsilon=0):
         

### 2019-02-25 sisyphus/envs/_choice.py

@@ -3,14 +3,14 @@ from scipy.stats import norm
 from ._base import GraphWorld
 
 class FreeChoice(GraphWorld):
-    """Instrumental variant of the Free Choice task.
+    """Instrumental variant of the free choice task.
     
     Parameters
     ----------
     rewards : array
         Outcome values.
     probs : array
-        Probability of rewards.
+        Probability of rewards. Defaults is uniform probability.
     
     Attributes
     ----------
@@ -23,7 +23,7 @@ class FreeChoice(GraphWorld):
     n_viable_states : int
         Number of viable states.
     info : DataFrame
-        Pandas DataFrame  storing the dynamics of the Markov decision process.
+        Pandas DataFramestoring the dynamics of the Markov decision process.
         Rows correspond to each viable Q-value, whereas each column contains
         its associated information.
                 
@@ -59,7 +59,7 @@ class FreeChoice(GraphWorld):
         terminal = np.arange(5,n+5)
 
         ## Initialize GraphWorld.
-        GraphWorld.__init__(self, T, R, start, terminal, 0)
+        GraphWorld.__init__(self, T, R, start, terminal, epsilon=0)
         
         ## Update info.
         for s in [2,3,4]:

### 2019-02-25 sisyphus/envs/_cliff.py

@@ -4,13 +4,33 @@ from ._base import GraphWorld, grid_to_adj
 class CliffWalking(GraphWorld):
     """Cliff-walking task environment.
     
+    Parameters
+    ----------
+    cliff : float
+        Value of falling off cliff.
+    
+    Attributes
+    ----------
+    states : array, shape = (n,)
+        Indices of states.
+    n_states : int
+        Total number of states.
+    viable_states : array
+        Indices of viable states.
+    n_viable_states : int
+        Number of viable states.
+    info : DataFrame
+        Pandas DataFrame storing the dynamics of the Markov decision process.
+        Rows correspond to each viable Q-value, whereas each column contains
+        its associated information.
+    
     References
     ----------
-    1. Sutton, R. S., & Barto, A. G. (1998). Reinforcement learning: An introduction. MIT press.
+    1. Sutton, R. S., & Barto, A. G. (2018). Reinforcement learning: An introduction. MIT press.
     2. Gaskett, C. (2003). Reinforcement learning under circumstances beyond its control.
     """
     
-    def __init__(self, epsilon=0):
+    def __init__(self, cliff=-100):
     
         ## Define gridworld.
         self.grid = np.arange(11 * 12, dtype=int).reshape(11,12)
@@ -25,13 +45,13 @@ class CliffWalking(GraphWorld):
         
         ## Define rewards.
         R = -1 * np.ones_like(T)              # Majority transitions
-        R[:,terminal[:-1]] = -100             # Cliff transitions
+        R[:,terminal[:-1]] = cliff            # Cliff transitions
         R[:,terminal[-1]] = 0                 # Safety transitions
         R[terminal,terminal] = 0              # Terminal transitions
         R *= T
             
         ## Initialize GraphWorld.
-        GraphWorld.__init__(self, T, R, start, terminal, epsilon)
+        GraphWorld.__init__(self, T, R, start, terminal, epsilon=0)
         
     def __repr__(self):
         return '<GraphWorld | Cliff-Walking Task>'
@@ -106,8 +126,8 @@ class CliffWalking(GraphWorld):
             Axes in which to draw the plot.
         pi : array
             Agent policy, i.e. ordered visitation of states.
-        color : str (default = white)
-            Color of arrow.
+        color : str, list
+            Color(s) of arrow.
         head_width : float (default=0.25)
             Width of the arrow head.
[…]

### 2019-02-25 sisyphus/envs/_field.py

@@ -2,9 +2,33 @@ import numpy as np
 from ._base import GraphWorld, grid_to_adj
 
 class OpenField(GraphWorld):
-    """Open field task environment."""
+    """Open field task environment.
     
-    def __init__(self, epsilon=0):
+    Parameters
+    ----------
+    reward : float
+        Value of reward.
+    punishment : float
+        Value of punishment.
+    
+    Attributes
+    ----------
+    states : array, shape = (n,)
+        Indices of states.
+    n_states : int
+        Total number of states.
+    viable_states : array
+        Indices of viable states.
+    n_viable_states : int
+        Number of viable states.
+    info : DataFrame
+        Pandas DataFrame storing the dynamics of the Markov decision process.
+        Rows correspond to each viable Q-value, whereas each column contains
+        its associated information.
+    
+    """
+    
+    def __init__(self, reward=10, punishment=-10):
     
         ## Define gridworld.
         self.grid = np.arange(11 * 11, dtype=int).reshape(11,11)
@@ -19,19 +43,19 @@ class OpenField(GraphWorld):
 
         ## Define rewards.
         R = 0 * np.ones_like(T)               # Majority transitions
-        R[:,13] =  10                         # Reward transition
-        R[:,19] = -10                         # Punishment transition
+        R[:,13] = reward                      # Reward transition
+        R[:,19] = punishment                  # Punishment transition
         R[terminal,terminal] = 0              # Terminal transitions
         R *= T
 
         ## Initialize GridWorld.
-        GraphWorld.__init__(self, T, R, start, terminal, epsilon)
+        GraphWorld.__init__(self, T, R, start, terminal, epsilon=0)
         
     def __repr__(self):
         return '<GraphWorld | Open Field Task>'
     
-    def plot_field(self, reward=10, shock=-10, annot=True, grid_color='0.8',  
-                   reward_color='#f3e1db', shock_color='#1c142a', 
+    def plot_field(self, reward=10, punishment=-10, annot=True, grid_color='0.8',  
+                   reward_color='#f3e1db', punishment_color='#1c142a', 
                    cbar=False, annot_kws=None, ax=None):
         """Plot open field environment.
 
@@ -39,15 +63,15 @@ class OpenField(GraphWorld):
         ----------
         reward : float
             Reward value.
-        shock : float
-            Shock value.
+        punishment : float
+            Punishment value.
[…]

### 2019-02-25 sisyphus/envs/_lh.py

@@ -8,10 +8,8 @@ class Helplessness(GraphWorld):
     ----------
     reward : float
         Value of reward.
-    shock : float
+    punishment : float
         Value of punishment.
-    epsilon : float
-        Randomness of state transitions (deterministic if 0).
     
     Attributes
     ----------
@@ -29,7 +27,7 @@ class Helplessness(GraphWorld):
         its associated information.
     """
     
-    def __init__(self, reward=10, shock=-10, epsilon=0):
+    def __init__(self, reward=10, punishment=-10):
         
         ## Define gridworld.
         grid = np.zeros((5,16),dtype=float)
@@ -47,21 +45,21 @@ class Helplessness(GraphWorld):
         self.T = T
         
         ## Define rewards.
-        outcomes = [reward, shock, 0]
+        outcomes = [reward, punishment, 0]
         R = np.zeros_like(T) 
         for s, r in zip(terminal, outcomes): R[:,s] = r
         R[terminal,terminal] = 0
         R *= T
         
         ## Initialize GraphWorld.
-        GraphWorld.__init__(self, T, R, start, terminal, epsilon)
+        GraphWorld.__init__(self, T, R, start, terminal, epsilon=0)
         self.R = R
         
     def __repr__(self):
         return '<GraphWorld | Learned Helplessness>'
     
-    def plot_lh(self, reward=10, shock=-10, annot=True, grid_color='0.8',  
-                   reward_color='#f3e1db', shock_color='#1c142a', 
+    def plot_lh(self, reward=10, punishment=-10, annot=True, grid_color='0.8',  
+                   reward_color='#f3e1db', punishment_color='#1c142a', 
                    cbar=False, annot_kws=None, ax=None):
         """Plot cliff-walking environment.
 
@@ -69,15 +67,15 @@ class Helplessness(GraphWorld):
         ----------
         reward : float
             Reward value.
-        shock : float
-            Shock value.
+        punishment : float
+            Punishment value.
         annot : bool
             Annotate states.
         grid_color : str
             Color of grid tiles.
         reward_color : str
             Color of rewarding tile.
-        shock_color : str
+        punishment_color : str
             Color of punishing tile.
         cbar : bool
             Whether to draw a colorbar.
@@ -101,10 +99,10 @@ class Helplessness(GraphWorld):
 
         ## Define grid.
[…]

### 2019-02-25 sisyphus/envs/_tree.py

@@ -3,14 +3,7 @@ from scipy.stats import norm
 from ._base import GraphWorld
 
 class DecisionTree(GraphWorld):
-    """Variant of the decision tree game.
-    
-    Parameters
-    ----------
-    rewards : array
-        Outcome values.
-    probs : array
-        Probability of rewards.
+    """Decision tree from aversive pruning experiments.
     
     Attributes
     ----------
@@ -66,7 +59,7 @@ class DecisionTree(GraphWorld):
         terminal = np.arange(7,15)
 
         ## Initialize GraphWorld.
-        GraphWorld.__init__(self, T, R, start, terminal, 0)
+        GraphWorld.__init__(self, T, R, start, terminal, epsilon=0)
         
     def __repr__(self):
         return '<GraphWorld | Decision Tree>'

### 2019-02-25 sisyphus/mdp/_dp.py

@@ -1,7 +1,8 @@
 """Dynamic programming module"""
 
 import numpy as np
-from ._misc import check_params, softmax, betamax
+from copy import deepcopy
+from ._misc import check_params, softmax, pessimism
 from warnings import warn
 
 class ValueIteration(object):
@@ -9,12 +10,14 @@ class ValueIteration(object):
     
     Parameters
     ----------
-    policy : max | min | softmax | betamax (default = betamax)
-        Choice policy.
-    beta : float
-        Inverse temperature (ignored if policy not softmax).
-    gamma : float
-        Discount factor.
+    policy : max | min | softmax | pessimism (default = pessimism)
+        Learning rule.
+    gamma : float (default = 0.9)
+        Temporal discounting factor.
+    beta : float (default = 10.0)
+        Inverse temperature for future choice (ignored if policy not softmax).
+    w : float (default = 1.0)
+        Pessimism weight (ignored if policy not pessimism).
     tol : float, default: 1e-4
         Tolerance for stopping criteria.
     max_iter : int, default: 100
@@ -22,29 +25,38 @@ class ValueIteration(object):
 
     References
     ----------
-    1. Sutton, R. S., & Barto, A. G. (1998). Reinforcement learning: An introduction. MIT press.
+    1. Sutton, R. S., & Barto, A. G. (2018). Reinforcement learning: An introduction. MIT press.
     """
     
-    def __init__(self, policy='betamax', beta=10, gamma=0.9, tol=0.0001, max_iter=100):
+    def __init__(self, policy='pessimism', gamma=0.9, beta=10.0, w=1.0, tol=0.0001, max_iter=100):
 
         ## Define choice policy.
         self.policy = policy
         if policy == 'max': self._policy = np.max
         elif policy == 'min': self._policy = np.min
         elif policy == 'softmax': self._policy = lambda arr: arr @ softmax(arr * self.beta)
-        elif policy == 'betamax': self._policy = lambda arr: betamax(arr, self.beta)
+        elif policy == 'pessimism': self._policy = lambda arr: pessimism(arr, self.w)
         else: raise ValueError('Policy "%s" not valid!' %self.policy)
         
         ## Check parameters.
-        self.beta = beta
         self.gamma = gamma
-        check_params(beta=self.beta, gamma=self.gamma)
+        self.beta = beta
+        self.w = w
+        check_params(gamma=self.gamma, beta=self.beta, w=self.w)
         
         ## Set convergence criteria.
[…]

### 2019-02-25 sisyphus/mdp/_misc.py

@@ -6,28 +6,28 @@ from warnings import warn
 
 @jit
 def softmax(arr):
-    """Scale-robust softmax."""
+    """Scale-robust softmax choice rule."""
     arr = np.exp(arr - np.max(arr))
     return arr / arr.sum()
 
 @jit
-def betamax(arr, beta):
-    """Beta-pessimism."""
-    return beta * np.max(arr) + (1 - beta) * np.min(arr)
+def pessimism(arr, w):
+    """Pessimistic learning rule."""
+    return w * np.max(arr) + (1 - w) * np.min(arr)
 
 def categorical(arr):
-    """Categorical distribution RNG."""
+    """Categorical distribution rng."""
     return np.argmax(np.random.multinomial(1,arr))
 
-def check_params(beta=None, eta=None, gamma=None, tau=None, epsilon=None):
-    
+def check_params(beta=None, eta=None, gamma=None, w=None, epsilon=None):
+    """Internal convenience function for sanity checking parameter values."""
     if beta is not None and abs(beta) > 50: 
         warn('Parameter "beta" set very large.')
     if eta is not None and (eta < 0 or eta > 1): 
         raise ValueError('Parameter "eta" must be in range [0,1].')        
     if gamma is not None and (gamma < 0 or gamma > 1): 
         raise ValueError('Parameter "gamma" must be in range [0,1].') 
-    if tau is not None and abs(tau) > 50: 
-        warn('Parameter "tau" set very large.')
+    if w is not None and (w < 0 or w > 1): 
+        raise ValueError('Parameter "w" must be in range [0,1].')       
     if epsilon is not None and (epsilon < 0 or epsilon > 1): 
         raise ValueError('Parameter "epsilon" must be in range [0,1].') 

### 2019-02-25 sisyphus/mdp/_td.py

@@ -2,14 +2,16 @@
 
 import numpy as np
 from copy import deepcopy
-from ._misc import check_params, betamax, categorical
+from ._misc import check_params, pessimism, categorical
 from ._misc import softmax as _softmax
 
-def egreedy(arr, epsilon):
+def epsilon_greedy(arr, epsilon):
+    """Epsilon-greedy choice rule."""
     if np.random.binomial(1,1-epsilon): return np.argmax(arr)
     else: return np.random.choice(np.arange(len(arr)),1,replace=False)[0]
     
 def softmax(arr, beta):
+    """Softmax choice rule."""
     theta = _softmax(arr * beta)
     return categorical(theta)  
 
@@ -18,35 +20,40 @@ class ModelFree(object):
     
     Parameters
     ----------
-    policy : max | min | softmax | betamax (default = softmax)
-        Choice policy.
-    beta : float
-      Inverse temperature for choice.
-    eta : float
-      Learning rate.
+    policy : max | min | softmax | pessimism (default = pessimism)
+        Learning rule.
+    eta : float (default = 0.1)
+        Learning rate.
+    gamma : float (default = 0.9)
+        Temporal discounting factor.
+    beta : float (default = 10.0)
+        Inverse temperature for future choice (ignored if policy not softmax).
+    w : float (default = 1.0)
+        Pessimism weight (ignored if policy not pessimism).
     gamma : float
       Discount factor.
 
     References
     ----------
-    1. Sutton, R. S., & Barto, A. G. (1998). Reinforcement learning: An introduction. MIT press.
+    1. Sutton, R. S., & Barto, A. G. (2018). Reinforcement learning: An introduction. MIT press.
     '''
     
-    def __init__(self, policy, beta=10, eta=0.2, gamma=0.9):
+    def __init__(self, policy='pessimism', eta=0.1, gamma=0.9, beta=10.0, w=1.0):
         
         ## Define choice policy.
         self.policy = policy
         if policy == 'max': self._policy = np.max
         elif policy == 'min': self._policy = np.min
         elif policy == 'softmax': self._policy = lambda arr: arr @ _softmax(arr * self.beta)
-        elif policy == 'betamax': self._policy = lambda arr: betamax(arr, self.beta)
+        elif policy == 'pessimism': self._policy = lambda arr: pessimism(arr, self.w)
         else: raise ValueError('Policy "%s" not valid!' %self.policy)
         
         ## Check parameters.
         self.beta = beta
         self.eta = eta
         self.gamma = gamma
[…]

### 2019-02-25 manuscript/manuscript.tex

@@
 
-We first briefly review the formalism of Markov decision processes, which will provide the framework for our results.
-For complete treatments, see \cite{SuttonBarto1998, SuttonBarto2018} and \cite{bertsekas2005}.
+We first briefly review the formalism of Markov decision processes, which will provide the framework for our results. For complete treatments, see \cite{SuttonBarto2018} and \cite{bertsekas2005}.
 
@@
 
-In the following sections, we show that our model is capable of recreating the results of several empirical studies of anxious cognition in sequential choice environments.
+In the following sections, we show that our model can account for the majority of extant empirical studies of anxious cognition in sequential choice environments.
 We represent these experiments again using the formalism of MDPs.
-In all simulations, the values of environment states under the optimal policy were solved for directly through Q-value iteration \citep{SuttonBarto1998, SuttonBarto1998, bertsekas2005}.
-The details of this algorithm are presented in the appendix.
-All simulations were carried out with the python programming language and are publicly available on Github (https://github.com/szorowi1/SecretFunTimes).
+In all simulations, the values of environment states under the optimal policy were solved for directly through Q-value iteration \citep{bertsekas2005, SuttonBarto2018}.
+All simulations were implemented using the python programming language and are publicly available on Github (https://github.com/szorowi1/SecretFunTimes).
 
@@
 
-Next, we turn our attention to a related phenomenon observed in anxiety: a bias towards avoidance during approach-avoidance conflict \citep{aupperle2010}.
-In situations where the potential for reward and harm are correlated,
-anxiety predicts the tendency to forego reward in favor of minimizing the risk of threat.
-As a consequence, individuals with anxiety may experience fewer rewards in their environment as compared to non-anxious individuals.
-For example, an individual with social anxiety may avoid a party for fear of social embarrassment at the expense of positive social interactions with friends.
-This is a primary mechanism by which clinical anxiety can disrupt everyday functioning.
+In the broadest of terms, the motivations for action can be characterized with respect to two fundamental drives: the drive to approach rewards and the drive to avoid harm \citep{corr2013}.
+Though usually aligned, these motivations are necessarily antithetical wherever reward and risk of harm are correlated.
+In approach-avoidance conflict, an agent must integrate information regarding potential gains and losses in order to decide whether to perform an approach or avoidance action.
+Unsurprisingly, approach-avoidance conflict processing been suggested as integral to understanding the anxiety disorders \citep{aupperle2010} insofar that many of the disruptions they cause to everyday functioning (e.g.
+avoiding social obligations for fear of possible social humiliation) can be recast as biased arbitration of approach-avoidance conflict.
 
-A relationship between anxiety and biased decision making in approach-avoidance conflict is well-documented in the empirical literature.
-One common measure is the balloon analog risk task (BART) \citep{Lejuez2002}.
-The task is straightforward: a participant inflates a virtual balloon, one pump at a time.
-The more inflated the balloon, the more points the participant earns.
-With increasing pumps, however, comes the risk of the balloon popping and the loss of all earned points so far.
-Thus, there is a conflict: a participant can choose to pump the balloon more to earn more points, at the risk of the balloon popping; alternately, the participant can choose to end the trial earlier with fewer points earned.
-Previous studies have found that anxiety is correlated with earlier leaving times, or a greater tendency to avoid risk \cite{Maner2007, Giorgetta2012}.
+Many experimental protocols exist for studying approach-avoidance conflict behavior.
+One popular paradigm is the balloon analog risk task (BART) \citep{Lejuez2002}.
+In the BART, participants are offered the opportunity to win money by pumping virtual balloons.
+With each successful pump, the balloon incrementally inflates and money is added to the participant's till.
+Following one too many pumps, however, the balloon bursts and all money earned on that trial is lost.
+At any point during a trial, a participant may also choose to cash out, thereby banking the money earned up to that point and ending the trial.
+Thus, the BART pits approach and avoidance motivations against each other: with each pump, the reward earned increases but so too does the risk of losing it all.
+Importantly, previous studies have found that anxiety is correlated with fewer pumps and earlier cash-out times \citep{Maner2007, ramirez2015}.
 
@@
   \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/03_fid.png}}%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/02_bart.png}}%
   }
-  \par \textbf{Figure 3:} The Balloon Analog Risk Task (BART).
-Risk of the balloon popping increases with each pump (top row).
-An optimistic agent ($w=1$) continues to pump until the risk of losing all points earned outweighs the marginal value of one additional pump (second row).
-A pessimistic agent ($w=0.8$) is less confident that it will make the optimal decision in the future, and thus leaves earlier (third row).
-Leaving time becomes sooner with increasing pessimism ($w=0.6$) (bottom row).
-(Parameters: $\gamma = 1$.)
+  \par \textbf{Figure 2:} The Balloon Analog Risk Task (BART).
+The risk of balloon burst (point loss) increases with each pump (A).
+The optimal policy (number of pumps) under increasingly pessimistic learning is presented for low risk (B) and high risk (C) balloons.
+The optimistic agent ($w=1$) learns a policy reflecting the objective environmental risk.
+The moderate ($w=0.6$) and strongly ($w=0.2$) pessimistic agents learn to cash-out earlier, as is observed in anxious individuals.
+(Parameters: $\gamma = 1.0$)
 \end{figure}
 
-This result can be accounted for by a loss of confidence in future reliability.
-A simulation of performance on the BART is shown in Figure 3.
-For an optimistic agent ($w=1$), the optimal policy is to continue pumping until the risk of popping is all but guaranteed.
-The optimal policy for pessimistic agents differ.
-Under a loss of confidence for future action, pessimistic agents cannot be certain they will make optimal choices in the future.
-As such, negative value propagates backwards resulting in earlier leaving times.
-The latency of leaving is dependent on the strength of the pessimistic belief.
+This empirical result is easily accounted for by our model.
+A simulation of performance on the BART is presented in Figure 2.
+For an optimistic agent ($w=1.0$), the optimal policy accurately reflects the statistics of the task.
+The optimal agent continues pumping the balloon until the marginal gain from one additional pump no longer exceeds the likelihood of of the balloon bursting.
+Thus, the optimistic agent continues pumping until reaching the (unobserved) mean number of pumps before the balloon bursts.
+The same is not true for the pessimistic agents.
+The moderately ($w=0.6$) and strongly ($w=0.2$) pessimistic agents exhibit behavioral policies with earlier cash-out times.
+The reason for this is that anticipated future harmful actions are able to exert influence on earlier states; risky pumps resulting in loss are more salient to these agents earlier in the trial.
+As such, the pessimistic agents perform similarly to what is observed empirically in anxious individuals.
 
-Interestingly, these results are similar to recent findings regarding flight initiation distance and anxiety \citep{Mobbs2018, Mobbs2019}.
-Flight initiation distance refers to a measure from behavioral ecology that assesses the leaving time of a foraging animal at risk of predation.
-In classic flight initiation scenarios, an animal can continue to forage and exploit resources, but the longer it does so the greater the risk it will encounter a predator.
-Thus, flight initiation scenarios are similar in structure to approach-avoidance conflict and the BART.
-Recently, \cite{Mobbs2019} found that anxiety is correlated with earlier leaving times in virtual flight initiation tasks.
-Interestingly, the authors make a distinction between "fast-" and "slow-attacking" predators, and find that trait anxiety is only correlated only in behavioral response to the latter.
-The findings here draw upon an emerging literature of distinct neural systems mediating "cognitive" vs.
-"reactive" threat responses, where anxiety evidently impacts the deliberative, cognitive responses.
-Thus, it is important to note the present theory can account for anxious approach-avoidance behavior in some, but certainly not all, threatening scenarios.
+It is worth noting that these results are similar to recent findings regarding flight initiation distance (FID) and anxiety \citep{Mobbs2018, Mobbs2019}.
+FID refers to a concept from behavioral ecology that measures the leaving time of a foraging animal at risk of predation.
+In the prototypical FID scenario, an animal can choose either to continue exploiting resources (i.e.
+accrue reward) or flee an approaching predator.
+Similar to the BART, the longer an animal chooses to exploit the greater reward and risk it experiences.
+In a recent study, \cite{Mobbs2019} investigated the relationship between sub-clinical levels of anxiety and leaving times during a virtual FID task involving both fast- and slow-acting predators.
+Interestingly, sub-clinical anxiety did not correlate with leaving times for fast predators, but was correlated with earlier leaving times for slow predators.
+(Sub-clinical anxiety also correlated with BOLD activity in brain structures including the ventral hippocampus, medial prefrontal cortex, amygdala and insula.) The authors use these results to support an argument distinguishing between "reactive" responses associated with immediate threat and "cognitive" responses associated with impending danger.
+Similar to the BART results, our model can readily account for the "cognitive"-type threat response reported in \cite{Mobbs2019}; importantly though, these findings suggest that our model may not be applicable to all types of threat processing, such as those with little reliance on future expectations (such as in immediate threat).
 
-These results may also be related to recent work on the computational mechanisms underlying behavioral inhibition, a temperament linked to anxiety \citep{bach2015, khemka2017}.
-\cite{bach2015} argues that delayed behavioral responses under threat are normative under the assumption that a pause in behavior (e.g.
-freezing) can be useful for surveying a threatening environment, avoiding detection by predators, and potentially waiting out a nearby threat.
-Behavioral inhibition, it is argued, is a useful Pavlovian heuristic that has evolved over time.
-The current models may predict similar results.
-Insofar that a loss of confidence in future action yields more negative estimates of action values, then the best action may often be brought closer in value to its alternatives.
-As such, a pessimistic agent may face increased decision conflict as compared to a non-anxious counterpart and therefore show increased response times.
-This is not to argue against this particular model of behavioral inhibition, but just to say that our account may yield similar behaviors.
+Similarly related are recent studies on behavioral inhibition, a behavioral style characterized by a temporal delay to initiate approach behavior associated with the development of anxiety disorders \citep{bach2015, khemka2017}.
+\cite{bach2015} argues that behaviorally inhibited responses (e.g.
+slowed reaction times under threat) reflect a normative Pavlovian heuristic, developed over evolutionary time, that is instrumental in surveying the environment, avoiding detection by predators, and possibly even waiting out nearby threat.
+Evidence for this heuristic comes from a sequential approach-avoidance conflict task, wherein a participant may choose to exit a safe zone to collect valuable tokens adjacent to a "sleeping" virtual predator.
+With each token collected, the probability of the predator "waking up", unbeknownst to the participant, increases.
+Similar to the BART, if the predator catches the participant, all tokens accrued during that trial are lost.
+Importantly, \cite{bach2015} (and later \cite{khemka2017}) find that response times for choosing to leave the safe zone increase as a function of predation threat, even though response times have no bearing on whether the predator "awakes", in line with the interpretation of behavioral inhibition as Pavlovian heuristic.
 
-\subsection{Relation to Depression}
+Our model can offer an alternative interpretation.
+As observed above, one consequence of uncertainty about avoiding harmful future actions is that antecedent actions become associated with threat.
+In the context of the sequential approach-avoidance task, our pessimistic bias should therefore reduce the value of the choice to leave the safe zone to collect tokens, making it closer in value to the choice to stay and forego future point collection.
+Insofar that choices between similarly valued options result in prolonged response times by virtue of increased decision conflict, then we should predict that anxious individuals would exhibit, in addition to earlier leaving times, slower responses in sequential approach-avoidance task independent of a Pavlovian heuristic.
+Further work is necessary to disentangle these two interpretations.
 
-As an extension of approach-avoidance conflict, we next want to discuss the relationship between anxiety and depression.
+\subsection{Anxiety-to-Depression Transition}
+
+As will become clear, the previous section concerning approach-avoidance conflict is a natural interlude to discussing the relationship between anxiety and depression.
 Anxiety and depression are highly comorbid, with roughly 45\% of individuals with a lifetime depression diagnosis also diagnosed with at least one anxiety disorder \citep{kessler2015}.
-Given this, many have clinical theories assume that depression and anxiety are slightly variant manifestations of the same underlying mechanisms involving negative affect (for review, see \cite{jacobson2014}).
-
-Where the two disorders can be distinguished is in relation to rewards and anhedonia; depressive disorders, but not anxiety disorders, are associated with disrupted reward processing.
-One interesting proposal in the literature is that at least certain types of anhedonic depression are preceded by anxiety \citep{alloy1990, moitra2008, jacobson2014}.
-Specifically, persistent avoidance has the long-run effect of reducing the likelihood of anxious individuals encountering positive events and activities in their everyday lives.
-As a consequence, anxious individuals come to expect an absence of reward in their environments and subsequently develop diminished reward motivation and behavioral activation.
-Recent large-scale epidemiological studies have provided  some empirical support for this claim, finding that anxiety symptoms and behavioral avoidance precede and predict the longitudinal onset of depression (\cite{mathew2011, jacobson2014, kessler2015}; though also see \cite{jacobson2017, plana2019}).
+Because of their frequent co-occurrence, many clinical theories treat anxiety and depressive disorders as variant manifestations of the same latent mechanisms causing disruption to negative affective processes (for review, see \cite{jacobson2014}).
+Where the two families of disorders can be distinguished is in their relation to positive affect and reward processing.
+Some have argued that whereas both anxiety and depression both involve exaggerated threat appraisals and pessimistic future expectations, only depression involves the certain belief about the absence of future rewards.
 
@@
   \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/06_lh.png}}%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/03_lh.png}}%
   }
-  \par \textbf{Figure 4:} Progression from anxiety to depression.
+  \par \textbf{Figure 3:} Progression from anxiety to depression.
 In this open environment, an agent starts opposite a reward (blue) with harm separating them (A).
@@
 
-This prediction is easily accommodated by the present model.
-Simulations of how this might come back is presented in Figure 4.
-In this environment, there is only one rewarding and harmful state.
-Just as in the open field environment, the world is deterministic and the harmful state poses no true threat to an agent.
-An optimistic agent ($w=1$), with a belief in its ability to maintain an optimistic policy, is unfazed by the presence of the harmful state and develops over time an overall positive representation of the value of the environment (Figure 4b-d).
+One interesting proposal in the clinical literature attempting to explain this divergence appeals to longitudinal association between anxiety disorders and depression.
+Specifically, it has been proposed that the onset of at least certain types of anhedonic depression are preceded first by the onset of an anxiety disorder \citep{alloy1990, moitra2008, jacobson2014}.
+The idea according to this view is that because anxiety is associated with a bias towards avoidance behavior, then anxious individuals should be more likely to forego reward and other positive events in their everyday lives.
+Over time, it is argued, this should translate to a stable belief that, in addition to threat being unavoidable, reward is unobtainable thereby resulting in reduced behavioral activation and increased motor inhibition.
+Recent large-scale epidemiological studies have provided some empirical support for this claim, finding that anxiety symptoms often precede depression (\cite{mathew2011, jacobson2014, kessler2015}; though also see \cite{jacobson2017, plana2019}).
+Interestingly, one study event found that anxious avoidance behaviors were predictive future depression onset over a decade later \citep{jacobson2014}.
 
-In contrast, a pessimistic agent ($w=0.6$) cannot be certain of its ability to steer clear of future harm.
-As such, negative value spreads from the harmful state to its antecedent states (Figure 4e-g).
-Because in this environment risk and reward are correlated (i.e.
-the agent must approach the harmful state in order to reach the reward given its starting position), the consequence of negatively spreading value is that the optimal policy is to stay in the initial position.
-Thus, the pessimistic agent adopts a behavioral policy akin to depressive immobility.
-Moreover, reward is inaccessible to this agent such that its positive expectations of the environment are also diminished.
-In summary, a loss of confidence in future reliability can result in behavioral policies that resemble depressive-like states in environments with correlated risk and reward.
+Though currently speculative, this prediction is accommodated by our model.
+Simulations of how this might come about is presented in Figure 3.
+We assume an open grid world, similar to that presented in Figure 1, except that rewards and harms are now (spatially) correlated.
+Figure 3a shows that, in order to reach the reward, an agent must approach (but need not encounter) danger.
+As above, the world is deterministic such that the presence of the harmful state poses no objective danger to the agent.
+The learning progression of an optimistic agent ($w=1$) reflects this (Figure 3b-d).
+An optimistic agent believes it will be able to act in accordance to its reward-maximizing preferences in the future; as such, it does not worry that it may inadvertently encounter danger in the future.
+As a consequence, only reward opportunity propagates from the source to its antecedent states whereas harm is effectively sequestered.
+The optimistic agent thereby learns a path from the start to reward in spite of the presence of a dangerous state.
 
-\subsection{Planning}
+In contrast, a pessimistic agent ($w=0.6$) is uncertain about its ability to effectively avoid future danger.
+As a consequence, reward and threat values propagate from their respective sources to their antecedent states (Figure 3e-g).
+Because reward and threat are (spatially) correlated in this environment, however, subjective threat increases with proximity to reward.
+As such, the pessimistic agent learns a behavioral policy distinct from the optimistic agent.
+Whereas the optimistic agent learns a path which circumnavigates danger, the pessimistic agent learns that any step toward reward is also a step towards danger.
+Consequently, the pessimistic agent learns to not move.
+In effect, the pessimistic agent exhibits a form of motor inhibition and immobility like that observed in depression.
+Moreover, as a function of its learned behavioral policy, the agent should not expect reward from the environment.
+In summary, our model can accommodate previous clinical theories about the transition from anxiety to depression; specifically, our model predicts that when reward and danger are correlated, pessimistic expectations and anxious avoidance can result over time in depression-like motor inhibition.
 
-In this next section, we review the literature on aversive pruning.
-In large, multi-step sequential decision environments, the space of all possible sequences of choices grows exponentially larger with increasing sequence length.
-As such, it is infeasible for decision making agents to exhaustively explore all possible sequences select that which maximizes reward from the entire set.
-Heuristics for narrowing the search space then are not only plausible but beneficial for a decision making agent.
+\subsection{Planning \& Aversive Pruning}
 
-One such heuristic that has been proposed in the literature is aversive pruning \citep{Huys2012}.
-Aversive pruning is defined as a Pavlovian response to encountering a large loss in planning such that sequences involving large losses are discarded from further evaluation.
-An example of such a scenario is presented in Figure 5.
-The prediction in this decision tree environment is that an agent would discard the branch of the decision tree involving an immediate large negative loss, even though this branch objectively contains the reward maximizing (loss minimizing) sequence of choices.
+In large sequential choice environments, the total number of possible choice sequences grows exponentially with sequence length.
+As such, it is infeasible (if not impossible) for an agent to exhaustively evaluate all possible sequences of choices in order to carry out the one which maximizes reward.
+It is plausible then that rational agents may develop useful heuristics for winnowing out likely sub-optimal choice sequences.
+One such heuristic that has been proposed is aversive pruning \citep{Huys2012}, which is characterized as a Pavlovian response to losses such that choice sequences involving large losses are discarded from further evaluation during planning.
 
-\cite{Huys2012} (and later \cite{Lally2017}) tested this prediction in this particular choice environment.
-Participants received extensive training in this decision tree environment (first without outcomes and then later with rewards) in order to facilitate planning.
-In free-planning trials, many participants exhibited this Pavlovian bias, rejecting choice sequences which involved the large loss even when the sequence was objectively reward maximizing (loss minimizing).
-Importantly \cite{Huys2012} and \cite{Lally2017} found the degree of aversive pruning was correlated with depressive and anxiety symptoms, respectively.
+\cite{Huys2012} (and later \cite{Lally2017}) sought evidence for aversive pruning in the decision tree environment presented in Figure 4a.
+Participants in these experiments received extensive training in this environment such that participants reportedly knew well both the transition and reward structure of the tree.
+During testing, participants completed a number of free-planning trials in which participants started at various states within the tree and selected paths along the tree so as to maximize reward.
+\cite{Huys2012} and \cite{Lally2017} predicted that participants starting at the top of the tree (as depicted in Figure 4a) would be more likely to discard the left branch as it involved an immediate large loss and would likely be discarded through aversive pruning (and despite the fact that it was the parent path to the reward-maximizing choice sequence).
+Both \cite{Huys2012} and \cite{Lally2017} reported that many participants exhibited this pattern of behavior in evidence of an aversive pruning heuristic.
+Most importantly for present purposes, \cite{Huys2012} and \cite{Lally2017} found the degree of aversive pruning was correlated with depressive and anxiety symptoms, respectively.
 
@@
   }
-  \par \textbf{Figure 5:} The decision tree environment from \cite{Huys2012} and \cite{Lally2017} (A).
+  \par \textbf{Figure 4:} The decision tree environment from \cite{Huys2012} and \cite{Lally2017} (A).
 An optimistic agent ($w=1$) learns the optimal loss-minimizing policy through the initial large loss (B).
@@
 
-The aversive pruning principle makes tremendous sense and draws on a robust literature of heuristics and computational shortcuts for circumventing computationally intractable cognitive processes.
-It is important to note, however, that a control interpretation also makes similar predictions.
-Figure 5b-d shows the preferred routes of simulated agents in the same environments under increasingly pessimistic expectations of future actions.
-As can be observed, pessimistic agents are similarly likely to avoid the branch with the large loss, though for different reasons.
-Under the pessimistic case, the agent is increasingly less confident that the large gains later in sequence will be realized; as such, the agent is less confident that the initial large loss will be offset.
-Consequently, pessimistic agents would prefer the branch with smaller losses (even if this is objectively suboptimal).
-This prediction is similarly consistent with the finding that avoidance of the initial large loss is correlated with anxiety \cite{Lally2017}.
+The aversive pruning heuristic is an intuitive and sensible strategy for circumventing exhaustive search, which draws upon a rich literature of Pavlovian biases.
+We note, however, that our model makes similar predictions in the decision tree environment employed by \cite{Huys2012} and \cite{Lally2017}.
+Assuming exhaustive training, the optimal policy for an optimistic agent ($w=1.0$) beginning at the top of the tree is presented in Figure 4b.
+Note the optimistic agent traverses through the initial large loss in order to complete the sequence that maximizes reward (minimizes loss).
+Different patterns of behavior are exhibited by moderate ($w=0.75$, Figure 4c) and strongly ($w=0.50$, Figure 4d) pessimistic agents.
+For these agents, the possibility of inadvertently choosing a future sub-optimal path, and accruing a large loss, are non-negligible.
+As such, preference for the objectively sub-optimal path not involving the large loss becomes increasingly preferred.
+This preference is reminiscent of what is observed for the subclinically depressed and anxious participants in the original studies, suggesting that our model may also describe anxious decision making in this task.
+This is not implausible insofar that working memory ability is often impaired in anxiety disorders \citep{Moran2016}, and pessimistic expectations about future choice may be warranted as such.
 
-%% An interpretation like this would be consistent with reduced beliefs in one's own competency in such an environment.
-This is not strictly surprising.
-Planning multi-choice sequences even in relatively simple environments like those used in these experiments still ostensibly require working memory ability to keep in mind the possible consequences along each branch.
-Insofar that working memory ability is disrupted in anxiety disorders \citep{Moran2016}, a pessimistic belief may be justified and possibly optimal.
-
-It is worth noting that these two alternate accounts can be disentangled through slight manipulations of the choice environment presented above.
-Aversive pruning and loss of confidence make similar predictions when a large loss begins a sequence, but not necessarily when a large potential loss comes later in a sequence.
-In the loss of confidence view, the potential for large loss late in a sequence negatively impacts the estimated value of actions early in the sequence; this is not true for aversive pruning.
-Identifying such choice sequences in the decision tree task may help identify whether this sort of inference is present in individuals with anxiety performing the task.
+We note that this reinterpretation of results is not meant to replace the aversive pruning heuristic.
+Our model makes no strong claims about mechanistic interpretations, and thus does not address the issue of computational intractability whereas the aversive pruning heuristic does.
+We are simply noting that they make similar behavior predictions for anxious decision making in the decision tree environment used thus far.
+Interestingly, these accounts could be disentangled through a minor manipulation of the choice environment presented above.
+One would simply need to move the large loss to later in the tree.
+Under the aversive pruning view, large losses only impact choice behavior when encountered in sequence.
+Under pessimistic sequential choice inference, however, future large losses impact antecedent states, making their selection less likely.
+Thus, a slight modification to the decision tree structure presented in Figure 4a might be used to disentangle which of the two theories better accounts for anxious decision making in the task.
 
@@
 
-In this final section, we propose based on the present theory that anxiety should be inversely related to the free choice bias.
-If an agent is confident in their future decision making abilities, then it follows that situations in which an individual is able to make a future choice should be preferable to situations in which an individual is not.
-At worst, an agent is no better off than if they had not made a choice; at best, an agent is better able to steer themselves toward desirable outcomes.
+In this final section, we offer a novel prediction regarding the free choice premium.
+If an agent assumes they will be able to act in accordance to their (reward-maximizing) preferences in the future, then it follows that choices that enable an individual to make a future choice should be preferable to situations in which an individual is not.
+At worst, an agent is no better off than if they had not made a choice; at best, an agent is better able to steer themselves towards a desired outcome.
 Such is the logic that underlies an emerging literature suggesting that choice is inherently valuable \citep{Leotti2010}.
 
-This free choice premium has been observed across several behavioral experiments \citep{Suzuki1997, Leotti2011, Leotti2014, Cockburn2014} using a variety of decision paradigms.
-For the present purposes, we will describe the experiments presented in \citep{Leotti2011, Leotti2014}.
-In these experiments, participants complete a two-stage decision making task (Figure 6).
-In the first stage, participants are make a choice between two cues: one leading to free choice and the other leading to a fixed choice.
-In the second stage, participants are able to choose between a second set of cues (free choice) or are forced to choose a cue (fixed choice).
-Unbeknownst to participants, all cues yield reward following the same outcome distribution; thus, no cue is advantageous in the long-run.
+A free choice premium (i.e.
+choice is inherently valuable, all else being equal) has been observed across several behavioral experiments \citep{Suzuki1997, Leotti2011, Leotti2014, Cockburn2014} using several different experimental protocols.
+For present purposes, we will describe the protocol used by \cite{Leotti2011, Leotti2014}.
+Here, participants complete a two-stage decision making task (Figure 5a).
+In the first stage, participants choose between two cues: one yielding a future free choice and the other yielding a future fixed choice.
+In the second stage, participants either make a second choice (free choice) or have a choice made for them (fixed choice).
+Unbeknownst to participants, all choices in the second stage yield a reward drawn identical reward distributions; thus, no choice is actually advantageous in the long-run.
 The free choice premium is measured as a preference for the cue leading to free choice.
-\cite{Leotti2011, Leotti2014} find participants still prefer the free choice cue, despite this conferring no objective benefit to participants.
+\cite{Leotti2011, Leotti2014} find that participants prefer the cue leading to future choice, despite this conferring no objective benefit to participants.
 
-Insofar that the free choice premimum is contingent on an agent's belief in its future ability, it stands to reason that this free choice premium should be absent in individuals with anxiety and a loss of confidence.
-Simulations of this effect are presented in Figure 6.
-Simulated agents complete an experiment similar to that used in \cite{Leotti2011, Leotti2014}.
-In contrast to optimistic agents, the value of the worst action in the free choice arm back-propagated to the free-choice cue in pessimistic agents.
-With increasing pessimism then, the free choice option is less preferable than the fixed choice arm.
-As a result, pessimistic learners, on average, do not exhibit a bias towards the free choice arm.
-To our knowledge this prediction has not been empirically tested.
+We hypothesize that the free choice premium is contingent on an agent's positive expectations about its future actions.
+If this is true, it stands to reason that the free choice premium should be absent in anxious individuals.
+We present simulations of this effect in Figure 5b using a virtual experiment similar to that used in \cite{Leotti2011, Leotti2014}.
+As expected, optimistic agents ($w=1.0$) exhibit the free choice premium, measured as a preference for the cue leading to future choice.
+In contrast, moderately ($w=0.5$) and strongly ($w=0.0$) pessimistic agents exhibit a diminished free choice bias.
+In these agents, the value of worst action in the free choice is able to propagate backwards to the free choice action in stage 1.
+As such, the free choice confers no subjective benefit to the simulated agents and thus the free choice premium disappears.
+We plan on explicitly testing this hypothesis in subclinical and clinical anxiety in the near future.
 
@@
   }
-  \par \textbf{Figure 6:} Schematic of the free choice bias tasks used in \cite{Leotti2011, Leotti2014} (A). The free choice bias is diminished in increasingly pessimistic agents (B).
+  \par \textbf{Figure 5:} Schematic of the free choice bias tasks used in \cite{Leotti2011, Leotti2014} (A). The free choice bias is diminished in increasingly pessimistic agents (B).
 \end{figure}
@@
 
-\section{Appendix}
-
-\begin{algorithm}
-  \caption{Value Iteration}
-
-  \State Algorithm parameter: a small threshold $\theta > 0$ determining accuracy of estimation
-  \State Initialize $V(s)$, for all $s \in S$ arbitrarily, except that $V(terminal) = 0$
-  \State
-  \While{$\Delta > \theta$}
-    \State $\Delta \leftarrow 0$
-    \Loop \ for each $s \in S$
-      \State $v \leftarrow V(s)$
-      \State $ V(s) = \max_a \sum_{s',r} p(s',r|s,a) \left[ r + \gamma V(s') \right] $
-      \State $\Delta \leftarrow \max(\Delta, |v - V(s)|)$
-    \EndLoop
-  \EndWhile
-
-\end{algorithm}
-
 \end{document}

### 2019-02-25 manuscripts/rldm2019/template.tex
reorganization of manuscripts.

--- 2019-01-04 manuscript/template.tex
@@
 \documentclass[11pt]{article} % For LaTeX2e
-\usepackage{manuscript,palatino}
+\usepackage{rldmsubmit,palatino}
 \usepackage{graphicx}
 
-\title{something something about anxiety}
+\title{Formatting Instructions for RLDM}
 
 \author{
-Samuel Zorowitz \\
-Princeton Neuroscience Institute\\
-Princeton University\\
-Princeton, NJ 08540 \\
-\texttt{zorowitz@princeton.edu} \\
+Dorsal S.~Raphe\thanks{ Use footnote for providing further information
+about author (webpage, alternative address)---\emph{not} for acknowledging
+funding agencies.} \\
+Department of Computer Science\\
+Cranberry-Lemon University\\
+Pittsburgh, PA 15213 \\
+\texttt{raphe@cs.cranberry-lemon.edu} \\
 \And
-Nathaniel Daw \\
-Princeton Neuroscience Institute\\
-Princeton University\\
-Princeton, NJ 08540 \\
-\texttt{ndaw@princeton.edu} \\
+Coauthor \\
+Affiliation \\
+Address \\
+\texttt{email} \\
+\AND
+Coauthor \\
+Affiliation \\
+Address \\
+\texttt{email} \\
+\And
+Coauthor \\
+Affiliation \\
+Address \\
+\texttt{email} \\
+\And
+Coauthor \\
+Affiliation \\
+Address \\
+\texttt{email} \\
+(if needed)\\
 }
@@
 \begin{abstract}
-To be filled in
+The \emph{title} should be a maximum of 100 characters.
+
+The \emph{abstract} should be a maximum of 2000 characters of text,
+including spaces (no figure is allowed). You will be asked to copy
+this into a text-only box; and it will appear as such in the
+conference booklet. Use 11~point type, with a vertical spacing of
+12~points.  The word \textbf{Abstract} must be centered, bold, and in
+point size 12. Two line spaces precede the abstract.
 \end{abstract}
@@
 \keywords{
-reinforcement learning; avoidance; anxiety
+enter key words here
 }
+
+\acknowledgements{We are deeply indebted to Google DeepMind and
+the Weinberg Institute for Cognitive Science for
+their generous support of RLDM2017.}
 

### 2019-02-27 manuscripts/rldm2019/rldm2019.tex
first (near complete) draft of rldm 2019

\documentclass[11pt]{article} % For LaTeX2e
\usepackage{rldm2019, palatino}
\usepackage{graphicx}
\usepackage{amsfonts, amsmath}
\usepackage{algorithm, algpseudocode}%
\usepackage[numbers,super]{natbib}

\title{Anxiety, avoidance, and sequential evaluation}

\author{
Samuel Zorowitz \\
Princeton Neuroscience Institute\\
Princeton University\\
Princeton, NJ 08540 \\
\texttt{zorowitz@princeton.edu} \\
\And
Ida Momennejad \\
Columbia University\\
New York, NY 10027 \\
\texttt{ida.m@columbia.edu} \\
\And
Nathaniel D. Daw \\
Princeton Neuroscience Institute\\
and Department of Psychology\\
Princeton University\\
Princeton, NJ 08540 \\
\texttt{ndaw@princeton.edu} \\
}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\begin{document}

\maketitle

\begin{abstract}
Anxiety disorders are characterized by aberrations in the processing and response to threat, involving symptoms like exaggerated threat appraisal, fear generalization, and persistent avoidance. These symptoms are puzzling from a decision theoretic perspective. The generalization of fear and avoidance to situations far antecedent of threat violates the core logic of evaluation over sequential trajectories of action. This is because avoidance is by nature protective: hence, the ability to avoid in future means I need not also do so now. This is an instance of a general property of evaluating actions in sequential tasks: such evaluation turns fundamentally on assumptions about subsequent events, including the agent's own subsequent choices. Typically, it is appropriate to assume that the agent will also make good choices down the line. This line of reasoning, then, hints that a core deficit in anxious disorders may relate to this assumption, which otherwise should preclude the spread of avoidance. Indeed, anxiety disorders are associated with pessimistic beliefs about the future. Here we develop this idea into a simple model of evaluation under pessimistic assumptions about future choices. We show that this single, minor deviation can explain a surprising range of features of anxious behavior  including exaggerated threat appraisals, fear generalization, and persistent avoidance. This model also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety. This perspective is importantly different from previous computational accounts of beliefs about control in mood disorders, which neglected the sequential aspects of choice. 
\end{abstract}

\keywords{
anxiety; avoidance; fear generalization; decision theory; computational psychiatry
}

\acknowledgements{lalalalala.}

\startmain

\section{Introduction}

Anxiety disorders are characterized by aberrations in the processing of and response to threat \citep{dsm5, ClarkBeck2011}. Specifically, anxiety is associated with exaggerated threat appraisal, or the tendency to perceive threat as disproportionately greater in likelihood and severity than is warranted; fear generalization, or the spread of fear to situations distantly associated with the primary threat \citep{dymond2015}; and persistent avoidance behavior, which can be disruptive to everyday function \citep{Arnaudova2017}. Though laboratory studies of learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.

These symptoms are particularly puzzling from a decision theoretic perspective \citep{huys2015}. Here we identify one anomaly, which we suggest offers fundamental insights into the core of the dysfunction: The generalization of fear and avoidance to situations far antecedent of threat violates the core logic of evaluation over sequential trajectories of action. This is because avoidance is by nature protective: the ability to avoid in future means I need not also do so now. For instance, cars endanger pedestrians but can be avoided by following traffic signals; given that, staying indoors offers no additional protection from accidents. This is an instance of a general property of evaluating actions in sequential tasks: such evaluation turns fundamentally on assumptions about subsequent events, including the agent's own subsequent choices. Typically, it is appropriate to assume that the agent will also make good choices down the line.

This line of reasoning, then, hints that a core deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of avoidance. Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}. For instance, anxious individuals are more likely to endorse beliefs like that they are likely to encounter danger in the future and that they will be unable to mitigate it when they do. Here we develop this idea into a simple model of evaluation under pessimistic assumptions about future choices. We show that this single, minor deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance. This model also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety. This perspective is importantly different from previous computational accounts of beliefs about control in mood disorders (e.g. \cite{HuysDayan2009}), which neglected the sequential aspects of choice.

\section{Model description}

We model anxious decision making in the context of Markov decision processes (MDPs). A standard normative assumption is that agents attempt to optimize the expected cumulative discounted reward
\begin{equation*}
Q^\pi(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \sum_{a'} \pi(a' \mid s') Q^\pi(s',a'),
\end{equation*}
but for any particular state-action $a$, this is necessarily defined only relative to a policy $\pi(a' \mid s')$ specifying the assumed distribution of \emph{future} choices. The return can be optimized self-consistently assuming the agent makes the return-maximizing choice at each step in the future, leading to the familiar expression for the optimal values
\begin{equation}\label{eq:bellman}
 Q^*(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \max_{a'} Q^*(s',a').
\end{equation}
The ``max'' operator in Eq.~\ref{eq:bellman} yields a fundamental asymmetry between approach and avoidance, as illustrated in Fig.~1. The assumption that the agent makes the return-optimizing choice at each step implies that opportunities propagate recursively to earlier steps, but avoidable dangers do not. This principle is highlighted in a toy MDP: a deterministic open gridworld with terminal states containing a single reward and punishment (Fig.~1a). The optimal state values $V^* = \max_a Q^*(s,a)$ (Fig.~1b) reflect a ``mountain'' of opportunity propagating recursively from the reward. Conversely, because harm is avoidable in this environment, its negative value is contained: all states (even those adjacent to threat) represent the opportunity for reward.

Consider an agent who instead has pessimistic expectations about their future actions. Such returns can be adaptive to achieve robustness such as under adversarial scenarios \cite{Garcia2015}; here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms. We adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define
\begin{equation*}
Q^w(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \left( w \max_{a'} Q^w(s',a') + (1 - w) \min_{a'} Q^w(s',a') \right)
\end{equation*}
Here, the parameter $w$ controls the degree of pessimism. An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences. Fig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in the example gridworld; here, the expectation that the agent may fail to avoid in future causes the value of threat to permeate the state space.

\begin{figure}
  \centerline{%
    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/01_field.png}}%
  }
  \par \textbf{Figure 1:} A simple deterministic gridworld with two terminal states: one rewarding (blue) and aversive (red) state (A). For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value (B). For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states (C). With increasing pessimism ($w=0$), the extent of the spread grows worse (D). (Parameters: $\gamma = 0.95$)
\end{figure}

This simple simulation --- reflecting a small violation of a core decision theoretic assumption --- already echoes several core symptoms of anxiety disorders. The pessimistic agents in Fig~1c,d exhibit exaggerated threat appraisals (otherwise neutral states signal danger); generalization of fear (threat value spreads across the gridworld); and persistent avoidance (the agent chooses paths which maintain increasing distance from threat).

\section{Simulations}
\textbf{Approach-avoid conflict:} One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}. Many of the disruptions anxiety causes to everyday functioning (e.g. avoiding social obligations for fear of possible social humiliation) can be understood in these terms. An analogous phenomenon can be measured in the laboratory. For instance, in the balloon analog risk task (BART) \citep{Lejuez2002}, participants can win money by pumping virtual balloons. With each pump, the balloon inflates and money is earned, but so too does the chance that the balloon pops (and money is lost). At any point, a participant may cash out, banking the money earned and ending the trial. Anxiety is correlated with fewer pumps and earlier cash-outs in the BART \citep{Maner2007, ramirez2015}.

As shown in Fig.~2, our model easily accommodates this result. Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier. This is because it anticipates and avoids future errors in choice resulting in the balloon popping. Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety\cite{bach2015,Mobbs2019}. One unique prediction of the model is that this bias should arise only when beliefs about \emph{future} avoidance are involved, rather than direct conflict between immediate impulses. Recent data \cite{Mobbs2019} using a predator avoidance task analogous to the BART, support this view: sub-clinical anxiety predicted earlier escape (analogous to cash-out in the BART) for slow predators (for whom future decisions to escape were a relevant consideration) but not fast ones (who would attack immediately, negating consideration of future steps).

\textbf{Aversive pruning:} Another empirical phenomenon associated with anxiety is ``aversive pruning'' in planning \cite{Huys2012, Lally2017}. This refers to the idea that when evaluating future trajectories in a sequential task like chess, people must selectively consider certain paths and neglect others. One proposal for how people accomplish this is aversive pruning \citep{Huys2012}: choice sequences involving large losses are discarded from further evaluation. For instance, although the optimal choice in the decision tree from Fig.~3a is to weather an initial large loss in order to reap the large gain that follows, people tend to disfavor this path suggesting they neglect the later gain. The degree of such pruning correlates with depressive \cite{Huys2012} and anxiety \cite{Lally2017} symptoms. Our model predicts this result (Fig.3~b), though for a somewhat different reason: in our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them, but because with increasing anxiety they increasingly predict the possibility of choosing incorrectly afterward, and failing to recoup the loss. Future research could use slight variants in the decision trees could be used to tease apart these different interpretations.

\begin{figure}
  \centerline{%
    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/02_bart.png}}%
  }
  \par \textbf{Figure 2:} The Balloon Analog Risk Task (BART). The risk of balloon burst (point loss) increases with each pump (A). The optimal policy (number of pumps) under increasingly pessimistic valuation is presented for low (B) and high risk (C) balloons. The optimistic agent ($w=1$) prefers a policy reflecting the true environmental risk. The moderate ($w=0.6$) and strongly ($w=0.2$) pessimistic agents cash-out earlier, as is observed in anxious individuals. (Parameters: $\gamma = 1.0$)
\end{figure}

\textbf{The anxiety-depression transition:} So far, we have considered the asymptotic preferences implied the pessimistic value function. But we can also consider learning under this value function (e.g., by Q-learning or DYNA using the $\beta$-pessimistic return), the dynamics of which may speak to the progression of symptoms.

Of note in this respect, anxiety and depression are highly comorbid, with roughly 45\% of individuals with a lifetime depression diagnosis also diagnosed with an anxiety disorder \citep{kessler2015}. One interesting proposal is that this association often arises longitudinally: in particular, that anxiety precedes certain types of depression \citep{alloy1990, jacobson2014}. The idea, in brief, is that anxiety begets avoidance and foregone reward, leading ultimately to a belief that reward is unattainable and subsequently depression. This story is borne out by simulations of learning in our model (not shown, for space) in environments like that of Fig.~1: the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment with the progress of learning, ultimately leading the agent to expect no reward and, echoing anergic symptoms of depression, forego action altogether.

\section{Discussion}

Central to anxiety disorders are symptoms including pessimistic inference, threat generalization, and excessive avoidance. We have presented a simple computational account suggesting how a single underlying misbelief can give rise to a range of these aberrations in learning and choice. Specifically, we showed how a failure to believe in the reliability of future self-action can effectively backpropagate negative value across states of the environment. This process results in a range of inferences and behaviors resembling those observed in clinical anxiety. Though this is by no means a complete account of anxiety, it ties together a surprisingly wide range of its symptoms.

One important ambiguity is that we have formalized pessimistic assumptions in terms of beliefs about the agent's own future actions: expecting failure, for instance, to avoid in future. Since Eq.~1 is also computed in expectation over the anticipated future environmental dynamics $p(s' \mid s,a)$, pessimism can alternatively be encoded in this distribution (e.g., a belief that the world's response to one's choices is unpredictable or adversarial). Since the Bellman equation averages out both distributions, either formulation can produce ultimately similar results. However from the perspective of cognitive theories of anxiety, these represent quite different maladaptive beliefs, which may be relevant, for instance, in guiding treatment.

\begin{figure}[!b]
  \centerline{%
    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_tree_rldm.png}}%
  }
  \par \textbf{Figure 3:} The decision tree environment from \cite{Huys2012} and \cite{Lally2017} (A). An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (B). A pessimistic agent ($w=0.5$) comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss. The net value $Q$ on each path is shown numerically.
\end{figure}

Indeed, our account formalizes a longstanding strand of theory on the role of control in anxiety. Central to many prominent cognitive theories of anxiety in the psychiatric literature is a perceived lack of control. For example, learned helplessness theory claims that pathological anxiety results from a belief that the environment is uncontrollable, such that threat cannot be effectively mitigated \cite{alloy1990}. In contrast, Bandura's self-efficacy theory of anxiety posited that a lack of confidence in one's own ability to effectively handle threat was instead central \cite{bandura1977}. As we note above, the present model (though stated here in terms echoing Bandura's) can accommodate either variant, and they are by no means mutually exclusive.

Finally, we are not the first to propose a formal theory of control in psychiatry using MDPs. Huys\cite{HuysDayan2009} provided a computational account of learned helplessness through simple models of one-step action-outcome contingencies. Our accounts differ particularly in our exclusive focus on control in the sequential setting, which Huys did not address. Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future events.

\bibliographystyle{naturemag}
\small{\bibliography{rldm2019}}

\end{document}

### 2019-02-28 manuscripts/rldm2019/rldm2019.tex
further rldm2019 updates

@@
 \begin{abstract}
-Anxiety disorders are characterized by aberrations in the processing and response to threat, involving symptoms like exaggerated threat appraisal, fear generalization, and persistent avoidance.
-These symptoms are puzzling from a decision theoretic perspective.
-The generalization of fear and avoidance to situations far antecedent of threat violates the core logic of evaluation over sequential trajectories of action.
-This is because avoidance is by nature protective: hence, the ability to avoid in future means I need not also do so now.
-This is an instance of a general property of evaluating actions in sequential tasks: such evaluation turns fundamentally on assumptions about subsequent events, including the agent's own subsequent choices.
-Typically, it is appropriate to assume that the agent will also make good choices down the line.
-This line of reasoning, then, hints that a core deficit in anxious disorders may relate to this assumption, which otherwise should preclude the spread of avoidance.
-Indeed, anxiety disorders are associated with pessimistic beliefs about the future.
-Here we develop this idea into a simple model of evaluation under pessimistic assumptions about future choices.
-We show that this single, minor deviation can explain a surprising range of features of anxious behavior  including exaggerated threat appraisals, fear generalization, and persistent avoidance.
-This model also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety.
+Anxiety disorders are characterized by a range of aberrations in the processing and response to threat, but there is little clarity what core pathogenesis might underlie these symptoms.
+Here we propose a decision theoretic analysis of maladaptive avoidance and embody it in a reinforcement learning model, which shows how a localized bias in beliefs can formally explain a range of phenomena related to anxiety.
+The core observation, implicit in standard decision theoretic accounts of sequential evaluation, is that avoidance should be protective: if danger can be avoided later, it poses no threat now.
+We show how a violation of this assumption --- a pessimistic, false belief that later avoidance will be unsuccessful --- leads to a characteristic propagation of fear and avoidance to situations far antecedent of threat.
+This single deviation can explain a surprising range of features of anxious behavior, including exaggerated threat appraisals, fear generalization, and persistent avoidance.
+Simulations of the model reproduce laboratory demonstrations of abnormal decision making in anxiety, including in situations of approach-avoid conflict and planning to avoid losses.
+The model also ties together a number of other seemingly disjoint issues in anxious disorders.
+For instance, learning under the pessimistic bias captures a hypothesis about the role of anxiety in the later development of depression.
+The bias itself offers a new formalization of classic insights from the psychiatric literature about the central role of maladaptive beliefs about control and self-efficacy in anxiety.
 This perspective is importantly different from previous computational accounts of beliefs about control in mood disorders, which neglected the sequential aspects of choice.
@@
 
-\acknowledgements{lalalalala.}
+\acknowledgements{This research was supported in part by NIH grant MH10917, part of the CRCNS program.}
 
@@
 
-These symptoms are particularly puzzling from a decision theoretic perspective \citep{huys2015}.
-Here we identify one anomaly, which we suggest offers fundamental insights into the core of the dysfunction: The generalization of fear and avoidance to situations far antecedent of threat violates the core logic of evaluation over sequential trajectories of action.
+The puzzle of these symptoms is particularly sharp from a decision theoretic perspective \citep{huys2015}.
+Here we isolate and address a  decision theoretic paradox that, we argue, clarifies the core psychopathology: The generalization of fear and avoidance to situations far antecedent of threat violates the basic logic of evaluation over sequential trajectories of action.
 This is because avoidance is by nature protective: the ability to avoid in future means I need not also do so now.
-For instance, cars endanger pedestrians but can be avoided by following traffic signals; given that, staying indoors offers no additional protection from accidents.
+For instance, cars endanger pedestrians but can be reliably avoided by following traffic signals; given that, staying indoors offers no additional protection from accidents.
 This is an instance of a general property of evaluating actions in sequential tasks: such evaluation turns fundamentally on assumptions about subsequent events, including the agent's own subsequent choices.
@@
 
-This line of reasoning, then, hints that a core deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of avoidance.
+This line of reasoning, then, hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of avoidance.
 Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
-For instance, anxious individuals are more likely to endorse beliefs like that they are likely to encounter danger in the future and that they will be unable to mitigate it when they do.
-Here we develop this idea into a simple model of evaluation under pessimistic assumptions about future choices.
-We show that this single, minor deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance.
+For instance, anxious individuals are more likely to endorse beliefs like that they are likely to encounter danger in the future and, notably, that they will be unable to mitigate it when they do.
+Here we develop this idea into a model of evaluation under pessimistic assumptions about future choices.
+We show that this single, localized deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance.
 This model also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety.
-This perspective is importantly different from previous computational accounts of beliefs about control in mood disorders (e.g.
+Our perspective is importantly different from previous computational accounts of beliefs about control in mood disorders (e.g.
 \cite{HuysDayan2009}), which neglected the sequential aspects of choice.
@@
 Consider an agent who instead has pessimistic expectations about their future actions.
-Such returns can be adaptive to achieve robustness such as under adversarial scenarios \cite{Garcia2015}; here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms.
+Such returns can be adaptive to ensure robustness such as under adversarial scenarios \cite{Garcia2015}; here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms.
 We adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define
@@
 
-This simple simulation --- reflecting a small violation of a core decision theoretic assumption --- already echoes several core symptoms of anxiety disorders.
+This simple simulation --- reflecting a localized violation of a core decision theoretic assumption --- already echoes several core symptoms of anxiety disorders.
 The pessimistic agents in Fig~1c,d exhibit exaggerated threat appraisals (otherwise neutral states signal danger); generalization of fear (threat value spreads across the gridworld); and persistent avoidance (the agent chooses paths which maintain increasing distance from threat).
@@
 The degree of such pruning correlates with depressive \cite{Huys2012} and anxiety \cite{Lally2017} symptoms.
-Our model predicts this result (Fig.3~b), though for a somewhat different reason: in our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them, but because with increasing anxiety they increasingly predict the possibility of choosing incorrectly afterward, and failing to recoup the loss.
-Future research could use slight variants in the decision trees could be used to tease apart these different interpretations.
+Our model predicts this result (Fig.~3b), though for a somewhat different reason: in our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them, but because with increasing anxiety they increasingly predict the possibility of choosing incorrectly afterward, and failing to recoup the loss.
+Future research could use slight variants in the decision trees to tease apart these different interpretations.
 
@@
 
+A second point about learning in this model is that, because the pessimistic return ultimately spreads biased value estimates to states and actions far distant from the primary danger (e.g., Fig.~1d), it would accordingly also take a great many steps of iterative learning to correct all these misattributions, even given direct experience that the danger itself has passed.
+This may offer at least a partial answer to a classic puzzle in pathological avoidance, i.e.
+why it is so resistant to extinction \cite{moutoussis2018}.
+
+\begin{figure}[!b]
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_tree_rldm.png}}%
+  }
+  \par \textbf{Figure 3:} The decision tree environment from \cite{Huys2012} and \cite{Lally2017} (A).
+An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (B).
+A pessimistic agent ($w=0.5$) comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss.
+One-step rewards are presented in each state; the net value $Q$ on each path is shown numerically.
+\end{figure}
+
 \section{Discussion}
@@
 Since Eq.~1 is also computed in expectation over the anticipated future environmental dynamics $p(s' \mid s,a)$, pessimism can alternatively be encoded in this distribution (e.g., a belief that the world's response to one's choices is unpredictable or adversarial).
-Since the Bellman equation averages out both distributions, either formulation can produce ultimately similar results.
+Because the Bellman equation averages out both distributions,  and because an unpredictable environment also reduces the efficacy of avoidance, either formulation can produce ultimately similar results.
 However from the perspective of cognitive theories of anxiety, these represent quite different maladaptive beliefs, which may be relevant, for instance, in guiding treatment.
-
-\begin{figure}[!b]
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_tree_rldm.png}}%
-  }
-  \par \textbf{Figure 3:} The decision tree environment from \cite{Huys2012} and \cite{Lally2017} (A).
-An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (B).
-A pessimistic agent ($w=0.5$) comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss.
-The net value $Q$ on each path is shown numerically.
-\end{figure}
 

### 2019-02-28 sisyphus/envs/_tree.py

@@ -99,11 +99,19 @@ class DecisionTree(GraphWorld):
 
         return ax
 
-    def _draw_node_labels(self, ax, xpos, ypos, fontsize=14):
-        """Draw decision tree node labels. See plot_decision tree for details."""
+    def _draw_node_labels(self, ax, fontsize=14, color='w'):
+        """Draws one-step reward values as node labels."""
 
-        for i, (x, y) in enumerate(zip(xpos, ypos)):
-            ax.text(x, y, '%0.0f' %(i+1), ha='center', va='center', fontsize=fontsize)
+        ## Define node positions.
+        xpos = [-1,1,-1.5,-0.5,0.5,1.5,-1.75,-1.25,-0.75,-0.25,0.25,0.75,1.25,1.75]
+        ypos = [2, 2, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0]
+        
+        ## Define node values.
+        labels = [-70, -20, -20, -70, -20,-70, -20, 20, 20, 140, -20, 20, -20, 20]
+        
+        for x, y, label in zip(xpos, ypos, labels):
+            ax.text(x, y, '%0.0f' %label, ha='center', va='center', fontsize=fontsize, 
+                    fontweight='heavy', color=color)
 
         return ax
 
@@ -224,7 +232,7 @@ class DecisionTree(GraphWorld):
 
         ## Optional details.
         if path_sums: ax = self._draw_path_sums(ax, xpos[-8:], alpha=alpha)
-        if node_labels: ax = self._draw_node_labels(ax, xpos, ypos)
+        if node_labels: ax = self._draw_node_labels(ax)
         if isinstance(edge_labels, (list, tuple, np.ndarray)): 
             ax = self._draw_edge_labels(ax, edge_labels, alpha=edge_label_alpha)
         elif np.equal(edge_labels, True):

### 2019-02-28 manuscripts/rldm2019/rldm2019.tex
rldm2019 submitted

@@
 
-This line of reasoning, then, hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of avoidance.
+This line of reasoning hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of avoidance.
 Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
-For instance, anxious individuals are more likely to endorse beliefs like that they are likely to encounter danger in the future and, notably, that they will be unable to mitigate it when they do.
-Here we develop this idea into a model of evaluation under pessimistic assumptions about future choices.
+Anxious individuals are more likely to endorse beliefs like that they are likely to encounter danger in the future and, notably, that they will be unable to mitigate it when they do.
+We develop this idea into a model of evaluation under pessimistic assumptions about future choices.
 We show that this single, localized deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance.
@@
 
-Consider an agent who instead has pessimistic expectations about their future actions.
+Consider an agent who instead has pessimistic expectations about events at future steps.
 Such returns can be adaptive to ensure robustness such as under adversarial scenarios \cite{Garcia2015}; here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms.
-We adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define
+Although such pessimism can be encoded either in $P(s' \mid s,a)$ or $\pi(a \mid s)$, here for concreteness we adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define
 \begin{equation*}
@@
   }
-  \par \textbf{Figure 1:} A simple deterministic gridworld with two terminal states: one rewarding (blue) and aversive (red) state (A).
-For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value (B).
-For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states (C).
-With increasing pessimism ($w=0$), the extent of the spread grows worse (D).
+  \par \textbf{Figure 1:} (A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and aversive (red) state.
+(B) For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value.
+(C) For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states.
+(D) With increasing pessimism ($w=0$), the extent of the spread grows worse.
 (Parameters: $\gamma = 0.95$)
@@
 \section{Simulations}
-\textbf{Approach-avoid conflict:} One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}.
+\textbf{Approach-avoidance conflict:} One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}.
 Many of the disruptions anxiety causes to everyday functioning (e.g.
@@
   }
-  \par \textbf{Figure 2:} The Balloon Analog Risk Task (BART).
-The risk of balloon burst (point loss) increases with each pump (A).
+  \par \textbf{Figure 2:} (A) The Balloon Analog Risk Task (BART).
+The risk of balloon burst (point loss) increases with each pump.
 The optimal policy (number of pumps) under increasingly pessimistic valuation is presented for low (B) and high risk (C) balloons.
@@
   }
-  \par \textbf{Figure 3:} The decision tree environment from \cite{Huys2012} and \cite{Lally2017} (A).
-An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (B).
-A pessimistic agent ($w=0.5$) comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss.
+  \par \textbf{Figure 3:} (A) The decision tree environment from \cite{Huys2012} and \cite{Lally2017}.
+An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (left branch).
+(B) A pessimistic agent ($w=0.5$) comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss.
 One-step rewards are presented in each state; the net value $Q$ on each path is shown numerically.
+(Parameters: $\gamma = 1.0$)
 \end{figure}

### 2019-02-28 sisyphus/tests/test_dp.py
integrating travis-ci

@@ -1,28 +1,28 @@
 import numpy as np
-from sisyphus import ValueIteration
-from sisyphus.envs.base import GraphWorld
+from sisyphus.mdp import ValueIteration
+from sisyphus.envs._base import GraphWorld
 from sisyphus.tests.common import test_world
 
 def test_value_iteration():
     "Test value iteration dynamic programming algorithm."
-    
+
     ## Generate test gym.
     gym = GraphWorld(*test_world())
-    
+
     ## Test best-case learning.
     qvi = ValueIteration(policy='max', gamma=0.9)
     qvi = qvi.fit(gym)
     assert np.array_equal(qvi.Q, [ 0.9,  1. , -1. ,  0. ,  0. ])
     assert np.array_equal(qvi.V, [ 0.9,  1. ,  0. ,  0. ])
     assert np.array_equal(qvi.pi, np.arange(3))
-    
+
     ## Test worst-case learning.
     qvi = ValueIteration(policy='min', gamma=0.9)
     qvi = qvi.fit(gym)
     assert np.array_equal(qvi.Q, [-0.9,  1. , -1. ,  0. ,  0. ])
     assert np.array_equal(qvi.V, [-0.9,  1. ,  0. ,  0. ])
     assert np.array_equal(qvi.pi, np.arange(3))
-    
+
     ## Test betamax learning.
     qvi = ValueIteration(policy='betamax', gamma=0.9, beta=0.5)
     qvi = qvi.fit(gym)

### 2019-02-28 sisyphus/tests/test_envs.py

@@ -1,19 +1,19 @@
 import numpy as np
-from sisyphus.envs.base import GraphWorld
+from sisyphus.envs._base import GraphWorld
 from sisyphus.tests.common import test_world
 
 def test_graph_world():
     """Test GraphWorld environment initialization."""
-    
+
     ## Generate test gym.
     gym = GraphWorld(*test_world())
-    
+
     ## Tests of states.
     assert np.array_equal(gym.states, np.arange(4))
     assert np.equal(gym.n_states, 4)
     assert np.array_equal(gym.viable_states, np.arange(2))
     assert np.equal(gym.n_viable_states, 2)
-    
+
     ## Tests of info.
     assert np.array_equal(gym.info.shape, [5,4])
     assert np.array_equal(gym.info["S"].values,           [0, 1, 1, 2, 3])

### 2019-02-28 sisyphus/tests/test_td.py

@@ -1,32 +1,32 @@
 import numpy as np
-from sisyphus import ModelFree
-from sisyphus.envs.base import GraphWorld
+from sisyphus.mdp import ModelFree
+from sisyphus.envs._base import GraphWorld
 from sisyphus.tests.common import test_world
 
 def test_model_free():
     """Test model free temporal difference learning algorithm."""
     np.random.seed(47404)
-    
+
     ## Generate test gym.
     gym = GraphWorld(*test_world())
-    
+
     ## Define exploration schedule.
     schedule = np.zeros(100)
-    
+
     ## Test best-case learning.
     agent = ModelFree(policy='max', eta=0.2, gamma=0.9)
     agent = agent.fit(gym, choice='softmax', schedule=schedule)
     assert np.allclose(agent.Q, [ 0.9,  1. , -1. ,  0. ,  0. ], atol=1e-3, rtol=0)
     assert np.allclose(agent.V, [ 0.9,  1. ,  0. ,  0. ], atol=1e-3, rtol=0)
     assert np.array_equal(agent.pi, np.arange(3))
-    
+
     ## Test worst-case learning.
     agent = ModelFree(policy='min', eta=0.2, gamma=0.9)
     agent = agent.fit(gym, choice='softmax', schedule=schedule)
     assert np.allclose(agent.Q, [-0.9,  1. , -1. ,  0. ,  0. ], atol=1e-3, rtol=0)
     assert np.allclose(agent.V, [-0.9,  1. ,  0. ,  0. ], atol=1e-3, rtol=0)
     assert np.array_equal(agent.pi, np.arange(3))
-    
+
     ## Test betamax learning.
     agent = ModelFree(policy='betamax', eta=0.2, gamma=0.9, beta=0.5)
     agent = agent.fit(gym, choice='softmax', schedule=schedule)

### 2019-02-28 sisyphus/tests/test_dp.py
fixed tests to reflect new jargon

@@ -24,7 +24,7 @@ def test_value_iteration():
     assert np.array_equal(qvi.pi, np.arange(3))
 
     ## Test betamax learning.
-    qvi = ValueIteration(policy='betamax', gamma=0.9, beta=0.5)
+    qvi = ValueIteration(policy='pessimism', gamma=0.9, w=0.5)
     qvi = qvi.fit(gym)
     assert np.array_equal(qvi.Q, [ 0.0,  1. , -1. ,  0. ,  0. ])
     assert np.array_equal(qvi.V, [ 0.0,  1. ,  0. ,  0. ])

### 2019-02-28 sisyphus/tests/test_td.py

@@ -28,7 +28,7 @@ def test_model_free():
     assert np.array_equal(agent.pi, np.arange(3))
 
     ## Test betamax learning.
-    agent = ModelFree(policy='betamax', eta=0.2, gamma=0.9, beta=0.5)
+    agent = ModelFree(policy='pessimism', eta=0.2, gamma=0.9, w=0.5)
     agent = agent.fit(gym, choice='softmax', schedule=schedule)
     assert np.allclose(agent.Q, [ 0.0,  1. , -1. ,  0. ,  0. ], atol=1e-3, rtol=0)
     assert np.allclose(agent.V, [ 0.0,  1. ,  0. ,  0. ], atol=1e-3, rtol=0)

### 2019-04-14 manuscripts/manuscript/manuscript.tex
making ready for overleaf import

--- 2019-02-28 manuscripts/rldm2019/rldm2019.tex
@@
 
+Distributional RL, Plagiarized (Bishop 2018)
+"Hence, anxious individuals might have even weaker upper bounds than other individuals for probability estimates of rare, extremely negative events.
+One strategy for robustly avoiding catastrophic failures is to adopt a worse-case scenario (H 1 control) [9], that is, to rely on the upper bound as opposed to the mean or median of the distribution of outcome probabilities.
+Consistent with this, clinically anxious individuals are reported to engage in catastrophizing, focusing on worst case outcomes [10]"
+
+There is a third that we did not address here: that of the state itself.
+Here we assumed that states were fully observable; in other words, there was no uncertainty about the state an agent was currently occupying.
+In partially-observable Markov decision problems, the state itself must be inferred from external and internal variables.
+This is yet another means by which bad value can spread if undue likelihood is granted to undesirable or aversive states.
+This ideas has previously been proposed to explain some symptoms in anxiety \citep{Paulus2012}, and not without good reason.
+Anxiety disorders involve negatively biased interpretation of ambiguous percepts \citep{Hartley2012}, which may result in aberrant planning and increased avoidance for similar reasons to those described above.
+This is an area for future research.
+
+Model-based
+want to build an account about worry, thinking anxiety is correlated with more worry time, and worry is specifically carrying out these sorts of thought patterns.
+
+Berenbaum (2010):
+First, lower levels of self-esteem, which might be expected to be associated with lower levels of perceived competence, have been found to be associated with higher levels of worrying (Meyer, Miller, Metzger, & Borkovec, 1990).
+Second, examining the content of anxiety diaries, Breitholtz, Johansson and Ost (1999) found that individuals with GAD were much more likely to write about doubts concerning their competence than were individuals with panic disorder.
+Third, Davey and Levy (1999) found that individuals with higher levels of worrying reported that they were more likely to endorse statements concerning personal inadequacy/incompetence when thinking about “ a stressful, traumatic or diffi cult situation.” Fourth, Davey and Levy (1998) found that the more severe the participants' worrying, the more they couched their worries in terms of personal inadequacies.
+Finally, in two separate studies, Berenbaum, Thompson, and Bredemeier (2007) found that lower levels of perceived competence were associated with higher levels of worry- ing.
+Further, the link between perceived competence and worrying was mediated by perceived probabilities of undesirable outcomes.
+
+Replay, Plagiarized (Bishop 2018):
+"Off-line replay of prior experiences and simulation of counterfactual actions and associated outcomes provide a means to update action values following the occurrence of an extremely negative event [18].
+It has been argued that model-based previous states should be prioritized for replay based on how much that replay would change value estimates [34 ].
+One way to accomplish this is by tagging states based on how much their successors value has changed [35].
+If change in value estimates determines priority for replay, the astronomically large discrepancy in outcome value occasioned by the occurrence of a rare, extremely negative event would be expected to result in prioritized replay of that events antecedents.""
+
 Finally, we are not the first to propose a formal theory of control in psychiatry using MDPs.

### 2019-04-14 sisyphus/envs/_prey.py

@@ -6,12 +6,8 @@ class SleepingPredator(GraphWorld):
     
     Parameters
     ----------
-    pumps : int
-        Maximum number of balloon pumps.
-    mu : float
-        Average state at which balloon pops.
-    sd : float
-        Deviation around mean.
+    p : float
+        Probability of predation.
     
     Attributes
     ----------
@@ -36,7 +32,7 @@ class SleepingPredator(GraphWorld):
        J Exp Psychol Hum Percept Perform 43:18–29.
     """
     
-    def __init__(self, p=0.1, n_bins=1):
+    def __init__(self, p=0.1):
         
         ## Define one-step transition matrix.
         n = 7
@@ -69,11 +65,10 @@ class SleepingPredator(GraphWorld):
         self.info = self.info[sane_ix].reset_index(drop=True)
             
         ## Update probability of being eaten.  
-        pmf = p * np.sum([(1-p)**i for i in range(n_bins)])
         for i, row in self.info.iterrows():
             s, s_prime = row["S"], row["S'"][0]
             if not s_prime in self.terminal:
-                self.info.at[i,'T'] = np.array([1-pmf, 0, pmf])
+                self.info.at[i,'T'] = np.array([1-p, 0, p])
                 
     def __repr__(self):
         return '<GraphWorld | Sleeping Predator Task>'

### 2019-04-14 manuscripts/manuscript/manuscript.tex

--- 2019-02-28 manuscripts/rldm2019/rldm2019.tex
@@
 \documentclass[11pt]{article} % For LaTeX2e
-\usepackage{rldm2019, palatino}
+\usepackage{manuscript, palatino}
 \usepackage{graphicx}
@@
 
+\begin{figure}[!b]
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/05_tree_rldm.png}}%
+  }
+  \par \textbf{Figure 3:} (A) The decision tree environment from \cite{Huys2012} and \cite{Lally2017}.
+An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (left branch).
+(B) A pessimistic agent ($w=0.5$) comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss.
+One-step rewards are presented in each state; the net value $Q$ on each path is shown numerically.
+(Parameters: $\gamma = 1.0$)
+\end{figure}
+
 \textbf{The anxiety-depression transition:} So far, we have considered the asymptotic preferences implied the pessimistic value function.
@@
 The idea, in brief, is that anxiety begets avoidance and foregone reward, leading ultimately to a belief that reward is unattainable and subsequently depression.
-This story is borne out by simulations of learning in our model (not shown, for space) in environments like that of Fig.~1: the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment with the progress of learning, ultimately leading the agent to expect no reward and, echoing anergic symptoms of depression, forego action altogether.
+This story is borne out by simulations of learning in our model (Fig.~4) in environments like that of Fig.~1: the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment with the progress of learning, ultimately leading the agent to expect no reward and, echoing anergic symptoms of depression, forego action altogether.
 
@@
   \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_tree_rldm.png}}%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_lh.png}}%
   }
@@
 
+The two accounts described above are by no means the only possible reinforcement learning formalisms that can give rise to in threat-avoidance behavior.
+Traditionally, reinforcement learning algorithms aim to find reward-maximizing policies with respect to the expected (average) returns.
+However, in environments with the potential for catastrophic loss it may be preferable to optimize a policy with respect to an alternative objective function insofar that the expected return does not necessarily avoid rare yet large negative outcomes.
+Such risk-sensitive reinforcement learning algorithms have been the subject of much recent research \cite{geibel2005, morimura2012} and show threat-avoidance behaviors not unlike those presented here.
+An often reported clinical observation is that anxious individuals disproportionately dwell upon low chance yet high cost possibilities.
+\cite{Miceli2005} One speculative hypothesis is that anxious individuals optimize with respect to a more conservative risk criterion.
+
+A related possibility concerns the notion of state itself.
+Here we assumed that states were fully observable, such that there was no uncertainty regarding the state an agent was currently occupying.
+In real-world environments, however, the state must be inferred from external and internal variables.
+As has been proposed elsewhere \cite{Paulus2012}, one possibility is that anxiety biases state inference resulting in erroneous beliefs about likely future events.
+There is evidence that anxious individuals are biased in their sampling and interpretation of incoming threat information \cite{Paulus2012}, suggesting that the avoidance behaviors may be a function of pessimistic beliefs about the current state of the environment.
+
+Of course, none of the above constitutes a claim about the biological mechanisms underlying pessimistic sequential evaluation.
+One possibility is that these symptoms reflect biased model-based planning.
+Pathological worry is strongly associated with anxiety disorders and may be viewed as a biased form of future action planning.
+In line with the present results, higher levels of worry is associated with reduced self-efficacy beliefs, greater endorsement of statements concerning personal inadequacies in the face of threat, and that the association between perceived competence and worry is mediated by exaggerated threat appraisal (for review, see \cite{Berenbaum2010}).
+As such, the behaviors observed above may reflect biased sampling of possible future outcomes as a function of pessimistic beliefs about self-efficacy.
+
+As alluded to above, a second possibility is that biased sequential evaluation may arise as a function of biased offline replay of events.
+During periods of rest and sleep, the hippocampus replays recent experiences.
+Recent modeling has shown that hippocampal replay is in part driven by gain from propagating newly encountered information to preceding locations.
+Biases in the tagging of memories for replay, as a function of either pessimistic expected future state occupancy or biased learning of information, may yield the propagation of threat value similarly as has been observed above.
+
 Finally, we are not the first to propose a formal theory of control in psychiatry using MDPs.
@@
 \bibliographystyle{naturemag}
-\small{\bibliography{rldm2019}}
+\small{\bibliography{manuscript}}
 

### 2019-06-24 manuscripts/manuscript/manuscript.tex
pushing previous changes

@@
 Specifically, anxiety is associated with exaggerated threat appraisal, or the tendency to perceive threat as disproportionately greater in likelihood and severity than is warranted; fear generalization, or the spread of fear to situations distantly associated with the primary threat \citep{dymond2015}; and persistent avoidance behavior, which can be disruptive to everyday function \citep{Arnaudova2017}.
-Though laboratory studies of learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.
+Though laboratory studies of decision making and learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.
+
+% Can probably fill in some of the above. Specifically flesh out descriptions of symptoms. It's rather short.
 
@@
 This is because avoidance is by nature protective: the ability to avoid in future means I need not also do so now.
-For instance, cars endanger pedestrians but can be reliably avoided by following traffic signals; given that, staying indoors offers no additional protection from accidents.
+For instance, cars endanger pedestrians but can be reliably avoided by following traffic signals; given that, staying indoors offers little or no additional protection from accidents.
 This is an instance of a general property of evaluating actions in sequential tasks: such evaluation turns fundamentally on assumptions about subsequent events, including the agent's own subsequent choices.
 Typically, it is appropriate to assume that the agent will also make good choices down the line.
+
+% might want to clean up language for clarity, but this is right. It's the "generalization of fear..." sentence that should be clarified. It's opaque.
 
@@
 We develop this idea into a model of evaluation under pessimistic assumptions about future choices.
-We show that this single, localized deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance.
+We show that a single, localized deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance.
 This model also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety.
@@
 \cite{HuysDayan2009}), which neglected the sequential aspects of choice.
+
+% "more likely to endorse beliefs..." should also be cleaned up. Just say threat is more likely
 
@@
 
-Consider an agent who instead has pessimistic expectations about events at future steps.
-Such returns can be adaptive to ensure robustness such as under adversarial scenarios \cite{Garcia2015}; here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms.
-Although such pessimism can be encoded either in $P(s' \mid s,a)$ or $\pi(a \mid s)$, here for concreteness we adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define
+An agent could expect to encounter dangerous future events under different assumptions about the future: either if the agent expects to fail to take the correct avoidance actions (i.e., expects to use a nonoptimal $\pi(a \mid s)$), or if the world's dynamics do not guarantee reliable avoidance even so (e.g., under stochastic or adversarial transition dynamics $P(s' \mid s,a)$).
+Consider an agent who has such pessimistic expectations about such events at future steps.
+Computing returns with such pessimistic predictions can help to quantify uncertainty in outcomes (i.e.
+to learn different points in the distribution of possible returns) \cite{deepmindstuff}, and choosing on this basis can be adaptive to ensure robustness such as under adversarial scenarios \cite{Garcia2015}.
+
+Here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms.
+Although such pessimism can be encoded either in $P(s' \mid s,a)$ or $\pi(a \mid s)$, here for concreteness we work with the policy, and adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define
 \begin{equation*}
@@
 \end{equation*}
-Here, the parameter $w$ controls the degree of pessimism.
+The parameter $w$ controls the degree of pessimism.
 An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences.
-Fig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in the example gridworld; here, the expectation that the agent may fail to avoid in future causes the value of threat to permeate the state space.
+(We return to pessimism about state transitions in the discussion.) åFig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in the example gridworld; here, the expectation that the agent may fail to avoid in future causes the value of threat to permeate the state space.
 
@@
   \par \textbf{Figure 1:} (A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and aversive (red) state.
+(B-D) States colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
 (B) For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value.
@@
 \section{Simulations}
+
 \textbf{Approach-avoidance conflict:} One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}.
@@
 
+\begin{figure}
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/02_appavo.png}}%
+  }
+  \par \textbf{Figure 2:} (A) The Balloon Analog Risk Task (BART).
+The risk of balloon burst (point loss) increases with each pump.
+The optimal policy (number of pumps) under increasingly pessimistic valuation is presented for low (B) and high risk (C) balloons.
+The optimistic agent ($w=1$) prefers a policy reflecting the true environmental risk.
+The moderate ($w=0.6$) and strongly ($w=0.2$) pessimistic agents cash-out earlier, as is observed in anxious individuals.
+(D/E) By contrast, in the sleeping predator task the risk of loss is constant but the cost of loss increases.
+The value of reward pursuit under increasingly pessimistic valuation is presented for low (D) and high risk (E) scenarios.
+(Parameters: $\gamma = 1.0$)
+\end{figure}
+
 As shown in Fig.~2, our model easily accommodates this result.
@@
 This is because it anticipates and avoids future errors in choice resulting in the balloon popping.
-Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety\cite{bach2015,Mobbs2019}.
+Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety\cite{Mobbs2019}.
 One unique prediction of the model is that this bias should arise only when beliefs about \emph{future} avoidance are involved, rather than direct conflict between immediate impulses.
 Recent data \cite{Mobbs2019} using a predator avoidance task analogous to the BART, support this view: sub-clinical anxiety predicted earlier escape (analogous to cash-out in the BART) for slow predators (for whom future decisions to escape were a relevant consideration) but not fast ones (who would attack immediately, negating consideration of future steps).
+
+The model can also explain findings of increased behavioral inhibition, measured as slower response times, in anxious individuals under threat.
+\cite{bach2015} In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator".
+Though the risk of predation is fixed, the potential loss from capture increases with each token collected.
+Bach \cite{bach2015} finds that participants are slower to collect tokens as loss increases, and that this is modulated by trait anxiety.
+The model recreates this phenomenon (Fig.~2d/e) as the relative value of approach as compared to avoidance is reduced under pessimistic assumptions about future action.
 
@@
 
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/02_bart.png}}%
-  }
-  \par \textbf{Figure 2:} (A) The Balloon Analog Risk Task (BART).
-The risk of balloon burst (point loss) increases with each pump.
-The optimal policy (number of pumps) under increasingly pessimistic valuation is presented for low (B) and high risk (C) balloons.
-The optimistic agent ($w=1$) prefers a policy reflecting the true environmental risk.
-The moderate ($w=0.6$) and strongly ($w=0.2$) pessimistic agents cash-out earlier, as is observed in anxious individuals.
-(Parameters: $\gamma = 1.0$)
-\end{figure}
-
 \begin{figure}[!b]
   \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/05_tree_rldm.png}}%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_tree_rldm.png}}%
   }
@@
   \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_lh.png}}%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/03_lh.png}}%
   }
@@
 \end{figure}
+
+\textbf{Free choice premium:} In this final section, we offer a novel prediction regarding the free choice premium.
+Simply put, the free choice premium predicts that, all else being equal, opportunities that beget more choice opportunities in the future are more valuable than those that do not.
+A free choice premium has been observed in multiple behavioral experiments.
+\cite{Suzuki1997, Leotti2011, Leotti2014, Cockburn2014} A variant of a free choice premium paradigm is presented in Fig.
+5a.
+In the task, participants repeated choose between a free choice option, allowing for an additional future choice, and a fixed choice option; importantly, both choices lead to identical stochastic outcomes (50/50 chance of [1, -1]).
+
+\begin{figure}[!b]
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/05_choice.png}}%
+  }
+  \par \textbf{Figure 5:} (A) The free choice premium paradigm from \cite{Leotti2011} and \cite{Leotti2014}.
+Non-anxious participants exhibit a preference for the free choice option (blue) despite it conferring no benefit over the fixed choice option (grey).
+(B) Pessimistic agents show an attenuated free choice bias.
+(Parameters: $\gamma = 1.0$, Inverse temperature, $\beta$ increased from 1 to 10 over 100 trials.
+Simulation of 500 agents.)
+\end{figure}
+
+If the free choice bias is a function of an individual's optimistic expectations about future choice opportunities, then we predict that anxious individuals should exhibit a reduced free choice bias.
+Indeed, the model makes this prediction (Fig.
+~5b).
+We plan on explicitly testing this hypothesis in subclinical and clinical anxiety in the near future.
 
@@
 
+%% Can probably flesh this out. Cite problem orientation literature as evidence.
+
 The two accounts described above are by no means the only possible reinforcement learning formalisms that can give rise to in threat-avoidance behavior.
@@
 
+% Citing the Wu stuff may be helpful here
+
 Of course, none of the above constitutes a claim about the biological mechanisms underlying pessimistic sequential evaluation.
@@
 During periods of rest and sleep, the hippocampus replays recent experiences.
-Recent modeling has shown that hippocampal replay is in part driven by gain from propagating newly encountered information to preceding locations.
-Biases in the tagging of memories for replay, as a function of either pessimistic expected future state occupancy or biased learning of information, may yield the propagation of threat value similarly as has been observed above.
+Recent work has shown that episodes are prioritized for replay based according to their likelihood in changing behavior, and these replay events iteratively backpropagate value to antecedent states.
+Biases in the tagging of memories for replay, as a function of either biased expectations of future state occupancy (need) or biased estimates in the value to be learned, may yield the propagation of threat value similarly as has been observed above.
+
+It is important to note that we do not believe that this model of anxiety is equally explanatory across the family of anxiety disorders.
+Indeed, those disorders more strongly associated with disordered prospective cognition, such as generalized anxiety disorder and social anxiety disorder, are more likely to be better described by biased sequential evaluation.
+This account will have trouble explaining other features of anxiety disorders, such as the compulsive behaviors of obsessive-compulsive disorder and the memory disturbancse associated with post-traumatic stress disorder.
+Importantly, other psychiatric disorders with avoidance behaviors, such as eating disorders, may also potentially be well described by this model.
+As such, this account is transdiagnostic.
+
+% this is probably fine?
 

### 2019-06-25 manuscripts/manuscript/manuscript.tex
reworked intro

@@
 
-Anxiety disorders are characterized by aberrations in the processing of and response to threat \citep{dsm5, ClarkBeck2011}.
-Specifically, anxiety is associated with exaggerated threat appraisal, or the tendency to perceive threat as disproportionately greater in likelihood and severity than is warranted; fear generalization, or the spread of fear to situations distantly associated with the primary threat \citep{dymond2015}; and persistent avoidance behavior, which can be disruptive to everyday function \citep{Arnaudova2017}.
+Though the anxiety disorders differ in their particular symptomology, and in the content and situations which elicit anxious responding, they all are characterized by aberrations in the processing of and response to threat\citep{dsm5}.
+In particular, at least three symptoms seem to manifest across anxiety disorders.
+First, anxiety is associated with exaggerated threat appraisal, or a bias towards evaluating threat as disproportionately greater in likelihood and severity than is warranted\citep{ClarkBeck2011}.
+Second, anxiety is also associated with fear generalization, wherein the primary threat becomes associated with increasingly distal locations, situations, and events\citep{dymond2015}.
+Finally, anxiety is associated with persistent avoidance behavior, which indirectly works to maintain pathology by preventing the experience of non-outcomes of expected bad events\citep{Arnaudova2017}.
 Though laboratory studies of decision making and learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.
 
-% Can probably fill in some of the above. Specifically flesh out descriptions of symptoms. It's rather short.
+These symptoms are particularly puzzling from a decision theoretic perspective\citep{huys2015}.
+Fear and avoidance of situations far antecedent of threat violates the basic logic of evaluation over sequential trajectories of action.
+Put another way, future threat need not necessarily influence decision making in the present.
+This is because avoidance is by nature protective: the ability to avoid in future means an agent need not also do so now.
+For instance, cars endanger pedestrians but can be reliably avoided by following traffic signals; given that, staying indoors offers little or no additional protection from accidents.
+This is an instance of a general property of evaluating actions in sequential decision making: such evaluation turns fundamentally on assumptions about subsequent events, including the agent's own subsequent choices.
+Typically, it is appropriate to assume that an agent will also make reward-maximizing (and harm-minimizing) choices down the line.
 
-The puzzle of these symptoms is particularly sharp from a decision theoretic perspective \citep{huys2015}.
-Here we isolate and address a decision theoretic paradox that, we argue, clarifies the core psychopathology: The generalization of fear and avoidance to situations far antecedent of threat violates the basic logic of evaluation over sequential trajectories of action.
-This is because avoidance is by nature protective: the ability to avoid in future means I need not also do so now.
-For instance, cars endanger pedestrians but can be reliably avoided by following traffic signals; given that, staying indoors offers little or no additional protection from accidents.
-This is an instance of a general property of evaluating actions in sequential tasks: such evaluation turns fundamentally on assumptions about subsequent events, including the agent's own subsequent choices.
-Typically, it is appropriate to assume that the agent will also make good choices down the line.
+This line of reasoning hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of threat and excessive avoidance.
+Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
+Trait and clinically anxious individuals rate the likelihood of future threatening scenarios higher than do non-anxious individuals\citep{???}.
+Most importantly, anxious individuals are more likely to believe that they are unable to handle and mitigate future threat\citep{???}.
 
-% might want to clean up language for clarity, but this is right. It's the "generalization of fear..." sentence that should be clarified. It's opaque.
-
-This line of reasoning hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of avoidance.
-Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
-Anxious individuals are more likely to endorse beliefs like that they are likely to encounter danger in the future and, notably, that they will be unable to mitigate it when they do.
-We develop this idea into a model of evaluation under pessimistic assumptions about future choices.
+Here, we develop this idea into a model of evaluation under pessimistic assumptions about future choices.
 We show that a single, localized deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance.
-This model also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety.
+This account also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety.
+Specifically, we demonstrate a model which distinctly lacks a belief in the efficacy of future action gives rise to the aforementioned symptoms.
 Our perspective is importantly different from previous computational accounts of beliefs about control in mood disorders (e.g.
 \cite{HuysDayan2009}), which neglected the sequential aspects of choice.
-
-% "more likely to endorse beliefs..." should also be cleaned up. Just say threat is more likely
 

### 2019-06-30 manuscripts/manuscript/manuscript.tex
intro completed. almost done discussion.

@@
 
-Though the anxiety disorders differ in their particular symptomology, and in the content and situations which elicit anxious responding, they all are characterized by aberrations in the processing of and response to threat\citep{dsm5}.
-In particular, at least three symptoms seem to manifest across anxiety disorders.
+Though anxiety disorders differ in their particular symptomology, and in the content and situations which elicit anxious responding, they all are similarly characterized by aberrations in the processing of and response to threat\citep{dsm5}.
+In particular, at least three symptoms are manifest across the anxiety disorders.
 First, anxiety is associated with exaggerated threat appraisal, or a bias towards evaluating threat as disproportionately greater in likelihood and severity than is warranted\citep{ClarkBeck2011}.
-Second, anxiety is also associated with fear generalization, wherein the primary threat becomes associated with increasingly distal locations, situations, and events\citep{dymond2015}.
-Finally, anxiety is associated with persistent avoidance behavior, which indirectly works to maintain pathology by preventing the experience of non-outcomes of expected bad events\citep{Arnaudova2017}.
+Second, anxiety is also associated with fear generalization, wherein the primary threat becomes associated with increasingly distal locations, events, and thoughts\citep{dymond2015}.
+Finally, anxiety is associated with persistent avoidance behavior, which often occurs well in advance of the materialization of actual threat\citep{Arnaudova2017}.
+Avoidance behaviors in anxiety are especially pernicious as they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
 Though laboratory studies of decision making and learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.
@@
 These symptoms are particularly puzzling from a decision theoretic perspective\citep{huys2015}.
-Fear and avoidance of situations far antecedent of threat violates the basic logic of evaluation over sequential trajectories of action.
-Put another way, future threat need not necessarily influence decision making in the present.
-This is because avoidance is by nature protective: the ability to avoid in future means an agent need not also do so now.
+Fear and avoidance of situations far in the future violates the basic logic of evaluation over sequential trajectories of action.
+Put another way, distant threat need not necessarily impinge on decision making in the present.
+This is because avoidance is by nature protective: the ability to successfully avoid danger in the future means an agent need not also do so now.
 For instance, cars endanger pedestrians but can be reliably avoided by following traffic signals; given that, staying indoors offers little or no additional protection from accidents.
-This is an instance of a general property of evaluating actions in sequential decision making: such evaluation turns fundamentally on assumptions about subsequent events, including the agent's own subsequent choices.
-Typically, it is appropriate to assume that an agent will also make reward-maximizing (and harm-minimizing) choices down the line.
+This is an instance of a general property of evaluation in sequential decision making: the value of present action turns fundamentally on assumptions about subsequent events, which include the agent's own subsequent choices.
+Typically, it is appropriate to assume that an agent will continue make good (i.e.
+reward-maximizing/harm-minimizing) choices down the line.
 
@@
 Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
-Trait and clinically anxious individuals rate the likelihood of future threatening scenarios higher than do non-anxious individuals\citep{???}.
-Most importantly, anxious individuals are more likely to believe that they are unable to handle and mitigate future threat\citep{???}.
+Trait and pathologically anxious individuals judge future threat as more likely than do non-anxious individuals\citep{ButlerMathews1983, ButlerMathews1987, MacleodByrne1996}.
+Importantly, the development and maintenance of clinical anxiety is strongly tied to diminished perceived control\citep{bandura1977, barlow2002, gallagher2014a}, such that anxious individuals are more likely to endorse the belief that they are unlikely or unable to mitigate future threat.
+Indeed, a belief (or lack thereof) of one's ability to successfully navigate future danger is associated with anxiety\citep{davey1996, dugas1997}, and an increased belief in perceived control over threat is correlated with symptom reduction across the family of anxiety disorders\citep{gallagher2014b}.
 
@@
 We show that a single, localized deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance.
-This account also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety.
-Specifically, we demonstrate a model which distinctly lacks a belief in the efficacy of future action gives rise to the aforementioned symptoms.
+This account also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety \citep{bandura1977, barlow2002}.
+Specifically, we demonstrate a model with a misbelief in the reliability of future self-action gives rise to the anxious symptoms.
 Our perspective is importantly different from previous computational accounts of beliefs about control in mood disorders (e.g.
@@
 Computing returns with such pessimistic predictions can help to quantify uncertainty in outcomes (i.e.
-to learn different points in the distribution of possible returns) \cite{deepmindstuff}, and choosing on this basis can be adaptive to ensure robustness such as under adversarial scenarios \cite{Garcia2015}.
+to learn different points in the distribution of possible returns)\cite{bellemare2017}, and choosing on this basis can be adaptive to ensure robustness such as under adversarial scenarios\cite{Garcia2015}.
 
@@
 
-Central to anxiety disorders are symptoms including pessimistic inference, threat generalization, and excessive avoidance.
-We have presented a simple computational account suggesting how a single underlying misbelief can give rise to a range of these aberrations in learning and choice.
+Central to anxiety disorders are symptoms including exaggerated threat appraisal, threat generalization, and excessive avoidance.
+We have presented a simple computational account suggesting how a single underlying misbelief can give rise to these aberrations in learning and choice.
 Specifically, we showed how a failure to believe in the reliability of future self-action can effectively backpropagate negative value across states of the environment.
@@
 
-One important ambiguity is that we have formalized pessimistic assumptions in terms of beliefs about the agent's own future actions: expecting failure, for instance, to avoid in future.
+One important ambiguity is that we have formalized pessimistic assumptions in terms of beliefs about the agent's own future actions: expecting failure in handling or avoiding future threat.
 Since Eq.~1 is also computed in expectation over the anticipated future environmental dynamics $p(s' \mid s,a)$, pessimism can alternatively be encoded in this distribution (e.g., a belief that the world's response to one's choices is unpredictable or adversarial).
@@
 Central to many prominent cognitive theories of anxiety in the psychiatric literature is a perceived lack of control.
-For example, learned helplessness theory claims that pathological anxiety results from a belief that the environment is uncontrollable, such that threat cannot be effectively mitigated \cite{alloy1990}.
-In contrast, Bandura's self-efficacy theory of anxiety posited that a lack of confidence in one's own ability to effectively handle threat was instead central \cite{bandura1977}.
-As we note above, the present model (though stated here in terms echoing Bandura's) can accommodate either variant, and they are by no means mutually exclusive.
+For example, self-efficacy theory\citep{bandura1977} and the triple vulnerabilities model\citep{barlow2002} both posit that a reduced belief in the ability to effectively respond to future threat is involved in the genesis and maintenance of clinical anxiety.
+Similarly, the learned helplessness theories of anxiety\cite{alloy1990} claims pathological anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
+As we note above, the present model (though stated here in terms of self-efficacy) can accommodate either variant, and they are by no means mutually exclusive.
 
-%% Can probably flesh this out. Cite problem orientation literature as evidence.
+Distortions in beliefs about future action or transition probabilities are in no way the only means through which anxious behaviors may arise in a decision theoretic account.
+Traditionally, the goal of reinforcement learning algorithms is to a find reward-maximizing policy with respect to the expectation (average) of returns.
+However, in environments where the distribution of returns has high variance, such that there is potential for catastrophic loss, it may be instead preferable to learn a policy with respect to an alternative value function.
+The same is true when there is uncertainty in the transition probabilities, such that an agent may inadvertently step towards undesirable and/or dangerous segments of the environment.
+How to learn risk-sensitive and robust policies is the focus of much recent research \cite{morimura2012, chow2015, bellemare2017}; interestingly, these studies have reported agent behaviors not unlike those presented here.
+It is important to note that anxiety is correlated with disproportionate worry about low chance but high cost events\cite{Miceli2005}.
+One speculative hypothesis is that anxious individuals learn and evaluate action with respect  to a more pessimistic point in the distribution of expected returns.
 
-The two accounts described above are by no means the only possible reinforcement learning formalisms that can give rise to in threat-avoidance behavior.
-Traditionally, reinforcement learning algorithms aim to find reward-maximizing policies with respect to the expected (average) returns.
-However, in environments with the potential for catastrophic loss it may be preferable to optimize a policy with respect to an alternative objective function insofar that the expected return does not necessarily avoid rare yet large negative outcomes.
-Such risk-sensitive reinforcement learning algorithms have been the subject of much recent research \cite{geibel2005, morimura2012} and show threat-avoidance behaviors not unlike those presented here.
-An often reported clinical observation is that anxious individuals disproportionately dwell upon low chance yet high cost possibilities.
-\cite{Miceli2005} One speculative hypothesis is that anxious individuals optimize with respect to a more conservative risk criterion.
+A separate possibility concerns the notion of state.
+In the present work, we considered only environments with fully observable states; in other words, there was no uncertainty regard the state an agent was currently occupying.
+In the real-world, however, the state must be inferred from external and internal observations.
+One possibility is that anxiety is a function of biases in state inference\cite{Paulus2012}.
+For example, if an agent erroneously comes to the conclusion that she is more proximal to threat than is actually the case, then symptoms like exaggerated threat appraisal and avoidance should arise.
+Indeed, a bias in value functions could result in the diffuse spread of threat across objects\citep{norbury2018} and location\citep{schulz2018}, thereby causing otherwise benign states to be perceived as threatening.
 
-A related possibility concerns the notion of state itself.
-Here we assumed that states were fully observable, such that there was no uncertainty regarding the state an agent was currently occupying.
-In real-world environments, however, the state must be inferred from external and internal variables.
-As has been proposed elsewhere \cite{Paulus2012}, one possibility is that anxiety biases state inference resulting in erroneous beliefs about likely future events.
-There is evidence that anxious individuals are biased in their sampling and interpretation of incoming threat information \cite{Paulus2012}, suggesting that the avoidance behaviors may be a function of pessimistic beliefs about the current state of the environment.
+None of the previous discussion constitutes a claim about the algorithmic and neural bases of pessimistic sequential evaluation.
+One possibility is that these symptoms reflect aberrations in model-based planning.
+Chronic worry is a symptom frequently observed in pathological anxiety, and could be construed as a deviant form of future action planning wherein the roll-out of future actions is biased towards bad- or worst-case contingencies.
+In line with the present results, higher levels of worry is correlated with reduced perceived control, increased doubt in self-efficacy in response to threat, and exaggerated threat appraisal\cite{Berenbaum2010}.
+In sum, anxious symptoms may in part result from biased planning where cost-minimizing action is discounted and
 
-% Citing the Wu stuff may be helpful here
+%% Need to clean the above up
 
-Of course, none of the above constitutes a claim about the biological mechanisms underlying pessimistic sequential evaluation.
-One possibility is that these symptoms reflect biased model-based planning.
-Pathological worry is strongly associated with anxiety disorders and may be viewed as a biased form of future action planning.
-In line with the present results, higher levels of worry is associated with reduced self-efficacy beliefs, greater endorsement of statements concerning personal inadequacies in the face of threat, and that the association between perceived competence and worry is mediated by exaggerated threat appraisal (for review, see \cite{Berenbaum2010}).
-As such, the behaviors observed above may reflect biased sampling of possible future outcomes as a function of pessimistic beliefs about self-efficacy.
+A second possibility is that pessimistic evaluation may arise as a result of biased offline replay of events.
+During periods of rest and sleep, the hippocampus replays recent experiences\citep{whats a good ref?}.
+A recent theoretical model of replay suggests that these replay events are able to iteratively backpropagate value from a focal point to antecedent states, and that episodes are prioritized for replay according to their likelihood in changing behavior.
+Biases in replay behavior, or in the mechanisms initiate replay, might result in repeated offline reinforcement of negative episodes thereby exaggerating the value of threat and diffusing its influence across an agent's internal cognitive map.
+Few studies have investigated replay behavior in aversive environments (e.g.
+\citep{wu2017}), so this hypothesis remains speculative at this point in time.
 
-As alluded to above, a second possibility is that biased sequential evaluation may arise as a function of biased offline replay of events.
-During periods of rest and sleep, the hippocampus replays recent experiences.
-Recent work has shown that episodes are prioritized for replay based according to their likelihood in changing behavior, and these replay events iteratively backpropagate value to antecedent states.
-Biases in the tagging of memories for replay, as a function of either biased expectations of future state occupancy (need) or biased estimates in the value to be learned, may yield the propagation of threat value similarly as has been observed above.
+%% Want to add anything about transition probability bias?
 
-It is important to note that we do not believe that this model of anxiety is equally explanatory across the family of anxiety disorders.
-Indeed, those disorders more strongly associated with disordered prospective cognition, such as generalized anxiety disorder and social anxiety disorder, are more likely to be better described by biased sequential evaluation.
-This account will have trouble explaining other features of anxiety disorders, such as the compulsive behaviors of obsessive-compulsive disorder and the memory disturbancse associated with post-traumatic stress disorder.
-Importantly, other psychiatric disorders with avoidance behaviors, such as eating disorders, may also potentially be well described by this model.
-As such, this account is transdiagnostic.
-
-% this is probably fine?
+It is important to note that the present model may not describe all anxiety disorders with equal accuracy.
+Indeed, our model of pessimistic sequential evaluation is, by definition, a model of prospective cognition.
+As such, the present results are more likely to accurately describe the anxiety disorders which primarily involve aberrations in future-oriented cognitive processes, such as generalized and social anxiety disorders.
+Naturally, the present model can neither account for the compulsive behaviors of obsessive-compulsive disorder nor the memory disturbances of post-traumatic stress disorder (though recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD\citep{gallagher2014a, gallagher2014b}).
+Importantly, other psychiatric disorders that involve future-oriented misbeliefs, worry, and avoidance behaviors (e.g.
+eating disorders\citep{konstantellou2011}) may similarly be well-described by the current account.
+As such, the present results are transdiagnostic and not limited to one particular diagnosis.
 
 Finally, we are not the first to propose a formal theory of control in psychiatry using MDPs.
-Huys\cite{HuysDayan2009} provided a computational account of learned helplessness through simple models of one-step action-outcome contingencies.
-Our accounts differ particularly in our exclusive focus on control in the sequential setting, which Huys did not address.
-Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future events.
+Huys \& Dayan \cite{HuysDayan2009} also provided a computational account of learned helplessness through simple models of one-step action-outcome contingencies.
+Our accounts differ particularly in our exclusive focus on control in the sequential setting, which Huys \& Dayan did not address.
+Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future choice and events.
 

### 2019-06-30 manuscripts/manuscript/manuscript.tex
working on results.

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 \usepackage{amsfonts, amsmath}
@@
 
-We model anxious decision making in the context of Markov decision processes (MDPs). A standard normative assumption is that agents attempt to optimize the expected cumulative discounted reward
+We model anxious decision making in the context of Markov decision processes (MDPs). A standard normative assumption is that agents attempt to optimize the expected cumulative discounted reward,
+
 \begin{equation*}
@@
 \end{equation*}
+
 but for any particular state-action $a$, this is necessarily defined only relative to a policy $\pi(a' \mid s')$ specifying the assumed distribution of \emph{future} choices.
-The return can be optimized self-consistently assuming the agent makes the return-maximizing choice at each step in the future, leading to the familiar expression for the optimal values
+The return can be optimized self-consistently under the assumption that the agent makes the return-maximizing choice at each step in the future, leading to the familiar expression for the optimal values,
+
 \begin{equation}\label{eq:bellman}
@@
 \end{equation}
+
 The ``max'' operator in Eq.~\ref{eq:bellman} yields a fundamental asymmetry between approach and avoidance, as illustrated in Fig.~1.
-The assumption that the agent makes the return-optimizing choice at each step implies that opportunities propagate recursively to earlier steps, but avoidable dangers do not.
+The assumption that the agent makes the return-maximizing choice at each step implies that opportunities for reward propagate recursively to earlier steps, but avoidable dangers do not.
 This principle is highlighted in a toy MDP: a deterministic open gridworld with terminal states containing a single reward and punishment (Fig.~1a).
@@
 
-An agent could expect to encounter dangerous future events under different assumptions about the future: either if the agent expects to fail to take the correct avoidance actions (i.e., expects to use a nonoptimal $\pi(a \mid s)$), or if the world's dynamics do not guarantee reliable avoidance even so (e.g., under stochastic or adversarial transition dynamics $P(s' \mid s,a)$).
-Consider an agent who has such pessimistic expectations about such events at future steps.
-Computing returns with such pessimistic predictions can help to quantify uncertainty in outcomes (i.e.
-to learn different points in the distribution of possible returns)\cite{bellemare2017}, and choosing on this basis can be adaptive to ensure robustness such as under adversarial scenarios\cite{Garcia2015}.
+An agent is not restricted to the return-maximizing assumption, and could anticipate encountering danger under different assumptions about the future.
+For example, an agent may expect to fail to take the correct protective actions (i.e., expects to use a nonoptimal $\pi(a \mid s)$), or may expect the world's future dynamics do not guarantee reliable avoidance even so (i.e., under stochastic or adversarial transition dynamics $P(s' \mid s,a)$).
+Consider an agent who has such pessimistic expectations about dangerous events at future steps.
+Computing returns with such pessimistic predictions might help to quantify uncertainty in outcomes (i.e.
+to learn different points in the distribution of possible returns)\cite{bellemare2017}, and choosing on this basis can be adaptive in that it increases the chances of safety in uncertain or even adversarial scenarios\cite{Garcia2015}.
 
 Here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms.
-Although such pessimism can be encoded either in $P(s' \mid s,a)$ or $\pi(a \mid s)$, here for concreteness we work with the policy, and adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define
+Although such pessimism can be encoded either in the policy, $P(s' \mid s,a)$, or transition probabilities, $\pi(a \mid s)$, here for concreteness we work with distortions in the policy.
+In particular, we adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define an expectation of state-action value,
+
 \begin{equation*}
@@
 \end{equation*}
+
 The parameter $w$ controls the degree of pessimism.
 An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences.
-(We return to pessimism about state transitions in the discussion.) åFig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in the example gridworld; here, the expectation that the agent may fail to avoid in future causes the value of threat to permeate the state space.
+(We return to pessimism about state transitions in the discussion.) åFig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in the example gridworld.
+An increasing expectation that the agent may fail to avoid in future causes the value of threat to permeate with increasing distance across the state space.
 
@@
   }
-  \par \textbf{Figure 1:} (A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and aversive (red) state.
+  \par \textbf{Figure 1:} (A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
 (B-D) States colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
@@
 
-This simple simulation --- reflecting a localized violation of a core decision theoretic assumption --- already echoes several core symptoms of anxiety disorders.
-The pessimistic agents in Fig~1c,d exhibit exaggerated threat appraisals (otherwise neutral states signal danger); generalization of fear (threat value spreads across the gridworld); and persistent avoidance (the agent chooses paths which maintain increasing distance from threat).
+This simple simulation --- reflecting a localized violation of the core decision theoretic assumption of future return-optimizing action --- already echoes several core symptoms of anxiety disorders.
+Namely, the pessimistic agents in Fig~1c,d exhibit exaggerated threat appraisals (otherwise neutral states unrealistically signal danger); generalization of fear (threat value spreads across the gridworld); and persistent avoidance (early on, the agent takes paths which maintain increasing distances from threat).
+Importantly, as will be elaborated on below, this deviation from the usual assumptions is supported by prominent clinical theories of anxiety.
 
@@
 
-\textbf{Approach-avoidance conflict:} One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}.
+In the following, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of learning and decision making.
+We will also show that our model is consistent with clinical theory describing the transition from pathological anxiety to depression.
+Unless otherwise noted, state and action values under varying degrees of pessimism were solved for using the value iteration dynamic programming method\citep{SuttonBarto2018}.
+All simulations were implemented in the Python programming language and the code is publicly available at \url{https://github.com/ndawlab/seqanx}.
+
+\subsection{Approach-avoidance conflict}
+
+One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}; anxious individuals are more likely to forgo potential gains in order to avoid potential danger.
 Many of the disruptions anxiety causes to everyday functioning (e.g.
 avoiding social obligations for fear of possible social humiliation) can be understood in these terms.
-An analogous phenomenon can be measured in the laboratory.
-For instance, in the balloon analog risk task (BART) \citep{Lejuez2002}, participants can win money by pumping virtual balloons.
-With each pump, the balloon inflates and money is earned, but so too does the chance that the balloon pops (and money is lost).
-At any point, a participant may cash out, banking the money earned and ending the trial.
-Anxiety is correlated with fewer pumps and earlier cash-outs in the BART \citep{Maner2007, ramirez2015}.
+As such, many have sought to probe and measure this behavior in the laboratory.
+For instance, in the balloon analog risk task (BART) \citep{Lejuez2002}, participants attempt to earn money by pumping virtual balloons.
+With each pump, the balloon inflates and money is earned, but so too does the chance that the balloon pops and the accumulated earnings are lost.
+At any point in a trial, a participant may cash out, banking the money earned and ending the trial.
+Anxiety is correlated with fewer pumps of the balloon and earlier cash-outs in the BART \citep{Maner2007, ramirez2015}.
 
@@
 
-\textbf{Aversive pruning:} Another empirical phenomenon associated with anxiety is ``aversive pruning'' in planning \cite{Huys2012, Lally2017}.
+\subsection{Aversive pruning} 
+Another empirical phenomenon associated with anxiety is ``aversive pruning'' in planning \cite{Huys2012, Lally2017}.
 This refers to the idea that when evaluating future trajectories in a sequential task like chess, people must selectively consider certain paths and neglect others.
@@
 
-\textbf{The anxiety-depression transition:} So far, we have considered the asymptotic preferences implied the pessimistic value function.
+\subsection{The anxiety-depression transition}
+So far, we have considered the asymptotic preferences implied the pessimistic value function.
 But we can also consider learning under this value function (e.g., by Q-learning or DYNA using the $\beta$-pessimistic return), the dynamics of which may speak to the progression of symptoms.
@@
 
-\textbf{Free choice premium:} In this final section, we offer a novel prediction regarding the free choice premium.
+\subsection{Free choice premium}
+In this final section, we offer a novel prediction regarding the free choice premium.
 Simply put, the free choice premium predicts that, all else being equal, opportunities that beget more choice opportunities in the future are more valuable than those that do not.

### 2019-06-30 manuscripts/manuscript/manuscript.tex
finished results

@@
 Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
-Trait and pathologically anxious individuals judge future threat as more likely than do non-anxious individuals\citep{ButlerMathews1983, ButlerMathews1987, MacleodByrne1996}.
+Clinically and subclinically anxious individuals judge future threat as more likely than do non-anxious individuals\citep{ButlerMathews1983, ButlerMathews1987, MacleodByrne1996}.
 Importantly, the development and maintenance of clinical anxiety is strongly tied to diminished perceived control\citep{bandura1977, barlow2002, gallagher2014a}, such that anxious individuals are more likely to endorse the belief that they are unlikely or unable to mitigate future threat.
@@
 
-In the following, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of learning and decision making.
+In the following, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of sequential learning and decision making.
 We will also show that our model is consistent with clinical theory describing the transition from pathological anxiety to depression.
@@
 \subsection{Approach-avoidance conflict}
+
+% Fix subsection spacing?
 
@@
 Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier.
-This is because it anticipates and avoids future errors in choice resulting in the balloon popping.
+This is because it anticipates and avoids future errors in choice which would otherwise result in the balloon popping.
 Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety\cite{Mobbs2019}.
 One unique prediction of the model is that this bias should arise only when beliefs about \emph{future} avoidance are involved, rather than direct conflict between immediate impulses.
-Recent data \cite{Mobbs2019} using a predator avoidance task analogous to the BART, support this view: sub-clinical anxiety predicted earlier escape (analogous to cash-out in the BART) for slow predators (for whom future decisions to escape were a relevant consideration) but not fast ones (who would attack immediately, negating consideration of future steps).
+Recent data \cite{Mobbs2019} using a predator avoidance task, analogous to the BART, support this view.
+In their task, trait anxiety predicted earlier escape (analogous to cash-out in the BART) for slow predators (for whom future decisions to escape were a relevant consideration) but not fast ones (who would attack immediately, negating consideration of future steps).
 
-The model can also explain findings of increased behavioral inhibition, measured as slower response times, in anxious individuals under threat.
-\cite{bach2015} In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator".
-Though the risk of predation is fixed, the potential loss from capture increases with each token collected.
-Bach \cite{bach2015} finds that participants are slower to collect tokens as loss increases, and that this is modulated by trait anxiety.
-The model recreates this phenomenon (Fig.~2d/e) as the relative value of approach as compared to avoidance is reduced under pessimistic assumptions about future action.
+The model can also explain findings of increased behavioral inhibition, measured as prolonged response times, in anxious individuals under threat\cite{bach2015}.
+In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator".
+Though the risk of predation is constant throughout a trial, the potential loss from capture increases with each token collected.
+Bach \cite{bach2015} finds that participants are slower to collect tokens as loss increases, and that this effect is modulated by subclinical anxiety.
+Our model recreates this phenomenon (Fig.~2d/e).
+Insofar that value and response time are negatively related \citep{need citation?}, and the relative value of approach as compared to avoidance is reduced under pessimistic assumptions about future action, then prolonged response times naturally follow.
 
 \subsection{Aversive pruning}
+
 Another empirical phenomenon associated with anxiety is ``aversive pruning'' in planning \cite{Huys2012, Lally2017}.
-This refers to the idea that when evaluating future trajectories in a sequential task like chess, people must selectively consider certain paths and neglect others.
-One proposal for how people accomplish this is aversive pruning \citep{Huys2012}: choice sequences involving large losses are discarded from further evaluation.
-For instance, although the optimal choice in the decision tree from Fig.~3a is to weather an initial large loss in order to reap the large gain that follows, people tend to disfavor this path suggesting they neglect the later gain.
-The degree of such pruning correlates with depressive \cite{Huys2012} and anxiety \cite{Lally2017} symptoms.
-Our model predicts this result (Fig.~3b), though for a somewhat different reason: in our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them, but because with increasing anxiety they increasingly predict the possibility of choosing incorrectly afterward, and failing to recoup the loss.
+This refers to the idea that when evaluating future action trajectories in a sequential task like chess, people are resource-limited, cannot evaluate all possible options, and must selectively consider certain paths and neglect others.
+One proposal for how people accomplish this is aversive pruning \citep{Huys2012}: choice sequences which involve large losses are discarded from further evaluation.
+An example of aversive pruning is shown in Fig.~3a.
+Although the optimal choice in the decision tree from is to weather an initial large loss (e.g., -70) in order to reap the large gain that follows, people tend to disfavor this path suggesting they prune it and consequently neglect the later gain.
+The degree of such pruning correlates with subclinical depressive \cite{Huys2012} and anxiety \cite{Lally2017} symptoms.
+
+Our model predicts this result (Fig.~3b), though for a somewhat different reason.
+In our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them, but because with increasing anxiety they increasingly expect the potential of choosing incorrectly afterward, thus failing to recoup the loss.
+This reinterpretation is consistent with the structure of the original task\citep{Huys2012, Lally2017}, where participants were required to plan in advance action sequences of two- to eight-steps (requiring the consideration of upwards of $2^8 = 256$ options).
+Uncertainty in one's ability to plan a sequence of reward-maximizing choices could similarly bias a participants towards smaller cost actions.
 Future research could use slight variants in the decision trees to tease apart these different interpretations.
+
+% Include the larger, four-panel aversive pruning figure instead?
 
@@
 \subsection{The anxiety-depression transition}
-So far, we have considered the asymptotic preferences implied the pessimistic value function.
-But we can also consider learning under this value function (e.g., by Q-learning or DYNA using the $\beta$-pessimistic return), the dynamics of which may speak to the progression of symptoms.
 
-Of note in this respect, anxiety and depression are highly comorbid, with roughly 45\% of individuals with a lifetime depression diagnosis also diagnosed with an anxiety disorder \citep{kessler2015}.
-One interesting proposal is that this association often arises longitudinally: in particular, that anxiety precedes certain types of depression \citep{alloy1990, jacobson2014}.
-The idea, in brief, is that anxiety begets avoidance and foregone reward, leading ultimately to a belief that reward is unattainable and subsequently depression.
-This story is borne out by simulations of learning in our model (Fig.~4) in environments like that of Fig.~1: the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment with the progress of learning, ultimately leading the agent to expect no reward and, echoing anergic symptoms of depression, forego action altogether.
+So far, we have considered the asymptotic preferences implied the pessimistic value function (solved for through value iteration).
+But we can also consider learning under this value function (e.g., by Q-learning\citep{SuttonBarto2018} or DYNA\citep{sutton1991} using the $\beta$-pessimistic return), the dynamics of which may speak to the progression of symptoms.
 
-A second point about learning in this model is that, because the pessimistic return ultimately spreads biased value estimates to states and actions far distant from the primary danger (e.g., Fig.~1d), it would accordingly also take a great many steps of iterative learning to correct all these misattributions, even given direct experience that the danger itself has passed.
+Of note in this respect, anxiety and depression are highly comorbid, with almost half of individuals with a lifetime depression diagnosis also diagnosed with an anxiety disorder \citep{kessler2015}.
+One notable proposal is that this association often arises longitudinally: in particular, that clinical anxiety precedes certain types of depression \citep{alloy1990, jacobson2014}.
+The idea, in brief, is that uncertainty in one's ability in the face of future threat results in anxiety and avoidance behaviors.
+Persistent avoidance, in turn, begets foregone reward, and ultimately to a belief that reward is unobtainable and subsequently depression.
+This story is borne out by simulations of learning in our model (Fig.~4) in environments like that of Fig.~1.
+As we observed before, the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment with the progress of learning.
+This in turn leads the agent to expect no reward and, echoing the anergic symptoms of depression, forego action altogether.
+
+In addition to accounting for comorbidity of anxiety and depression, our model also hints at a reason for the longevity and recurrence of anxiety disorders even with treatment.
+Because pessimistic expectations allow for threat value to spread to states and actions far antecedent of the primary danger (e.g., Fig.~1d), it would accordingly also take a great many steps of iterative learning to correct all these exaggerated appraisals of threat.
+Frustratingly, these biased estimates of value may still remain even after a misbelief in the efficacy of future action is corrected for in a course of therapy.
 This may offer at least a partial answer to a classic puzzle in pathological avoidance, i.e.
-why it is so resistant to extinction \cite{moutoussis2018}.
+why it is so resistant to extinction \cite{moutoussis2018}, and to the unfortunately high rates of anxiety recurrence following treatment\citep{pittig2018}.
 
@@
   }
-  \par \textbf{Figure 3:} (A) The decision tree environment from \cite{Huys2012} and \cite{Lally2017}.
-An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (left branch).
-(B) A pessimistic agent ($w=0.5$) comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss.
-One-step rewards are presented in each state; the net value $Q$ on each path is shown numerically.
-(Parameters: $\gamma = 1.0$)
+  \par \textbf{Figure 4:} A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
+(B-D) States colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
+For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value with learning.
+(E-G) For a pessimistic agent ($w=0.6$), negative value spreads from the source to antecedent states.
+As a result of avoidance, the agent learns reward is unobtainable and develops anergic symptoms (i.e.
+foregoes action).
+(Parameters: $\gamma = 0.95$)
 \end{figure}
@@
 \subsection{Free choice premium}
+
 In this final section, we offer a novel prediction regarding the free choice premium.
-Simply put, the free choice premium predicts that, all else being equal, opportunities that beget more choice opportunities in the future are more valuable than those that do not.
-A free choice premium has been observed in multiple behavioral experiments.
-\cite{Suzuki1997, Leotti2011, Leotti2014, Cockburn2014} A variant of a free choice premium paradigm is presented in Fig.
+Simply put, the free choice premium phenomenon predicts that, all else being equal, options which lead to more choice opportunities in the future are more valuable than those that lead to fewer future choice opportunities.
+A free choice premium has been observed in multiple behavioral experiments \cite{Leotti2010, ly2019}.
+A variant of a free choice premium paradigm from \cite{Leotti2011} and \cite{Leotti2014} is presented in Fig.
 5a.
-In the task, participants repeated choose between a free choice option, allowing for an additional future choice, and a fixed choice option; importantly, both choices lead to identical stochastic outcomes (50/50 chance of [1, -1]).
+In the task, participants repeated choose between a free choice option, allowing for an additional future choice, and a fixed choice option; importantly, both choices lead to identical, stochastic outcomes (e.g.
+50/50 chance of [1, -1]).
+Previous studies have found a preference for the free choice option despite it conferring no additional benefits relative to the fixed choice option.
 
@@
   }
-  \par \textbf{Figure 5:} (A) The free choice premium paradigm from \cite{Leotti2011} and \cite{Leotti2014}.
+  \par \textbf{Figure 4:} (A) The free choice premium paradigm from \cite{Leotti2011} and \cite{Leotti2014} with equal chance of outcome $R \in [1, -1]$.
 Non-anxious participants exhibit a preference for the free choice option (blue) despite it conferring no benefit over the fixed choice option (grey).
 (B) Pessimistic agents show an attenuated free choice bias.
-(Parameters: $\gamma = 1.0$, Inverse temperature, $\beta$ increased from 1 to 10 over 100 trials.
-Simulation of 500 agents.)
+(Parameters: $\gamma = 1.0$, Inverse temperature, $\beta$, increased from 1 to 10 over 100 episodes.
+Each dot represents one of 500 simulated agents.)
 \end{figure}
 
-If the free choice bias is a function of an individual's optimistic expectations about future choice opportunities, then we predict that anxious individuals should exhibit a reduced free choice bias.
+We (and others)\citep{ly2019} suspect the free choice premium reflects the usual assumption about choice (i.e.
+an agent will continue to make reward-maximizing choices in the future).
+Under such an optimistic assumptions, additional options are valuable insofar that they allow for additional future reward opportunities (or, at the worst, pose no possible harm).
+On the contrary, if anxiety reflects a violation of this typical assumption, then we should suspect anxious individuals will exhibit a diminished free choice bias.
 Indeed, the model makes this prediction (Fig.

### 2019-06-30 manuscripts/manuscript/manuscript.tex
minor changes.

@@
 In the following, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of sequential learning and decision making.
-We will also show that our model is consistent with clinical theory describing the transition from pathological anxiety to depression.
+We will also show that our model is consistent with clinical theory describing the transition from clinical anxiety to depression.
 Unless otherwise noted, state and action values under varying degrees of pessimism were solved for using the value iteration dynamic programming method\citep{SuttonBarto2018}.
@@
 \subsection{Approach-avoidance conflict}
-
-% Fix subsection spacing?
 
@@
 (B) Pessimistic agents show an attenuated free choice bias.
-(Parameters: $\gamma = 1.0$, Inverse temperature, $\beta$, increased from 1 to 10 over 100 episodes.
+(Parameters: Q-learning with $\gamma = 1.0$, and inverse temperature, $\beta$, increased from 1 to 10 over 100 episodes.
 Each dot represents one of 500 simulated agents.)
@@
 For example, self-efficacy theory\citep{bandura1977} and the triple vulnerabilities model\citep{barlow2002} both posit that a reduced belief in the ability to effectively respond to future threat is involved in the genesis and maintenance of clinical anxiety.
-Similarly, the learned helplessness theories of anxiety\cite{alloy1990} claims pathological anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
+Similarly, the learned helplessness theories of anxiety\cite{alloy1990} claims clinical anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
 As we note above, the present model (though stated here in terms of self-efficacy) can accommodate either variant, and they are by no means mutually exclusive.
@@
 
+% not sure if this paragraph is good
+
 None of the previous discussion constitutes a claim about the algorithmic and neural bases of pessimistic sequential evaluation.
 One possibility is that these symptoms reflect aberrations in model-based planning.
-Chronic worry is a symptom frequently observed in pathological anxiety, and could be construed as a deviant form of future action planning wherein the roll-out of future actions is biased towards bad- or worst-case contingencies.
-In line with the present results, higher levels of worry is correlated with reduced perceived control, increased doubt in self-efficacy in response to threat, and exaggerated threat appraisal\cite{Berenbaum2010}.
-In sum, anxious symptoms may in part result from biased planning where cost-minimizing action is discounted and
-
-%% Need to clean the above up
+Model-based planning refers to the process of sequentially evaluation of trajectories of action, which may go awry in anxiety if successful avoidance actions are discounted or otherwise downweighted.
+Chronic worry is a symptom frequently observed in clinical anxiety, and bears resemblance to aberrant model-based planning wherein the trajectories are biased towards bad- or worst-case contingencies.
+In line with the present results, chronic worry is associated with reduced perceived control, diminished belief in self-efficacy in response to threat, and exaggerated threat appraisal\cite{Berenbaum2010}.
+This suggests that clinical anxiety may in part result from planning processes gone awry.
 
@@
 As such, the present results are more likely to accurately describe the anxiety disorders which primarily involve aberrations in future-oriented cognitive processes, such as generalized and social anxiety disorders.
-Naturally, the present model can neither account for the compulsive behaviors of obsessive-compulsive disorder nor the memory disturbances of post-traumatic stress disorder (though recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD\citep{gallagher2014a, gallagher2014b}).
+Naturally, the present model can neither account for the compulsive behaviors of obsessive-compulsive disorder nor the memory disturbances of post-traumatic stress disorder (though recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD)\citep{gallagher2014a, gallagher2014b}.
 Importantly, other psychiatric disorders that involve future-oriented misbeliefs, worry, and avoidance behaviors (e.g.

### 2019-07-02 manuscripts/manuscript/manuscript.tex
co-author edits

@@
 \begin{abstract}
-Anxiety disorders are characterized by a range of aberrations in the processing and response to threat, but there is little clarity what core pathogenesis might underlie these symptoms.
+Anxiety disorders are characterized by a range of aberrations in the processing of and response to threat, but there is little clarity what core pathogenesis might underlie these symptoms.
 Here we propose a decision theoretic analysis of maladaptive avoidance and embody it in a reinforcement learning model, which shows how a localized bias in beliefs can formally explain a range of phenomena related to anxiety.
-The core observation, implicit in standard decision theoretic accounts of sequential evaluation, is that avoidance should be protective: if danger can be avoided later, it poses no threat now.
+The core observation, implicit in standard decision theoretic accounts of sequential evaluation, is that the potential for avoidance should be protective: if danger can be avoided later, it poses no threat now.
 We show how a violation of this assumption --- a pessimistic, false belief that later avoidance will be unsuccessful --- leads to a characteristic propagation of fear and avoidance to situations far antecedent of threat.
@@
 Simulations of the model reproduce laboratory demonstrations of abnormal decision making in anxiety, including in situations of approach-avoid conflict and planning to avoid losses.
-The model also ties together a number of other seemingly disjoint issues in anxious disorders.
+The model also ties together a number of other seemingly disjoint phenomena in anxious disorders.
 For instance, learning under the pessimistic bias captures a hypothesis about the role of anxiety in the later development of depression.
 The bias itself offers a new formalization of classic insights from the psychiatric literature about the central role of maladaptive beliefs about control and self-efficacy in anxiety.
-This perspective is importantly different from previous computational accounts of beliefs about control in mood disorders, which neglected the sequential aspects of choice.
+This perspective also extends previous computational accounts of beliefs about control in mood disorders, which neglected the sequential aspects of choice.
 \end{abstract}
@@
 
-Though anxiety disorders differ in their particular symptomology, and in the content and situations which elicit anxious responding, they all are similarly characterized by aberrations in the processing of and response to threat\citep{dsm5}.
+Though anxiety disorders differ in their particular symptomology, and in the content and situations which elicit symptoms, they all are similarly characterized by aberrations in the processing of and response to threat\citep{dsm5}.
 In particular, at least three symptoms are manifest across the anxiety disorders.
@@
 Finally, anxiety is associated with persistent avoidance behavior, which often occurs well in advance of the materialization of actual threat\citep{Arnaudova2017}.
-Avoidance behaviors in anxiety are especially pernicious as they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
+Such inapposite avoidance behaviors are especially deleterious aspects of the disoders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
 Though laboratory studies of decision making and learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.
@@
 These symptoms are particularly puzzling from a decision theoretic perspective\citep{huys2015}.
-Fear and avoidance of situations far in the future violates the basic logic of evaluation over sequential trajectories of action.
-Put another way, distant threat need not necessarily impinge on decision making in the present.
+In many circumstances, distant threat should not impinge upon decision making in the present.
+Indeed, fear and avoidance of situations far in the future violates the basic logic of evaluation over sequential trajectories of action.
 This is because avoidance is by nature protective: the ability to successfully avoid danger in the future means an agent need not also do so now.
 For instance, cars endanger pedestrians but can be reliably avoided by following traffic signals; given that, staying indoors offers little or no additional protection from accidents.
-This is an instance of a general property of evaluation in sequential decision making: the value of present action turns fundamentally on assumptions about subsequent events, which include the agent's own subsequent choices.
+This is an instance of a fundamental property of evaluation in sequential decision making: the value of present action turns fundamentally on assumptions about subsequent events, which include the agent's own subsequent choices.
 Typically, it is appropriate to assume that an agent will continue make good (i.e.
-reward-maximizing/harm-minimizing) choices down the line.
+reward-maximizing/harm-minimizing) choices down the line, and that choices at the current stage should reflect this.
 
-This line of reasoning hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of threat and excessive avoidance.
+This line of reasoning hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of threat to antecedent situations, and excessive avoidance.
 Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
@@
 
-Here, we develop this idea into a model of evaluation under pessimistic assumptions about future choices.
+Here, we develop this idea into a formal model of evaluation under pessimistic assumptions about future choices.
 We show that a single, localized deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance.
 This account also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety \citep{bandura1977, barlow2002}.
-Specifically, we demonstrate a model with a misbelief in the reliability of future self-action gives rise to the anxious symptoms.
-Our perspective is importantly different from previous computational accounts of beliefs about control in mood disorders (e.g.
+Specifically, we demonstrate a model with a misbelief about the reliability of future self-action gives rise to the anxious symptoms.
+In this way, our perspective also extends previous computational accounts of beliefs about control in mood disorders (e.g.
 \cite{HuysDayan2009}), which neglected the sequential aspects of choice.
@@
 
-We model anxious decision making in the context of Markov decision processes (MDPs). A standard normative assumption is that agents attempt to optimize the expected cumulative discounted reward,
+We model anxious decision making in the context of Markov decision processes (MDPs). A standard normative assumption is that agents attempt to optimize the expected cumulative discounted reward:
 
 \begin{equation*}
-Q^\pi(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \sum_{a'} \pi(a' \mid s') Q^\pi(s',a'),
+Q^\pi(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \sum_{a'} \pi(a' \mid s') Q^\pi(s',a').
 \end{equation*}
 
-but for any particular state-action $a$, this is necessarily defined only relative to a policy $\pi(a' \mid s')$ specifying the assumed distribution of \emph{future} choices.
+For any particular state-action $(s,a)$, this is necessarily defined only relative to a policy $\pi(a' \mid s')$ specifying the assumed distribution of \emph{future} choices.
 The return can be optimized self-consistently under the assumption that the agent makes the return-maximizing choice at each step in the future, leading to the familiar expression for the optimal values,
@@
 The assumption that the agent makes the return-maximizing choice at each step implies that opportunities for reward propagate recursively to earlier steps, but avoidable dangers do not.
-This principle is highlighted in a toy MDP: a deterministic open gridworld with terminal states containing a single reward and punishment (Fig.~1a).
+(To the extent obtaining reward or avoiding harm are only imperfectly attainable, this propagation and attenuation at each step are graded according to success probability, but the basic asymmetry remains.) This principle is highlighted in a toy MDP: a deterministic open gridworld with two terminal states, a rewarding state and a punishing one (Fig.~1a).
 The optimal state values $V^* = \max_a Q^*(s,a)$ (Fig.~1b) reflect a ``mountain'' of opportunity propagating recursively from the reward.
@@
 
-An agent is not restricted to the return-maximizing assumption, and could anticipate encountering danger under different assumptions about the future.
+An agent is not restricted to the return-maximizing assumption, and could anticipate encountering danger under different assumptions about the future, including of course incorrect or pathological beliefs.
 For example, an agent may expect to fail to take the correct protective actions (i.e., expects to use a nonoptimal $\pi(a \mid s)$), or may expect the world's future dynamics do not guarantee reliable avoidance even so (i.e., under stochastic or adversarial transition dynamics $P(s' \mid s,a)$).
 Consider an agent who has such pessimistic expectations about dangerous events at future steps.
-Computing returns with such pessimistic predictions might help to quantify uncertainty in outcomes (i.e.
-to learn different points in the distribution of possible returns)\cite{bellemare2017}, and choosing on this basis can be adaptive in that it increases the chances of safety in uncertain or even adversarial scenarios\cite{Garcia2015}.
+Note that assumptions of this sort, even if incorrect, can serve adaptive purposes.
+It is common in machine learning theory, for instance, to optimize outcomes under worst-case assumptions.
+In general, this can help to ensure robustness and safety under uncertain or even adversarial scenarios \cite{Garcia2015}; related work in RL shows how computing returns with pessimistic predictions can help to quantify variability in outcomes (i.e.
+to learn different points in the distribution of possible returns)\cite{bellemare2017}, and explicitly take account of risk tolerance.
 
 Here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms.
-Although such pessimism can be encoded either in the policy, $P(s' \mid s,a)$, or transition probabilities, $\pi(a \mid s)$, here for concreteness we work with distortions in the policy.
-In particular, we adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define an expectation of state-action value,
+Although such pessimism can be encoded either in the policy, $P(s' \mid s,a)$, or transition probabilities, $\pi(a \mid s)$ (i.e.
+misbeliefs about either one's own avoidance actions or the environment's responses to them, a point we return to in discussion), here for concreteness we work with distortions in the policy.
+In particular, we adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define state-action value in expectation over a mixture of the best and worst action:
 
-\begin{equation*}
+\begin{equation}\label{eq:pessimism}
 Q^w(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \left( w \max_{a'} Q^w(s',a') + (1 - w) \min_{a'} Q^w(s',a') \right)
-\end{equation*}
+\end{equation}
 
 The parameter $w$ controls the degree of pessimism.
-An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences.
-(We return to pessimism about state transitions in the discussion.) åFig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in the example gridworld.
-An increasing expectation that the agent may fail to avoid in future causes the value of threat to permeate with increasing distance across the state space.
+An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences, whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences.
+Fig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in the example gridworld.
+An increasing expectation that the agent may fail to avoid in future causes the value of threat to propagate with increasing distance across the state space.
 
@@
 (C) For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states.
-(D) With increasing pessimism ($w=0$), the extent of the spread grows worse.
+(D) With increasing pessimism ($w=0$), the extent of the spread grows worse, and the return-optimizing trajectory becomes more distorted and avoidant.
 (Parameters: $\gamma = 0.95$)
@@
 The moderate ($w=0.6$) and strongly ($w=0.2$) pessimistic agents cash-out earlier, as is observed in anxious individuals.
-(D/E) By contrast, in the sleeping predator task the risk of loss is constant but the cost of loss increases.
-The value of reward pursuit under increasingly pessimistic valuation is presented for low (D) and high risk (E) scenarios.
+(D/E) In the sleeping predator task, the risk of loss is constant but the cost of loss still increases as more rewards are gathered.
+The value of reward pursuit under increasingly pessimistic valuation is presented for scenarios with low (D) and high risk (E)  of predator awakening.
+The relative value of approach (vs.
+avoid) decreases with loss amount and threat level, and moreso under pessimistic assumptions.
 (Parameters: $\gamma = 1.0$)
@@
 
-The model can also explain findings of increased behavioral inhibition, measured as prolonged response times, in anxious individuals under threat\cite{bach2015}.
+The model can also capture findings of increased behavioral inhibition, measured as prolonged response times, in anxious individuals under threat\cite{bach2015}.
 In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator".
-Though the risk of predation is constant throughout a trial, the potential loss from capture increases with each token collected.
-Bach \cite{bach2015} finds that participants are slower to collect tokens as loss increases, and that this effect is modulated by subclinical anxiety.
-Our model recreates this phenomenon (Fig.~2d/e).
-Insofar that value and response time are negatively related \citep{need citation?}, and the relative value of approach as compared to avoidance is reduced under pessimistic assumptions about future action, then prolonged response times naturally follow.
+Though in this variant the risk of predation is constant throughout a trial, the potential loss from capture still increases with each token collected.
+Bach \cite{bach2015} finds that participants are slower to collect tokens as loss increases, and that this effect is enhanced by subclinical anxiety.
+We can capture this effect in our model by noting that the relative value of approach compared to avoidance is reduced as the potential loss, and the risk the predator awakes, increases (Fig.~2d/e).
+These effects are amplified under more pessimistic (i.e., on our view, anxious) assumptions about future action.
+Thus, as before, anxiety in our model produces greater and earlier choice of avoidance (analogous to earlier cash-out in the BART).
+To extend this effect to reaction times, we adopt the further, standard assumption that actions (here, approach) are slower when their relative value compared to alternatives is lower \citep{oud2016}.
+In this case, the model captures behavioral inhibition (slower responses as threat increases), and its enhancement by anxiety, as measured by Bach.
+(Note that we need not assume the coupling of reaction times to action value spread is due to difficulty in decision formation per se, which Bach argues against: it may, for instance, reflect Pavlovian initiation biases \cite{niv2007}.)
 
@@
 Our model predicts this result (Fig.~3b), though for a somewhat different reason.
-In our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them, but because with increasing anxiety they increasingly expect the potential of choosing incorrectly afterward, thus failing to recoup the loss.
+In our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them (here we assume full evaluation of the Bellman equation), but because with increasing anxiety they increasingly expect the potential of choosing incorrectly afterward, thus failing to recoup the loss.
 This reinterpretation is consistent with the structure of the original task\citep{Huys2012, Lally2017}, where participants were required to plan in advance action sequences of two- to eight-steps (requiring the consideration of upwards of $2^8 = 256$ options).
@@
 % Include the larger, four-panel aversive pruning figure instead?
+%%% I think this figure is OK but shouldnt it include optimal Q values for paths on the left also, showing why the best path is favored?
+%%% also the graying out of the unchosen options on the right is maybe a little strong, hard to read, maybe darken these a bit (then you can also gray out 
+%%% suboptimal paths on the left to make them really parallel?)
 
@@
 
-So far, we have considered the asymptotic preferences implied the pessimistic value function (solved for through value iteration).
-But we can also consider learning under this value function (e.g., by Q-learning\citep{SuttonBarto2018} or DYNA\citep{sutton1991} using the $\beta$-pessimistic return), the dynamics of which may speak to the progression of symptoms.
+So far, we have considered the asymptotic preferences implied the pessimistic value function (solved for directly by using value iteration to compute the return).
+But we can also consider learning under this value function (e.g., by variants of Q-learning\citep{SuttonBarto2018} or DYNA\citep{sutton1991} using the $\beta$-pessimistic return), the dynamics of which may speak to the progression of symptoms.
 
@@
 Frustratingly, these biased estimates of value may still remain even after a misbelief in the efficacy of future action is corrected for in a course of therapy.
-This may offer at least a partial answer to a classic puzzle in pathological avoidance, i.e.
+This phenomenon (similar to failures of model-free RL algorithms to adjust to reward revaluation without extensive relearning; \cite{daw2005}) may offer at least a partial answer to a classic puzzle in pathological avoidance, i.e.
 why it is so resistant to extinction \cite{moutoussis2018}, and to the unfortunately high rates of anxiety recurrence following treatment\citep{pittig2018}.
@@
   \par \textbf{Figure 4:} A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
-(B-D) States colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
-For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value with learning.
+(B-D, E-G) The development of value expectancies over three steps of learning, for two levels of pessimism.
+  States are colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
+(B-D) For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value with learning.
 (E-G) For a pessimistic agent ($w=0.6$), negative value spreads from the source to antecedent states.
@@
 
-In this final section, we offer a novel prediction regarding the free choice premium.
-Simply put, the free choice premium phenomenon predicts that, all else being equal, options which lead to more choice opportunities in the future are more valuable than those that lead to fewer future choice opportunities.
+Finally, the model also offers a novel prediction tying anxious beliefs to a classic, but hitherto separate, phenomenon known as the free choice premium.
+This refers to the finding that, all else being equal, people tend to treat choice as itself valuable: i.e., choices which lead to more choice opportunities in the future are preferred to those that lead to fewer future choice opportunities.
 A free choice premium has been observed in multiple behavioral experiments \cite{Leotti2010, ly2019}.
@@
 5a.
-In the task, participants repeated choose between a free choice option, allowing for an additional future choice, and a fixed choice option; importantly, both choices lead to identical, stochastic outcomes (e.g.
+In the task, participants repeatedly choose between a free choice option, allowing for an additional future choice, and a fixed choice option; importantly, both choices lead to identical, stochastic outcomes (e.g.
 50/50 chance of [1, -1]).
-Previous studies have found a preference for the free choice option despite it conferring no additional benefits relative to the fixed choice option.
+Previous studies have found that human subjects (from a general, healthy population) prefer the free choice option despite it conferring no additional benefits relative to the fixed choice option.
 
@@
 (B) Pessimistic agents show an attenuated free choice bias.
-(Parameters: Q-learning with $\gamma = 1.0$, and inverse temperature, $\beta$, increased from 1 to 10 over 100 episodes.
-Each dot represents one of 500 simulated agents.)
+The fractional preference for the free choice option over the simulated experiment is shown for three populations of subjects with levels of pessimism $w$ (y-axis; each dot represents an individual simulated agent, and the smoothed density of free choice bias for each pessimism level is also shown.)
+  (Parameters: Q-learning with $\gamma = 1.0$, and inverse temperature, $\beta$, increased from 1 to 10 over 100 episodes.)
 \end{figure}
 
-We (and others)\citep{ly2019} suspect the free choice premium reflects the usual assumption about choice (i.e.
-an agent will continue to make reward-maximizing choices in the future).
-Under such an optimistic assumptions, additional options are valuable insofar that they allow for additional future reward opportunities (or, at the worst, pose no possible harm).
-On the contrary, if anxiety reflects a violation of this typical assumption, then we should suspect anxious individuals will exhibit a diminished free choice bias.
-Indeed, the model makes this prediction (Fig.
-~5b).
-We plan on explicitly testing this hypothesis in subclinical and clinical anxiety in the near future.
+On one account (e.g., \citep{ly2019}) this free choice preference directly and specifically reflects the assumption about sequential choice whose violation we argue is core to anxiety: i.e.
+that the agent will make reward-maximizing choices in the future.
+Under such an optimistic assumption (and given noisy and imperfect knowledge about the value of each option, due to learning), additional options are valuable in the sense that free choice can be expected to exploit the best among them: the maximum over several noisy values is, in general, larger than a single option from the same distribution.
+Our theory makes the novel prediction that, if as we hypothesize anxiety reflects a violation of this optimistic assumption, then anxious individuals will exhibit a diminished or reversed free choice bias, as shown in simulation in Fig.
+~5b.
+Future research will be required to test this prediction.
 
@@
 We have presented a simple computational account suggesting how a single underlying misbelief can give rise to these aberrations in learning and choice.
-Specifically, we showed how a failure to believe in the reliability of future self-action can effectively backpropagate negative value across states of the environment.
+Specifically, we showed how a failure to believe in the reliability of one's future actions can effectively backpropagate negative value across states of the environment.
 This process results in a range of inferences and behaviors resembling those observed in clinical anxiety.
@@
 
-One important ambiguity is that we have formalized pessimistic assumptions in terms of beliefs about the agent's own future actions: expecting failure in handling or avoiding future threat.
+One important ambiguity is that, for concreteness we have formalized pessimistic assumptions in terms of beliefs about the agent's own future actions: expecting failure in handling or avoiding future threat.
 Since Eq.~1 is also computed in expectation over the anticipated future environmental dynamics $p(s' \mid s,a)$, pessimism can alternatively be encoded in this distribution (e.g., a belief that the world's response to one's choices is unpredictable or adversarial).
-Because the Bellman equation averages out both distributions, and because an unpredictable environment also reduces the efficacy of avoidance, either formulation can produce ultimately similar results.
+Because the Bellman equation for the return averages over both distributions at each step, and because an unpredictable environment also reduces the efficacy of avoidance, either formulation can produce ultimately similar results in our simulations here.
+Other sorts of uncertainty (e.g.
+about the state of the environment, discussed below) can also produce similar effects for analogous reasons.
 However from the perspective of cognitive theories of anxiety, these represent quite different maladaptive beliefs, which may be relevant, for instance, in guiding treatment.
@@
 For example, self-efficacy theory\citep{bandura1977} and the triple vulnerabilities model\citep{barlow2002} both posit that a reduced belief in the ability to effectively respond to future threat is involved in the genesis and maintenance of clinical anxiety.
-Similarly, the learned helplessness theories of anxiety\cite{alloy1990} claims clinical anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
-As we note above, the present model (though stated here in terms of self-efficacy) can accommodate either variant, and they are by no means mutually exclusive.
+Analogously, but focused less on the self, the learned helplessness theories of anxiety\cite{alloy1990} claims clinical anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
+As we note above, the present model and analysis (though stated here in terms of self-efficacy) can accommodate either variant.
+Also, they are by no means mutually exclusive, though they might also reflect multiple distinct (and differently treatable) routes to similar symptomatology.
 
-Distortions in beliefs about future action or transition probabilities are in no way the only means through which anxious behaviors may arise in a decision theoretic account.
-Traditionally, the goal of reinforcement learning algorithms is to a find reward-maximizing policy with respect to the expectation (average) of returns.
-However, in environments where the distribution of returns has high variance, such that there is potential for catastrophic loss, it may be instead preferable to learn a policy with respect to an alternative value function.
-The same is true when there is uncertainty in the transition probabilities, such that an agent may inadvertently step towards undesirable and/or dangerous segments of the environment.
-How to learn risk-sensitive and robust policies is the focus of much recent research \cite{morimura2012, chow2015, bellemare2017}; interestingly, these studies have reported agent behaviors not unlike those presented here.
-It is important to note that anxiety is correlated with disproportionate worry about low chance but high cost events\cite{Miceli2005}.
-One speculative hypothesis is that anxious individuals learn and evaluate action with respect  to a more pessimistic point in the distribution of expected returns.
-
-A separate possibility concerns the notion of state.
-In the present work, we considered only environments with fully observable states; in other words, there was no uncertainty regard the state an agent was currently occupying.
-In the real-world, however, the state must be inferred from external and internal observations.
-One possibility is that anxiety is a function of biases in state inference\cite{Paulus2012}.
-For example, if an agent erroneously comes to the conclusion that she is more proximal to threat than is actually the case, then symptoms like exaggerated threat appraisal and avoidance should arise.
+Indeed, yet a third variant of our idea concerns the notion of state.
+In the present work, we considered only environments with fully observable states; in other words, there was no uncertainty regarding the state (e.g.
+grid world location) an agent currently occupied.
+In the real world, however, the state must be inferred from external and internal observations.
+This provides yet another source of uncertainty similar to the future policy and transition dynamics, which must be averaged over in evaluating actions \citep{kaelbling1998}.
+Thus another variant of our approach is to view anxiety as potentially also arising due to pessimistic biases in state inference\cite{Paulus2012}, which would again have similar effects on the resulting value function.
+For example, if an agent is erroneously biased toward believing herself more proximal to threat than is actually the case, then symptoms like exaggerated threat appraisal and avoidance should arise.
 Indeed, a bias in value functions could result in the diffuse spread of threat across objects\citep{norbury2018} and location\citep{schulz2018}, thereby causing otherwise benign states to be perceived as threatening.
 
-% not sure if this paragraph is good
+Speculatively, all these distorted beliefs may reflect pathological cases of an otherwise adaptive evaluation strategy.
+Traditionally, the goal of reinforcement learning algorithms is to a find reward-maximizing policy with respect to the expectation (average) of returns.
+However, depending on one's risk attitude and state of uncertainty (e.g.
+if there is potential for catastrophic loss), it may be instead preferable to learn a policy with respect to an alternative and more pessimistic objective function, similar to the one considered here.
+How to learn risk-sensitive and robust policies is the focus of much recent research \cite{morimura2012, chow2015, bellemare2017}; interestingly, different algorithmic variants in this literature produce agent behaviors not unlike those presented here.
+Thus another perspective on our hypothesis is that anxious individuals, in effect, evaluate actions with respect to a more pessimistic point in the distribution of expected returns.
+Suggestively, anxiety is correlated with disproportionate worry about low chance but high cost events\cite{Miceli2005}.
 
-None of the previous discussion constitutes a claim about the algorithmic and neural bases of pessimistic sequential evaluation.
-One possibility is that these symptoms reflect aberrations in model-based planning.
-Model-based planning refers to the process of sequentially evaluation of trajectories of action, which may go awry in anxiety if successful avoidance actions are discounted or otherwise downweighted.
-Chronic worry is a symptom frequently observed in clinical anxiety, and bears resemblance to aberrant model-based planning wherein the trajectories are biased towards bad- or worst-case contingencies.
-In line with the present results, chronic worry is associated with reduced perceived control, diminished belief in self-efficacy in response to threat, and exaggerated threat appraisal\cite{Berenbaum2010}.
+We have centered our discussion at Marr's \cite{marr1982} computational level: on beliefs and their consequences in terms of action values, while remaining agnostic as to how, algorithmically or mechanistically, these misbeliefs are evaluated.
+Of course, the brain is believed to contain multiple distinct mechanisms for evaluating actions (e.g, model-based and model-free; learning; \cite{daw2005, huys2015}), and pessimisitic beliefs might play out either differentially or similarly through each of these mechanisms.
+One particular possibility is that these symptoms mainly reflect aberrations in model-based planning \cite{huys2015}, i.e.
+explicitly evaluating valuations by mentally simulating possible trajectories.
+Recent work suggests this process may be accomplished by mentally ``replaying'' potential trajectories successively (e.g.
+in hippocampus, \cite{mattar2018}).
+In this setting the biases we suggest would amount to over-contemplating, or over-weighting, certain pessimistic trajectories \cite{hunter2019}.
+This, in turn, may relate to worry and rumination; indeed, in line with the present results, chronic worry is associated with reduced perceived control, diminished belief in self-efficacy in response to threat, and exaggerated threat appraisal\cite{Berenbaum2010}.
 This suggests that clinical anxiety may in part result from planning processes gone awry.
-
-A second possibility is that pessimistic evaluation may arise as a result of biased offline replay of events.
-During periods of rest and sleep, the hippocampus replays recent experiences\citep{whats a good ref?}.
-A recent theoretical model of replay suggests that these replay events are able to iteratively backpropagate value from a focal point to antecedent states, and that episodes are prioritized for replay according to their likelihood in changing behavior.
-Biases in replay behavior, or in the mechanisms initiate replay, might result in repeated offline reinforcement of negative episodes thereby exaggerating the value of threat and diffusing its influence across an agent's internal cognitive map.
-Few studies have investigated replay behavior in aversive environments (e.g.
-\citep{wu2017}), so this hypothesis remains speculative at this point in time.
-
-%% Want to add anything about transition probability bias?
 
@@
 As such, the present results are more likely to accurately describe the anxiety disorders which primarily involve aberrations in future-oriented cognitive processes, such as generalized and social anxiety disorders.
-Naturally, the present model can neither account for the compulsive behaviors of obsessive-compulsive disorder nor the memory disturbances of post-traumatic stress disorder (though recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD)\citep{gallagher2014a, gallagher2014b}.
+Naturally, the present model can account neither for the compulsive behaviors of obsessive-compulsive disorder nor the memory disturbances of post-traumatic stress disorder (though recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD)\citep{gallagher2014a, gallagher2014b}.
 Importantly, other psychiatric disorders that involve future-oriented misbeliefs, worry, and avoidance behaviors (e.g.
@@
 
+\section{Methods}
+
+We model anxious decision making in the context of Markov decision processes (MDPs).
+Tasks were modeled as deterministic, infinite horizon, discrete time environments.
+Some (detailed below) were modeled with discounted returns $\gamma < 1$.
+All simulations were implemented in the Python programming language and the code is publicly available at \url{https://github.com/ndawlab/seqanx}.
+
+For all but the free choice premium task, we defined state-action values, $Q(s,a)$, in accordance with our modified, pessimistic Bellman equation Eq.~\ref{eq:pessimism}.
+Here, $Q$-values were solved for directly through value iteration.
+\citep{SuttonBarto2018} By contrast, $Q$-values in the free choice task were computed using the $\beta$-pessimism temporal-difference learning algorithm \citep{Gaskett2003}, where the reward prediction error is defined as:
+
+\begin{equation*}
+\delta^w = r(s,a) + \gamma \left( w \max_{a'} Q^w(s',a') + (1 - w) \min_{a'} Q^w(s',a') \right) - Q^w(s,a), 
+\end{equation*}
+
+and the update rule is defined as:
+
+\begin{equation*}
+Q^w(s,a) \leftarrow Q^w(s,a) + \eta \cdot \delta^w
+\end{equation*}
+
+where $\eta$ is a learning rate. The parameterizations of the MDP environments and learning algorithms are next specified in turn.
+
+\textbf{Toy MDP / Anxiety-to-Depression Transition} Both the Toy MDP and anxiety-to-depression transition simulations were performed using simple gridworlds.
+Both environments involved only two non-zero states, one rewarding ($r=10$) and one aversive ($r=-10$).
+In both environments, we solved for the discounted, asymptomatic $Q$-values using value iteration with $\gamma = 0.95$.
+In the toy gridworld, we performed simulations under pessimistic assumptions, $w \in [1.0, 0.5, 0.0]$.
+In the transition gridworld, we performed simulations under pessimistic assumptions, $w \in [1.0, 0.6]$.
+To highlight the effects of learned, we took "snapshots" of $Q$-values prior to asymtote, three and five steps into value iteration.
+
+\textbf{Balloon Analog Risk Task / Predator Avoidance Task} The balloon analog risk task \citep{Lejuez2002} has participants inflate a virtual balloon for points.
+Earnings rise with each pump, but so too does the risk of the balloon popping and subsequent point loss.
+Unbeknownst to participants, the number of pumps before balloon pop is predefined and drawn randomly from some distribution (e.g.
+uniform, normal, exponential), where the mean controls the risk (i.e.
+average number of pumps before point loss).
+
+Here, we modeled the BART as an undiscounted MDP with 20 states, where transitioning to each successive state yielded $r=1$.
+The only available actions were to transition to the next state (e.g.
+$S_1 \rightarrow S_2, S_2 \rightarrow S_3$) or end the episode (i.e.
+cash-out).
+With each act to move to the next state, there was some probability of transitioning to a bad terminal state (i.e.
+balloon pop) with reward equal to the negative equivalent of accumulated gain thus far.
+The probability of this bad transition was modeled using normal density function, with parameters $\mathcal{N}(16,0.5)$ for low risk and $\mathcal{N}(8,0.5)$ high risk.
+The asymptomatic $Q$-values were solved for using value iteration for both the low and high risk conditions under pessimistic assumptions, $w \in [1.0, 0.6, 0.2]$.
+
+The predator avoidance task \citep{Mobbs2019} can be analogously modeled.
+There, a virtual predator approaches participants over discrete time steps while participants "forage".
+For every time step the participant remains (i.e.
+does not flee), points are accumulated; if the participant is caught, however, all earnings were lost and an additional penalty was received.
+Thus, the predator avoidance task bears striking resemblance to the BART; foraging is equivalent to virtual pumps, and fleeing is equivalent to cashing-out.
+
+\textbf{Behavioral Inhibition Task} In the behavioral inhibition task \citep{bach2015} (or sleeping predator task), participants collect virtual tokens while evading capture from a virtual predator.
+Unlike the BART, the risk in the behavioral inhibition task (i.e.
+predator "waking up") is constant.
+However, the cost of capture increases as sequential tokens are collected.
+
+True to the original, we model the task as an undiscounted MDP with 6 states where transitioning to each successive state yielded $r=1$.
+Identical to the BART, the only available actions were to transition to the next state or end the episode (i.e.
+avoid predator).
+With each act to move to the next state, there was a constant probability of transitioning to a bad terminal state (i.e.
+capture by predator) with reward equal to the negative equivalent of accumulated gain thus far.
+The probability of bad transition was defined as $p = 0.10$ for low risk and $p = 0.15$ high risk, based on the objective risk probabilities in \citep{bach2015}.
+The asymptomatic $Q$-values were solved for using value iteration for both the low and high risk conditions under pessimistic assumptions, $w \in [1.0, 0.6, 0.2]$.
+
+\textbf{Averse Pruning} In the aversive pruning task \citep{Huys2012, Lally2017}, participants learn to navigate a six-state graphworld, where each state is directly connected to only two other states.
+Each state is associated with some reward (cost), and participants must plan trajectories through the state-space so as to maximize reward (minimize cost).
+
+We model the task as an undiscounted MDP where the original 6-node network has been "unravelled" into a 15-state decision tree.
+This is equivalent to having the participant start in one state and planning to make three actions.
+The rewards associated with transitioning to each state were taken directly from \citep{Huys2012, Lally2017}.
+The asymptomatic $Q$-values were solved for using value iteration for both the low and high risk conditions under pessimistic assumptions, $w \in [1.0, 0.75, 0.5]$.
+
+\textbf{Free Choice Task} In the free choice task \citep{Leotti2011}, participants complete a series of two-stage trials.
+In the first stage, they select between a free choice option, allowing them to make an additional choice in the second stage, and a fixed choice, permitting no choice in the second stage.
+In the second stage, participants select between one of two bandits (free choice) or are randomly assigned a bandit (fixed choice).
+Importantly, all bandits pay out under an identical reward distribution.
+
+We model the task as an undiscounted MDP with 6 state decision tree structure.
+In the free choice branch, agents are able to select between two terminal bandits; in the fixed choice branch, agents can only choose one terminal bandit.
+All bandits pay out identically, in this case randomly in the set, $r \in [-1, 1]$.
+Learned $Q$-values were computed using pessimistic temporal difference learning algorithm, with learning rate $\eta = 0.1$.
+Each simulated agent learned the values of each action over 100 trials, with an increasing, logarithmically spaced inverse temperature in the range of $\beta \in [0, 1]$.
+(Inverse temperature was gradually increased over learning to facilitate exploration of choice options.) The resulting $Q$-values were stored for 500 simulated agents, run separately for pessimistic assumption, $w \in [1.0, 0.5, 0.0]$.
+
 \bibliographystyle{naturemag}

### 2019-07-02 manuscripts/manuscript/manuscript.tex
Updates from Overleaf

@@
 This is because it anticipates and avoids future errors in choice which would otherwise result in the balloon popping.
-Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety\cite{Mobbs2019}.
+Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety\cite{fung2019}.
 One unique prediction of the model is that this bias should arise only when beliefs about \emph{future} avoidance are involved, rather than direct conflict between immediate impulses.
-Recent data \cite{Mobbs2019} using a predator avoidance task, analogous to the BART, support this view.
+Recent data \cite{fung2019} using a predator avoidance task, analogous to the BART, support this view.
 In their task, trait anxiety predicted earlier escape (analogous to cash-out in the BART) for slow predators (for whom future decisions to escape were a relevant consideration) but not fast ones (who would attack immediately, negating consideration of future steps).
@@
 
-The predator avoidance task \citep{Mobbs2019} can be analogously modeled.
+The predator avoidance task \citep{fung2019} can be analogously modeled.
 There, a virtual predator approaches participants over discrete time steps while participants "forage".

### 2019-07-02 sisyphus/envs/_tree.py
fixed decision tree figure

@@ -103,8 +103,8 @@ class DecisionTree(GraphWorld):
         """Draws one-step reward values as node labels."""
 
         ## Define node positions.
-        xpos = [-1,1,-1.5,-0.5,0.5,1.5,-1.75,-1.25,-0.75,-0.25,0.25,0.75,1.25,1.75]
-        ypos = [2, 2, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0]
+        xpos = [-1,1,-1.5,-0.5,0.5,1.5,-1.75,-1.25,-0.75,-0.25,0.25,0.765,1.25,1.77]
+        ypos = [1.98, 1.98, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0]
         
         ## Define node values.
         labels = [-70, -20, -20, -70, -20,-70, -20, 20, 20, 140, -20, 20, -20, 20]

### 2019-07-02 manuscripts/manuscript/manuscript.tex
Updates from Overleaf

@@
 \usepackage{amsfonts, amsmath}
-\usepackage{algorithm, algpseudocode}%
 \usepackage[numbers,super]{natbib}
@@
 
-Though anxiety disorders differ in their particular symptomology, and in the content and situations which elicit symptoms, they all are similarly characterized by aberrations in the processing of and response to threat\citep{dsm5}.
+Though anxiety disorders differ in their particular symptomology, and in the content and situations which elicit symptoms, they all are similarly characterized by aberrations in the processing of and response to threat \citep{dsm5}.
 In particular, at least three symptoms are manifest across the anxiety disorders.
-First, anxiety is associated with exaggerated threat appraisal, or a bias towards evaluating threat as disproportionately greater in likelihood and severity than is warranted\citep{ClarkBeck2011}.
-Second, anxiety is also associated with fear generalization, wherein the primary threat becomes associated with increasingly distal locations, events, and thoughts\citep{dymond2015}.
-Finally, anxiety is associated with persistent avoidance behavior, which often occurs well in advance of the materialization of actual threat\citep{Arnaudova2017}.
+First, anxiety is associated with exaggerated threat appraisal, or a bias towards evaluating threat as disproportionately greater in likelihood and severity than is warranted \citep{ClarkBeck2011}.
+Second, anxiety is also associated with fear generalization, wherein the primary threat becomes associated with increasingly distal locations, events, and thoughts \citep{dymond2015}.
+Finally, anxiety is associated with persistent avoidance behavior, which often occurs well in advance of the materialization of actual threat \citep{Arnaudova2017}.
 Such inapposite avoidance behaviors are especially deleterious aspects of the disoders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
@@
 Central to many prominent cognitive theories of anxiety in the psychiatric literature is a perceived lack of control.
-For example, self-efficacy theory\citep{bandura1977} and the triple vulnerabilities model\citep{barlow2002} both posit that a reduced belief in the ability to effectively respond to future threat is involved in the genesis and maintenance of clinical anxiety.
-Analogously, but focused less on the self, the learned helplessness theories of anxiety\cite{alloy1990} claims clinical anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
+For example, self-efficacy theory \citep{bandura1977} and the triple vulnerabilities model \citep{barlow2002} both posit that a reduced belief in the ability to effectively respond to future threat is involved in the genesis and maintenance of clinical anxiety.
+Analogously, but focused less on the self, the learned helplessness theories of anxiety \cite{alloy1990} claims clinical anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
 As we note above, the present model and analysis (though stated here in terms of self-efficacy) can accommodate either variant.
@@
 
-For all but the free choice premium task, we defined state-action values, $Q(s,a)$, in accordance with our modified, pessimistic Bellman equation Eq.~\ref{eq:pessimism}.
+For all but the free choice premium task, we defined state-action values, $Q(s,a)$, in accordance with our modified, pessimistic Bellman equation Eq.~\ref{eq:pessimism}, reproduced below for convenience:
+
+\begin{equation*}
+Q^w(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \left( w \max_{a'} Q^w(s',a') + (1 - w) \min_{a'} Q^w(s',a') \right)
+\end{equation*}
+
 Here, $Q$-values were solved for directly through value iteration.
@@
 The rewards associated with transitioning to each state were taken directly from \citep{Huys2012, Lally2017}.
-The asymptomatic $Q$-values were solved for using value iteration for both the low and high risk conditions under pessimistic assumptions, $w \in [1.0, 0.75, 0.5]$.
+The asymptomatic $Q$-values were solved for using value iteration for both the low and high risk conditions under pessimistic assumptions, $w \in [1.0, 0.5]$.
 

### 2019-08-01 manuscripts/manuscript/manuscript.tex
Updates from Overleaf

@@
 Ida Momennejad \\
+Department of Biomedical Engineering\\
 Columbia University\\
@@
 Anxiety disorders are characterized by a range of aberrations in the processing of and response to threat, but there is little clarity what core pathogenesis might underlie these symptoms.
-Here we propose a decision theoretic analysis of maladaptive avoidance and embody it in a reinforcement learning model, which shows how a localized bias in beliefs can formally explain a range of phenomena related to anxiety.
-The core observation, implicit in standard decision theoretic accounts of sequential evaluation, is that the potential for avoidance should be protective: if danger can be avoided later, it poses no threat now.
-We show how a violation of this assumption --- a pessimistic, false belief that later avoidance will be unsuccessful --- leads to a characteristic propagation of fear and avoidance to situations far antecedent of threat.
-This single deviation can explain a surprising range of features of anxious behavior, including exaggerated threat appraisals, fear generalization, and persistent avoidance.
+Here we propose that a particular set of unrealistically pessimistic assumptions can distort an agent's behavior and underlie a host of seemingly disparate anxiety symptoms.
+We formalize this hypothesis in a decision theoretic analysis of maladaptive avoidance and a reinforcement learning model, which shows how a localized bias in beliefs can formally explain a range of phenomena related to anxiety.
+The core observation, implicit in standard decision theoretic accounts of sequential evaluation, is that the potential for avoidance should be protective: if danger can be avoided later, it poses less threat now.
+We show how a violation of this assumption --- via a pessimistic, false belief that later avoidance will be unsuccessful --- leads to a characteristic, excessive propagation of fear and avoidance to situations far antecedent of threat.
+This single deviation can explain a range of features of anxious behavior, including exaggerated threat appraisals, fear generalization, and persistent avoidance.
 Simulations of the model reproduce laboratory demonstrations of abnormal decision making in anxiety, including in situations of approach-avoid conflict and planning to avoid losses.
@@
 
-\acknowledgements{This research was supported in part by NIH grant MH10917, part of the CRCNS program.}
+\acknowledgements{This research was supported in part by NIH grant MH10917, part of the CRCNS program, and by an NSF Graduate Research Fellowship.}
 
@@
 
-Though anxiety disorders differ in their particular symptomology, and in the content and situations which elicit symptoms, they all are similarly characterized by aberrations in the processing of and response to threat \citep{dsm5}.
-In particular, at least three symptoms are manifest across the anxiety disorders.
+Though anxiety disorders differ in their particular symptomology, and in the content and situations that elicit symptoms, they all are similarly characterized by aberrations in the processing of and response to threat \citep{dsm5}.
+In particular, at least three symptoms manifest across the anxiety disorders.
 First, anxiety is associated with exaggerated threat appraisal, or a bias towards evaluating threat as disproportionately greater in likelihood and severity than is warranted \citep{ClarkBeck2011}.
@@
 Finally, anxiety is associated with persistent avoidance behavior, which often occurs well in advance of the materialization of actual threat \citep{Arnaudova2017}.
-Such inapposite avoidance behaviors are especially deleterious aspects of the disoders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
+Such inappropriate avoidance behaviors are especially harmful aspects of the disorders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
 Though laboratory studies of decision making and learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.
 
-These symptoms are particularly puzzling from a decision theoretic perspective\citep{huys2015}.
+These symptoms are particularly puzzling from a decision theoretic perspective \citep{huys2015}.
 In many circumstances, distant threat should not impinge upon decision making in the present.
-Indeed, fear and avoidance of situations far in the future violates the basic logic of evaluation over sequential trajectories of action.
-This is because avoidance is by nature protective: the ability to successfully avoid danger in the future means an agent need not also do so now.
+Indeed, we argue that fear and avoidance of situations far in the future violates the basic logic of evaluation over sequential trajectories of action.
+This is because avoidance is by nature protective: the ability successfully to avoid danger in the future means an agent need not also do so now.
 For instance, cars endanger pedestrians but can be reliably avoided by following traffic signals; given that, staying indoors offers little or no additional protection from accidents.
-This is an instance of a fundamental property of evaluation in sequential decision making: the value of present action turns fundamentally on assumptions about subsequent events, which include the agent's own subsequent choices.
-Typically, it is appropriate to assume that an agent will continue make good (i.e.
-reward-maximizing/harm-minimizing) choices down the line, and that choices at the current stage should reflect this.
+This is an instance of a fundamental property of evaluation in sequential decision making: the value of present action turns fundamentally on assumptions about subsequent events, which importantly include the agent's own subsequent choices.
+Typically, it is appropriate to assume that an agent will continue to make good (i.e.
+reward-maximizing/harm-minimizing) choices down the line, and that good choices at the current stage should therefore anticipate this.
 
-This line of reasoning hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of threat to antecedent situations, and excessive avoidance.
+This line of reasoning hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of threat to antecedent situations and subsequent excessive avoidance.
 Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
@@
 Importantly, the development and maintenance of clinical anxiety is strongly tied to diminished perceived control\citep{bandura1977, barlow2002, gallagher2014a}, such that anxious individuals are more likely to endorse the belief that they are unlikely or unable to mitigate future threat.
-Indeed, a belief (or lack thereof) of one's ability to successfully navigate future danger is associated with anxiety\citep{davey1996, dugas1997}, and an increased belief in perceived control over threat is correlated with symptom reduction across the family of anxiety disorders\citep{gallagher2014b}.
+Indeed, lack of belief in one's ability to successfully navigate future danger is associated with anxiety \citep{davey1996, dugas1997}, and an increased belief in perceived control over threat is correlated with symptom reduction across the family of anxiety disorders\citep{gallagher2014b}.
 
-Here, we develop this idea into a formal model of evaluation under pessimistic assumptions about future choices.
-We show that a single, localized deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance.
+Here, we develop this idea, that symptoms of anxiety may arise from misbeliefs about future avoidance, into a formal model of evaluation under pessimistic assumptions about future choices.
+We show that a single, localized deviation from normative evaluation can explain a surprising range of features of anxious behavior, including exaggerated threat appraisal, fear generalization, and persistent avoidance.
 This account also offers a new formalization of classic insights from the psychiatric literature about the central role of beliefs about control and self-efficacy in anxiety \citep{bandura1977, barlow2002}.
-Specifically, we demonstrate a model with a misbelief about the reliability of future self-action gives rise to the anxious symptoms.
-In this way, our perspective also extends previous computational accounts of beliefs about control in mood disorders (e.g.
+Specifically, we show through simulation that a model with a misbelief about the reliability of future self-action gives rise to a number of characteristic symptoms and laboratory results concerning anxiety.
+Our perspective also extends previous computational accounts of beliefs about control in mood disorders (e.g.
 \cite{HuysDayan2009}), which neglected the sequential aspects of choice.
@@
 The ``max'' operator in Eq.~\ref{eq:bellman} yields a fundamental asymmetry between approach and avoidance, as illustrated in Fig.~1.
-The assumption that the agent makes the return-maximizing choice at each step implies that opportunities for reward propagate recursively to earlier steps, but avoidable dangers do not.
-(To the extent obtaining reward or avoiding harm are only imperfectly attainable, this propagation and attenuation at each step are graded according to success probability, but the basic asymmetry remains.) This principle is highlighted in a toy MDP: a deterministic open gridworld with two terminal states, a rewarding state and a punishing one (Fig.~1a).
+It formalizes the assumption that the agent makes the return-maximizing choice at each step.
+Through this operation, opportunities for reward propagate (which maximize the argument) recursively to earlier steps, but avoidable dangers do not, because the return-maximizing action is to avoid.
+(To the extent obtaining reward or avoiding harm are only imperfectly achievable, this propagation and attenuation at each step are graded according to success probability, but the basic asymmetry remains.) This principle is highlighted in a toy MDP (Fig.~1): a deterministic open gridworld with two terminal states, a rewarding state and a punishing one (Fig.~1a).
 The optimal state values $V^* = \max_a Q^*(s,a)$ (Fig.~1b) reflect a ``mountain'' of opportunity propagating recursively from the reward.
-Conversely, because harm is avoidable in this environment, its negative value is contained: all states (even those adjacent to threat) represent the opportunity for reward.
+Conversely, because harm is avoidable in this environment, its negative value is contained: all states (even those adjacent to threat) represent the positive opportunity for reward.
 
-An agent is not restricted to the return-maximizing assumption, and could anticipate encountering danger under different assumptions about the future, including of course incorrect or pathological beliefs.
-For example, an agent may expect to fail to take the correct protective actions (i.e., expects to use a nonoptimal $\pi(a \mid s)$), or may expect the world's future dynamics do not guarantee reliable avoidance even so (i.e., under stochastic or adversarial transition dynamics $P(s' \mid s,a)$).
+Although the return-maximizing assumption self-consistently defines optimal behavior, an agent need not be restricted to it \citep{symmonds2010}, and might in principle anticipate encountering danger under different (e.g., pathological) assumptions about the future.
+For example, an agent may expect to fail to take the correct protective actions in later states (i.e., to use a suboptimal $\pi(a \mid s)$), or may expect the world's future dynamics do not guarantee reliable avoidance even so (i.e., under stochastic or adversarial transition dynamics $P(s' \mid s,a)$).
 Consider an agent who has such pessimistic expectations about dangerous events at future steps.
 Note that assumptions of this sort, even if incorrect, can serve adaptive purposes.
-It is common in machine learning theory, for instance, to optimize outcomes under worst-case assumptions.
-In general, this can help to ensure robustness and safety under uncertain or even adversarial scenarios \cite{Garcia2015}; related work in RL shows how computing returns with pessimistic predictions can help to quantify variability in outcomes (i.e.
-to learn different points in the distribution of possible returns)\cite{bellemare2017}, and explicitly take account of risk tolerance.
+In general, pessimistic assumptions can help to ensure robustness and safety under uncertain or even adversarial scenarios \cite{Garcia2015}.
+Related work in RL shows how computing returns with pessimistic predictions can help to quantify variability in outcomes (i.e., to learn different points in the distribution of possible returns)\cite{bellemare2017}, which is one way to explicitly and tunably take account of risk tolerance.
+Indeed, it is common in machine learning theory to optimize outcomes under worst-case assumptions.
 
 Here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms.
-Although such pessimism can be encoded either in the policy, $P(s' \mid s,a)$, or transition probabilities, $\pi(a \mid s)$ (i.e.
-misbeliefs about either one's own avoidance actions or the environment's responses to them, a point we return to in discussion), here for concreteness we work with distortions in the policy.
+Such pessimism can be encoded either in the policy, $\pi(a \mid s)$, or transition probabilities , $P(s' \mid s,a)$.
+These, respectively, correspond to misbeliefs about either one's own avoidance actions or the environment's responses to them, a point we return to in discussion.
+Here for concreteness we focus on distortions in the policy.
 In particular, we adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define state-action value in expectation over a mixture of the best and worst action:
@@
 An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences, whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences.
-Fig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in the example gridworld.
-An increasing expectation that the agent may fail to avoid in future causes the value of threat to propagate with increasing distance across the state space.
+\Importantly, this belief is false under the model: that is, we assume, at each step, that the agent actually chooses by maximizing the values given by Equation~\ref{eq:pessimism}.
+But these values are always computed under the assumption that the agent will then not maximize at all later steps.
+(Another way to say this is that we assume choices at each step are optimal, but only under the assumption that later choices will not be.) Fig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in an example gridworld.
+The value of threat propagates, with increasing distance across the state space, due to an increasing expectation that the agent may fail to avoid threat in the future.
 
@@
 
-This simple simulation --- reflecting a localized violation of the core decision theoretic assumption of future return-optimizing action --- already echoes several core symptoms of anxiety disorders.
+This simple simulation reflects a localized violation of the core decision theoretic assumption of future return-optimizing action.
+Therefore, the model's behavior already echoes several core symptoms of anxiety disorders.
 Namely, the pessimistic agents in Fig~1c,d exhibit exaggerated threat appraisals (otherwise neutral states unrealistically signal danger); generalization of fear (threat value spreads across the gridworld); and persistent avoidance (early on, the agent takes paths which maintain increasing distances from threat).
-Importantly, as will be elaborated on below, this deviation from the usual assumptions is supported by prominent clinical theories of anxiety.
+Importantly, as we will elaborate below, this deviation from the usual assumptions is supported by prominent clinical theories of anxiety.
 
@@
 
-In the following, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of sequential learning and decision making.
+In what follows, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of sequential learning and decision making.
 We will also show that our model is consistent with clinical theory describing the transition from clinical anxiety to depression.
@@
 
-One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}; anxious individuals are more likely to forgo potential gains in order to avoid potential danger.
+One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}.
+Anxious individuals are more likely to forgo potential gains in order to avoid potential danger.
 Many of the disruptions anxiety causes to everyday functioning (e.g.
@@
   }
-  \par \textbf{Figure 2:} (A) The Balloon Analog Risk Task (BART).
+  \par \textbf{Figure 2:} (A) The Balloon Analog Risk Task (BART)\citep{Lejuez2002,Maner2007, ramirez2015}.
 The risk of balloon burst (point loss) increases with each pump.
@@
 As shown in Fig.~2, our model easily accommodates this result.
-Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier.
-This is because it anticipates and avoids future errors in choice which would otherwise result in the balloon popping.
+Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier -- similar to empirical findings\citep{Maner2007, ramirez2015}.
+This is because it anticipates and avoids future errors in choice, which would otherwise result in the balloon popping.
 Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety\cite{fung2019}.
-One unique prediction of the model is that this bias should arise only when beliefs about \emph{future} avoidance are involved, rather than direct conflict between immediate impulses.
+
+A unique prediction of the model (because we assume optimal choice under the assumption of \emph{future} suboptimality) is that bias should arise only when beliefs about \emph{future} avoidance are involved, rather than direct conflict between immediate impulses.
 Recent data \cite{fung2019} using a predator avoidance task, analogous to the BART, support this view.
-In their task, trait anxiety predicted earlier escape (analogous to cash-out in the BART) for slow predators (for whom future decisions to escape were a relevant consideration) but not fast ones (who would attack immediately, negating consideration of future steps).
+In this task, increasing trait anxiety predicted earlier escape (analogous to cash-out in the BART) for slow predators (for whom future decisions to escape were a relevant consideration) but not for fast ones (who would attack immediately, mooting consideration of future steps).
 
-The model can also capture findings of increased behavioral inhibition, measured as prolonged response times, in anxious individuals under threat\cite{bach2015}.
-In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator".
-Though in this variant the risk of predation is constant throughout a trial, the potential loss from capture still increases with each token collected.
-Bach \cite{bach2015} finds that participants are slower to collect tokens as loss increases, and that this effect is enhanced by subclinical anxiety.
-We can capture this effect in our model by noting that the relative value of approach compared to avoidance is reduced as the potential loss, and the risk the predator awakes, increases (Fig.~2d/e).
-These effects are amplified under more pessimistic (i.e., on our view, anxious) assumptions about future action.
-Thus, as before, anxiety in our model produces greater and earlier choice of avoidance (analogous to earlier cash-out in the BART).
+Relatedly, the model can also capture findings of increased behavioral inhibition, measured as prolonged response times, in anxious individuals under threat\cite{bach2015}.
+In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator," which are all lost if the predator awakes.
+Though in this variant the risk of predation is constant throughout a trial (rather than increasing, as in the BART) the potential loss from capture still increases with each token collected.
+Bach \cite{bach2015} finds that participants are slower to collect tokens as this potential loss increases, and that this slowing is enhanced by subclinical anxiety.
+We can capture this effect in our model by noting that the relative value of approach compared to avoidance is reduced as potential loss and the risk that the predator awakes increase (Fig.~2d/e).
+These effects are amplified under more pessimistic (i.e., in our model, more anxious) assumptions about future actions.
+Thus, as before, anxiety in our model produces greater and earlier choice of avoidance, analogous to earlier cash-out in the BART.
 To extend this effect to reaction times, we adopt the further, standard assumption that actions (here, approach) are slower when their relative value compared to alternatives is lower \citep{oud2016}.
@@
 
-Another empirical phenomenon associated with anxiety is ``aversive pruning'' in planning \cite{Huys2012, Lally2017}.
+Another laboratory phenomenon associated with anxiety is ``aversive pruning'' in planning \cite{Huys2012, Lally2017}.
 This refers to the idea that when evaluating future action trajectories in a sequential task like chess, people are resource-limited, cannot evaluate all possible options, and must selectively consider certain paths and neglect others.
-One proposal for how people accomplish this is aversive pruning \citep{Huys2012}: choice sequences which involve large losses are discarded from further evaluation.
+One proposal for how people accomplish this is aversive pruning \citep{Huys2012}, wherein choice sequences involving large losses are discarded from further evaluation.
 An example of aversive pruning is shown in Fig.~3a.
@@
 Our model predicts this result (Fig.~3b), though for a somewhat different reason.
-In our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them (here we assume full evaluation of the Bellman equation), but because with increasing anxiety they increasingly expect the potential of choosing incorrectly afterward, thus failing to recoup the loss.
-This reinterpretation is consistent with the structure of the original task\citep{Huys2012, Lally2017}, where participants were required to plan in advance action sequences of two- to eight-steps (requiring the consideration of upwards of $2^8 = 256$ options).
-Uncertainty in one's ability to plan a sequence of reward-maximizing choices could similarly bias a participants towards smaller cost actions.
+In our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them (here we assume full evaluation of the Bellman equation), but because with increasing anxiety they increasingly expect the potential of choosing incorrectly afterward, thus failing to recoup the loss (and, mathematically, probabilistically pruning the better branches).
+%% I'm not sure if this is an attempt to make the story here intuitive or what but this seems wrong to me?
+Our model doesn't actually invoke uncertainty about one's own ability to plan -- it's not that meta -- instead, it plans perfectly but under particular assumptions about one's later behavior.
+%This reinterpretation is consistent with the structure of the original task\citep{Huys2012, Lally2017}, where participants were required to plan in advance action sequences of two- to eight-steps (requiring the consideration of upwards of $2^8 = 256$ options).
+Uncertainty in one's ability to plan a sequence of reward-maximizing choices (see discussion) could similarly bias a participant towards smaller cost actions.
 Future research could use slight variants in the decision trees to tease apart these different interpretations.
-
-% Include the larger, four-panel aversive pruning figure instead?
-%%% I think this figure is OK but shouldnt it include optimal Q values for paths on the left also, showing why the best path is favored?
-%%% also the graying out of the unchosen options on the right is maybe a little strong, hard to read, maybe darken these a bit (then you can also gray out 
-%%% suboptimal paths on the left to make them really parallel?)
 
@@
   \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_tree_rldm.png}}%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_tree.png}}%
   }
-  \par \textbf{Figure 3:} (A) The decision tree environment from \cite{Huys2012} and \cite{Lally2017}.
+  \par \textbf{Figure 3:} (A) The decision tree environment \cite{Huys2012} and \cite{Lally2017}.
 An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (left branch).
 (B) A pessimistic agent ($w=0.5$) comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss.
-One-step rewards are presented in each state; the net value $Q$ on each path is shown numerically.
+One-step rewards (or costs) are presented in each state; the net value $Q$ on each path is shown numerically.
 (Parameters: $\gamma = 1.0$)
@@
 
-So far, we have considered the asymptotic preferences implied the pessimistic value function (solved for directly by using value iteration to compute the return).
-But we can also consider learning under this value function (e.g., by variants of Q-learning\citep{SuttonBarto2018} or DYNA\citep{sutton1991} using the $\beta$-pessimistic return), the dynamics of which may speak to the progression of symptoms.
+So far, we have considered only the asymptotic preferences implied by our pessimistic value function, which we computed directly value iteration.
+But we can also consider the process of learning under this value function (e.g., by variants of Q-learning\citep{SuttonBarto2018} or DYNA\citep{sutton1991, russek_predictive_2017} using the $\beta$-pessimistic return).
+The dynamics of such learning may speak to the progression of symptoms.
 
@@
 This story is borne out by simulations of learning in our model (Fig.~4) in environments like that of Fig.~1.
-As we observed before, the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment with the progress of learning.
-This in turn leads the agent to expect no reward and, echoing the anergic symptoms of depression, forego action altogether.
+Over the course of learning, the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment.
+This can in turn lead the agent to expect no reward and, also echoing the anergic symptoms of depression, forego action altogether.
 
-In addition to accounting for comorbidity of anxiety and depression, our model also hints at a reason for the longevity and recurrence of anxiety disorders even with treatment.
+In addition to accounting for the comorbidity of anxiety and depression, our model also hints at a reason for the longevity and recurrence of anxiety disorders even with treatment.
 Because pessimistic expectations allow for threat value to spread to states and actions far antecedent of the primary danger (e.g., Fig.~1d), it would accordingly also take a great many steps of iterative learning to correct all these exaggerated appraisals of threat.
@@
 A free choice premium has been observed in multiple behavioral experiments \cite{Leotti2010, ly2019}.
-A variant of a free choice premium paradigm from \cite{Leotti2011} and \cite{Leotti2014} is presented in Fig.
+A variant of a free choice premium paradigm from two previous studies\cite{Leotti2011,Leotti2014} is presented in Fig.
 5a.
-In the task, participants repeatedly choose between a free choice option, allowing for an additional future choice, and a fixed choice option; importantly, both choices lead to identical, stochastic outcomes (e.g.
+In the task, participants repeatedly choose between a free choice option, allowing for an additional future choice, and a fixed choice option.
+Importantly, both choices lead to identical, stochastic outcomes (e.g.
 50/50 chance of [1, -1]).
-Previous studies have found that human subjects (from a general, healthy population) prefer the free choice option despite it conferring no additional benefits relative to the fixed choice option.
+Empirical studies have found that human subjects (from a general, healthy population) prefer the free choice option despite it conferring no additional benefits relative to the fixed choice option.
 
@@
   }
-  \par \textbf{Figure 4:} (A) The free choice premium paradigm from \cite{Leotti2011} and \cite{Leotti2014} with equal chance of outcome $R \in [1, -1]$.
+  \par \textbf{Figure 4:} (A) The free choice premium paradigm\cite{Leotti2011,Leotti2014} with equal chance of outcome $R \in [1, -1]$.
 Non-anxious participants exhibit a preference for the free choice option (blue) despite it conferring no benefit over the fixed choice option (grey).
@@
 
-On one account (e.g., \citep{ly2019}) this free choice preference directly and specifically reflects the assumption about sequential choice whose violation we argue is core to anxiety: i.e.
+On one account\citep{ly2019}, this free choice preference directly and specifically reflects the assumption about sequential choice whose violation we argue is core to anxiety: i.e.
 that the agent will make reward-maximizing choices in the future.
-Under such an optimistic assumption (and given noisy and imperfect knowledge about the value of each option, due to learning), additional options are valuable in the sense that free choice can be expected to exploit the best among them: the maximum over several noisy values is, in general, larger than a single option from the same distribution.
-Our theory makes the novel prediction that, if as we hypothesize anxiety reflects a violation of this optimistic assumption, then anxious individuals will exhibit a diminished or reversed free choice bias, as shown in simulation in Fig.
+Under such an optimistic assumption (and given noisy and imperfect knowledge about the value of each option, due to learning), additional options are valuable in the sense that free choice can be expected to exploit the best among them.
+Namely, the maximum over several noisy values is, in general, larger than a single option from the same distribution.
+Our proposal makes the novel prediction that, if as we hypothesize anxiety reflects a violation of this optimistic assumption, then anxious individuals will exhibit a diminished or reversed free choice bias, as shown in simulation in Fig.
 ~5b.
-Future research will be required to test this prediction.
+Future empirical research will be required to test this prediction.
 
@@
 
-Central to anxiety disorders are symptoms including exaggerated threat appraisal, threat generalization, and excessive avoidance.
-We have presented a simple computational account suggesting how a single underlying misbelief can give rise to these aberrations in learning and choice.
+Central to anxiety disorders are symptoms including exaggerated threat appraisal, threat generalization, and excessive avoidance \citep{ClarkBeck2011, dymond2015, Arnaudova2017}.
+We have presented a simple computational account suggesting how a single underlying pessimistic misbelief can give rise to these aberrations in learning and choice.
+We use a reinforcement learning approach in which undue pessimism results in maladaptive policy.
 Specifically, we showed how a failure to believe in the reliability of one's future actions can effectively backpropagate negative value across states of the environment.
 This process results in a range of inferences and behaviors resembling those observed in clinical anxiety.
-Though this is by no means a complete account of anxiety, it ties together a surprisingly wide range of its symptoms.
+Though it is by no means a complete account of anxiety, our account ties together a surprisingly wide range of symptoms of anxiety disorders.
 
-One important ambiguity is that, for concreteness we have formalized pessimistic assumptions in terms of beliefs about the agent's own future actions: expecting failure in handling or avoiding future threat.
-Since Eq.~1 is also computed in expectation over the anticipated future environmental dynamics $p(s' \mid s,a)$, pessimism can alternatively be encoded in this distribution (e.g., a belief that the world's response to one's choices is unpredictable or adversarial).
-Because the Bellman equation for the return averages over both distributions at each step, and because an unpredictable environment also reduces the efficacy of avoidance, either formulation can produce ultimately similar results in our simulations here.
-Other sorts of uncertainty (e.g.
-about the state of the environment, discussed below) can also produce similar effects for analogous reasons.
-However from the perspective of cognitive theories of anxiety, these represent quite different maladaptive beliefs, which may be relevant, for instance, in guiding treatment.
+We are not the first to propose a formal theory of control in psychiatry using MDPs.
+Huys \& Dayan \cite{HuysDayan2009} also provided a computational account of learned helplessness through simple models of one-step action-outcome contingencies.
+Our accounts differ particularly in our exclusive focus on control in the sequential setting, which Huys \& Dayan did not address.
+Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future choice and events.
+Similarly, research and modeling by Bishop and colleagues \citep{XXX} also taking a decision theoretic approach has stressed the importance of uncertainty, and aversion to it, as a core feature of anxiety.
+As for control, we extend this view to focus on how  uncertainty is resolved in the sequential setting, and also to zero in on particular instances of uncertainty (about future actions, and some other options discussed next) and misbeliefs about them that, we argue, are particularly consequential.
 
-Indeed, our account formalizes a longstanding strand of theory on the role of control in anxiety.
+For concreteness, we formalized pessimistic assumptions in terms of only one of several variants of a more general family of models, but we do not mean this restriction as a substantive claim.
+In particular, we focused on the agent's beliefs about their own future actions: expecting failure in handling or avoiding future threat.
+However, this is just one of several different pessimistic misbeliefs that could satisfy the basic logic of our model and produce similar symptoms.
+These other beliefs need not be mutually exclusive, though they might reflect different cognitive routes to symptoms that would, in turn, imply different psychotherapeutic strategies.
+
+For instance, one variant of the model is suggested by the observation that Eq.~1 is also computed in expectation over the anticipated future environmental dynamics $p(s' \mid s,a)$.
+Thus, pessimism can alternatively be encoded in this distribution: e.g., a false belief that the world's response to one's choices is unpredictable or adversarial.
+Because the Bellman equation for the return averages over this distribution in addition to the choice policy at each step, and because an unpredictable environment also reduces the efficacy of avoidance, either formulation can produce ultimately similar results in our simulations here.
+A third variant of our model arises from uncertainty about the current state $s$ of the environment.
+Although we have taken it as fully observed, if the world state is only partly known, then this distribution too must be averaged out in evaluating each action \citep{kaelbling1998}, and here also a pessimistic skew will propagate the expectation of danger and result in exaggerated avoidance.
+In summary, pessimistic resolution of several different varieties of uncertainty (e.g., about future action, environmental dynamics, or environmental state) could each produce similar symptoms for analogous reasons.
+However, from the perspective of cognitive theories of anxiety, these represent quite different maladaptive beliefs: a key difference that may be relevant in guiding treatment (especially cognitive psychotherapies aimed at ameliorating the false beliefs) of a host of anxiety disorders.
+
+Particularly due to the way it encompasses several such variants, our account formalizes a longstanding range of theory on the role of control in anxiety.
 Central to many prominent cognitive theories of anxiety in the psychiatric literature is a perceived lack of control.
 For example, self-efficacy theory \citep{bandura1977} and the triple vulnerabilities model \citep{barlow2002} both posit that a reduced belief in the ability to effectively respond to future threat is involved in the genesis and maintenance of clinical anxiety.
-Analogously, but focused less on the self, the learned helplessness theories of anxiety \cite{alloy1990} claims clinical anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
-As we note above, the present model and analysis (though stated here in terms of self-efficacy) can accommodate either variant.
-Also, they are by no means mutually exclusive, though they might also reflect multiple distinct (and differently treatable) routes to similar symptomatology.
+In contrast, and focused less on the self, the learned helplessness theories of anxiety \cite{alloy1990} claims clinical anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
+As we note above, the present model and analysis (though simulated above in terms of self-efficacy) can accommodate either variant, and shows how they relate to one another as members of a more general family of accounts.
+%Also, they are by no means mutually exclusive, though they might also reflect multiple distinct (and differently treatable) routes to similar symptomatology.
 
-Indeed, yet a third variant of our idea concerns the notion of state.
+%An alternative variant of our proposal concerns the representation of states in the environment.
 In the present work, we considered only environments with fully observable states; in other words, there was no uncertainty regarding the state (e.g.
 grid world location) an agent currently occupied.
-In the real world, however, the state must be inferred from external and internal observations.
+In the real world, however, states must commonly be inferred from external and internal observations.
 This provides yet another source of uncertainty similar to the future policy and transition dynamics, which must be averaged over in evaluating actions \citep{kaelbling1998}.
@@
 For example, if an agent is erroneously biased toward believing herself more proximal to threat than is actually the case, then symptoms like exaggerated threat appraisal and avoidance should arise.
-Indeed, a bias in value functions could result in the diffuse spread of threat across objects\citep{norbury2018} and location\citep{schulz2018}, thereby causing otherwise benign states to be perceived as threatening.
+Indeed, such a bias could result in the diffuse spread of threat across objects\citep{norbury2018} and location\citep{schulz2018}, thereby causing otherwise benign states to be perceived as threatening.
 
-Speculatively, all these distorted beliefs may reflect pathological cases of an otherwise adaptive evaluation strategy.
+Importantly, although we have considered the most pathological cases of pessimism, these may reflect the exaggerative extremes of an otherwise adaptive evaluation strategy.
 Traditionally, the goal of reinforcement learning algorithms is to a find reward-maximizing policy with respect to the expectation (average) of returns.
-However, depending on one's risk attitude and state of uncertainty (e.g.
+However, depending on one's risk attitude and uncertainty about the environment (e.g.
 if there is potential for catastrophic loss), it may be instead preferable to learn a policy with respect to an alternative and more pessimistic objective function, similar to the one considered here.
-How to learn risk-sensitive and robust policies is the focus of much recent research \cite{morimura2012, chow2015, bellemare2017}; interestingly, different algorithmic variants in this literature produce agent behaviors not unlike those presented here.
-Thus another perspective on our hypothesis is that anxious individuals, in effect, evaluate actions with respect to a more pessimistic point in the distribution of expected returns.
+Accordingly, returns and agent behaviors similar to ours arise in previous research on learning risk-sensitive and robust policies \cite{morimura2012, chow2015, bellemare2017}.
+%% I think the following sentence is what made it seem like we had too many model variants, so I tried to refocus on the usefulness of this type of thing and leave it at that.
+%Thus another perspective on our hypothesis is that anxious individuals, in effect, evaluate actions with respect to a more pessimistic point in the distribution of expected returns.
 Suggestively, anxiety is correlated with disproportionate worry about low chance but high cost events\cite{Miceli2005}.
 
-We have centered our discussion at Marr's \cite{marr1982} computational level: on beliefs and their consequences in terms of action values, while remaining agnostic as to how, algorithmically or mechanistically, these misbeliefs are evaluated.
-Of course, the brain is believed to contain multiple distinct mechanisms for evaluating actions (e.g, model-based and model-free; learning; \cite{daw2005, huys2015}), and pessimisitic beliefs might play out either differentially or similarly through each of these mechanisms.
-One particular possibility is that these symptoms mainly reflect aberrations in model-based planning \cite{huys2015}, i.e.
-explicitly evaluating valuations by mentally simulating possible trajectories.
-Recent work suggests this process may be accomplished by mentally ``replaying'' potential trajectories successively (e.g.
-in hippocampus, \cite{mattar2018}).
-In this setting the biases we suggest would amount to over-contemplating, or over-weighting, certain pessimistic trajectories \cite{hunter2019}.
-This, in turn, may relate to worry and rumination; indeed, in line with the present results, chronic worry is associated with reduced perceived control, diminished belief in self-efficacy in response to threat, and exaggerated threat appraisal\cite{Berenbaum2010}.
+We have centered our discussion at Marr's \cite{marr1982} computational level: on beliefs and their consequences in terms of action values.
+We have so far remained agnostic as to how, algorithmically or mechanistically, these misbeliefs are implemented in the brain.
+Importantly, the brain is believed to contain multiple distinct mechanisms for evaluating actions (e.g, model-based and model-free learning; \cite{daw2005, huys2015}), and pessimisitic beliefs might play out either differentially or similarly through each of these mechanisms.
+One promising possibility is that these symptoms mainly reflect aberrations in model-based planning \cite{huys2015}, i.e.
+explicitly evaluating actions by mentally simulating possible trajectories.
+Recent work suggests this process may be accomplished by mentally ``replaying'' individual potential trajectories \cite{momennejad_offline_2018,mattar2018}.
+In this setting, the biases we suggest would amount to over-contemplating, or over-weighting, certain pessimistic trajectories \cite{hunter2019}.
+Such a bias might be detectable using neuroimaging, as a change in which types of events that tend to be replayed \cite{momennejad_offline_2018, ambrose2016}.
+Such a biased replay process, in turn, may also correspond to worry and rumination.
+Indeed, in line with the present results, chronic worry is associated with reduced perceived control, diminished belief in self-efficacy in response to threat, and exaggerated threat appraisal\cite{Berenbaum2010}.
 This suggests that clinical anxiety may in part result from planning processes gone awry.
@@
 It is important to note that the present model may not describe all anxiety disorders with equal accuracy.
-Indeed, our model of pessimistic sequential evaluation is, by definition, a model of prospective cognition.
-As such, the present results are more likely to accurately describe the anxiety disorders which primarily involve aberrations in future-oriented cognitive processes, such as generalized and social anxiety disorders.
-Naturally, the present model can account neither for the compulsive behaviors of obsessive-compulsive disorder nor the memory disturbances of post-traumatic stress disorder (though recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD)\citep{gallagher2014a, gallagher2014b}.
-Importantly, other psychiatric disorders that involve future-oriented misbeliefs, worry, and avoidance behaviors (e.g.
+Indeed, our analysis of pessimistic sequential evaluation is, by definition, a model of prospective cognition.
+Thus, the present results are more likely to accurately describe the anxiety disorders which primarily involve aberrations in future-oriented cognitive processes, such as generalized and social anxiety disorders.
+Naturally, the present model can account neither for the compulsive behaviors of obsessive-compulsive disorder (OCD) nor the memory disturbances of post-traumatic stress disorder (PTSD).
+That said, recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD \citep{gallagher2014a, gallagher2014b}.
+Importantly, aspects of other psychiatric disorders that involve future-oriented misbeliefs, worry, and avoidance behaviors (e.g.
 eating disorders\citep{konstantellou2011}) may similarly be well-described by the current account.
 As such, the present results are transdiagnostic and not limited to one particular diagnosis.
+Finally, the relationship between evaluation, planning, and neural replay discussed above suggests potential future work that might help to bring this model into contact with the more memory-related aspects of PTSD.
+For instance, the same replay processes that can be used to evaluate actions can also, in theory, update predictive representations, cognitive maps, or models of the environment \citep{russek_predictive_2017}, such as the successor representation \cite{dayan1992,momennejad_successor_2017}.
+If so, similar biases in replay could result in not just abberant avoidance behavior, but also progressive, abberant remodeling of world models or cognitive maps, an observation which may connect to the rich and complex set of issues on memory involvement in PTSD.
 
-Finally, we are not the first to propose a formal theory of control in psychiatry using MDPs.
-Huys \& Dayan \cite{HuysDayan2009} also provided a computational account of learned helplessness through simple models of one-step action-outcome contingencies.
-Our accounts differ particularly in our exclusive focus on control in the sequential setting, which Huys \& Dayan did not address.
-Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future choice and events.
+%Note that in the present work, pessimism leads to distorted policy.
+Future work is required to show how, first, such pessimism can also bias offline replay with negative prediction errors (PE) rather than prioritizing surprising events based on their unsigned PE \citep{momennejad_offline_2018}.
+
+%Recent research using human experiments and reinforcement learning models has shown that offline replay can update predictive representations of the environment \citep{momennejad_successor_2017; russek_predictive_2017}.
+Second, and sometimes as a result of biased replay, pessimism can be encoded in state representations or cognitive maps, e.g., transition probabilities, or multi-step successor representations \citep{momennejad_successor_2017}, that the agent uses to make decisions.
+Maladaptive effects of pessimistic bias on replay and predictive representations of the environment may underlie, and explain, behavioral symptoms of memory-related anxiety disorders such as post traumatic stress disorder (PTSD).
 
@@
 
-Here, $Q$-values were solved for directly through value iteration.
-\citep{SuttonBarto2018} By contrast, $Q$-values in the free choice task were computed using the $\beta$-pessimism temporal-difference learning algorithm \citep{Gaskett2003}, where the reward prediction error is defined as:
+Here, $Q$-values were solved for directly through value iteration \citep{SuttonBarto2018}.
+By contrast, $Q$-values in the free choice task were computed using the $\beta$-pessimism temporal-difference learning algorithm \citep{Gaskett2003}, where the reward prediction error is defined as:
 
@@
 There, a virtual predator approaches participants over discrete time steps while participants "forage".
-For every time step the participant remains (i.e.
-does not flee), points are accumulated; if the participant is caught, however, all earnings were lost and an additional penalty was received.
+For every time step the participant remained (i.e.
+does not flee), points were accumulated.
+However, if the participant is caught all earnings were lost and an additional penalty was received.
 Thus, the predator avoidance task bears striking resemblance to the BART; foraging is equivalent to virtual pumps, and fleeing is equivalent to cashing-out.
@@
 capture by predator) with reward equal to the negative equivalent of accumulated gain thus far.
-The probability of bad transition was defined as $p = 0.10$ for low risk and $p = 0.15$ high risk, based on the objective risk probabilities in \citep{bach2015}.
+The probability of bad transition was defined as $p = 0.10$ for low risk and $p = 0.15$ high risk, based on the objective risk probabilities in the emepirical experiment \citep{bach2015}.
 The asymptomatic $Q$-values were solved for using value iteration for both the low and high risk conditions under pessimistic assumptions, $w \in [1.0, 0.6, 0.2]$.
@@
 We model the task as an undiscounted MDP where the original 6-node network has been "unravelled" into a 15-state decision tree.
-This is equivalent to having the participant start in one state and planning to make three actions.
+This is equivalent to having the participant start in one state and plan to make three actions.
 The rewards associated with transitioning to each state were taken directly from \citep{Huys2012, Lally2017}.

### 2019-08-01 manuscripts/manuscript/manuscript_old.tex

--- 2019-07-02 manuscripts/manuscript/manuscript.tex
@@
 Anxiety disorders are characterized by a range of aberrations in the processing of and response to threat, but there is little clarity what core pathogenesis might underlie these symptoms.
-Here we propose a decision theoretic analysis of maladaptive avoidance and embody it in a reinforcement learning model, which shows how a localized bias in beliefs can formally explain a range of phenomena related to anxiety.
+Here we propose that unrealistically pessimistic assumptions can distort an action policy and underlie a host of anxiety symptoms.
+We adopt a decision theoretic analysis of maladaptive avoidance and embody it in a reinforcement learning model.
+This model shows how a localized bias in beliefs can formally explain a range of phenomena related to anxiety.
 The core observation, implicit in standard decision theoretic accounts of sequential evaluation, is that the potential for avoidance should be protective: if danger can be avoided later, it poses no threat now.
@@
 Finally, anxiety is associated with persistent avoidance behavior, which often occurs well in advance of the materialization of actual threat \citep{Arnaudova2017}.
-Such inapposite avoidance behaviors are especially deleterious aspects of the disoders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
+Such inappropriate avoidance behaviors are especially harmful aspects of the disorders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
 Though laboratory studies of decision making and learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.
@@
 In many circumstances, distant threat should not impinge upon decision making in the present.
-Indeed, fear and avoidance of situations far in the future violates the basic logic of evaluation over sequential trajectories of action.
+Indeed, we argue that fear and avoidance of situations far in the future violates the basic logic of evaluation over sequential trajectories of action.
 This is because avoidance is by nature protective: the ability to successfully avoid danger in the future means an agent need not also do so now.
@@
 
-This line of reasoning hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of threat to antecedent situations, and excessive avoidance.
+This line of reasoning hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of threat to antecedent situations and subsequent excessive avoidance.
 Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
@@
 
-Here, we develop this idea into a formal model of evaluation under pessimistic assumptions about future choices.
+Here, we develop this idea, of construing anxiety as maladaptive avoidance strategies, into a formal model of evaluation under pessimistic assumptions about future choices.
 We show that a single, localized deviation can explain a surprising range of features of anxious behavior including exaggerated threat appraisal, fear generalization, and persistent avoidance.
@@
 The ``max'' operator in Eq.~\ref{eq:bellman} yields a fundamental asymmetry between approach and avoidance, as illustrated in Fig.~1.
-The assumption that the agent makes the return-maximizing choice at each step implies that opportunities for reward propagate recursively to earlier steps, but avoidable dangers do not.
-(To the extent obtaining reward or avoiding harm are only imperfectly attainable, this propagation and attenuation at each step are graded according to success probability, but the basic asymmetry remains.) This principle is highlighted in a toy MDP: a deterministic open gridworld with two terminal states, a rewarding state and a punishing one (Fig.~1a).
+The assumption is that the agent makes the return-maximizing choice at each step.
+Therefore, opportunities for reward propagate recursively to earlier steps, but avoidable dangers do not.
+(To the extent obtaining reward or avoiding harm are only imperfectly attainable, this propagation and attenuation at each step are graded according to success probability, but the basic asymmetry remains.) This principle is highlighted in a toy MDP (Fig.~1): a deterministic open gridworld with two terminal states, a rewarding state and a punishing one (Fig.~1a).
 The optimal state values $V^* = \max_a Q^*(s,a)$ (Fig.~1b) reflect a ``mountain'' of opportunity propagating recursively from the reward.
@@
 
-An agent is not restricted to the return-maximizing assumption, and could anticipate encountering danger under different assumptions about the future, including of course incorrect or pathological beliefs.
-For example, an agent may expect to fail to take the correct protective actions (i.e., expects to use a nonoptimal $\pi(a \mid s)$), or may expect the world's future dynamics do not guarantee reliable avoidance even so (i.e., under stochastic or adversarial transition dynamics $P(s' \mid s,a)$).
+An agent need not be restricted to the return-maximizing assumption \citep{symmonds2010}, and might anticipate encountering danger under different (e.g.
+pathological) assumptions about the future.
+For example, an agent may expect to fail to take the correct protective actions (i.e., expects to use a suboptimal $\pi(a \mid s)$), or may expect the world's future dynamics do not guarantee reliable avoidance even so (i.e., under stochastic or adversarial transition dynamics $P(s' \mid s,a)$).
 Consider an agent who has such pessimistic expectations about dangerous events at future steps.
 Note that assumptions of this sort, even if incorrect, can serve adaptive purposes.
-It is common in machine learning theory, for instance, to optimize outcomes under worst-case assumptions.
-In general, this can help to ensure robustness and safety under uncertain or even adversarial scenarios \cite{Garcia2015}; related work in RL shows how computing returns with pessimistic predictions can help to quantify variability in outcomes (i.e.
+In general, pessimistic assumptions can help to ensure robustness and safety under uncertain or even adversarial scenarios \cite{Garcia2015}.
+Related work in RL shows how computing returns with pessimistic predictions can help to quantify variability in outcomes (i.e.
 to learn different points in the distribution of possible returns)\cite{bellemare2017}, and explicitly take account of risk tolerance.
+Indeed, it is common in machine learning theory to optimize outcomes under worst-case assumptions.
 
 Here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms.
-Although such pessimism can be encoded either in the policy, $P(s' \mid s,a)$, or transition probabilities, $\pi(a \mid s)$ (i.e.
-misbeliefs about either one's own avoidance actions or the environment's responses to them, a point we return to in discussion), here for concreteness we work with distortions in the policy.
+Although such pessimism can be encoded either in the policy, $\pi(a \mid s)$, or transition probabilities , $P(s' \mid s,a)$.
+These, respectively, correspond to misbeliefs about either one's own avoidance actions or the environment's responses to them, a point we return to in discussion.
+Here for concreteness we work with distortions in the policy.
 In particular, we adopt the $\beta$-pessimistic value function from \cite{Gaskett2003}, to define state-action value in expectation over a mixture of the best and worst action:
@@
 An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences, whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences.
-Fig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in the example gridworld.
-An increasing expectation that the agent may fail to avoid in future causes the value of threat to propagate with increasing distance across the state space.
+Fig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in an example gridworld.
+The value of threat propagates, with increasing distance across the state space, due to an increasing expectation that the agent may fail to avoid threat in the future.
 
@@
 
-This simple simulation --- reflecting a localized violation of the core decision theoretic assumption of future return-optimizing action --- already echoes several core symptoms of anxiety disorders.
+This simple simulation reflects a localized violation of the core decision theoretic assumption of future return-optimizing action.
+Therefore, the model's behavior already echoes several core symptoms of anxiety disorders.
 Namely, the pessimistic agents in Fig~1c,d exhibit exaggerated threat appraisals (otherwise neutral states unrealistically signal danger); generalization of fear (threat value spreads across the gridworld); and persistent avoidance (early on, the agent takes paths which maintain increasing distances from threat).
@@
 As shown in Fig.~2, our model easily accommodates this result.
-Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier.
-This is because it anticipates and avoids future errors in choice which would otherwise result in the balloon popping.
+Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier -- similar to empirical findings\citep{Maner2007, ramirez2015}.
+This is because it anticipates and avoids future errors in choice, which would otherwise result in the balloon popping.
 Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety\cite{fung2019}.
-One unique prediction of the model is that this bias should arise only when beliefs about \emph{future} avoidance are involved, rather than direct conflict between immediate impulses.
+
+A unique prediction of the model is that bias should arise only when beliefs about \emph{future} avoidance are involved, rather than direct conflict between immediate impulses.
 Recent data \cite{fung2019} using a predator avoidance task, analogous to the BART, support this view.
@@
 Bach \cite{bach2015} finds that participants are slower to collect tokens as loss increases, and that this effect is enhanced by subclinical anxiety.
-We can capture this effect in our model by noting that the relative value of approach compared to avoidance is reduced as the potential loss, and the risk the predator awakes, increases (Fig.~2d/e).
+We can capture this effect in our model by noting that the relative value of approach compared to avoidance is reduced as potential loss and the risk that the predator awakes increase (Fig.~2d/e).
 These effects are amplified under more pessimistic (i.e., on our view, anxious) assumptions about future action.
@@
 This refers to the idea that when evaluating future action trajectories in a sequential task like chess, people are resource-limited, cannot evaluate all possible options, and must selectively consider certain paths and neglect others.
-One proposal for how people accomplish this is aversive pruning \citep{Huys2012}: choice sequences which involve large losses are discarded from further evaluation.
+One proposal for how people accomplish this is aversive pruning \citep{Huys2012} wherein choice sequences involving large losses are discarded from further evaluation.
 An example of aversive pruning is shown in Fig.~3a.
@@
 This reinterpretation is consistent with the structure of the original task\citep{Huys2012, Lally2017}, where participants were required to plan in advance action sequences of two- to eight-steps (requiring the consideration of upwards of $2^8 = 256$ options).
-Uncertainty in one's ability to plan a sequence of reward-maximizing choices could similarly bias a participants towards smaller cost actions.
+Uncertainty in one's ability to plan a sequence of reward-maximizing choices (see discussion) could similarly bias a participants towards smaller cost actions.
 Future research could use slight variants in the decision trees to tease apart these different interpretations.
-
-% Include the larger, four-panel aversive pruning figure instead?
-%%% I think this figure is OK but shouldnt it include optimal Q values for paths on the left also, showing why the best path is favored?
-%%% also the graying out of the unchosen options on the right is maybe a little strong, hard to read, maybe darken these a bit (then you can also gray out 
-%%% suboptimal paths on the left to make them really parallel?)
 
@@
 
-So far, we have considered the asymptotic preferences implied the pessimistic value function (solved for directly by using value iteration to compute the return).
-But we can also consider learning under this value function (e.g., by variants of Q-learning\citep{SuttonBarto2018} or DYNA\citep{sutton1991} using the $\beta$-pessimistic return), the dynamics of which may speak to the progression of symptoms.
+So far, we have considered only the asymptotic preferences returned by our pessimistic value function, solved for directly through value iteration.
+But we can also consider learning under this value function (e.g., by variants of Q-learning\citep{SuttonBarto2018} or DYNA\citep{sutton1991, momennejad_successor_2017, russek_predictive_2017} using the $\beta$-pessimistic return).
+The dynamics of such learning may speak to the progression of symptoms.
 
@@
 As we observed before, the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment with the progress of learning.
-This in turn leads the agent to expect no reward and, echoing the anergic symptoms of depression, forego action altogether.
+This can in turn lead the agent to expect no reward and, echoing the anergic symptoms of depression, forego action altogether.
 
-In addition to accounting for comorbidity of anxiety and depression, our model also hints at a reason for the longevity and recurrence of anxiety disorders even with treatment.
+In addition to accounting for the comorbidity of anxiety and depression, our model also hints at a reason for the longevity and recurrence of anxiety disorders even with treatment.
 Because pessimistic expectations allow for threat value to spread to states and actions far antecedent of the primary danger (e.g., Fig.~1d), it would accordingly also take a great many steps of iterative learning to correct all these exaggerated appraisals of threat.
@@
 
-Central to anxiety disorders are symptoms including exaggerated threat appraisal, threat generalization, and excessive avoidance.
-We have presented a simple computational account suggesting how a single underlying misbelief can give rise to these aberrations in learning and choice.
+Central to anxiety disorders are symptoms including exaggerated threat appraisal, threat generalization, and excessive avoidance \citep{ClarkBeck2011, dymond2015, Arnaudova2017}.
+We have presented a simple computational account suggesting how a single underlying pessimistic misbelief can give rise to these aberrations in learning and choice.
+We use a reinforcement learning approach in which undue pessimism results in maladaptive policy.
 Specifically, we showed how a failure to believe in the reliability of one's future actions can effectively backpropagate negative value across states of the environment.
 This process results in a range of inferences and behaviors resembling those observed in clinical anxiety.
-Though this is by no means a complete account of anxiety, it ties together a surprisingly wide range of its symptoms.
+Though it is by no means a complete account of anxiety, our account ties together a surprisingly wide range of symptoms of anxiety disorders.
 
-One important ambiguity is that, for concreteness we have formalized pessimistic assumptions in terms of beliefs about the agent's own future actions: expecting failure in handling or avoiding future threat.
-Since Eq.~1 is also computed in expectation over the anticipated future environmental dynamics $p(s' \mid s,a)$, pessimism can alternatively be encoded in this distribution (e.g., a belief that the world's response to one's choices is unpredictable or adversarial).
+One important ambiguity is that we, for concreteness, have formalized pessimistic assumptions in terms of beliefs about the agent's own future actions: expecting failure in handling or avoiding future threat.
+However, note that Eq.~1 is also computed in expectation over the anticipated future environmental dynamics $p(s' \mid s,a)$.
+Therefore, pessimism can alternatively be encoded in this distribution (e.g., a belief that the world's response to one's choices is unpredictable or adversarial).
 Because the Bellman equation for the return averages over both distributions at each step, and because an unpredictable environment also reduces the efficacy of avoidance, either formulation can produce ultimately similar results in our simulations here.
-Other sorts of uncertainty (e.g.
+Consequently, other varieties of uncertainty (e.g.
 about the state of the environment, discussed below) can also produce similar effects for analogous reasons.
-However from the perspective of cognitive theories of anxiety, these represent quite different maladaptive beliefs, which may be relevant, for instance, in guiding treatment.
+However, from the perspective of cognitive theories of anxiety, these represent quite different maladaptive beliefs: a key difference that may be relevant in guiding treatment of a host of anxiety disorders.
 
@@
 
-Indeed, yet a third variant of our idea concerns the notion of state.
+An alternative variant of our proposal concerns the representation of states in the environment.
 In the present work, we considered only environments with fully observable states; in other words, there was no uncertainty regarding the state (e.g.
 grid world location) an agent currently occupied.
-In the real world, however, the state must be inferred from external and internal observations.
+In the real world, however, states must commonly be inferred from external and internal observations.
 This provides yet another source of uncertainty similar to the future policy and transition dynamics, which must be averaged over in evaluating actions \citep{kaelbling1998}.
@@
 For example, if an agent is erroneously biased toward believing herself more proximal to threat than is actually the case, then symptoms like exaggerated threat appraisal and avoidance should arise.
-Indeed, a bias in value functions could result in the diffuse spread of threat across objects\citep{norbury2018} and location\citep{schulz2018}, thereby causing otherwise benign states to be perceived as threatening.
+Indeed, such a bias could result in the diffuse spread of threat across objects\citep{norbury2018} and location\citep{schulz2018}, thereby causing otherwise benign states to be perceived as threatening.
 
-Speculatively, all these distorted beliefs may reflect pathological cases of an otherwise adaptive evaluation strategy.
+All of the varieties of distorted beliefs mentioned above may reflect pathological cases of an otherwise adaptive evaluation strategy.
 Traditionally, the goal of reinforcement learning algorithms is to a find reward-maximizing policy with respect to the expectation (average) of returns.
-However, depending on one's risk attitude and state of uncertainty (e.g.
+However, depending on one's risk attitude and uncertainty about the environment (e.g.
 if there is potential for catastrophic loss), it may be instead preferable to learn a policy with respect to an alternative and more pessimistic objective function, similar to the one considered here.
-How to learn risk-sensitive and robust policies is the focus of much recent research \cite{morimura2012, chow2015, bellemare2017}; interestingly, different algorithmic variants in this literature produce agent behaviors not unlike those presented here.
+Interestingly, previous research on learning risk-sensitive and robust policies \cite{morimura2012, chow2015, bellemare2017} have produced agent behaviors not unlike those presented here.
 Thus another perspective on our hypothesis is that anxious individuals, in effect, evaluate actions with respect to a more pessimistic point in the distribution of expected returns.
@@
 explicitly evaluating valuations by mentally simulating possible trajectories.
-Recent work suggests this process may be accomplished by mentally ``replaying'' potential trajectories successively (e.g.
-in hippocampus, \cite{mattar2018}).
+Recent work suggests this process may be accomplished by mentally ``replaying'' potential trajectories that are prioritized by uncertainty or prediction errors in cortical and sub-cortical regions \cite{momennejad_offline_2018}, and this prioritized trajectory replay is mediated by multi-step computations in the hippocampus \cite{mattar2018}).
 In this setting the biases we suggest would amount to over-contemplating, or over-weighting, certain pessimistic trajectories \cite{hunter2019}.
@@
 Indeed, our model of pessimistic sequential evaluation is, by definition, a model of prospective cognition.
-As such, the present results are more likely to accurately describe the anxiety disorders which primarily involve aberrations in future-oriented cognitive processes, such as generalized and social anxiety disorders.
-Naturally, the present model can account neither for the compulsive behaviors of obsessive-compulsive disorder nor the memory disturbances of post-traumatic stress disorder (though recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD)\citep{gallagher2014a, gallagher2014b}.
+Thus, the present results are more likely to accurately describe the anxiety disorders which primarily involve aberrations in future-oriented cognitive processes, such as generalized and social anxiety disorders.
+Naturally, the present model can account neither for the compulsive behaviors of obsessive-compulsive disorder (OCD) nor the memory disturbances of post-traumatic stress disorder (PTSD).
+That said, recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD \citep{gallagher2014a, gallagher2014b}.
 Importantly, other psychiatric disorders that involve future-oriented misbeliefs, worry, and avoidance behaviors (e.g.
@@
 As such, the present results are transdiagnostic and not limited to one particular diagnosis.
+
+Note that in the present work, pessimism leads to distorted policy.
+Future work is required to show how, first, such pessimism can also bias offline replay with negative prediction errors (PE) rather than prioritizing surprising events based on their unsigned PE \citep{momennejad_offline_2018}.
+Recent research using human experiments and reinforcement learning models has shown that offline replay can update predictive representations of the environment \citep{momennejad_successor_2017; russek_predictive_2017}.
+Second, and sometimes as a result of biased replay, pessimism can be encoded in state representations or cognitive maps, e.g., transition probabilities, or multi-step successor representations \citep{momennejad_successor_2017}, that the agent uses to make decisions.
+Maladaptive effects of pessimistic bias on replay and predictive representations of the environment may underlie, and explain, behavioral symptoms of memory-related anxiety disorders such as post traumatic stress disorder (PTSD).
 
@@
 Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future choice and events.
+
+<<<2-3 sentence summary paragraph?>>>
 

### 2019-08-01 sisyphus/envs/_choice.py
improved figures

@@ -102,9 +102,9 @@ class FreeChoice(GraphWorld):
 
         ax.text(0.5,  0.15, '0', ha='center', va='bottom', fontsize=fontsize)
         ax.text(0.5, -0.15, '0', ha='center', va='top', fontsize=fontsize)
-        ax.text(1.5,  0.40, r'$R$', ha='center', va='bottom', fontsize=fontsize)
-        ax.text(1.5,  0.10, r'$R$', ha='center', va='top', fontsize=fontsize)
-        ax.text(1.5, -0.40, r'$R$', ha='center', va='top', fontsize=fontsize)
+        ax.text(1.5,  0.35, r'$[-1, 1]$', ha='center', va='bottom', fontsize=fontsize, rotation=15)
+        ax.text(1.5,  0.15, r'$[-1, 1]$', ha='center', va='top', fontsize=fontsize, rotation=-15)
+        ax.text(1.5, -0.35, r'$[-1, 1]$', ha='center', va='top', fontsize=fontsize, rotation=-15)
         return ax
 
     def plot_free_choice(self, ax, s=50, free_color='#834c7d', fixed_color='0.9',

### 2019-08-01 manuscripts/manuscript/manuscript.tex
Updates from Overleaf

@@
 Finally, anxiety is associated with persistent avoidance behavior, which often occurs well in advance of the materialization of actual threat \citep{Arnaudova2017}.
-Such inappropriate avoidance behaviors are especially harmful aspects of the disorders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
+Excessive avoidance behaviors are an especially harmful aspect of anxiety disorders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
 Though laboratory studies of decision making and learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.
@@
 Although the return-maximizing assumption self-consistently defines optimal behavior, an agent need not be restricted to it \citep{symmonds2010}, and might in principle anticipate encountering danger under different (e.g., pathological) assumptions about the future.
-For example, an agent may expect to fail to take the correct protective actions in later states (i.e., to use a suboptimal $\pi(a \mid s)$), or may expect the world's future dynamics do not guarantee reliable avoidance even so (i.e., under stochastic or adversarial transition dynamics $P(s' \mid s,a)$).
+For example, an agent may expect to fail to take the correct protective actions in later states (i.e., to use a suboptimal $\pi(a \mid s)$), or may believe the world's future dynamics do not guarantee reliable avoidance even so (i.e., under stochastic or adversarial transition dynamics $P(s' \mid s,a)$).
 Consider an agent who has such pessimistic expectations about dangerous events at future steps.
@@
 An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences, whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences.
-\Importantly, this belief is false under the model: that is, we assume, at each step, that the agent actually chooses by maximizing the values given by Equation~\ref{eq:pessimism}.
+Importantly, this belief is false under the model: that is, we assume, at each step, that the agent actually chooses by maximizing the values given by Equation~\ref{eq:pessimism}.
 But these values are always computed under the assumption that the agent will then not maximize at all later steps.
@@
 
-One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}.
-Anxious individuals are more likely to forgo potential gains in order to avoid potential danger.
-Many of the disruptions anxiety causes to everyday functioning (e.g.
-avoiding social obligations for fear of possible social humiliation) can be understood in these terms.
-As such, many have sought to probe and measure this behavior in the laboratory.
-For instance, in the balloon analog risk task (BART) \citep{Lejuez2002}, participants attempt to earn money by pumping virtual balloons.
-With each pump, the balloon inflates and money is earned, but so too does the chance that the balloon pops and the accumulated earnings are lost.
-At any point in a trial, a participant may cash out, banking the money earned and ending the trial.
-Anxiety is correlated with fewer pumps of the balloon and earlier cash-outs in the BART \citep{Maner2007, ramirez2015}.
-
 \begin{figure}
   \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/02_appavo.png}}%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/02_appavo.png}}
   }
@@
 \end{figure}
+
+One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}.
+Anxious individuals are more likely to forgo potential gains in order to avoid potential danger.
+Many of the disruptions anxiety causes to everyday functioning (e.g.
+avoiding social obligations for fear of possible social humiliation) can be understood in these terms.
+As such, many have sought to probe and measure this behavior in the laboratory.
+For instance, in the balloon analog risk task (BART) \citep{Lejuez2002}, participants attempt to earn money by pumping virtual balloons.
+With each pump, the balloon inflates and money is earned, but so too does the chance that the balloon pops and the accumulated earnings are lost.
+At any point in a trial, a participant may cash out, banking the money earned and ending the trial.
+Anxiety is correlated with fewer pumps of the balloon and earlier cash-outs in the BART \citep{Maner2007, ramirez2015}.
 
@@
 
+\begin{figure}[!b]
+  \centerline{%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_tree.png}}%
+  }
+  \par \textbf{Figure 3:} (A) The decision tree environment \cite{Huys2012} and \cite{Lally2017}.
+An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (left branch).
+(B) A pessimistic agent ($w=0.5$) comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss.
+One-step rewards (or costs) are presented in each state; the net value $Q$ on each path is shown numerically.
+(Parameters: $\gamma = 1.0$)
+\end{figure}
+
 Another laboratory phenomenon associated with anxiety is ``aversive pruning'' in planning \cite{Huys2012, Lally2017}.
@@
 
+\subsection{The anxiety-depression transition}
+
 \begin{figure}[!b]
   \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_tree.png}}%
+    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/03_lh.png}}%
   }
-  \par \textbf{Figure 3:} (A) The decision tree environment \cite{Huys2012} and \cite{Lally2017}.
-An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (left branch).
-(B) A pessimistic agent ($w=0.5$) comes to prefer the branch without the large loss so as to avoid being unable to recoup the large initial loss.
-One-step rewards (or costs) are presented in each state; the net value $Q$ on each path is shown numerically.
-(Parameters: $\gamma = 1.0$)
+  \par \textbf{Figure 4:} A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
+(B-D, E-G) The development of value expectancies over three steps of learning, for two levels of pessimism.
+  States are colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
+(B-D) For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value with learning.
+(E-G) For a pessimistic agent ($w=0.6$), negative value spreads from the source to antecedent states.
+As a result of avoidance, the agent learns reward is unobtainable and develops anergic symptoms (i.e.
+foregoes action).
+(Parameters: $\gamma = 0.95$)
 \end{figure}
 
-\subsection{The anxiety-depression transition}
-
-So far, we have considered only the asymptotic preferences implied by our pessimistic value function, which we computed directly value iteration.
+So far, we have considered only the asymptotic preferences implied by our pessimistic value function, which we computed directly through value iteration.
 But we can also consider the process of learning under this value function (e.g., by variants of Q-learning\citep{SuttonBarto2018} or DYNA\citep{sutton1991, russek_predictive_2017} using the $\beta$-pessimistic return).
@@
 
-\begin{figure}[!b]
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/03_lh.png}}%
-  }
-  \par \textbf{Figure 4:} A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
-(B-D, E-G) The development of value expectancies over three steps of learning, for two levels of pessimism.
-  States are colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
-(B-D) For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value with learning.
-(E-G) For a pessimistic agent ($w=0.6$), negative value spreads from the source to antecedent states.
-As a result of avoidance, the agent learns reward is unobtainable and develops anergic symptoms (i.e.
-foregoes action).
-(Parameters: $\gamma = 0.95$)
-\end{figure}
-
 \subsection{Free choice premium}
@@
   }
-  \par \textbf{Figure 4:} (A) The free choice premium paradigm\cite{Leotti2011,Leotti2014} with equal chance of outcome $R \in [1, -1]$.
+  \par \textbf{Figure 5:} (A) The free choice premium paradigm\cite{Leotti2011,Leotti2014} with equal chance of outcome $R \in [1, -1]$.
 Non-anxious participants exhibit a preference for the free choice option (blue) despite it conferring no benefit over the fixed choice option (grey).
@@
 A third variant of our model arises from uncertainty about the current state $s$ of the environment.
-Although we have taken it as fully observed, if the world state is only partly known, then this distribution too must be averaged out in evaluating each action \citep{kaelbling1998}, and here also a pessimistic skew will propagate the expectation of danger and result in exaggerated avoidance.
+Although we have taken it as fully observed, if the world state is only partly known, then this distribution too must be averaged out in evaluating each action \citep{kaelbling1998}, and here also a pessimistic skew will propagate the expectation of danger and result in exaggerated avoidance \citep{Paulus2012}.
 In summary, pessimistic resolution of several different varieties of uncertainty (e.g., about future action, environmental dynamics, or environmental state) could each produce similar symptoms for analogous reasons.
@@
 As we note above, the present model and analysis (though simulated above in terms of self-efficacy) can accommodate either variant, and shows how they relate to one another as members of a more general family of accounts.
-%Also, they are by no means mutually exclusive, though they might also reflect multiple distinct (and differently treatable) routes to similar symptomatology.
-
-%An alternative variant of our proposal concerns the representation of states in the environment.
-In the present work, we considered only environments with fully observable states; in other words, there was no uncertainty regarding the state (e.g.
-grid world location) an agent currently occupied.
-In the real world, however, states must commonly be inferred from external and internal observations.
-This provides yet another source of uncertainty similar to the future policy and transition dynamics, which must be averaged over in evaluating actions \citep{kaelbling1998}.
-Thus another variant of our approach is to view anxiety as potentially also arising due to pessimistic biases in state inference\cite{Paulus2012}, which would again have similar effects on the resulting value function.
-For example, if an agent is erroneously biased toward believing herself more proximal to threat than is actually the case, then symptoms like exaggerated threat appraisal and avoidance should arise.
-Indeed, such a bias could result in the diffuse spread of threat across objects\citep{norbury2018} and location\citep{schulz2018}, thereby causing otherwise benign states to be perceived as threatening.
 
@@
 Finally, the relationship between evaluation, planning, and neural replay discussed above suggests potential future work that might help to bring this model into contact with the more memory-related aspects of PTSD.
-For instance, the same replay processes that can be used to evaluate actions can also, in theory, update predictive representations, cognitive maps, or models of the environment \citep{russek_predictive_2017}, such as the successor representation \cite{dayan1992,momennejad_successor_2017}.
-If so, similar biases in replay could result in not just abberant avoidance behavior, but also progressive, abberant remodeling of world models or cognitive maps, an observation which may connect to the rich and complex set of issues on memory involvement in PTSD.
-
-%Note that in the present work, pessimism leads to distorted policy.
-Future work is required to show how, first, such pessimism can also bias offline replay with negative prediction errors (PE) rather than prioritizing surprising events based on their unsigned PE \citep{momennejad_offline_2018}.
-
-%Recent research using human experiments and reinforcement learning models has shown that offline replay can update predictive representations of the environment \citep{momennejad_successor_2017; russek_predictive_2017}.
-Second, and sometimes as a result of biased replay, pessimism can be encoded in state representations or cognitive maps, e.g., transition probabilities, or multi-step successor representations \citep{momennejad_successor_2017}, that the agent uses to make decisions.
-Maladaptive effects of pessimistic bias on replay and predictive representations of the environment may underlie, and explain, behavioral symptoms of memory-related anxiety disorders such as post traumatic stress disorder (PTSD).
+For instance, the same replay processes that can be used to evaluate actions can also, in theory, update predictive representations, cognitive maps, or models of the environment \citep{russek_predictive_2017}, such as the successor representation \cite{dayan1993, momennejad_successor_2017}.
+If so, similar biases in replay could result in not just aberrant avoidance behavior, but also progressive, aberrant remodeling of world models or cognitive maps, an observation which may connect to the rich and complex set of issues on memory involvement in PTSD.
 

### 2019-08-01 sisyphus/envs/_choice.py
minor figure fixes

@@ -98,13 +98,13 @@ class FreeChoice(GraphWorld):
 
         return ax
 
-    def _draw_edge_labels(self, ax, fontsize=14):
+    def _draw_edge_labels(self, ax, fontsize):
 
-        ax.text(0.5,  0.15, '0', ha='center', va='bottom', fontsize=fontsize)
-        ax.text(0.5, -0.15, '0', ha='center', va='top', fontsize=fontsize)
-        ax.text(1.5,  0.35, r'$[-1, 1]$', ha='center', va='bottom', fontsize=fontsize, rotation=15)
-        ax.text(1.5,  0.15, r'$[-1, 1]$', ha='center', va='top', fontsize=fontsize, rotation=-15)
-        ax.text(1.5, -0.35, r'$[-1, 1]$', ha='center', va='top', fontsize=fontsize, rotation=-15)
+        ax.text(0.5,  0.150, '0', ha='center', va='bottom', fontsize=fontsize)
+        ax.text(0.5, -0.150, '0', ha='center', va='top', fontsize=fontsize)
+        ax.text(1.5,  0.350, r'$[-1, 1]$', ha='center', va='bottom', fontsize=fontsize, rotation=15)
+        ax.text(1.5,  0.145, r'$[-1, 1]$', ha='center', va='top', fontsize=fontsize, rotation=-15)
+        ax.text(1.5, -0.355, r'$[-1, 1]$', ha='center', va='top', fontsize=fontsize, rotation=-15)
         return ax
 
     def plot_free_choice(self, ax, s=50, free_color='#834c7d', fixed_color='0.9',
@@ -138,7 +138,7 @@ class FreeChoice(GraphWorld):
         ax = self._draw_nodes(ax, s=50, free_color=free_color, fixed_color=fixed_color)
 
         ## Optional details.
-        if edge_labels: ax = self._draw_edge_labels(ax, fontsize=18)
+        if edge_labels: ax = self._draw_edge_labels(ax, fontsize=14)
 
         ## Clean up.
         ax.set(xlim=(-0.5,2.5), xticks=[], ylim=(-0.7,0.7), yticks=[])

### 2019-08-01 manuscripts/manuscript/manuscript.tex
almost ready

@@
   \par \textbf{Figure 4:} A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
-(B-D, E-G) The development of value expectancies over three steps of learning, for two levels of pessimism.
+(B, c) The development of value expectancies over three steps of learning, for two levels of pessimism.
   States are colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
-(B-D) For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value with learning.
-(E-G) For a pessimistic agent ($w=0.6$), negative value spreads from the source to antecedent states.
+(B) For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value with learning.
+(C) For a pessimistic agent ($w=0.6$), negative value spreads from the source to antecedent states.
 As a result of avoidance, the agent learns reward is unobtainable and develops anergic symptoms (i.e.
@@
 
-\begin{figure}[!b]
+\begin{figure}[t]
   \centerline{%

### 2019-08-02 manuscripts/manuscript/manuscript.tex
completed draft

@@
 The moderate ($w=0.6$) and strongly ($w=0.2$) pessimistic agents cash-out earlier, as is observed in anxious individuals.
-(D/E) In the sleeping predator task, the risk of loss is constant but the cost of loss still increases as more rewards are gathered.
+(D,E) In the sleeping predator task, the risk of loss is constant but the cost of loss still increases as more rewards are gathered.
 The value of reward pursuit under increasingly pessimistic valuation is presented for scenarios with low (D) and high risk (E)  of predator awakening.
@@
   }
-  \par \textbf{Figure 3:} (A) The decision tree environment \cite{Huys2012} and \cite{Lally2017}.
+  \par \textbf{Figure 3:} (A) The decision tree environment \cite{Huys2012,Lally2017}.
 An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (left branch).
@@
 In our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them (here we assume full evaluation of the Bellman equation), but because with increasing anxiety they increasingly expect the potential of choosing incorrectly afterward, thus failing to recoup the loss (and, mathematically, probabilistically pruning the better branches).
-%% I'm not sure if this is an attempt to make the story here intuitive or what but this seems wrong to me?
-Our model doesn't actually invoke uncertainty about one's own ability to plan -- it's not that meta -- instead, it plans perfectly but under particular assumptions about one's later behavior.
-%This reinterpretation is consistent with the structure of the original task\citep{Huys2012, Lally2017}, where participants were required to plan in advance action sequences of two- to eight-steps (requiring the consideration of upwards of $2^8 = 256$ options).
-Uncertainty in one's ability to plan a sequence of reward-maximizing choices (see discussion) could similarly bias a participant towards smaller cost actions.
 Future research could use slight variants in the decision trees to tease apart these different interpretations.
@@
   }
-  \par \textbf{Figure 4:} A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
+  \par \textbf{Figure 4:} (A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
 (B, c) The development of value expectancies over three steps of learning, for two levels of pessimism.
@@
   }
-  \par \textbf{Figure 5:} (A) The free choice premium paradigm\cite{Leotti2011,Leotti2014} with equal chance of outcome $R \in [1, -1]$.
+  \par \textbf{Figure 5:} (A) The free choice premium task\cite{Leotti2011,Leotti2014} with equal chance of outcome $R \in [1, -1]$.
 Non-anxious participants exhibit a preference for the free choice option (blue) despite it conferring no benefit over the fixed choice option (grey).
@@
 Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future choice and events.
-Similarly, research and modeling by Bishop and colleagues \citep{XXX} also taking a decision theoretic approach has stressed the importance of uncertainty, and aversion to it, as a core feature of anxiety.
+Similarly, research and modeling by Bishop and colleagues \citep{browning2015, gagne2018} also taking a decision theoretic approach has stressed the importance of uncertainty, and aversion to it, as a core feature of anxiety.
 As for control, we extend this view to focus on how  uncertainty is resolved in the sequential setting, and also to zero in on particular instances of uncertainty (about future actions, and some other options discussed next) and misbeliefs about them that, we argue, are particularly consequential.
@@
 Accordingly, returns and agent behaviors similar to ours arise in previous research on learning risk-sensitive and robust policies \cite{morimura2012, chow2015, bellemare2017}.
-%% I think the following sentence is what made it seem like we had too many model variants, so I tried to refocus on the usefulness of this type of thing and leave it at that.
-%Thus another perspective on our hypothesis is that anxious individuals, in effect, evaluate actions with respect to a more pessimistic point in the distribution of expected returns.
-Suggestively, anxiety is correlated with disproportionate worry about low chance but high cost events\cite{Miceli2005}.
 

### 2019-08-04 manuscripts/manuscript/manuscript.tex
submitted :)

@@
 
-\acknowledgements{This research was supported in part by NIH grant MH10917, part of the CRCNS program, and by an NSF Graduate Research Fellowship.}
+\acknowledgements{The authors are grateful to Zeb Kurth-Nelson and Will Dabney for helpful discussion.
+This work was funded in part by NIDA 1R01DA038891, part of the CRCNS program, and by an NSF Graduate Research Fellowship.}
 
@@
 We have so far remained agnostic as to how, algorithmically or mechanistically, these misbeliefs are implemented in the brain.
-Importantly, the brain is believed to contain multiple distinct mechanisms for evaluating actions (e.g, model-based and model-free learning; \cite{daw2005, huys2015}), and pessimisitic beliefs might play out either differentially or similarly through each of these mechanisms.
+\cite{friston2014computational} Importantly, the brain is believed to contain multiple distinct mechanisms for evaluating actions (e.g, model-based and model-free learning; \cite{daw2005, huys2015}), and pessimisitic beliefs might play out either differentially or similarly through each of these mechanisms.
 One promising possibility is that these symptoms mainly reflect aberrations in model-based planning \cite{huys2015}, i.e.

### 2019-11-17 manuscripts/manuscript/manuscript.tex

@@
 Though anxiety disorders differ in their particular symptomology, and in the content and situations that elicit symptoms, they all are similarly characterized by aberrations in the processing of and response to threat \citep{dsm5}.
-In particular, at least three symptoms manifest across the anxiety disorders.
+In particular, at least three symptoms manifest across many of the anxiety disorders.
 First, anxiety is associated with exaggerated threat appraisal, or a bias towards evaluating threat as disproportionately greater in likelihood and severity than is warranted \citep{ClarkBeck2011}.
@@
 
-This line of reasoning hints that a fundamental deficit in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of threat to antecedent situations and subsequent excessive avoidance.
+This line of reasoning hints that a fundamental aberration in anxiety disorders may relate to this assumption, which otherwise should preclude the spread of threat to antecedent situations and subsequent excessive avoidance.
 Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
@@
 
-\begin{equation*}
+\begin{equation}\label{eq:bellman_full}
 Q^\pi(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \sum_{a'} \pi(a' \mid s') Q^\pi(s',a').
-\end{equation*}
+\end{equation}
 
@@
 Relatedly, the model can also capture findings of increased behavioral inhibition, measured as prolonged response times, in anxious individuals under threat\cite{bach2015}.
-In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator," which are all lost if the predator awakes.
+In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator", which are all lost if the predator awakes.
 Though in this variant the risk of predation is constant throughout a trial (rather than increasing, as in the BART) the potential loss from capture still increases with each token collected.
@@
 Future research could use slight variants in the decision trees to tease apart these different interpretations.
+Tree order matters in delineating between these theories.
+What's one good example?
 
@@
 
+Insert paragraph here about testing. Delgado experiment for one. Learning transitions another?
+
 Particularly due to the way it encompasses several such variants, our account formalizes a longstanding range of theory on the role of control in anxiety.
@@
 We have so far remained agnostic as to how, algorithmically or mechanistically, these misbeliefs are implemented in the brain.
-\cite{friston2014computational} Importantly, the brain is believed to contain multiple distinct mechanisms for evaluating actions (e.g, model-based and model-free learning; \cite{daw2005, huys2015}), and pessimisitic beliefs might play out either differentially or similarly through each of these mechanisms.
+\cite{friston2014computational} Importantly, the brain is believed to contain multiple distinct mechanisms for evaluating actions (e.g, model-based and model-free learning; \cite{daw2005, huys2015}), and pessimistic beliefs might play out either differentially or similarly through each of these mechanisms.
 One promising possibility is that these symptoms mainly reflect aberrations in model-based planning \cite{huys2015}, i.e.
@@
 eating disorders\citep{konstantellou2011}) may similarly be well-described by the current account.
+Speculatively, the model may also have implications for understanding dysregulated goal pursuit in mania and bipolar disorder.
+Manic symptoms are associated with increases in perceived control \citep{
+implications for understanding increased perceived control in mania \cite{}.
 As such, the present results are transdiagnostic and not limited to one particular diagnosis.

### 2019-12-05 sisyphus/mdp/_td.py
fix figure 5

@@ -68,6 +68,9 @@ class ModelFree(object):
         ## Define starting state.
         copy = gym.info.copy()
         s = gym.start  
+        
+        ## Initialize action list.
+        actions = []
 
         for _ in np.arange(n_steps):
 
@@ -78,7 +81,8 @@ class ModelFree(object):
             copy['Q'] = Q.copy()
             i = choice(copy.loc[copy.S==s,'Q'].values, epsilon)
             a = copy[copy.S==s].index[i]
-                        
+            actions.append(a)
+                
             ## Observe next state and reward.
             i = categorical(copy.loc[a,'T'])
             s_prime = copy.loc[a,"S'"][i]
@@ -92,7 +96,7 @@ class ModelFree(object):
             ## Update state.
             s = s_prime
 
-        return Q
+        return Q, actions
     
     def _v_solve(self, info):
         """Compute state value from Q-table."""
@@ -132,7 +136,7 @@ class ModelFree(object):
                 
         return policy
         
-    def fit(self, gym, choice='softmax', schedule=None, n_steps=100, overwrite=False):
+    def fit(self, gym, choice='softmax', schedule=None, n_steps=100, overwrite=False, return_actions=False):
         '''Run a single test episode (i.e. Q-values not updated).
         
         Parameters
@@ -148,6 +152,8 @@ class ModelFree(object):
             Maximum number of steps allowed in a single episode.
         overwrite : True | False
             If true, overwrite previously stored Q-values (if any).
+        return_actions : True | False
+            If true, return all choices made during training.
             
         Returns
         -------
@@ -185,7 +191,11 @@ class ModelFree(object):
             Q = self.Q.copy()
             
         ## Solve for Q-values.
-        for e in schedule: Q = self._run_episode(Q, gym, choice, e, n_steps)
+        actions = []
+        for e in schedule: 
+            Q, a = self._run_episode(Q, gym, choice, e, n_steps)
+            actions.append(a)
+            
         self.Q = Q
         
         ## Solve for values.
@@ -194,4 +204,5 @@ class ModelFree(object):
         ## Compute policy.
         self.pi = self._pi_solve(gym)
                 
-        return self
+        if return_actions: return self, actions
+        else: return self

### 2019-12-05 manuscripts/manuscript/manuscript.tex
Updates from Overleaf

@@
 Finally, anxiety is associated with persistent avoidance behavior, which often occurs well in advance of the materialization of actual threat \citep{Arnaudova2017}.
-Excessive avoidance behaviors are an especially harmful aspect of anxiety disorders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
+(Here we distinguish between avoidance and escape, where the former describes actions taken to prevent the onset of threats whereas the latter describes defensive responses to proximal threat.) Excessive avoidance behaviors are an especially harmful aspect of anxiety disorders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats.
 Though laboratory studies of decision making and learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.
@@
 
-For any particular state-action $(s,a)$, this is necessarily defined only relative to a policy $\pi(a' \mid s')$ specifying the assumed distribution of \emph{future} choices.
-The return can be optimized self-consistently under the assumption that the agent makes the return-maximizing choice at each step in the future, leading to the familiar expression for the optimal values,
+For any particular state-action $(s,a)$, this is necessarily defined relative to a policy $\pi(a' \mid s')$ specifying the assumed distribution of \emph{future} choices.
+(It is also defined relative to a discount rate $\gamma$, controlling the present value of future outcomes.) The return can be optimized self-consistently under the assumption that the agent makes the return-maximizing choice at each step in the future, leading to the familiar expression for the optimal values,
 
@@
 It formalizes the assumption that the agent makes the return-maximizing choice at each step.
-Through this operation, opportunities for reward propagate (which maximize the argument) recursively to earlier steps, but avoidable dangers do not, because the return-maximizing action is to avoid.
+Through this operation, opportunities for reward (which maximize the argument) propagate recursively to earlier steps, but avoidable dangers do not, because the return-maximizing action is to avoid.
 (To the extent obtaining reward or avoiding harm are only imperfectly achievable, this propagation and attenuation at each step are graded according to success probability, but the basic asymmetry remains.) This principle is highlighted in a toy MDP (Fig.~1): a deterministic open gridworld with two terminal states, a rewarding state and a punishing one (Fig.~1a).
@@
 
-The parameter $w$ controls the degree of pessimism.
+The weight parameter $w$ controls the degree of pessimism.
 An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences, whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences.
@@
 
-In what follows, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of sequential learning and decision making.
-We will also show that our model is consistent with clinical theory describing the transition from clinical anxiety to depression.
+In what follows, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of \emph{sequential learning}  and decision making.
+(Because in our model anxiety arises through biased sequential evaluation, we will not address one-step bandit tasks where others have reported learning deficits associated with anxiety \citep{Harle2017, Aylward2019}.) We will also show that our model is consistent with clinical theory describing the transition from clinical anxiety to depression.
 Unless otherwise noted, state and action values under varying degrees of pessimism were solved for using the value iteration dynamic programming method\citep{SuttonBarto2018}.
@@
   \par \textbf{Figure 2:} (A) The Balloon Analog Risk Task (BART)\citep{Lejuez2002,Maner2007, ramirez2015}.
-The risk of balloon burst (point loss) increases with each pump.
-The optimal policy (number of pumps) under increasingly pessimistic valuation is presented for low (B) and high risk (C) balloons.
+The risk of balloon burst (point loss) increases with each pump and does so earlier for the high risk (red) than low risk (blue) balloons.
+(For full rules of the task, see Methods.) The optimal policy (number of pumps) under increasingly pessimistic valuation is presented for low (B) and high risk (C) balloons.
 The optimistic agent ($w=1$) prefers a policy reflecting the true environmental risk.
@@
 This is because it anticipates and avoids future errors in choice, which would otherwise result in the balloon popping.
-Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety\cite{fung2019}.
+Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety, such as in the predator avoidance task\cite{fung2019}.
 
@@
 These effects are amplified under more pessimistic (i.e., in our model, more anxious) assumptions about future actions.
-Thus, as before, anxiety in our model produces greater and earlier choice of avoidance, analogous to earlier cash-out in the BART.
+Thus, as before, anxious pessimism in our model produces greater and earlier choice of avoidance, analogous to earlier cash-out in the BART.
 To extend this effect to reaction times, we adopt the further, standard assumption that actions (here, approach) are slower when their relative value compared to alternatives is lower \citep{oud2016}.
@@
 Although the optimal choice in the decision tree from is to weather an initial large loss (e.g., -70) in order to reap the large gain that follows, people tend to disfavor this path suggesting they prune it and consequently neglect the later gain.
-The degree of such pruning correlates with subclinical depressive \cite{Huys2012} and anxiety \cite{Lally2017} symptoms.
+The degree of such pruning correlates, depending on the study, with subclinical depressive \cite{Huys2012} or anxiety \cite{Lally2017} symptoms.
 
-Our model predicts this result (Fig.~3b), though for a somewhat different reason.
+Our model predicts this result (Fig.~3b) as specifically linked to our model of anxious pessimism, though for a somewhat different reason than in Huys' original modeling.
 In our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them (here we assume full evaluation of the Bellman equation), but because with increasing anxiety they increasingly expect the potential of choosing incorrectly afterward, thus failing to recoup the loss (and, mathematically, probabilistically pruning the better branches).
-Future research could use slight variants in the decision trees to tease apart these different interpretations.
-Tree order matters in delineating between these theories.
-What's one good example?
+Future research could use slight variants in the decision trees to tease apart these different interpretations, for example, by comparing decision trees that differ only in what follows the large initial loss.
+Under a model of aversive pruning, such a change should not impact the proportion of agents selecting the left branch; in contrast, our model predicts choice should parametrically increase with the extent of the amelioration.
 
@@
 Of note in this respect, anxiety and depression are highly comorbid, with almost half of individuals with a lifetime depression diagnosis also diagnosed with an anxiety disorder \citep{kessler2015}.
-One notable proposal is that this association often arises longitudinally: in particular, that clinical anxiety precedes certain types of depression \citep{alloy1990, jacobson2014}.
+One notable proposal is that this association often (though by no means exclusively) arises longitudinally: in particular, that clinical anxiety precedes certain types of depression \citep{alloy1990, jacobson2014}.
 The idea, in brief, is that uncertainty in one's ability in the face of future threat results in anxiety and avoidance behaviors.
 Persistent avoidance, in turn, begets foregone reward, and ultimately to a belief that reward is unobtainable and subsequently depression.
-This story is borne out by simulations of learning in our model (Fig.~4) in environments like that of Fig.~1.
+This informal story can be captured by simulations of learning in our model (Fig.~4) in environments like that of Fig.~1.
 Over the course of learning, the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment.
 This can in turn lead the agent to expect no reward and, also echoing the anergic symptoms of depression, forego action altogether.
+This last point in particular dovetails nicely with theoretical accounts of the anergic aspects of depression \citep{huys2015}, which point out that low experienced reward rates should in decision theoretic accounts lead to reduced response vigor \citep{niv2007} leading to a potentially self-reinforcing downward spiral.
 
-In addition to accounting for the comorbidity of anxiety and depression, our model also hints at a reason for the longevity and recurrence of anxiety disorders even with treatment.
+In addition to suggesting one explanation for the comorbidity of anxiety and depression, our model also hints at a reason for the longevity and recurrence of anxiety disorders even with treatment.
 Because pessimistic expectations allow for threat value to spread to states and actions far antecedent of the primary danger (e.g., Fig.~1d), it would accordingly also take a great many steps of iterative learning to correct all these exaggerated appraisals of threat.
@@
 Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future choice and events.
-Similarly, research and modeling by Bishop and colleagues \citep{browning2015, gagne2018} also taking a decision theoretic approach has stressed the importance of uncertainty, and aversion to it, as a core feature of anxiety.
+Similarly, research and modeling by Bishop and colleagues \citep{browning2015, gagne2018} also taking a decision theoretic approach has stressed the importance of uncertainty as a core feature of anxiety.
+Specifically, they have described uncertainty as inherently aversive in anxiety, and have presented models of how uncertainty may be increased in anxiety (e.g.
+aberrant processing of environmental volatility).
+The present work is compatible with deficits in processing uncertainty, and might instead be viewed as an attempt to unpack why uncertainty is aversive: because, in our view, it is resolved (i.e.
+marginalized) under pessimistic distributional assumptions.
 As for control, we extend this view to focus on how uncertainty is resolved in the sequential setting, and also to zero in on particular instances of uncertainty (about future actions, and some other options discussed next) and misbeliefs about them that, we argue, are particularly consequential.
@@
 
-Insert paragraph here about testing. Delgado experiment for one. Learning transitions another?
-
 Particularly due to the way it encompasses several such variants, our account formalizes a longstanding range of theory on the role of control in anxiety.
@@
 As we note above, the present model and analysis (though simulated above in terms of self-efficacy) can accommodate either variant, and shows how they relate to one another as members of a more general family of accounts.
+
+The possibility of multiple anxious phenotypes, each characterized by unique but not mutually exclusive beliefs, suggests the need for behavioral assays designed to isolate and interrogate such biases.
+One such task is the free choice premium paradigm described above, which captures pessimism (or optimism) about one's own choices.
+An analogous task might measure pessimistic expectations about environmental state transition probabilities.
+For instance, this could be accomplished in a variant of sequential decision making tasks that require subjects to learn the transition structure of a multi-step decision tree \citep{glascher2010, lee2014} and make choices to gather rewards in it.
+Pessimistic expectations about environmental state transitions would bias choices in this type of task.
+Individually, these tasks could test our hypothesis that either sort of misbelief is associated with symptoms of anxiety; they could also be compared to one another (and to more detailed self-report assessments of beliefs about control or self-efficacy) to investigate potential heterogeneity across patients in the antecedent of anxiety.
 
@@
 eating disorders\citep{konstantellou2011}) may similarly be well-described by the current account.
-Speculatively, the model may also have implications for understanding dysregulated goal pursuit in mania and bipolar disorder.
-Manic symptoms are associated with increases in perceived control \citep{
-implications for understanding increased perceived control in mania \cite{}.
+Indeed, and much more speculatively, the model may also have implications for bipolar disorders.
+The onset of manic symptoms is associated with overoptimistic perceived control \cite{alloy2006}, and our same model and basic reasoning (though now envisioning excessively optimistic rather than excessively pessimistic beliefs) may help to explain how this bias may translate into risk-seeking behavior and dysreguated goal pursuit.
 As such, the present results are transdiagnostic and not limited to one particular diagnosis.
+
 Finally, the relationship between evaluation, planning, and neural replay discussed above suggests potential future work that might help to bring this model into contact with the more memory-related aspects of PTSD.
@@
 
-\textbf{Averse Pruning} In the aversive pruning task \citep{Huys2012, Lally2017}, participants learn to navigate a six-state graphworld, where each state is directly connected to only two other states.
+\textbf{Aversive Pruning} In the aversive pruning task \citep{Huys2012, Lally2017}, participants learn to navigate a six-state graphworld, where each state is directly connected to only two other states.
 Each state is associated with some reward (cost), and participants must plan trajectories through the state-space so as to maximize reward (minimize cost).

### 2019-12-05 manuscripts/manuscript/manuscript.tex
latest draft

@@
 In the transition gridworld, we performed simulations under pessimistic assumptions, $w \in [1.0, 0.6]$.
-To highlight the effects of learned, we took "snapshots" of $Q$-values prior to asymtote, three and five steps into value iteration.
+To highlight the effects of learned, we took "snapshots" of $Q$-values prior to asymptote, three and five steps into value iteration.
 
@@
 capture by predator) with reward equal to the negative equivalent of accumulated gain thus far.
-The probability of bad transition was defined as $p = 0.10$ for low risk and $p = 0.15$ high risk, based on the objective risk probabilities in the emepirical experiment \citep{bach2015}.
+The probability of bad transition was defined as $p = 0.10$ for low risk and $p = 0.15$ high risk, based on the objective risk probabilities in the empirical experiment \citep{bach2015}.
 The asymptomatic $Q$-values were solved for using value iteration for both the low and high risk conditions under pessimistic assumptions, $w \in [1.0, 0.6, 0.2]$.
@@
 All bandits pay out identically, in this case randomly in the set, $r \in [-1, 1]$.
-Learned $Q$-values were computed using pessimistic temporal difference learning algorithm, with learning rate $\eta = 0.1$.
-Each simulated agent learned the values of each action over 100 trials, with an increasing, logarithmically spaced inverse temperature in the range of $\beta \in [0, 1]$.
-(Inverse temperature was gradually increased over learning to facilitate exploration of choice options.) The resulting $Q$-values were stored for 500 simulated agents, run separately for pessimistic assumption, $w \in [1.0, 0.5, 0.0]$.
+Learned $Q$-values were computed using pessimistic temporal difference learning algorithm, with learning rate $\eta = 0.4$.
+Each simulated agent learned the values of each action over 100 trials, with an increasing, logarithmically spaced inverse temperature in the range of $\beta \in [0, 15]$.
+(Inverse temperature was gradually increased over learning to facilitate exploration of choice options.) The resulting fraction of free choices made over the last 50 trials were stored for 1000 simulated agents, run separately for pessimistic assumption, $w \in [1.0, 0.5, 0.0]$.
 

### 2019-12-06 manuscripts/manuscript/manuscript.tex
Updates from Overleaf

@@
 The fractional preference for the free choice option over the simulated experiment is shown for three populations of subjects with levels of pessimism $w$ (y-axis; each dot represents an individual simulated agent, and the smoothed density of free choice bias for each pessimism level is also shown.)
-  (Parameters: Q-learning with $\gamma = 1.0$, and inverse temperature, $\beta$, increased from 1 to 10 over 100 episodes.)
+  (Parameters: Q-learning with $\gamma = 1.0$, and inverse temperature, $\beta$, increased from 1 to 15 over 100 episodes.)
 \end{figure}

### 2019-12-29 manuscripts/cspy/CompPsychTemplate.tex
accepted :)

--- 2019-12-06 manuscripts/manuscript/manuscript.tex
@@
-\documentclass[11pt]{article} % For LaTeX2e
-\usepackage{manuscript, palatino}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{amsfonts, amsmath}
-\usepackage[numbers,super]{natbib}
+%% Documentclass:
+\documentclass[CompPsych]{stjour}
+
+%% Manuscript, for double spaced, larger fonts
+% \documentclass[manuscript]{stjour}
+%% Only needed if you use `manuscript' option
+% \journalname{Computational Psychiatry}
+
+%%%%%%%%%%% Article Set-Up %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Article Type:
+%% Default is Research Article.
+
+%% Or, choose one of these options:
+%% Commentary
+%% Review Article
+%% Modification
+%% Retraction
+
+\articletype{Research Article}
+
+%%%%%%%%%%% Please supply information %%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% For Computational Psychiatry, Supporting Info
+\supportinginfo{http://dx.doi.org/XXXXXX}
+
+%% If no conflicts, this command doesn't need to be used
+%% \conflictsofinterest{}
+
+%%%%%%%%%%% to be supplied by MIT Press, only %%%%%%%%%%%%%%%%%
+
+\citation{Zorowitz, S., Momennejad, I., Daw, N. D. (2019). Anxiety, avoidance, and sequential evaluation. XXXXXX}
+
+\received{4 August 2019}
+\accepted{29 December 2019}
+\published{XXXXXX}
+
+%% DOI address:
+\setdoi{10.1098/rsif.2013.0969}
+
+%%%%%%%% End MIT Press commands %%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% author definitions should be placed here:
+
+%% example definition
+\def\taupav{\tau_{\mathrm{Pav}}}
+\begin{document}
 
@@
 
-\author{
-Samuel Zorowitz \\
-Princeton Neuroscience Institute\\
-Princeton University\\
-Princeton, NJ 08540 \\
-\texttt{zorowitz@princeton.edu} \\
-\And
-Ida Momennejad \\
-Department of Biomedical Engineering\\
-Columbia University\\
-New York, NY 10027 \\
-\texttt{ida.m@columbia.edu} \\
-\And
-Nathaniel D. Daw \\
-Princeton Neuroscience Institute\\
-and Department of Psychology\\
-Princeton University\\
-Princeton, NJ 08540 \\
-\texttt{ndaw@princeton.edu} \\
-}
+%%% Author/Affil
+\author[Author Names]
+{Samuel Zorowitz\affil{1}, Ida Momennejad\affil{2}, \and Nathaniel D. Daw\affil{1,3}}
 
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
+\affiliation{1}{Princeton Neuroscience Institute,
+Princeton University, Princeton, NJ 08540}
 
-\begin{document}
+\affiliation{2}{Department of Biomedical Engineering, 
+Columbia University, New York, NY 10027}
 
-\maketitle
+\affiliation{3}{Department of Psychology,
+Princeton University, Princeton, NJ 08540}
+
+\correspondingauthor{Samuel Zorowitz}{szorowi1@gmail.com}
+
+\keywords{anxiety, avoidance, fear generalization, decision theory}
 
@@
 
-\keywords{
-anxiety; avoidance; fear generalization; decision theory; computational psychiatry
-}
-
-\acknowledgements{The authors are grateful to Zeb Kurth-Nelson and Will Dabney for helpful discussion.
-This work was funded in part by NIDA 1R01DA038891, part of the CRCNS program, and by an NSF Graduate Research Fellowship.}
-
-\startmain
+\begin{authorsummary}
+Author Summary Here, required for Computational Psychiatry
+\end{authorsummary}
 
@@
 Indeed, anxiety disorders are associated with pessimistic beliefs about the future \citep{ClarkBeck2011}.
-Clinically and subclinically anxious individuals judge future threat as more likely than do non-anxious individuals\citep{ButlerMathews1983, ButlerMathews1987, MacleodByrne1996}.
-Importantly, the development and maintenance of clinical anxiety is strongly tied to diminished perceived control\citep{bandura1977, barlow2002, gallagher2014a}, such that anxious individuals are more likely to endorse the belief that they are unlikely or unable to mitigate future threat.
-Indeed, lack of belief in one's ability to successfully navigate future danger is associated with anxiety \citep{davey1996, dugas1997}, and an increased belief in perceived control over threat is correlated with symptom reduction across the family of anxiety disorders\citep{gallagher2014b}.
+Clinically and subclinically anxious individuals judge future threat as more likely than do non-anxious individuals \citep{ButlerMathews1983, ButlerMathews1987, MacleodByrne1996}.
+Importantly, the development and maintenance of clinical anxiety is strongly tied to diminished perceived control \citep{bandura1977, barlow2002, gallagher2014a}, such that anxious individuals are more likely to endorse the belief that they are unlikely or unable to mitigate future threat.
+Indeed, lack of belief in one's ability to successfully navigate future danger is associated with anxiety \citep{davey1996, dugas1997}, and an increased belief in perceived control over threat is correlated with symptom reduction across the family of anxiety disorders \citep{gallagher2014b}.
 
@@
 Specifically, we show through simulation that a model with a misbelief about the reliability of future self-action gives rise to a number of characteristic symptoms and laboratory results concerning anxiety.
-Our perspective also extends previous computational accounts of beliefs about control in mood disorders (e.g.
-\cite{HuysDayan2009}), which neglected the sequential aspects of choice.
+Our perspective also extends previous computational accounts of beliefs about control in mood disorders (e.g., \citealp{HuysDayan2009}), which neglected the sequential aspects of choice.
 
@@
 \begin{equation}\label{eq:bellman_full}
-Q^\pi(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \sum_{a'} \pi(a' \mid s') Q^\pi(s',a').
+Q^\pi(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \sum_{a'} \pi(a' \mid s') Q^\pi(s',a')
 \end{equation}
@@
 \begin{equation}\label{eq:bellman}
- Q^*(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \max_{a'} Q^*(s',a').
+ Q^*(s,a) = r(s,a) + \gamma \sum_{s'} p(s' \mid s,a) \max_{a'} Q^*(s',a')
 \end{equation}
@@
 Note that assumptions of this sort, even if incorrect, can serve adaptive purposes.
-In general, pessimistic assumptions can help to ensure robustness and safety under uncertain or even adversarial scenarios \cite{Garcia2015}.
-Related work in RL shows how computing returns with pessimistic predictions can help to quantify variability in outcomes (i.e., to learn different points in the distribution of possible returns)\cite{bellemare2017}, which is one way to explicitly and tunably take account of risk tolerance.
+In general, pessimistic assumptions can help to ensure robustness and safety under uncertain or even adversarial scenarios \citep{Garcia2015}.
+Related work in RL shows how computing returns with pessimistic predictions can help to quantify variability in outcomes (i.e., to learn different points in the distribution of possible returns; \citealp{bellemare2017}), which is one way to explicitly and tunably take account of risk tolerance.
 Indeed, it is common in machine learning theory to optimize outcomes under worst-case assumptions.
@@
 
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/01_field.png}}%
-  }
-  \par \textbf{Figure 1:} (A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
+\begin{figure}[!b]
+\widefigure{\fullpagewidth}{../../figures/01_field.png}
+\caption{(A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
 (B-D) States colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
@@
 (D) With increasing pessimism ($w=0$), the extent of the spread grows worse, and the return-optimizing trajectory becomes more distorted and avoidant.
-(Parameters: $\gamma = 0.95$)
+(Parameters: $\gamma = 0.95$)}
+\label{fig:field}
 \end{figure}
@@
 In what follows, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of \emph{sequential learning}  and decision making.
-(Because in our model anxiety arises through biased sequential evaluation, we will not address one-step bandit tasks where others have reported learning deficits associated with anxiety \citep{Harle2017, Aylward2019}.) We will also show that our model is consistent with clinical theory describing the transition from clinical anxiety to depression.
-Unless otherwise noted, state and action values under varying degrees of pessimism were solved for using the value iteration dynamic programming method\citep{SuttonBarto2018}.
+(Because in our model anxiety arises through biased sequential evaluation, we will not address one-step bandit tasks where others have reported learning deficits associated with anxiety, e.g., \citealp{Harle2017, Aylward2019}.) We will also show that our model is consistent with clinical theory describing the transition from clinical anxiety to depression.
+Unless otherwise noted, state and action values under varying degrees of pessimism were solved for using the value iteration dynamic programming method \citep{SuttonBarto2018}.
 All simulations were implemented in the Python programming language and the code is publicly available at \url{https://github.com/ndawlab/seqanx}.
@@
 
-\begin{figure}
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/02_appavo.png}}
-  }
-  \par \textbf{Figure 2:} (A) The Balloon Analog Risk Task (BART)\citep{Lejuez2002,Maner2007, ramirez2015}.
+\begin{figure}[!t]
+\widefigure{\fullpagewidth}{../../figures/02_appavo.png}
+\caption{(A) The Balloon Analog Risk Task (BART; \citealp{Lejuez2002}).
 The risk of balloon burst (point loss) increases with each pump and does so earlier for the high risk (red) than low risk (blue) balloons.
@@
 avoid) decreases with loss amount and threat level, and moreso under pessimistic assumptions.
-(Parameters: $\gamma = 1.0$)
+(Parameters: $\gamma = 1.0$)}
+\label{fig:appavo}
 \end{figure}
@@
 Anxious individuals are more likely to forgo potential gains in order to avoid potential danger.
-Many of the disruptions anxiety causes to everyday functioning (e.g.
-avoiding social obligations for fear of possible social humiliation) can be understood in these terms.
+Many of the disruptions anxiety causes to everyday functioning (e.g., avoiding social obligations for fear of possible social humiliation) can be understood in these terms.
 As such, many have sought to probe and measure this behavior in the laboratory.
-For instance, in the balloon analog risk task (BART) \citep{Lejuez2002}, participants attempt to earn money by pumping virtual balloons.
+For instance, in the balloon analog risk task (BART; \citealp{Lejuez2002}), participants attempt to earn money by pumping virtual balloons.
 With each pump, the balloon inflates and money is earned, but so too does the chance that the balloon pops and the accumulated earnings are lost.
@@
 As shown in Fig.~2, our model easily accommodates this result.
-Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier -- similar to empirical findings\citep{Maner2007, ramirez2015}.
+Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier -- similar to empirical findings \citep{Maner2007, ramirez2015}.
 This is because it anticipates and avoids future errors in choice, which would otherwise result in the balloon popping.
-Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety, such as in the predator avoidance task\cite{fung2019}.
+Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety, such as in the predator avoidance task \citep{fung2019}.
 
 A unique prediction of the model (because we assume optimal choice under the assumption of \emph{future} suboptimality) is that bias should arise only when beliefs about \emph{future} avoidance are involved, rather than direct conflict between immediate impulses.
-Recent data \cite{fung2019} using a predator avoidance task, analogous to the BART, support this view.
+Recent data \citep{fung2019} using a predator avoidance task, analogous to the BART, support this view.
 In this task, increasing trait anxiety predicted earlier escape (analogous to cash-out in the BART) for slow predators (for whom future decisions to escape were a relevant consideration) but not for fast ones (who would attack immediately, mooting consideration of future steps).
 
-Relatedly, the model can also capture findings of increased behavioral inhibition, measured as prolonged response times, in anxious individuals under threat\cite{bach2015}.
+Relatedly, the model can also capture findings of increased behavioral inhibition, measured as prolonged response times, in anxious individuals under threat \citep{bach2015}.
 In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator", which are all lost if the predator awakes.
 Though in this variant the risk of predation is constant throughout a trial (rather than increasing, as in the BART) the potential loss from capture still increases with each token collected.
-Bach \cite{bach2015} finds that participants are slower to collect tokens as this potential loss increases, and that this slowing is enhanced by subclinical anxiety.
+\citeauthor{bach2015} finds that participants are slower to collect tokens as this potential loss increases, and that this slowing is enhanced by subclinical anxiety.
 We can capture this effect in our model by noting that the relative value of approach compared to avoidance is reduced as potential loss and the risk that the predator awakes increase (Fig.~2d/e).
@@
 In this case, the model captures behavioral inhibition (slower responses as threat increases), and its enhancement by anxiety, as measured by Bach.
-(Note that we need not assume the coupling of reaction times to action value spread is due to difficulty in decision formation per se, which Bach argues against: it may, for instance, reflect Pavlovian initiation biases \cite{niv2007}.)
+(Note that we need not assume the coupling of reaction times to action value spread is due to difficulty in decision formation per se, which Bach argues against: it may, for instance, reflect Pavlovian initiation biases; \citealp{niv2007}.)
 
@@
 \begin{figure}[!b]
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/04_tree.png}}%
-  }
-  \par \textbf{Figure 3:} (A) The decision tree environment \cite{Huys2012,Lally2017}.
+\widefigure{\fullpagewidth}{../../figures/04_tree.png}
+\caption{(A) The decision tree environment \citep{Huys2012, Lally2017}.
 An optimistic agent ($w=1$) prefers the optimal loss-minimizing policy through the initial large loss (left branch).
@@
 One-step rewards (or costs) are presented in each state; the net value $Q$ on each path is shown numerically.
-(Parameters: $\gamma = 1.0$)
+(Parameters: $\gamma = 1.0$)}
+\label{fig:tree}
 \end{figure}
 
-Another laboratory phenomenon associated with anxiety is ``aversive pruning'' in planning \cite{Huys2012, Lally2017}.
+Another laboratory phenomenon associated with anxiety is ``aversive pruning'' in planning \citep{Huys2012, Lally2017}.
 This refers to the idea that when evaluating future action trajectories in a sequential task like chess, people are resource-limited, cannot evaluate all possible options, and must selectively consider certain paths and neglect others.
@@
 Although the optimal choice in the decision tree from is to weather an initial large loss (e.g., -70) in order to reap the large gain that follows, people tend to disfavor this path suggesting they prune it and consequently neglect the later gain.
-The degree of such pruning correlates, depending on the study, with subclinical depressive \cite{Huys2012} or anxiety \cite{Lally2017} symptoms.
+The degree of such pruning correlates, depending on the study, with subclinical depressive \citep{Huys2012} or anxiety \citep{Lally2017} symptoms.
 
@@
 
+So far, we have considered only the asymptotic preferences implied by our pessimistic value function, which we computed directly through value iteration.
+But we can also consider the process of learning under this value function (e.g., by variants of Q-learning \citep{SuttonBarto2018} or DYNA \citep{sutton1991, russek_predictive_2017} using the $\beta$-pessimistic return).
+The dynamics of such learning may speak to the progression of symptoms.
+
 \begin{figure}[!b]
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/03_lh.png}}%
-  }
-  \par \textbf{Figure 4:} (A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
+\widefigure{\fullpagewidth}{../../figures/03_lh.png}
+\caption{(A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
 (B, c) The development of value expectancies over three steps of learning, for two levels of pessimism.
@@
 foregoes action).
-(Parameters: $\gamma = 0.95$)
+(Parameters: $\gamma = 0.95$)}
+\label{fig:lh}
 \end{figure}
-
-So far, we have considered only the asymptotic preferences implied by our pessimistic value function, which we computed directly through value iteration.
-But we can also consider the process of learning under this value function (e.g., by variants of Q-learning\citep{SuttonBarto2018} or DYNA\citep{sutton1991, russek_predictive_2017} using the $\beta$-pessimistic return).
-The dynamics of such learning may speak to the progression of symptoms.
 
@@
 Frustratingly, these biased estimates of value may still remain even after a misbelief in the efficacy of future action is corrected for in a course of therapy.
-This phenomenon (similar to failures of model-free RL algorithms to adjust to reward revaluation without extensive relearning; \cite{daw2005}) may offer at least a partial answer to a classic puzzle in pathological avoidance, i.e.
-why it is so resistant to extinction \cite{moutoussis2018}, and to the unfortunately high rates of anxiety recurrence following treatment\citep{pittig2018}.
+This phenomenon (similar to failures of model-free RL algorithms to adjust to reward revaluation without extensive relearning; \citealp{daw2005}) may offer at least a partial answer to a classic puzzle in pathological avoidance, i.e.
+why it is so resistant to extinction \citep{moutoussis2018}, and to the unfortunately high rates of anxiety recurrence following treatment \citep{pittig2018}.
 
@@
 This refers to the finding that, all else being equal, people tend to treat choice as itself valuable: i.e., choices which lead to more choice opportunities in the future are preferred to those that lead to fewer future choice opportunities.
-A free choice premium has been observed in multiple behavioral experiments \cite{Leotti2010, ly2019}.
-A variant of a free choice premium paradigm from two previous studies\cite{Leotti2011,Leotti2014} is presented in Fig.
+A free choice premium has been observed in multiple behavioral experiments \citep{Leotti2010, ly2019}.
+A variant of a free choice premium paradigm from two previous studies \citep{Leotti2011,Leotti2014} is presented in Fig.
 5a.
 In the task, participants repeatedly choose between a free choice option, allowing for an additional future choice, and a fixed choice option.
-Importantly, both choices lead to identical, stochastic outcomes (e.g.
-50/50 chance of [1, -1]).
+Importantly, both choices lead to identical, stochastic outcomes (e.g., 50/50 chance of [1, -1]).
 Empirical studies have found that human subjects (from a general, healthy population) prefer the free choice option despite it conferring no additional benefits relative to the fixed choice option.
 
-\begin{figure}[t]
-  \centerline{%
-    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/05_choice.png}}%
-  }
-  \par \textbf{Figure 5:} (A) The free choice premium task\cite{Leotti2011,Leotti2014} with equal chance of outcome $R \in [1, -1]$.
+\begin{figure}[!t]
+\widefigure{\fullpagewidth}{../../figures/05_choice.png}
+\caption{(A) The free choice premium task \citep{Leotti2011, Leotti2014} with equal chance of outcome $R \in [1, -1]$.
 Non-anxious participants exhibit a preference for the free choice option (blue) despite it conferring no benefit over the fixed choice option (grey).
 (B) Pessimistic agents show an attenuated free choice bias.
-The fractional preference for the free choice option over the simulated experiment is shown for three populations of subjects with levels of pessimism $w$ (y-axis; each dot represents an individual simulated agent, and the smoothed density of free choice bias for each pessimism level is also shown.)
-  (Parameters: Q-learning with $\gamma = 1.0$, and inverse temperature, $\beta$, increased from 1 to 15 over 100 episodes.)
+The fractional preference for the free choice option over the simulated experiment is shown for three populations of subjects with levels of pessimism $w$ (y-axis; each dot represents an individual simulated agent, and the smoothed density of free choice bias for each pessimism level is also shown.) (Parameters: Q-learning with $\gamma = 1.0$, and inverse temperature, $\beta$, increased from 1 to 15 over 100 episodes.)}
+\label{fig:choice}
 \end{figure}
 
-On one account\citep{ly2019}, this free choice preference directly and specifically reflects the assumption about sequential choice whose violation we argue is core to anxiety: i.e.
+On one account \citep{ly2019}, this free choice preference directly and specifically reflects the assumption about sequential choice whose violation we argue is core to anxiety: i.e.
 that the agent will make reward-maximizing choices in the future.
@@
 Our proposal makes the novel prediction that, if as we hypothesize anxiety reflects a violation of this optimistic assumption, then anxious individuals will exhibit a diminished or reversed free choice bias, as shown in simulation in Fig.
-~5b.
+5b.
 Future empirical research will be required to test this prediction.
@@
 We are not the first to propose a formal theory of control in psychiatry using MDPs.
-Huys \& Dayan \cite{HuysDayan2009} also provided a computational account of learned helplessness through simple models of one-step action-outcome contingencies.
-Our accounts differ particularly in our exclusive focus on control in the sequential setting, which Huys \& Dayan did not address.
+\cite{HuysDayan2009} also provided a computational account of learned helplessness through simple models of one-step action-outcome contingencies.
+Our accounts differ particularly in our exclusive focus on control in the sequential setting, which \citeauthor{HuysDayan2009} did not address.
 Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future choice and events.
 Similarly, research and modeling by Bishop and colleagues \citep{browning2015, gagne2018} also taking a decision theoretic approach has stressed the importance of uncertainty as a core feature of anxiety.
-Specifically, they have described uncertainty as inherently aversive in anxiety, and have presented models of how uncertainty may be increased in anxiety (e.g.
-aberrant processing of environmental volatility).
+Specifically, they have described uncertainty as inherently aversive in anxiety, and have presented models of how uncertainty may be increased in anxiety (e.g., aberrant processing of environmental volatility).
 The present work is compatible with deficits in processing uncertainty, and might instead be viewed as an attempt to unpack why uncertainty is aversive: because, in our view, it is resolved (i.e.
@@
 For example, self-efficacy theory \citep{bandura1977} and the triple vulnerabilities model \citep{barlow2002} both posit that a reduced belief in the ability to effectively respond to future threat is involved in the genesis and maintenance of clinical anxiety.
-In contrast, and focused less on the self, the learned helplessness theories of anxiety \cite{alloy1990} claims clinical anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
+In contrast, and focused less on the self, the learned helplessness theories of anxiety \citep{alloy1990} claims clinical anxiety results from an uncertain belief in the controllability of the environment, such that future threat cannot be effectively mitigated or avoided.
 As we note above, the present model and analysis (though simulated above in terms of self-efficacy) can accommodate either variant, and shows how they relate to one another as members of a more general family of accounts.
@@
 Traditionally, the goal of reinforcement learning algorithms is to a find reward-maximizing policy with respect to the expectation (average) of returns.
-However, depending on one's risk attitude and uncertainty about the environment (e.g.
-if there is potential for catastrophic loss), it may be instead preferable to learn a policy with respect to an alternative and more pessimistic objective function, similar to the one considered here.
-Accordingly, returns and agent behaviors similar to ours arise in previous research on learning risk-sensitive and robust policies \cite{morimura2012, chow2015, bellemare2017}.
+However, depending on one's risk attitude and uncertainty about the environment (e.g., if there is potential for catastrophic loss), it may be instead preferable to learn a policy with respect to an alternative and more pessimistic objective function, similar to the one considered here.
+Accordingly, returns and agent behaviors similar to ours arise in previous research on learning risk-sensitive and robust policies \citep{morimura2012, chow2015, bellemare2017}.
 
-We have centered our discussion at Marr's \cite{marr1982} computational level: on beliefs and their consequences in terms of action values.
-We have so far remained agnostic as to how, algorithmically or mechanistically, these misbeliefs are implemented in the brain.
-\cite{friston2014computational} Importantly, the brain is believed to contain multiple distinct mechanisms for evaluating actions (e.g, model-based and model-free learning; \cite{daw2005, huys2015}), and pessimistic beliefs might play out either differentially or similarly through each of these mechanisms.
-One promising possibility is that these symptoms mainly reflect aberrations in model-based planning \cite{huys2015}, i.e.
+We have centered our discussion at Marr's \citep{marr1982} computational level: on beliefs and their consequences in terms of action values.
+We have so far remained agnostic as to how, algorithmically or mechanistically, these misbeliefs are implemented in the brain \citep{friston2014computational}.
+Importantly, the brain is believed to contain multiple distinct mechanisms for evaluating actions (e.g, model-based and model-free learning; \citealp{daw2005, huys2015}), and pessimistic beliefs might play out either differentially or similarly through each of these mechanisms.
+One promising possibility is that these symptoms mainly reflect aberrations in model-based planning \citep{huys2015}, i.e.
 explicitly evaluating actions by mentally simulating possible trajectories.
-Recent work suggests this process may be accomplished by mentally ``replaying'' individual potential trajectories \cite{momennejad_offline_2018,mattar2018}.
-In this setting, the biases we suggest would amount to over-contemplating, or over-weighting, certain pessimistic trajectories \cite{hunter2019}.
-Such a bias might be detectable using neuroimaging, as a change in which types of events that tend to be replayed \cite{momennejad_offline_2018, ambrose2016}.
+Recent work suggests this process may be accomplished by mentally ``replaying'' individual potential trajectories \citep{momennejad_offline_2018, mattar2018}.
+In this setting, the biases we suggest would amount to over-contemplating, or over-weighting, certain pessimistic trajectories \citep{hunter2019}.
+Such a bias might be detectable using neuroimaging, as a change in which types of events that tend to be replayed \citep{momennejad_offline_2018, ambrose2016}.
 Such a biased replay process, in turn, may also correspond to worry and rumination.
-Indeed, in line with the present results, chronic worry is associated with reduced perceived control, diminished belief in self-efficacy in response to threat, and exaggerated threat appraisal\cite{Berenbaum2010}.
+Indeed, in line with the present results, chronic worry is associated with reduced perceived control, diminished belief in self-efficacy in response to threat, and exaggerated threat appraisal \citep{Berenbaum2010}.
 This suggests that clinical anxiety may in part result from planning processes gone awry.
@@
 That said, recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD \citep{gallagher2014a, gallagher2014b}.
-Importantly, aspects of other psychiatric disorders that involve future-oriented misbeliefs, worry, and avoidance behaviors (e.g.
-eating disorders\citep{konstantellou2011}) may similarly be well-described by the current account.
+Importantly, aspects of other psychiatric disorders that involve future-oriented misbeliefs, worry, and avoidance behaviors (e.g., eating disorders; \citealp{konstantellou2011}) may similarly be well-described by the current account.
 Indeed, and much more speculatively, the model may also have implications for bipolar disorders.
-The onset of manic symptoms is associated with overoptimistic perceived control \cite{alloy2006}, and our same model and basic reasoning (though now envisioning excessively optimistic rather than excessively pessimistic beliefs) may help to explain how this bias may translate into risk-seeking behavior and dysreguated goal pursuit.
+The onset of manic symptoms is associated with overoptimistic perceived control \citep{alloy2006}, and our same model and basic reasoning (though now envisioning excessively optimistic rather than excessively pessimistic beliefs) may help to explain how this bias may translate into risk-seeking behavior and dysreguated goal pursuit.
 As such, the present results are transdiagnostic and not limited to one particular diagnosis.
@@
 Finally, the relationship between evaluation, planning, and neural replay discussed above suggests potential future work that might help to bring this model into contact with the more memory-related aspects of PTSD.
-For instance, the same replay processes that can be used to evaluate actions can also, in theory, update predictive representations, cognitive maps, or models of the environment \citep{russek_predictive_2017}, such as the successor representation \cite{dayan1993, momennejad_successor_2017}.
+For instance, the same replay processes that can be used to evaluate actions can also, in theory, update predictive representations, cognitive maps, or models of the environment \citep{russek_predictive_2017}, such as the successor representation \citep{dayan1993, momennejad_successor_2017}.
 If so, similar biases in replay could result in not just aberrant avoidance behavior, but also progressive, aberrant remodeling of world models or cognitive maps, an observation which may connect to the rich and complex set of issues on memory involvement in PTSD.
@@
 
-\textbf{Toy MDP / Anxiety-to-Depression Transition} Both the Toy MDP and anxiety-to-depression transition simulations were performed using simple gridworlds.
+\subsection{Toy MDP / Anxiety-to-Depression Transition}
+Both the Toy MDP and anxiety-to-depression transition simulations were performed using simple gridworlds.
 Both environments involved only two non-zero states, one rewarding ($r=10$) and one aversive ($r=-10$).
@@
 
-\textbf{Balloon Analog Risk Task / Predator Avoidance Task} The balloon analog risk task \citep{Lejuez2002} has participants inflate a virtual balloon for points.
+\subsection{Balloon Analog Risk Task / Predator Avoidance Task}
+The balloon analog risk task \citep{Lejuez2002} has participants inflate a virtual balloon for points.
 Earnings rise with each pump, but so too does the risk of the balloon popping and subsequent point loss.
-Unbeknownst to participants, the number of pumps before balloon pop is predefined and drawn randomly from some distribution (e.g.
-uniform, normal, exponential), where the mean controls the risk (i.e.
+Unbeknownst to participants, the number of pumps before balloon pop is predefined and drawn randomly from some distribution (e.g., uniform, normal, exponential), where the mean controls the risk (i.e.
 average number of pumps before point loss).
@@
 Here, we modeled the BART as an undiscounted MDP with 20 states, where transitioning to each successive state yielded $r=1$.
-The only available actions were to transition to the next state (e.g.
-$S_1 \rightarrow S_2, S_2 \rightarrow S_3$) or end the episode (i.e.
+The only available actions were to transition to the next state (e.g., $S_1 \rightarrow S_2, S_2 \rightarrow S_3$) or end the episode (i.e.
 cash-out).
@@
 
-\textbf{Behavioral Inhibition Task} In the behavioral inhibition task \citep{bach2015} (or sleeping predator task), participants collect virtual tokens while evading capture from a virtual predator.
+\subsection{Behavioral Inhibition Task}
+In the behavioral inhibition task \citep{bach2015} (or sleeping predator task), participants collect virtual tokens while evading capture from a virtual predator.
 Unlike the BART, the risk in the behavioral inhibition task (i.e.
@@
 
-\textbf{Aversive Pruning} In the aversive pruning task \citep{Huys2012, Lally2017}, participants learn to navigate a six-state graphworld, where each state is directly connected to only two other states.
+\subsection{Aversive Pruning}
+In the aversive pruning task \citep{Huys2012, Lally2017}, participants learn to navigate a six-state graphworld, where each state is directly connected to only two other states.
 Each state is associated with some reward (cost), and participants must plan trajectories through the state-space so as to maximize reward (minimize cost).
@@
 
-\textbf{Free Choice Task} In the free choice task \citep{Leotti2011}, participants complete a series of two-stage trials.
+\subsection{Free Choice Task}
+In the free choice task \citep{Leotti2011}, participants complete a series of two-stage trials.
 In the first stage, they select between a free choice option, allowing them to make an additional choice in the second stage, and a fixed choice, permitting no choice in the second stage.
@@
 
-\bibliographystyle{naturemag}
-\small{\bibliography{manuscript}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% End of Article
+%%% Each of these elements should be used, with the exception of 
+%%% Appendix which is optional.
+
+%%% \section{Supportive Information}
+% \acknowledgments
+% \authorcontributions 
+% \bibliography{<name of .bib file>}
+% \appendix % optional
+
+\authorcontributions 
+Conceptualization: Samuel Zorowitz, Ida Momennejad, Nathaniel D. Daw; Formal analysis and software: Samuel Zorowitz; Writing and editing: Samuel Zorowitz, Ida Momennejad, Nathaniel D. Daw.
+
+\section{Funding Information}
+This work was funded in part by NIDA 1R01DA038891, part of the CRCNS program, and by an NSF Graduate Research Fellowship.
+
+\acknowledgments
+The authors are grateful to Zeb Kurth-Nelson and Will Dabney for helpful discussion.
+
+\bibliography{CompPsychRefs}
 
 \end{document}
+
+\section{Sample figures}
+
+\begin{figure}[h] 
+\centerline{\includegraphics[width=\textwidth]{Fig1}}
+\caption{(Colour online) \textbf{Task and key features of the
+ data.} \\
+ A) Cumulative handling time (CHT) task. Grey bars denote work
+(depressing a lever), white gaps show leisure. The subject must
+ accumulate work up to a total period of time called the
+\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
+intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
+$2$s each time the price is attained, during which the lever is retracted so it cannot
+work; not shown).
+}
+\label{fig:task_data}
+\end{figure}
+
+\begin{figure}[ht] 
+\widefigure{\fullpagewidth}{Fig1}
+\caption{(Colour online) \textbf{Task and key features of the
+ data.} \\
+A) Cumulative handling time (CHT) task. Grey bars denote work
+(depressing a lever), white gaps show leisure. The subject must
+accumulate work up to a total period of time called the
+\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
+intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
+$2$s each time the price is attained, during which the lever is retracted so it cannot
+work; not shown).
+}
+\label{newfig:task_data}
+\end{figure}
+
+\clearpage

### 2019-12-30 manuscripts/cspy/CompPsychManuscript.tex
apa format

--- 2019-12-29 manuscripts/cspy/CompPsychTemplate.tex
@@
 %% Documentclass:
-\documentclass[CompPsych]{stjour}
+\documentclass[manuscript]{stjour}
 
@@
 Indeed, we argue that fear and avoidance of situations far in the future violates the basic logic of evaluation over sequential trajectories of action.
-This is because avoidance is by nature protective: the ability successfully to avoid danger in the future means an agent need not also do so now.
+This is because avoidance is by nature protective: the ability to successfully to avoid danger in the future means an agent need not also do so now.
 For instance, cars endanger pedestrians but can be reliably avoided by following traffic signals; given that, staying indoors offers little or no additional protection from accidents.
@@
 
-We have centered our discussion at Marr's \citep{marr1982} computational level: on beliefs and their consequences in terms of action values.
+We have centered our discussion at Marr's (\citeyear{marr1982}) computational level: on beliefs and their consequences in terms of action values.
 We have so far remained agnostic as to how, algorithmically or mechanistically, these misbeliefs are implemented in the brain \citep{friston2014computational}.
@@
 \end{document}
-
-\section{Sample figures}
-
-\begin{figure}[h] 
-\centerline{\includegraphics[width=\textwidth]{Fig1}}
-\caption{(Colour online) \textbf{Task and key features of the
- data.} \\
- A) Cumulative handling time (CHT) task. Grey bars denote work
-(depressing a lever), white gaps show leisure. The subject must
- accumulate work up to a total period of time called the
-\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
-intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
-$2$s each time the price is attained, during which the lever is retracted so it cannot
-work; not shown).
-}
-\label{fig:task_data}
-\end{figure}
-
-\begin{figure}[ht] 
-\widefigure{\fullpagewidth}{Fig1}
-\caption{(Colour online) \textbf{Task and key features of the
- data.} \\
-A) Cumulative handling time (CHT) task. Grey bars denote work
-(depressing a lever), white gaps show leisure. The subject must
-accumulate work up to a total period of time called the
-\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
-intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
-$2$s each time the price is attained, during which the lever is retracted so it cannot
-work; not shown).
-}
-\label{newfig:task_data}
-\end{figure}
-
-\clearpage

### 2019-12-30 manuscripts/cspy/CompPsychManuscript.tex
ready for submit

@@
 %% Only needed if you use `manuscript' option
-% \journalname{Computational Psychiatry}
-
-%%%%%%%%%%% Article Set-Up %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Article Type:
-%% Default is Research Article.
-
-%% Or, choose one of these options:
-%% Commentary
-%% Review Article
-%% Modification
-%% Retraction
+\journalname{Computational Psychiatry}
 
@@
 
-\begin{authorsummary}
-Author Summary Here, required for Computational Psychiatry
-\end{authorsummary}
-
 \section{Introduction}
@@
 
+\begin{figure}[!b]
+\widefigure{\fullpagewidth}{../../figures/01_field.png}
+\caption{(A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
+(B-D) States colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
+(B) For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value.
+(C) For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states.
+(D) With increasing pessimism ($w=0$), the extent of the spread grows worse, and the return-optimizing trajectory becomes more distorted and avoidant.
+(Parameters: $\gamma = 0.95$)}
+\label{fig:field}
+\end{figure}
+
 The ``max'' operator in Eq.~\ref{eq:bellman} yields a fundamental asymmetry between approach and avoidance, as illustrated in Fig.~1.
@@
 Here we propose unrealistically pessimistic assumptions as a root cause of many anxious symptoms.
-Such pessimism can be encoded either in the policy, $\pi(a \mid s)$, or transition probabilities , $P(s' \mid s,a)$.
+Such pessimism can be encoded either in the policy, $\pi(a \mid s)$, or transition probabilities, $P(s' \mid s,a)$.
 These, respectively, correspond to misbeliefs about either one's own avoidance actions or the environment's responses to them, a point we return to in discussion.
@@
 
-\begin{figure}[!b]
-\widefigure{\fullpagewidth}{../../figures/01_field.png}
-\caption{(A) A simple deterministic gridworld with two terminal states: one rewarding (blue) and one aversive (red).
-(B-D) States colored by their value under different levels of pessimism, with arrows showing an optimal trajectory.
-(B) For an optimistic agent ($w=1$), all states (other than the harmful state) take on positive value.
-(C) For a pessimistic agent ($w=0.5$), negative value spreads from the source to antecedent states.
-(D) With increasing pessimism ($w=0$), the extent of the spread grows worse, and the return-optimizing trajectory becomes more distorted and avoidant.
-(Parameters: $\gamma = 0.95$)}
-\label{fig:field}
-\end{figure}
-
 This simple simulation reflects a localized violation of the core decision theoretic assumption of future return-optimizing action.
@@
 \subsection{Approach-avoidance conflict}
+
+One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}.
+Anxious individuals are more likely to forgo potential gains in order to avoid potential danger.
+Many of the disruptions anxiety causes to everyday functioning (e.g., avoiding social obligations for fear of possible social humiliation) can be understood in these terms.
+As such, many have sought to probe and measure this behavior in the laboratory.
+For instance, in the balloon analog risk task (BART; \citealp{Lejuez2002}), participants attempt to earn money by pumping virtual balloons.
+With each pump, the balloon inflates and money is earned, but so too does the chance that the balloon pops and the accumulated earnings are lost.
+At any point in a trial, a participant may cash out, banking the money earned and ending the trial.
+Anxiety is correlated with fewer pumps of the balloon and earlier cash-outs in the BART \citep{Maner2007, ramirez2015}.
+
+As shown in Fig.~2, our model easily accommodates this result.
+Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier -- similar to empirical findings \citep{Maner2007, ramirez2015}.
+This is because it anticipates and avoids future errors in choice, which would otherwise result in the balloon popping.
+Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety, such as in the predator avoidance task \citep{fung2019}.
 
@@
 \end{figure}
-
-One behavioral finding characteristic of anxiety disorders is unbalanced processing of approach-avoidance conflict \citep{aupperle2010}.
-Anxious individuals are more likely to forgo potential gains in order to avoid potential danger.
-Many of the disruptions anxiety causes to everyday functioning (e.g., avoiding social obligations for fear of possible social humiliation) can be understood in these terms.
-As such, many have sought to probe and measure this behavior in the laboratory.
-For instance, in the balloon analog risk task (BART; \citealp{Lejuez2002}), participants attempt to earn money by pumping virtual balloons.
-With each pump, the balloon inflates and money is earned, but so too does the chance that the balloon pops and the accumulated earnings are lost.
-At any point in a trial, a participant may cash out, banking the money earned and ending the trial.
-Anxiety is correlated with fewer pumps of the balloon and earlier cash-outs in the BART \citep{Maner2007, ramirez2015}.
-
-As shown in Fig.~2, our model easily accommodates this result.
-Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier -- similar to empirical findings \citep{Maner2007, ramirez2015}.
-This is because it anticipates and avoids future errors in choice, which would otherwise result in the balloon popping.
-Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety, such as in the predator avoidance task \citep{fung2019}.
 
@@
 
+\nocite{*}
 \bibliography{CompPsychRefs}

### 2019-12-30 manuscripts/cspy/CompPsychTemplate.tex
ready for submit

@@
 % \journalname{Computational Psychiatry}
-
-%%%%%%%%%%% Article Set-Up %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Article Type:
-%% Default is Research Article.
-
-%% Or, choose one of these options:
-%% Commentary
-%% Review Article
-%% Modification
-%% Retraction
 
